### 2015-08-14 paper.tex
added tex and sty files

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Draft}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
Your Abstract Text Goes Here.  Just a few facts.
Whet our appetites.

\section{Introduction}

A paragraph of text goes here.  Lots of text.  Plenty of interesting
text. \\

More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\

\section{This is Another Section}

Some embedded literal typset code might 
look like the following :

{\tt \small
\begin{verbatim}
int wrap_fact(ClientData clientData,
              Tcl_Interp *interp,
              int argc, char *argv[]) {
    int result;
    int arg0;
    if (argc != 2) {
        interp->result = "wrong # args";
        return TCL_ERROR;
    }
    arg0 = atoi(argv[1]);
    result = fact(arg0);
    sprintf(interp->result,"%d",result);
    return TCL_OK;
}
\end{verbatim}
}

Now we're going to cite somebody.  Watch for the cite tag.
Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
in the source means a non-breaking space.  This way, your reference will
always be attached to the word that preceded it, instead of going to the
next line.

\section{This Section has SubSections}
\subsection{First SubSection}

Here's a typical figure reference.  The figure is centered at the
top of the column.  It's scaled.  It's explicitly placed.  You'll
have to tweak the numbers to get what you want.\\

% you can also use the wonderful epsfig package...
\begin{figure}[t]
\begin{center}
\begin{picture}(300,150)(0,200)
\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
\end{picture}\\
\end{center}
\caption{Wonderful Flowchart}
\end{figure}

This text came after the figure, so we'll casually refer to Figure 1
as we go on our merry way.

\subsection{New Subsection}

It can get tricky typesetting Tcl and C code in LaTeX because they share
a lot of mystical feelings about certain magic characters.  You
will have to do a lot of escaping to typeset curly braces and percent
signs, for example, like this:
``The {\tt \%module} directive
sets the name of the initialization function.  This is optional, but is
recommended if building a Tcl 7.5 module.
Everything inside the {\tt \%\{, \%\}}
block is copied directly into the output. allowing the inclusion of
header files and additional C code." \\

Sometimes you want to really call attention to a piece of text.  You
can center it in the column like this:
\begin{center}
{\tt \_1008e614\_Vector\_p}
\end{center}
and people will really notice it.\\

\noindent
The noindent at the start of this paragraph makes it clear that it's
a continuation of the preceding text, not a new para in its own right.

Now this is an ingenious way to get a forced space.
{\tt Real~$*$} and {\tt double~$*$} are equivalent.

Now here is another way to call attention to a line of code, but instead
of centering it, we noindent and bold it.\\

\noindent
{\bf \tt size\_t : fread ptr size nobj stream } \\

And here we have made an indented para like a definition tag (dt)
in HTML.  You don't need a surrounding list macro pair.
\begin{itemize}
\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
the number of objects read. 
\end{itemize}
This concludes the definitions tag.

\subsection{How to Build Your Paper}

You have to run {\tt latex} once to prepare your references for
munging.  Then run {\tt bibtex} to build your bibliography metadata.
Then run {\tt latex} twice to ensure all references have been resolved.
If your source file is called {\tt usenixTemplate.tex} and your {\tt
  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
{\tt \small
\begin{verbatim}
latex usenixTemplate
bibtex usenixTemplate
latex usenixTemplate
latex usenixTemplate
\end{verbatim}
}

\subsection{Last SubSection}

Well, it's getting boring isn't it.  This is the last subsection
before we wrap it up.

\section{Acknowledgments}

A polite author always includes acknowledgments.  Thank everyone,
especially those who funded the work.

\section{Availability}

It's great when this section says that MyWonderfulApp is free software, 
available via anonymous FTP from

\begin{center}
{\tt ftp.site.dom/pub/myname/Wonderful}\\
\end{center}

Also, it's even greater when you can write that information is also 
available on the Wonderful homepage at

\begin{center}
{\tt http://www.site.dom/\~{}myname/SWIG}
\end{center}

Now we get serious and fill in those references.  Remember you will
have to run latex twice on the document in order to resolve those
cite tags you met earlier.  This is where they get resolved.
We've preserved some real ones in addition to the template-speak.
After the bibliography you are DONE.

{\footnotesize \bibliographystyle{acm}
\bibliography{../common/bibliography}}

\theendnotes

\end{document}

### 2015-08-16 paper.tex
added a very draft title and abstract

@@
-% TEMPLATE for Usenix papers, specifically to meet requirements of
-%  USENIX '05
-% originally a template for producing IEEE-format articles using LaTeX.
-%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
-% adapted by David Beazley for his excellent SWIG paper in Proceedings,
-%   Tcl 96
-% turned into a smartass generic template by De Clarke, with thanks to
-%   both the above pioneers
-% use at your own risk.  Complaints to /dev/null.
-% make it two column with no page numbering, default is 10 point
+% This is "sig-alternate.tex" V2.0 May 2012
+% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
+%
+% This example file demonstrates the use of the 'sig-alternate.cls'
+% V2.5 LaTeX2e document class file. It is for those submitting
+% articles to ACM Conference Proceedings WHO DO NOT WISH TO
+% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
+% The 'sig-alternate.cls' file will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+%
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V2.5) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) NO page numbers
+%
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'sig-alternate.cls' you have control, however, from within
+% the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the ACM Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% e.g.
+% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
+% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% ================= IF YOU HAVE QUESTIONS =======================
+% Questions regarding the SIGS styles, SIGS policies and
+% procedures, Conferences etc. should be sent to
+% Adrienne Griscti (griscti@acm.org)
+%
+% Technical questions _only_ to
+% Gerald Murray (murray@hq.acm.org)
+% ===============================================================
+%
+% For tracking purposes - this is V2.0 - May 2012
 
-% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
-% the .sty file from the LaTeX source template, so that people can
-% more easily include the .sty file into an existing document.  Also
-% changed to more closely follow the style guidelines as represented
-% by the Word sample file.
+\documentclass{sig-alternate}
 
-% Note that since 2010, USENIX does not require endnotes. If you want
-% foot of page notes, don't include the endnotes package in the 
-% usepackage command, below.
+\usepackage{xspace}
 
-% This version uses the latex2e styles, not the very ancient 2.09 stuff.
-\documentclass[letterpaper,twocolumn,10pt]{article}
-\usepackage{usenix,epsfig,endnotes}
+\newcommand{\psharp}{P\#\xspace}
+\newcommand{\csharp}{C\#\xspace}
+
 \begin{document}
+%
+% --- Author Metadata here ---
+\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
 
-%don't want date printed
-\date{}
+\title{Uncovering Distributed System Bugs\\ during Testing (not Production!) -- draft title}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
 
-%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf Draft}
+\numberofauthors{5} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
+       \affaddr{Institute for Clarity in Documentation}\\
+       \affaddr{1932 Wallamaloo Lane}\\
+       \affaddr{Wallamaloo, New Zealand}\\
+       \email{trovato@corporation.com}
+% 2nd. author
+\alignauthor
+G.K.M. Tobin\titlenote{The secretary disavows
+any knowledge of this author's actions.}\\
+       \affaddr{Institute for Clarity in Documentation}\\
+       \affaddr{P.O. Box 1212}\\
+       \affaddr{Dublin, Ohio 43017-6221}\\
+       \email{webmaster@marysville-ohio.com}
+% 3rd. author
+\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
+one who did all the really hard work.}\\
+       \affaddr{The Th{\o}rv{\"a}ld Group}\\
+       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
+       \affaddr{Hekla, Iceland}\\
+       \email{larst@affiliation.org}
+\and  % use '\and' if you need 'another row' of author names
+% 4th. author
+\alignauthor Lawrence P. Leipuner\\
+       \affaddr{Brookhaven Laboratories}\\
+       \affaddr{Brookhaven National Lab}\\
+       \affaddr{P.O. Box 5000}\\
+       \email{lleipuner@researchlabs.org}
+% 5th. author
+\alignauthor Sean Fogarty\\
+       \affaddr{NASA Ames Research Center}\\
+       \affaddr{Moffett Field}\\
+       \affaddr{California 94035}\\
+       \email{fogartys@amesres.org}
+}
 
-%for single author (just remove % characters)
-\author{
-{\rm Your N.\ Here}\\
-Your Institution
-\and
-{\rm Second Name}\\
-Second Institution
-% copy the following lines to add more authors
-% \and
-% {\rm Name}\\
-%Name Institution
-} % end author
+\date{15 August 2015}
+% Just remember to make sure that the TOTAL number of authors
+% is the number that will appear on the first page PLUS the
+% number that will appear in the \additionalauthors section.
 
 \maketitle
+\begin{abstract}
+Distributed systems are notoriously hard to test.
+This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
+Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
 
-% Use the following at camera-ready time to suppress page numbers.
-% Comment it out when you first submit the paper for review.
-\thispagestyle{empty}
+We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
+Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, to model the environment of a distributed system, and then capture and systematically explore all sources of nondeterminism.
+We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
+Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
+\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
 
-\subsection*{Abstract}
-Your Abstract Text Goes Here.  Just a few facts.
-Whet our appetites.
+\end{abstract}
+
+% A category with the (minimum) three required fields
+\category{H.4}{Information Systems Applications}{Miscellaneous}
+%A category including the fourth, optional field follows...
+\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+
+\terms{Theory}
+
+\keywords{ACM proceedings, \LaTeX, text tagging}
 
 \section{Introduction}
+The \textit{proceedings} are the records of a conference.
+ACM seeks to give these conference by-products a uniform,
+high-quality appearance.  To do this, ACM has some rigid
+requirements for the format of the proceedings documents: there
+is a specified format (balanced  double columns), a specified
+set of fonts (Arial or Helvetica and Times Roman) in
+certain specified sizes (for instance, 9 point for body copy),
+a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
+the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
+and (1.9 cm [.75"]) left and right; specified column width
+(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).
 
-A paragraph of text goes here.  Lots of text.  Plenty of interesting
-text. \\
+The good news is, with only a handful of manual
+settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
+and {\texttt{\char'134 alignauthor}} commands, you have
+already used; another, {\texttt{\char'134 balancecolumns}}, will
+be used in your very last run of \LaTeX\ to ensure
+balanced column heights on the last page.}, the \LaTeX\ document
+class file handles all of this for you.
 
-More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\
+The remainder of this document is concerned with showing, in
+the context of an ``actual'' document, the \LaTeX\ commands
+specifically available for denoting the structure of a
+proceedings paper, rather than with giving rigorous descriptions
+or explanations of such commands.
 
-\section{This is Another Section}
+\section{The {\secit Body} of The Paper}
+Typically, the body of a paper is organized
+into a hierarchical structure, with numbered or unnumbered
+headings for sections, subsections, sub-subsections, and even
+smaller sections.  The command \texttt{{\char'134}section} that
+precedes this paragraph is part of such a
+hierarchy.\footnote{This is the second footnote.  It
+starts a series of three footnotes that add nothing
+informational, but just give an idea of how footnotes work
+and look. It is a wordy one, just so you see
+how a longish one plays out.} \LaTeX\ handles the numbering
+and placement of these headings for you, when you use
+the appropriate heading commands around the titles
+of the headings.  If you want a sub-subsection or
+smaller part to be unnumbered in your output, simply append an
+asterisk to the command name.  Examples of both
+numbered and unnumbered headings will appear throughout the
+balance of this sample document.
 
-Some embedded literal typset code might 
-look like the following :
+Because the entire article is contained in
+the \textbf{document} environment, you can indicate the
+start of a new paragraph with a blank line in your
+input file; that is why this sentence forms a separate paragraph.
 
-{\tt \small
-\begin{verbatim}
-int wrap_fact(ClientData clientData,
-              Tcl_Interp *interp,
-              int argc, char *argv[]) {
-    int result;
-    int arg0;
-    if (argc != 2) {
-        interp->result = "wrong # args";
-        return TCL_ERROR;
-    }
-    arg0 = atoi(argv[1]);
-    result = fact(arg0);
-    sprintf(interp->result,"%d",result);
-    return TCL_OK;
-}
-\end{verbatim}
-}
+\subsection{Type Changes and {\subsecit Special} Characters}
+We have already seen several typeface changes in this sample.  You
+can indicate italicized words or phrases in your text with
+the command \texttt{{\char'134}textit}; emboldening with the
+command \texttt{{\char'134}textbf}
+and typewriter-style (for instance, for computer code) with
+\texttt{{\char'134}texttt}.  But remember, you do not
+have to indicate typestyle changes when such changes are
+part of the \textit{structural} elements of your
+article; for instance, the heading of this subsection will
+be in a sans serif\footnote{A third footnote, here.
+Let's make this a rather short one to
+see how it looks.} typeface, but that is handled by the
+document class file. Take care with the use
+of\footnote{A fourth, and last, footnote.}
+the curly braces in typeface changes; they mark
+the beginning and end of
+the text that is to be in the different typeface.
 
-Now we're going to cite somebody.  Watch for the cite tag.
-Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
-in the source means a non-breaking space.  This way, your reference will
-always be attached to the word that preceded it, instead of going to the
-next line.
+You can use whatever symbols, accented characters, or
+non-English characters you need anywhere in your document;
+you can find a complete list of what is
+available in the \textit{\LaTeX\
+User's Guide}\cite{Lamport:LaTeX}.
 
-\section{This Section has SubSections}
-\subsection{First SubSection}
+\subsection{Math Equations}
+You may want to display math equations in three distinct styles:
+inline, numbered or non-numbered display.  Each of
+the three are discussed in the next sections.
 
-Here's a typical figure reference.  The figure is centered at the
-top of the column.  It's scaled.  It's explicitly placed.  You'll
-have to tweak the numbers to get what you want.\\
+\subsubsection{Inline (In-text) Equations}
+A formula that appears in the running text is called an
+inline or in-text formula.  It is produced by the
+\textbf{math} environment, which can be
+invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
+construction or with the short form \texttt{\$. . .\$}. You
+can use any of the symbols and structures,
+from $\alpha$ to $\omega$, available in
+\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
+few examples of in-text equations in context. Notice how
+this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
+set here in in-line math style, looks slightly different when
+set in display style.  (See next section).
 
-% you can also use the wonderful epsfig package...
-\begin{figure}[t]
-\begin{center}
-\begin{picture}(300,150)(0,200)
-\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
-\end{picture}\\
-\end{center}
-\caption{Wonderful Flowchart}
-\end{figure}
+\subsubsection{Display Equations}
+A numbered display equation -- one set off by vertical space
+from the text and centered horizontally -- is produced
+by the \textbf{equation} environment. An unnumbered display
+equation is produced by the \textbf{displaymath} environment.
 
-This text came after the figure, so we'll casually refer to Figure 1
-as we go on our merry way.
+Again, in either environment, you can use any of the symbols
+and structures available in \LaTeX; this section will just
+give a couple of examples of display equations in context.
+First, consider the equation, shown as an inline equation above:
+\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
+Notice how it is formatted somewhat differently in
+the \textbf{displaymath}
+environment.  Now, we'll enter an unnumbered equation:
+\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
+and follow it with another numbered equation:
+\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
+just to demonstrate \LaTeX's able handling of numbering.
 
-\subsection{New Subsection}
+\subsection{Citations}
+Citations to articles \cite{bowman:reasoning,
+clark:pct, braams:babel, herlihy:methodology},
+conference proceedings \cite{clark:pct} or
+books \cite{salas:calculus, Lamport:LaTeX} listed
+in the Bibliography section of your
+article will occur throughout the text of your article.
+You should use BibTeX to automatically produce this bibliography;
+you simply need to insert one of several citation commands with
+a key of the item cited in the proper location in
+the \texttt{.tex} file \cite{Lamport:LaTeX}.
+The key is a short reference you invent to uniquely
+identify each work; in this sample document, the key is
+the first author's surname and a
+word from the title.  This identifying key is included
+with each item in the \texttt{.bib} file for your article.
 
-It can get tricky typesetting Tcl and C code in LaTeX because they share
-a lot of mystical feelings about certain magic characters.  You
-will have to do a lot of escaping to typeset curly braces and percent
-signs, for example, like this:
-``The {\tt \%module} directive
-sets the name of the initialization function.  This is optional, but is
-recommended if building a Tcl 7.5 module.
-Everything inside the {\tt \%\{, \%\}}
-block is copied directly into the output. allowing the inclusion of
-header files and additional C code." \\
+The details of the construction of the \texttt{.bib} file
+are beyond the scope of this sample document, but more
+information can be found in the \textit{Author's Guide},
+and exhaustive details in the \textit{\LaTeX\ User's
+Guide}\cite{Lamport:LaTeX}.
 
-Sometimes you want to really call attention to a piece of text.  You
-can center it in the column like this:
-\begin{center}
-{\tt \_1008e614\_Vector\_p}
-\end{center}
-and people will really notice it.\\
+This article shows only the plainest form
+of the citation command, using \texttt{{\char'134}cite}.
+This is what is stipulated in the SIGS style specifications.
+No other citation format is endorsed or supported.
 
-\noindent
-The noindent at the start of this paragraph makes it clear that it's
-a continuation of the preceding text, not a new para in its own right.
+\subsection{Tables}
+Because tables cannot be split across pages, the best
+placement for them is typically the top of the page
+nearest their initial cite.  To
+ensure this proper ``floating'' placement of tables, use the
+environment \textbf{table} to enclose the table's contents and
+the table caption.  The contents of the table itself must go
+in the \textbf{tabular} environment, to
+be aligned properly in rows and columns, with the desired
+horizontal and vertical rules.  Again, detailed instructions
+on \textbf{tabular} material
+is found in the \textit{\LaTeX\ User's Guide}.
 
-Now this is an ingenious way to get a forced space.
-{\tt Real~$*$} and {\tt double~$*$} are equivalent.
+Immediately following this sentence is the point at which
+Table 1 is included in the input file; compare the
+placement of the table here with the table in the printed
+dvi output of this document.
 
-Now here is another way to call attention to a line of code, but instead
-of centering it, we noindent and bold it.\\
+\begin{table}
+\centering
+\caption{Frequency of Special Characters}
+\begin{tabular}{|c|c|l|} \hline
+Non-English or Math&Frequency&Comments\\ \hline
+\O & 1 in 1,000& For Swedish names\\ \hline
+$\pi$ & 1 in 5& Common in math\\ \hline
+\$ & 4 in 5 & Used in business\\ \hline
+$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
+\hline\end{tabular}
+\end{table}
 
-\noindent
-{\bf \tt size\_t : fread ptr size nobj stream } \\
+To set a wider table, which takes up the whole width of
+the page's live area, use the environment
+\textbf{table*} to enclose the table's contents and
+the table caption.  As with a single-column table, this wide
+table will ``float" to a location deemed more desirable.
+Immediately following this sentence is the point at which
+Table 2 is included in the input file; again, it is
+instructive to compare the placement of the
+table here with the table in the printed dvi
+output of this document.
 
-And here we have made an indented para like a definition tag (dt)
-in HTML.  You don't need a surrounding list macro pair.
-\begin{itemize}
-\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
-most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
-the number of objects read. 
-\end{itemize}
-This concludes the definitions tag.
+\begin{table*}
+\centering
+\caption{Some Typical Commands}
+\begin{tabular}{|c|c|l|} \hline
+Command&A Number&Comments\\ \hline
+\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
+\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
+\texttt{{\char'134}table}& 300 & For tables\\ \hline
+\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
+\end{table*}
+% end the environment with {table*}, NOTE not {table}!
 
-\subsection{How to Build Your Paper}
+\subsection{Figures}
+Like tables, figures cannot be split across pages; the
+best placement for them
+is typically the top or the bottom of the page nearest
+their initial cite.  To ensure this proper ``floating'' placement
+of figures, use the environment
+\textbf{figure} to enclose the figure and its caption.
 
-You have to run {\tt latex} once to prepare your references for
-munging.  Then run {\tt bibtex} to build your bibliography metadata.
-Then run {\tt latex} twice to ensure all references have been resolved.
-If your source file is called {\tt usenixTemplate.tex} and your {\tt
-  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
-{\tt \small
-\begin{verbatim}
-latex usenixTemplate
-bibtex usenixTemplate
-latex usenixTemplate
-latex usenixTemplate
-\end{verbatim}
-}
+This sample document contains examples of \textbf{.eps}
+and \textbf{.ps} files to be displayable with \LaTeX.  More
+details on each of these is found in the \textit{Author's Guide}.
 
-\subsection{Last SubSection}
+As was the case with tables, you may want a figure
+that spans two columns.  To do this, and still to
+ensure proper ``floating'' placement of tables, use the environment
+\textbf{figure*} to enclose the figure and its caption.
+and don't forget to end the environment with
+{figure*}, not {figure}!
 
-Well, it's getting boring isn't it.  This is the last subsection
-before we wrap it up.
+Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
+used; use
+the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
+commands as appropriate for the different file types.
 
+\subsection{Theorem-like Constructs}
+Other common constructs that may occur in your article are
+the forms for logical constructs like theorems, axioms,
+corollaries and proofs.  There are
+two forms, one produced by the
+command \texttt{{\char'134}newtheorem} and the
+other by the command \texttt{{\char'134}newdef}; perhaps
+the clearest and easiest way to distinguish them is
+to compare the two in the output of this sample document:
+
+This uses the \textbf{theorem} environment, created by
+the\linebreak\texttt{{\char'134}newtheorem} command:
+\newtheorem{theorem}{Theorem}
+\begin{theorem}
+Let $f$ be continuous on $[a,b]$.  If $G$ is
+an antiderivative for $f$ on $[a,b]$, then
+\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
+\end{theorem}
+
+The other uses the \textbf{definition} environment, created
+by the \texttt{{\char'134}newdef} command:
+\newdef{definition}{Definition}
+\begin{definition}
+If $z$ is irrational, then by $e^z$ we mean the
+unique number which has
+logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
+\end{definition}
+
+Two lists of constructs that use one of these
+forms is given in the
+\textit{Author's  Guidelines}.
+
+There is one other similar construct environment, which is
+already set up
+for you; i.e. you must \textit{not} use
+a \texttt{{\char'134}newdef} command to
+create it: the \textbf{proof} environment.  Here
+is a example of its use:
+\begin{proof}
+Suppose on the contrary there exists a real number $L$ such that
+\begin{displaymath}
+\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
+\end{displaymath}
+Then
+\begin{displaymath}
+l=\lim_{x\rightarrow c} f(x)
+= \lim_{x\rightarrow c}
+\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
+= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
+\frac{f(x)}{g(x)} = 0\cdot L = 0,
+\end{displaymath}
+which contradicts our assumption that $l\neq 0$.
+\end{proof}
+
+Complete rules about using these environments and using the
+two different creation commands are in the
+\textit{Author's Guide}; please consult it for more
+detailed instructions.  If you need to use another construct,
+not listed therein, which you want to have the same
+formatting as the Theorem
+or the Definition\cite{salas:calculus} shown above,
+use the \texttt{{\char'134}newtheorem} or the
+\texttt{{\char'134}newdef} command,
+respectively, to create it.
+
+\subsection*{A {\secit Caveat} for the \TeX\ Expert}
+Because you have just been given permission to
+use the \texttt{{\char'134}newdef} command to create a
+new form, you might think you can
+use \TeX's \texttt{{\char'134}def} to create a
+new command: \textit{Please refrain from doing this!}
+Remember that your \LaTeX\ source code is primarily intended
+to create camera-ready copy, but may be converted
+to other forms -- e.g. HTML. If you inadvertently omit
+some or all of the \texttt{{\char'134}def}s recompilation will
+be, to say the least, problematic.
+
+\section{Conclusions}
+This paragraph will end the body of this sample document.
+Remember that you might still have Acknowledgments or
+Appendices; brief samples of these
+follow.  There is still the Bibliography to deal with; and
+we will make a disclaimer about that here: with the exception
+of the reference to the \LaTeX\ book, the citations in
+this paper are to articles which have nothing to
+do with the present subject and are used as
+examples only.
+%\end{document}  % This is where a 'short' article might terminate
+
+%ACKNOWLEDGMENTS are optional
 \section{Acknowledgments}
+This section is optional; it is a location for you
+to acknowledge grants, funding, editing assistance and
+what have you.  In the present case, for example, the
+authors would like to thank Gerald Murray of ACM for
+his help in codifying this \textit{Author's Guide}
+and the \textbf{.cls} and \textbf{.tex} files that it describes.
 
-A polite author always includes acknowledgments.  Thank everyone,
-especially those who funded the work.
-
-\section{Availability}
-
-It's great when this section says that MyWonderfulApp is free software, 
-available via anonymous FTP from
-
-\begin{center}
-{\tt ftp.site.dom/pub/myname/Wonderful}\\
-\end{center}
-
-Also, it's even greater when you can write that information is also 
-available on the Wonderful homepage at
-
-\begin{center}
-{\tt http://www.site.dom/\~{}myname/SWIG}
-\end{center}
-
-Now we get serious and fill in those references.  Remember you will
-have to run latex twice on the document in order to resolve those
-cite tags you met earlier.  This is where they get resolved.
-We've preserved some real ones in addition to the template-speak.
-After the bibliography you are DONE.
-
-{\footnotesize \bibliographystyle{acm}
-\bibliography{../common/bibliography}}
-
-\theendnotes
-
+%
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
+%
+% ACM needs 'a single self-contained file'!
+%
+%APPENDICES are optional
+%\balancecolumns
+\appendix
+%Appendix A
+\section{Headings in Appendices}
+The rules about hierarchical headings discussed above for
+the body of the article are different in the appendices.
+In the \textbf{appendix} environment, the command
+\textbf{section} is used to
+indicate the start of each Appendix, with alphabetic order
+designation (i.e. the first is A, the second B, etc.) and
+a title (if you include one).  So, if you need
+hierarchical structure
+\textit{within} an Appendix, start with \textbf{subsection} as the
+highest level. Here is an outline of the body of this
+document in Appendix-appropriate form:
+\subsection{Introduction}
+\subsection{The Body of the Paper}
+\subsubsection{Type Changes and  Special Characters}
+\subsubsection{Math Equations}
+\paragraph{Inline (In-text) Equations}
+\paragraph{Display Equations}
+\subsubsection{Citations}
+\subsubsection{Tables}
+\subsubsection{Figures}
+\subsubsection{Theorem-like Constructs}
+\subsubsection*{A Caveat for the \TeX\ Expert}
+\subsection{Conclusions}
+\subsection{Acknowledgments}
+\subsection{Additional Authors}
+This section is inserted by \LaTeX; you do not insert it.
+You just add the names and information in the
+\texttt{{\char'134}additionalauthors} command at the start
+of the document.
+\subsection{References}
+Generated by bibtex from your ~.bib file.  Run latex,
+then bibtex, then latex twice (to resolve references)
+to create the ~.bbl file.  Insert that ~.bbl file into
+the .tex source file and comment out
+the command \texttt{{\char'134}thebibliography}.
+% This next section command marks the start of
+% Appendix B, and does not continue the present hierarchy
+\section{More Help for the Hardy}
+The sig-alternate.cls file itself is chock-full of succinct
+and helpful comments.  If you consider yourself a moderately
+experienced to expert user of \LaTeX, you may find reading
+it useful but please remember not to change it.
+%\balancecolumns % GM June 2007
+% That's all folks!
 \end{document}

### 2015-08-16 paper.tex
author placeholders
placeholder sections

--- 2015-08-14 paper.tex
@@
-% TEMPLATE for Usenix papers, specifically to meet requirements of
-%  USENIX '05
-% originally a template for producing IEEE-format articles using LaTeX.
-%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
-% adapted by David Beazley for his excellent SWIG paper in Proceedings,
-%   Tcl 96
-% turned into a smartass generic template by De Clarke, with thanks to
-%   both the above pioneers
-% use at your own risk.  Complaints to /dev/null.
-% make it two column with no page numbering, default is 10 point
+% This is "sig-alternate.tex" V2.0 May 2012
+% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
+%
+% This example file demonstrates the use of the 'sig-alternate.cls'
+% V2.5 LaTeX2e document class file. It is for those submitting
+% articles to ACM Conference Proceedings WHO DO NOT WISH TO
+% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
+% The 'sig-alternate.cls' file will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+%
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V2.5) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) NO page numbers
+%
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'sig-alternate.cls' you have control, however, from within
+% the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the ACM Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% e.g.
+% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
+% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% ================= IF YOU HAVE QUESTIONS =======================
+% Questions regarding the SIGS styles, SIGS policies and
+% procedures, Conferences etc. should be sent to
+% Adrienne Griscti (griscti@acm.org)
+%
+% Technical questions _only_ to
+% Gerald Murray (murray@hq.acm.org)
+% ===============================================================
+%
+% For tracking purposes - this is V2.0 - May 2012
 
-% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
-% the .sty file from the LaTeX source template, so that people can
-% more easily include the .sty file into an existing document.  Also
-% changed to more closely follow the style guidelines as represented
-% by the Word sample file.
+\documentclass{sig-alternate}
 
-% Note that since 2010, USENIX does not require endnotes. If you want
-% foot of page notes, don't include the endnotes package in the 
-% usepackage command, below.
+\usepackage{xspace}
 
-% This version uses the latex2e styles, not the very ancient 2.09 stuff.
-\documentclass[letterpaper,twocolumn,10pt]{article}
-\usepackage{usenix,epsfig,endnotes}
+\newcommand{\psharp}{P\#\xspace}
+\newcommand{\csharp}{C\#\xspace}
+
 \begin{document}
+%
+% --- Author Metadata here ---
+%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
 
-%don't want date printed
-\date{}
+\title{Uncovering Distributed System Bugs\\ during Testing (not Production!) -- draft title}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
 
-%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf Draft}
+\numberofauthors{5} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor Draft\\
+       \affaddr{Draft}\\
+       \email{Draft}
+% 2nd. author
+\alignauthor Draft\\
+       \affaddr{Draft}\\
+       \email{Draft}
+% 3rd. author
+\alignauthor Draft\\
+       \affaddr{Draft}\\
+       \email{Draft}
+\and  % use '\and' if you need 'another row' of author names
+% 4th. author
+\alignauthor Draft\\
+       \affaddr{Draft}\\
+       \email{Draft}
+% 5th. author
+\alignauthor Draft\\
+       \affaddr{Draft}\\
+       \email{Draft}
+}
 
-%for single author (just remove % characters)
-\author{
-{\rm Your N.\ Here}\\
-Your Institution
-\and
-{\rm Second Name}\\
-Second Institution
-% copy the following lines to add more authors
-% \and
-% {\rm Name}\\
-%Name Institution
-} % end author
+\date{15 August 2015}
+% Just remember to make sure that the TOTAL number of authors
+% is the number that will appear on the first page PLUS the
+% number that will appear in the \additionalauthors section.
 
 \maketitle
+\begin{abstract}
+Distributed systems are notoriously hard to test.
+This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
+Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
 
-% Use the following at camera-ready time to suppress page numbers.
-% Comment it out when you first submit the paper for review.
-\thispagestyle{empty}
+We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
+Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, to model the environment of a distributed system, and then capture and systematically explore all sources of nondeterminism.
+We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
+Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
+\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
 
-\subsection*{Abstract}
-Your Abstract Text Goes Here.  Just a few facts.
-Whet our appetites.
+\end{abstract}
+
+% A category with the (minimum) three required fields
+\category{D.2.8}{Software Engineering}{Metrics}
+
+\keywords{distributed systems, testing}
 
 \section{Introduction}
+\label{sec:intro}
 
-A paragraph of text goes here.  Lots of text.  Plenty of interesting
-text. \\
+Draft
 
-More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\
+\section{Background}
+\label{sec:bg}
 
-\section{This is Another Section}
+Draft
 
-Some embedded literal typset code might 
-look like the following :
+\section{Evaluation}
+\label{sec:eval}
 
-{\tt \small
-\begin{verbatim}
-int wrap_fact(ClientData clientData,
-              Tcl_Interp *interp,
-              int argc, char *argv[]) {
-    int result;
-    int arg0;
-    if (argc != 2) {
-        interp->result = "wrong # args";
-        return TCL_ERROR;
-    }
-    arg0 = atoi(argv[1]);
-    result = fact(arg0);
-    sprintf(interp->result,"%d",result);
-    return TCL_OK;
-}
-\end{verbatim}
-}
+Draft
 
-Now we're going to cite somebody.  Watch for the cite tag.
-Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
-in the source means a non-breaking space.  This way, your reference will
-always be attached to the word that preceded it, instead of going to the
-next line.
+\section{Case studies}
+\label{sec:cases}
 
-\section{This Section has SubSections}
-\subsection{First SubSection}
+Draft
 
-Here's a typical figure reference.  The figure is centered at the
-top of the column.  It's scaled.  It's explicitly placed.  You'll
-have to tweak the numbers to get what you want.\\
+\section{Related Work}
+\label{sec:rw}
 
-% you can also use the wonderful epsfig package...
-\begin{figure}[t]
-\begin{center}
-\begin{picture}(300,150)(0,200)
-\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
-\end{picture}\\
-\end{center}
-\caption{Wonderful Flowchart}
-\end{figure}
+Draft
 
-This text came after the figure, so we'll casually refer to Figure 1
-as we go on our merry way.
+\section{Conclusions}
+\label{sec:concl}
 
-\subsection{New Subsection}
+Draft.
 
-It can get tricky typesetting Tcl and C code in LaTeX because they share
-a lot of mystical feelings about certain magic characters.  You
-will have to do a lot of escaping to typeset curly braces and percent
-signs, for example, like this:
-``The {\tt \%module} directive
-sets the name of the initialization function.  This is optional, but is
-recommended if building a Tcl 7.5 module.
-Everything inside the {\tt \%\{, \%\}}
-block is copied directly into the output. allowing the inclusion of
-header files and additional C code." \\
+%\section{Acknowledgments}
+%Draft.
 
-Sometimes you want to really call attention to a piece of text.  You
-can center it in the column like this:
-\begin{center}
-{\tt \_1008e614\_Vector\_p}
-\end{center}
-and people will really notice it.\\
-
-\noindent
-The noindent at the start of this paragraph makes it clear that it's
-a continuation of the preceding text, not a new para in its own right.
-
-Now this is an ingenious way to get a forced space.
-{\tt Real~$*$} and {\tt double~$*$} are equivalent.
-
-Now here is another way to call attention to a line of code, but instead
-of centering it, we noindent and bold it.\\
-
-\noindent
-{\bf \tt size\_t : fread ptr size nobj stream } \\
-
-And here we have made an indented para like a definition tag (dt)
-in HTML.  You don't need a surrounding list macro pair.
-\begin{itemize}
-\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
-most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
-the number of objects read. 
-\end{itemize}
-This concludes the definitions tag.
-
-\subsection{How to Build Your Paper}
-
-You have to run {\tt latex} once to prepare your references for
-munging.  Then run {\tt bibtex} to build your bibliography metadata.
-Then run {\tt latex} twice to ensure all references have been resolved.
-If your source file is called {\tt usenixTemplate.tex} and your {\tt
-  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
-{\tt \small
-\begin{verbatim}
-latex usenixTemplate
-bibtex usenixTemplate
-latex usenixTemplate
-latex usenixTemplate
-\end{verbatim}
-}
-
-\subsection{Last SubSection}
-
-Well, it's getting boring isn't it.  This is the last subsection
-before we wrap it up.
-
-\section{Acknowledgments}
-
-A polite author always includes acknowledgments.  Thank everyone,
-especially those who funded the work.
-
-\section{Availability}
-
-It's great when this section says that MyWonderfulApp is free software, 
-available via anonymous FTP from
-
-\begin{center}
-{\tt ftp.site.dom/pub/myname/Wonderful}\\
-\end{center}
-
-Also, it's even greater when you can write that information is also 
-available on the Wonderful homepage at
-
-\begin{center}
-{\tt http://www.site.dom/\~{}myname/SWIG}
-\end{center}
-
-Now we get serious and fill in those references.  Remember you will
-have to run latex twice on the document in order to resolve those
-cite tags you met earlier.  This is where they get resolved.
-We've preserved some real ones in addition to the template-speak.
-After the bibliography you are DONE.
-
-{\footnotesize \bibliographystyle{acm}
-\bibliography{../common/bibliography}}
-
-\theendnotes
+%
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
+%
+% ACM needs 'a single self-contained file'!
+%
+%APPENDICES are optional
+%\balancecolumns
 

### 2015-08-16 paper.tex
update in categories
some draft text

@@
 % --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
 %\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
-Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, to model the environment of a distributed system, and then capture and systematically explore all sources of nondeterminism.
+Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
 We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
@@
 
-% A category with the (minimum) three required fields
-\category{D.2.8}{Software Engineering}{Metrics}
+\category{D.2.4}{Software Engineering}{Software/Program Verification}
+\category{D.2.5}{Software Engineering}{Testing and Debugging}
 
@@
 
-Draft
+Distributed systems are notoriously hard to test.
+This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
+Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
+
+We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
+Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
+We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
+Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
+\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
 
@@
 \label{sec:bg}
+
+Easy to think you can do better until it happens to you :-)
+
+\section{Case studies}
+\label{sec:cases}
 
@@
 \label{sec:eval}
-
-Draft
-
-\section{Case studies}
-\label{sec:cases}
 

### 2015-08-16 background.tex
added background text on P#

Easy to think you can do better until it happens to you :-)

\subsection{Programming and testing with \psharp}
\label{bg:psharp}

The \psharp~\cite{deligiannis2015psharp} project provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} framework for developing highly-reliable distributed systems.

The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines. \psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems. This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g. Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).

A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods. Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler. This handler might update a field, create a new machine, or send an event to another machine. In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.

Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code. Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.

The second core component of the \psharp project is a concurrency unit testing framework that is embedded in the \psharp runtime. This framework allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.

\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.

### 2015-08-16 intro.tex
added background text on P#

--- 2015-08-16 paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-\documentclass{sig-alternate}
-
-\usepackage{xspace}
-
-\newcommand{\psharp}{P\#\xspace}
-\newcommand{\csharp}{C\#\xspace}
-
-\begin{document}
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
-\title{Uncovering Distributed System Bugs\\ during Testing (not Production!) -- draft title}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{5} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-% 2nd. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-% 3rd. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-% 5th. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-}
-
-\date{15 August 2015}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
 Distributed systems are notoriously hard to test.
@@
 We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
-Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, to model the environment of a distributed system, and then capture and systematically explore all sources of nondeterminism.
+Our approach involves the use of \psharp~\cite{deligiannis2015psharp}, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
 We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
@@
 \psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
-
-\end{abstract}
-
-% A category with the (minimum) three required fields
-\category{D.2.8}{Software Engineering}{Metrics}
-
-\keywords{distributed systems, testing}
-
-\section{Introduction}
-\label{sec:intro}
-
-Draft
-
-\section{Background}
-\label{sec:bg}
-
-Draft
-
-\section{Evaluation}
-\label{sec:eval}
-
-Draft
-
-\section{Case studies}
-\label{sec:cases}
-
-Draft
-
-\section{Related Work}
-\label{sec:rw}
-
-Draft
-
-\section{Conclusions}
-\label{sec:concl}
-
-Draft.
-
-%\section{Acknowledgments}
-%Draft.
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
-
-\end{document}

### 2015-08-16 background.tex
added a reference to AWS using TLA+
added some of the quotes from Cheng regarding top problems

@@
+\subsection{Finding bugs in distributed systems}
+\label{bg:bugs}
+
+Draft -- We asked engineers in the Microsoft Azure team to tell us their number one problem when developing distributed systems.
+We got the following answers: ``Due to limited testing coverage, many correctness problems are only exposed in production through live-sites''; ``Engineering overhead extremely high to identify problems''; ``Practical tools that can improve testing coverage highly appreciated!''.
+
 Easy to think you can do better until it happens to you :-)
@@
 \psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+
+%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

### 2015-08-16 paper.tex
added background text on P#
update in abstract

@@
 
+\usepackage{url}
 \usepackage{xspace}
@@
 This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
-Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
+Stress testing techniques, commonly used in industry today, are unable to capture and control all these sources of nondeterminism.
+This results in many tricky bugs be missed during testing and only get exposed after a system has been put in production.
 
@@
 
-Distributed systems are notoriously hard to test.
-This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
-Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
-
-We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
-Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
-We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
-\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
+\input{intro}
 
@@
 
-Easy to think you can do better until it happens to you :-)
+\input{background}
 
@@
 
-Draft
+\input{related}
 
-\section{Conclusions}
+\section{Conclusion}
 \label{sec:concl}
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{references}  % sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file

### 2015-08-16 paper.tex
draft text for Cheng's case study

@@
 
-We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
+We present a new methodology for programming, modeling and testing distributed systems, and uncovering bugs before these systems are released in the wild.
 Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
@@
 
-Draft
+\subsection{Distributed Extent Management System}
+
+We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system.
+This component is responsible for managing the partitioned extent metadata and works as follows.
+
+There are multiple \emph{extent managers} (EMgrs) that are in charge of managing a subset of the extents.
+Each EMgr communicates asynchronously (via remote procedure calls) with a number of \emph{extent nodes} (ENs) that store the extents.
+Each EN sends: (i) a \emph{heartbeat} every 5 seconds, to notify the EMgr that it is still available; and (ii) a \emph{sync report} every 5 minutes, to synchronize its extent with the rest of the nodes.
+The EMgr contains two data structures: an \emph{extent center} (ECtr), which is updated every time an EN syncs; and an EN map, which is updated every time it receives a heartbeat from an EN.
+The EN map runs an EN expiration loop that is responsible for removing ENs that have expired from the EN map and also delete them from the ECtr.
+Finally, the EMgr runs an extent repair loop, which examines all contents of all extents in the ECtr and schedules repairs of extents if they are out-of-date (via a repair request).
 

### 2015-08-16 icse_style/template.tex
added icse latex style

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2015-08-17 paper.tex
added author names in the paper (its currently in alphabetic order and kinda broken, we should discuss and fix)
Initial description of what live migration does.

@@
 
-\numberofauthors{5} %  in this sample file, there are a *total*
+\numberofauthors{7} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 % 1st. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
+\alignauthor Nikolaj Bjorner\\
+       \affaddr{Microsoft Research}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{nbjorner@microsoft.com}
 % 2nd. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-% 3rd. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
+\alignauthor Pantazis Deligiannis\\
+       \affaddr{Imperial College London}\\
+       \affaddr{London, UK}\\
+       \email{p.deligiannis@imperial.ac.uk}
+% 3th. author
+\alignauthor John Erickson\\
+       \affaddr{Microsoft}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{jerick@microsoft.com}
+% 4th. author
+\alignauthor Cheng Huang\\
+       \affaddr{Microsoft Research}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{cheng.huang@microsoft.com}
 \and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
-% 5th. author
-\alignauthor Draft\\
-       \affaddr{Draft}\\
-       \email{Draft}
+% 5rd. author
+\alignauthor Matt McCutchen\\
+       \affaddr{MIT}\\
+       \affaddr{Cambridge, MA, USA}\\
+       \email{rmccutch@mit.edu}
+% 6th. author
+\alignauthor Wolfram Schulte\\
+       \affaddr{Microsoft}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{schulte@microsoft.com}
+% 7th. author
+\alignauthor Shaz Qadeer\\
+       \affaddr{Microsoft Research}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{qadeer@microsoft.com}
 }
@@
 
+\subsection{Live Azure Table Migration}
+
+We also used \psharp to test the MigratingTable library, which can be integrated into an application that accesses data using the Windows Azure table storage service and is capable of migrating the data from one underlying Azure table to another, transparent to the application.
+MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table.
+Each read and write on the virtual table is translated to a sequence of reads and writes on the underlying tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the underlying tables respect their own linearizability guarantees.
+
+% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
+The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because of the limit on traffic supported by a single Azure storage account.
+As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
+During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration.
+MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
+
+Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
+
+\subsubsection{Modeling approach}
+% TBD where this goes ends up in the paper ~ Matt 2015-08-17
+
+TODO
+
 \section{Evaluation}

### 2015-08-17 paper.tex
update

@@
 
+\section{Overview}
+\label{sec:overview}
+
+How cheng was testing his code before and how after -- high level, pretty picture
+
+%\input{overview}
+
 \section{Background}
@@
 
+Dependency injection pattern
+
+\section{Methodology}
+\label{sec:methodology}
+
+%\input{methodology}
+
+Very simple running example (maybe some master and some worker nodes) and describe how its environment is modeled and its tested.
+
 \section{Case studies}
@@
 
+% cheng to review
 \subsection{Distributed Extent Management System}
@@
 
-Draft
+Paragraph that says what will be discussed in evaluation and what are the results (very short)
+
+\subsection{Experimental Setup}
+
+\subsection{Benchmarks}
+
+Benchmarks that are not production code (e.g. multipaxos) that can be used to evaluate testing
+
+Cheng's case study with the actual bug
+
+Matt's case study with injected bugs
+
+\subsection{Stressing testing vs Systematic testing}
+
+Comparison of traditional stress testing VS what we do with \psharp
+
+How long you had to run the test to find the bug (if you can find it) and how long are the traces
+
+How easy it is to pinpoint the error using the \psharp traces, comparing to traditional stress testing
+
+Effort required to use the system
 

### 2015-08-18 paper.tex
update

@@
 
-\title{Uncovering Distributed System Bugs\\ during Testing (not Production!) -- draft title}
+\title{Uncovering Distributed System Bugs\\ during Testing (not Production!)}
 %
@@
 % 1st. author
-\alignauthor Nikolaj Bjorner\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{nbjorner@microsoft.com}
-% 2nd. author
 \alignauthor Pantazis Deligiannis\\
@@
        \email{p.deligiannis@imperial.ac.uk}
-% 3th. author
+% 2nd. author
 \alignauthor John Erickson\\
@@
        \email{jerick@microsoft.com}
-% 4th. author
+% 3rd. author
 \alignauthor Cheng Huang\\
@@
 \and  % use '\and' if you need 'another row' of author names
-% 5rd. author
+% 4th. author
 \alignauthor Matt McCutchen\\
@@
        \email{rmccutch@mit.edu}
+% 5th. author
+\alignauthor Shaz Qadeer\\
+       \affaddr{Microsoft Research}\\
+       \affaddr{Redmond, WA, USA}\\
+       \email{qadeer@microsoft.com}
 % 6th. author
@@
        \email{schulte@microsoft.com}
-% 7th. author
-\alignauthor Shaz Qadeer\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{qadeer@microsoft.com}
 }
@@
 \begin{abstract}
-Distributed systems are notoriously hard to test.
-This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
-Stress testing techniques, commonly used in industry today, are unable to capture and control all these sources of nondeterminism.
-This results in many tricky bugs be missed during testing and only get exposed after a system has been put in production.
+Testing distributed systems is very challenging due to many sources of nondeterminism, such as races in the asynchronous communication between components of a system, and unexpected failures.
+Stress testing techniques, commonly used in industry today, are unable to capture and control all these sources of nondeterminism, which results in many bugs being missed during testing and only getting exposed in production.
 
-We present a new methodology for programming, modeling and testing distributed systems, and uncovering bugs before these systems are released in the wild.
-Our approach involves the use of \psharp, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
-We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
-\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
+We present a new methodology for systematically testing unmodified distributed systems.
+Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism.
+We present two case studies of using \psharp to test production distributed systems inside Microsoft.
+Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it.
+\psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
 

### 2015-08-18 paper.tex
High-level modeling approach for migration.  I haven't added the figure or explained async yet.

--- 2015-08-17 paper.tex
@@
 
-We also used \psharp to test the MigratingTable library, which can be integrated into an application that accesses data using the Windows Azure table storage service and is capable of migrating the data from one underlying Azure table to another, transparent to the application.
-MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table.
-Each read and write on the virtual table is translated to a sequence of reads and writes on the underlying tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the underlying tables respect their own linearizability guarantees.
+We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing it.
+MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
+It moves all data from the old table to the new table in the background.
+Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
 
@@
 \subsubsection{Modeling approach}
-% TBD where this goes ends up in the paper ~ Matt 2015-08-17
+% TBD where this ends up in the paper ~ Matt 2015-08-17
 
-TODO
+MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
+MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
+Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances backed by the same tables, the behavior is compliant with the specification of IChainTable2 for the combined input history.
+
+% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
+Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.
+Given a multi-page read, SpecTable can actually produce a list of all valid results.
+Our correctness property is then:
+\begin{quote}
+For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+\end{quote}
+%
+\def\term#1{\emph{#1}}
+We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
+Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
+This makes it possible to verify the correctness property as the system executes.
+We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
+Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
+When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.
+When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.
+\psharp controls the interleaving of the backend calls.
+After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
+We also use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp scheduling strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
+
+To do...
 

### 2015-08-18 intro.tex
added a reference to AWS using TLA+
updates in intro

@@
 Distributed systems are notoriously hard to test.
-This is due to many well-known sources of nondeterminism: races in the asynchronous interaction between system components; the use of multithreaded code inside individual components; unexpected node failures; message losses; and interaction with a complex external environment.
-Stress testing techniques that are commonly used in industry are unable to effectively capture and control all these sources of nondeterminism, which causes many tricky bugs to be missed during testing and only get exposed after a system has been put in production.
+This is due to many well-known sources of \emph{nondeterminism}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and message losses, and interaction with clients and external services.
+%
+All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
+A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases.
 
-We present a new methodology for testing distributed systems, and uncovering bugs before these systems are released in the wild.
-Our approach involves the use of \psharp~\cite{deligiannis2015psharp}, an asynchronous programming language and concurrency unit testing framework, which provides capabilities for modeling the environment of a distributed system, and then captures and systematically explores all sources of nondeterminism.
-We present two case studies of using \psharp to test production distributed systems inside Microsoft: a distributed storage management system for Windows Azure, and a live migration protocol.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it and nail down the culprit.
-\psharp uncovered the problem in a very small setting, which made it easy to examine traces and identify the problem.
+Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, stress testing, fault injection, and chaos testing) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky backs being missed during testing and only getting exposed after a system has been put in production.
+Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.
+
+We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the unified response was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
+
+Amazon recently published an article~\cite{newcombe2015aws} that describes their use of TLA+~\cite{lamport1994temporal} to detect distributed system bugs and prevent them from reaching production.
+TLA+ is a powerful specification language for verifying distributed protocols, but it is unable to verify the code that is actually being executed.
+The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system.
+Although many design bugs can be caught with this approach, there is no guarantee that the real distributed system will be free of bugs.
+
+In this work, our goal is to \emph{test what is being executed}.
+We present a new methodology for testing legacy distributed systems and uncovering bugs before these systems are released in the wild.
+We achieve this using \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides two key capabilities: (i) a flexible way of modeling the environment using simple language features; and (ii) a systematic concurrency testing framework that is able to capture and take control of all the nondeterminism in a real system (together with its modeled environment) and systematically explore execution paths to discover bugs.
+
+We present two case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol.
+Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time as they did not have an effective way to reproduce the bug and nail down the culprit.
+\psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.

### 2015-08-18 paper.tex
updates in intro

--- 2015-08-18 paper.tex
@@
 \begin{abstract}
-Testing distributed systems is very challenging due to many sources of nondeterminism, such as races in the asynchronous communication between components of a system, and unexpected failures.
-Stress testing techniques, commonly used in industry today, are unable to capture and control all these sources of nondeterminism, which results in many bugs being missed during testing and only getting exposed in production.
+Testing distributed systems is very challenging due to multiple sources of nondeterminism, such as arbitrary interleavings between event handlers and unexpected node failures.
+Stress testing, commonly used in industry today, is unable to deal with this kind of nondeterminism, which results in the most tricky bugs being missed during testing and only getting exposed in production.
 

### 2015-08-18 intro.tex
added some draft contributions

@@
 The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system.
-Although many design bugs can be caught with this approach, there is no guarantee that the real distributed system will be free of bugs.
+Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.
 
@@
 \psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.
+
+To summarize, our contributions are as follows:
+
+\begin{itemize}
+\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms.
+\item Our infrastructure can test production code written in \csharp, which is a mainstream language.
+\item We present two case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
+\item ...
+\end{itemize}

### 2015-08-18 intro.tex
some updates

@@
 Distributed systems are notoriously hard to test.
-This is due to many well-known sources of \emph{nondeterminism}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and message losses, and interaction with clients and external services.
+This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with clients and external services.
 %
@@
 
-Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, stress testing, fault injection, and chaos testing) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky backs being missed during testing and only getting exposed after a system has been put in production.
+Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
 Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.

### 2015-08-18 overview.tex
added missing bibliography and overview

Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages (in this paper we will refer them as events) through a network layer. When a component receives an event, it responds by executing an predefined event handler.

Figure~\ref{fig:azurestore} shows the high-level architecture of Azure Storage vNext, a distributed extent management system for Windows Azure. This system consists of multiple extent managers, extent nodes and a remote procedure call (RPC) communication engine. An extent manager is responsible for managing a subset of the extent nodes. An extent node is responsible for storing the extent in a local storage. Finally, the RPC communication engine is responsible for sending messages across the network, and enqueuing any received messages to the correct nodes. Azure Storage vNext is one of the two industrial case studies discussed in this paper and more details can be found in Section~\ref{sec:cases:azurestore}.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/azurestore}
\caption{High-level architecture of a distributed extent management system for Windows Azure.}
\label{fig:azurestore}
\end{figure}

How cheng was testing his code before and how after -- high level, pretty picture

### 2015-08-18 overview.tex
updates in overview

@@
-Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages (in this paper we will refer them as events) through a network layer.
-When a component receives an event, it responds by executing an predefined event handler.
+Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer.
+When a component receives a message, it responds by executing a predefined \emph{message handler}.
+This handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
 
-Figure~\ref{fig:azurestore} shows the high-level architecture of Azure Storage vNext, a distributed extent management system for Windows Azure.
-This system consists of multiple extent managers, extent nodes and a remote procedure call (RPC) communication engine.
-An extent manager is responsible for managing a subset of the extent nodes.
-An extent node is responsible for storing the extent in a local storage.
-Finally, the RPC communication engine is responsible for sending messages across the network, and enqueuing any received messages to the correct nodes.
-Azure Storage vNext is one of the two industrial case studies discussed in this paper and more details can be found in Section~\ref{sec:cases:azurestore}.
+As an example of a distributed system, Figure~\ref{fig:azurestore} shows the top-level components of Azure Storage vNext, a distributed extent management system for Windows Azure that is used in production inside Microsoft.
+Azure Storage vNext consists of multiple extent managers, extent nodes and a remote procedure call (RPC) communication engine.
+Each extent manager is responsible for managing a subset of the extent nodes.
+Each extent node is responsible for storing its corresponding extent in a local storage.
+Finally, the RPC engine is responsible for sending messages across the network, and enqueuing any received messages in the inbox queue of the corresponding node.
+This system is described in more details in Section~\ref{sec:cases:azurestore}.
+
+\subsection{Finding bugs in distributed systems}
+\label{sec:overview:bugs}
+
+Due to the asynchronous nature of message passing, message handlers can potentially interleave in arbitrary order.
+To complicate matters further, random failures are the norm in distributed systems that operate in production: nodes might fail at any moment in a cluster, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and appropriately recover the state of the system.
+Also, with multicore machines becoming a commodity nowadays, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
+
+To be able to efficiently find bugs in a distributed system and deterministically reproduce them, a testing tool must provide three key capabilities: capturing all sources of nondeterminism; injecting faults in arbitrary interleavings; and exploring all possible interleavings in a systematic fashion.
+Traditional testing techniques used by product groups, although focus on stressing the system and randomly injecting faults, they are typically unable to capture and control all sources of nondeterminism, which makes them ineffective in finding the most tricky of bugs.
 

### 2015-08-18 overview.tex
updates in overview

@@
 
-\subsection{Finding bugs in distributed systems}
+\subsection{Testing distributed systems}
 \label{sec:overview:bugs}
 
-Due to the asynchronous nature of message passing, message handlers can potentially interleave in arbitrary order.
-To complicate matters further, random failures are the norm in distributed systems that operate in production: nodes might fail at any moment in a cluster, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and appropriately recover the state of the system.
-Also, with multicore machines becoming a commodity nowadays, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
-
-To be able to efficiently find bugs in a distributed system and deterministically reproduce them, a testing tool must provide three key capabilities: capturing all sources of nondeterminism; injecting faults in arbitrary interleavings; and exploring all possible interleavings in a systematic fashion.
-Traditional testing techniques used by product groups, although focus on stressing the system and randomly injecting faults, they are typically unable to capture and control all sources of nondeterminism, which makes them ineffective in finding the most tricky of bugs.
+In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication.
+To complicate matters further, unexpected failures are the norm in production: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system.
+Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
 
@@
 \includegraphics[width=\linewidth]{img/azurestore}
-\caption{High-level architecture of a distributed extent management system for Windows Azure.}
+\caption{Top-level components of a distributed extent management system for Windows Azure.}
 \label{fig:azurestore}
@@
 
+All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+%
+The ideal testing tool should be able to work on unmodified systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
+However, this is easier said than done when testing production systems.
+
+\subsection{Bugs in distributed systems}
+\label{sec:overview:bugs}
+
+There are two types of properties that developers of distributed systems are mostly interested in testing: \emph{safety} and \emph{liveness} properties.
+A safety property holds if it is \emph{always} satisfied, i.e.
+holds invariantly in every program state during execution.
+Likewise, a liveness property holds if it becomes \emph{eventually} satisfied, i.e.
+it will hold at some program state during execution.
+A safety or liveness property violation means that there is a bug that should be fixed.
+
+Examples ...
+
 How cheng was testing his code before and how after -- high level, pretty picture

### 2015-08-18 overview.tex
merged bg with overview

@@
 How cheng was testing his code before and how after -- high level, pretty picture
+
+\subsection{The \psharp approach}
+\label{sec:overview:psharp}
+
+Insert fancy \psharp picture
+
+The \psharp~\cite{deligiannis2015psharp} project provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} framework for developing highly-reliable distributed systems.
+
+The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines.
+\psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems.
+This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
+Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+
+A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods.
+Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
+This handler might update a field, create a new machine, or send an event to another machine.
+In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+
+Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
+Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
+
+The second core component of the \psharp project is a concurrency unit testing framework that is embedded in the \psharp runtime.
+This framework allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
+
+\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+
+%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
+
+mention dependency injection pattern?

### 2015-08-18 overview.tex
added some text in overview and some cites

@@
 
-\subsection{Testing distributed systems}
+\subsection{Bugs in distributed systems}
 \label{sec:overview:bugs}
@@
 
-\subsection{Bugs in distributed systems}
-\label{sec:overview:bugs}
+There are two types of bugs that developers of distributed systems are mostly interested in finding: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
 
-There are two types of properties that developers of distributed systems are mostly interested in testing: \emph{safety} and \emph{liveness} properties.
-A safety property holds if it is \emph{always} satisfied, i.e.
-holds invariantly in every program state during execution.
-Likewise, a liveness property holds if it becomes \emph{eventually} satisfied, i.e.
-it will hold at some program state during execution.
-A safety or liveness property violation means that there is a bug that should be fixed.
+\begin{description}
+\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.
 
-Examples ...
+\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
+\end{description}
+
+\noindent
+A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
+As an example, in Azure Storage vNext we assert that whenever a message gets dequeued there must be an action that can handle the received message.
+Note that this safety property is not specific to Azure Storage vNext, and can be used in any system based on message passing.
+
+Example of liveness bug in Azure Storage vNext.
+
+Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
+Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
+Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
+Other researchers proposed the use of heuristics and exploring finite executions of an infinite state space system using random walks to identify if a liveness property is satisfied~\cite{killian2007life}.
 

### 2015-08-18 overview.tex
updates in overview

@@
 
-Example of liveness bug in Azure Storage vNext.
+Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
+The liveness property that is being checked in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
 
-Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
 Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
 Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
-Other researchers proposed the use of heuristics and exploring finite executions of an infinite state space system using random walks to identify if a liveness property is satisfied~\cite{killian2007life}.
+Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
+
+TODO: Quickly say how we found Cheng's bug?
+
+\subsection{Distributed systems testing today}
+\label{sec:overview:testing}
 

### 2015-08-18 overview.tex
added some cites

@@
-Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer.
+Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer~\cite{lamport1978time}.
 When a component receives a message, it responds by executing a predefined \emph{message handler}.
@@
 
-TODO: Quickly say how we found Cheng's bug?
+\PDComment{Quickly say how we found Cheng's bug?}
 
-\subsection{Distributed systems testing today}
+\subsection{Testing distributed systems today}
 \label{sec:overview:testing}
 
-How cheng was testing his code before and how after -- high level, pretty picture
+Although a significant amount of research has been conducting during the past few decades on how to test distributed systems,
+
+only a small amount of this research was successful enough to be picked up by industry.
+
+\PDComment{How cheng was testing his code before and how after -- high level, pretty picture}
 
@@
 
-Insert fancy \psharp picture
+\PDComment{Insert fancy \psharp picture}
 
@@
 
-mention dependency injection pattern?
+\PDComment{mention dependency injection pattern?}

### 2015-08-18 intro.tex
added some cites

@@
-Distributed systems are notoriously hard to test.
-This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with clients and external services.
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, maddox2015test}.
+This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
 %
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases.
+A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}.
 

### 2015-08-18 overview.tex
added some cites

@@
 
-All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
 %

### 2015-08-18 overview.tex
updates in overview

@@
 
-Although a significant amount of research has been conducting during the past few decades on how to test distributed systems,
+Although a significant amount of research has been conducted during the past few decades on how to test distributed systems~\cite{lamport1977proving, lamport1994temporal, schuppan2004efficient, killian2007life, yang2009modist}, it is still not clear how to successfully apply all these promising research techniques in a complex production environment.
 
-only a small amount of this research was successful enough to be picked up by industry.
+Product groups are still using techniques such as unit testing, integration testing and stress testing, which although can catch many bugs, they are not able to catch the most tricky ones.
 

### 2015-08-18 intro.tex
updates in overview

@@
 
-Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
+Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
 Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.

### 2015-08-18 overview.tex
updates in overview

--- 2015-08-18 overview.tex
@@
 
-All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
-%
-The ideal testing tool should be able to work on unmodified systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
-However, this is easier said than done when testing production systems.
-
-There are two types of bugs that developers of distributed systems are mostly interested in finding: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
+All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
+The majority of bugs in distributed systems can be classified in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
 
@@
 
-Although a significant amount of research has been conducting during the past few decades on how to test distributed systems,
+A significant amount of research has been conducted during the past few decades on how to test distributed systems~\cite{lamport1977proving, lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but it is still not clear how to successfully apply all these research techniques in a real production environment, which can be full of complications~\cite{cavage2013there, maddox2015test}.
 
-only a small amount of this research was successful enough to be picked up by industry.
+Product groups today rely on techniques such as unit testing, integration testing and stress testing, which although can catch many bugs, are unable to catch the most tricky ones, which can be caught only by using advanced \emph{asynchrony-aware} testing techniques.
+The ideal testing tool should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
+However, this is easier said than done when testing production systems.
 

### 2015-08-18 paper.tex
fix in additional authors
added a picture depicting the high level architecture of xstore vnext
merged bg with overview
added some cites
Starting to describe P# async integration for migration.

@@
 
-\numberofauthors{7} %  in this sample file, there are a *total*
+\numberofauthors{6} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 
-\date{15 August 2015}
+\date{17 August 2015}
 % Just remember to make sure that the TOTAL number of authors
@@
 
-How cheng was testing his code before and how after -- high level, pretty picture
-
-%\input{overview}
-
-\section{Background}
-\label{sec:bg}
-
-\input{background}
-
-Dependency injection pattern
+\input{overview}
 
@@
 \subsection{Distributed Extent Management System}
+\label{sec:cases:azurestore}
 
@@
 
-We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing it.
+We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
 MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
@@
 After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We also use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp scheduling strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
+We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
 
-To do...
+% Draft
+We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.
+Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.
+By default, all event processing occurs on the .NET thread pool.
+
+We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.
+Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.
 

### 2015-08-19 overview.tex
updates in overview

@@
 Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer~\cite{lamport1978time}.
-When a component receives a message, it responds by executing a predefined \emph{message handler}.
-This handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
+Each component has its own input message queue, and when a message arrives, the component responds by executing an appropriate \emph{message handler}.
+Such a handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
 
 As an example of a distributed system, Figure~\ref{fig:azurestore} shows the top-level components of Azure Storage vNext, a distributed extent management system for Windows Azure that is used in production inside Microsoft.
-Azure Storage vNext consists of multiple extent managers, extent nodes and a remote procedure call (RPC) communication engine.
+We will use this system as the main running example in this paper.
+The system is described in more details in Section~\ref{sec:cases:azurestore}, together with another case study from Microsoft.
+
+Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node.
 Each extent manager is responsible for managing a subset of the extent nodes.
-Each extent node is responsible for storing its corresponding extent in a local storage.
-Finally, the RPC engine is responsible for sending messages across the network, and enqueuing any received messages in the inbox queue of the corresponding node.
-This system is described in more details in Section~\ref{sec:cases:azurestore}.
+Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager.
+When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/azurestore}
+\caption{Top-level components of a distributed extent management system for Windows Azure.}
+\label{fig:azurestore}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/azurestore}
-\caption{Top-level components of a distributed extent management system for Windows Azure.}
-\label{fig:azurestore}
-\end{figure}
-
-All these sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
-The majority of bugs in distributed systems can be classified in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
+All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
+Bugs in distributed systems can be largely classified in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
 
@@
 
-\subsection{Testing distributed systems today}
+\subsection{Testing production systems}
 \label{sec:overview:testing}
 
-A significant amount of research has been conducted during the past few decades on how to test distributed systems~\cite{lamport1977proving, lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but it is still not clear how to successfully apply all these research techniques in a real production environment, which can be full of complications~\cite{cavage2013there, maddox2015test}.
+A significant amount of research has been conducted during the past few decades on how to test distributed systems~\cite{lamport1977proving, lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but it is still not clear how to apply all these research techniques in a real production environment, which can be full of complications~\cite{cavage2013there, maddox2015test}.
 
 Product groups today rely on techniques such as unit testing, integration testing and stress testing, which although can catch many bugs, are unable to catch the most tricky ones, which can be caught only by using advanced \emph{asynchrony-aware} testing techniques.
+
+without using advanced \emph{asynchrony-aware} testing techniques
+
 The ideal testing tool should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.

### 2015-08-19 overview.tex
updates in overview

@@
 
-All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
-Bugs in distributed systems can be largely classified in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
+All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+
+We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
 
@@
 
-A significant amount of research has been conducted during the past few decades on how to test distributed systems~\cite{lamport1977proving, lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but it is still not clear how to apply all these research techniques in a real production environment, which can be full of complications~\cite{cavage2013there, maddox2015test}.
+Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
+However, these techniques are not effective for testing distributed systems, which have many sources of nondeterminism.
 
-Product groups today rely on techniques such as unit testing, integration testing and stress testing, which although can catch many bugs, are unable to catch the most tricky ones, which can be caught only by using advanced \emph{asynchrony-aware} testing techniques.
+As an example, all the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
+However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions.
+The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied.
+The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
 
-without using advanced \emph{asynchrony-aware} testing techniques
-
-The ideal testing tool should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
+The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
 However, this is easier said than done when testing production systems.
 
-\PDComment{How cheng was testing his code before and how after -- high level, pretty picture}
+A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
+A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
+Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
+A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
+Even if the model is verified, there is no guarantee that the code that will actually execute is free of bugs.
+
+Our goal in this work instead is to \emph{test what is actually being executed}.
+To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework.
+The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.
 
@@
 
-\PDComment{Insert fancy \psharp picture}
-
-The \psharp~\cite{deligiannis2015psharp} project provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} framework for developing highly-reliable distributed systems.
+The \psharp~\cite{deligiannis2015psharp} framework provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} engine for developing highly-reliable distributed systems.
 
@@
 
-The second core component of the \psharp project is a concurrency unit testing framework that is embedded in the \psharp runtime.
-This framework allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
+The second core component of the \psharp project is a concurrency unit testing engine that is embedded in the \psharp runtime.
+This engine allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
 
@@
 %The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
-
-\PDComment{mention dependency injection pattern?}

### 2015-08-19 related.tex
updates in overview

A significant amount of research has been conducted on how to analyze and test distributed systems~\cite{lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but a lot of these techniques either have significant limitations, or cannot be easily applied in a production environment, due to many complexities that are outside the scope of a research project.

### 2015-08-19 overview.tex
updates in overview

@@
 
-\subsection{The \psharp approach}
+\subsection{The \psharp framework}
 \label{sec:overview:psharp}
 
-The \psharp~\cite{deligiannis2015psharp} framework provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} engine for developing highly-reliable distributed systems.
+\psharp~\cite{deligiannis2015psharp} provides an \emph{event-driven asynchronous programming} language and a \emph{systematic concurrency testing} engine for developing highly-reliable distributed systems.
 
-The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines.
-\psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems.
+The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
+\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
 This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
@@
 
-A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods.
+A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
 Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
@@
 In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
 
@@
 
-The second core component of the \psharp project is a concurrency unit testing engine that is embedded in the \psharp runtime.
-This engine allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where a embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
 
@@
 
+\PDComment{Do we want to actually mention (and compare with) P in this paper?}
 %The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

### 2015-08-19 methodology.tex
some draft stuff in methodology

The original approach to using \psharp was either to develop a system from scratch using the \psharp language, or to port (parts of) an existing system in \psharp and then test them.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/models_old}
\caption{Draft.}
\label{fig:oldapproach}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/models_new}
\caption{Draft.}
\label{fig:newapproach}
\end{figure}

\PDComment{mention dependency injection pattern?}

### 2015-08-19 methodology.tex
some draft stuff in methodology

@@
-The original approach to using \psharp was either to develop a system from scratch using the \psharp language, or to port (parts of) an existing system in \psharp and then test them.
+In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
+First, we ported the system to \psharp, we then modeled its environment as \psharp state machines, and then we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
+The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
+However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
+
+In order to solve this problem, and allow \psharp to be used for testing \emph{unmodified} legacy distributed systems, we decided to take a radically different modeling approach.
+Our approach is \emph{flexible}, involves the use of \emph{simple language features} that developers already use for testing their systems, and can be applied to production code.
+
+We only model the environment of the program that we would like to test, and then allow the developer to connect the original system with the environmental model using \emph{dynamic dispatch}, which is provided by the \csharp runtime.
+
+We now give an example of using dynamic dispatch to model the network engine of the extent manager in the Azure Storage vNext case study.
 

### 2015-08-19 paper.tex
updates in overview
- Minor edits to Matt's text - Draft some of the introduction to the evaluation.  No actual results yet.

@@
 \newcommand{\csharp}{C\#\xspace}
+
+\usepackage{color}
+\newcommand{\Comment}[1]{\textcolor{cyan}{#1}}
+\newcommand{\PDComment}[1]{\Comment{Pantazis: #1}}
 
@@
 
-%\input{methodology}
-
-Very simple running example (maybe some master and some worker nodes) and describe how its environment is modeled and its tested.
+\input{methodology}
 
@@
 % N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
-The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because of the limit on traffic supported by a single Azure storage account.
+The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
 As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
@@
 MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
-Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances backed by the same tables, the behavior is compliant with the specification of IChainTable2 for the combined input history.
+Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.
 
@@
 We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
-Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
+Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
 This makes it possible to verify the correctness property as the system executes.
-We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
+We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
 Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
@@
 \psharp controls the interleaving of the backend calls.
-After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
+To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
 We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
@@
 
-Paragraph that says what will be discussed in evaluation and what are the results (very short)
+We compared the bug-finding effectiveness of \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control on several benchmark bugs, described below.
+We report the length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
+\Comment{Overview of actual results?}
 
 \subsection{Experimental Setup}
+
+All experiments were run on \Comment{insert MSR machine, or perhaps CloudDevR1S04 if John/Matt are able to grant access to the rest of the collaborators}. \Comment{specs}
 
@@
 
-Matt's case study with injected bugs
+After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs:
+\begin{enumerate}
+\item Filtering on a user-defined property did not retrieve non-matching rows from the new table that might shadow a matching row from the old table.
+\item ...
+\end{enumerate}
 

### 2015-08-19 overview.tex
updates in overview

@@
 Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
-The liveness property that is being checked in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+As an example, the liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug.
+We discuss later in this paper, how we are able to detect and reproduce this bug using our approach.
 

### 2015-08-19 methodology.tex
updates in overview
some text about method dispatch

@@
 In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
-First, we ported the system to \psharp, we then modeled its environment as \psharp state machines, and then we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
+First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
 The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
 However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
+Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.
 
 In order to solve this problem, and allow \psharp to be used for testing \emph{unmodified} legacy distributed systems, we decided to take a radically different modeling approach.
-Our approach is \emph{flexible}, involves the use of \emph{simple language features} that developers already use for testing their systems, and can be applied to production code.
+We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
 
-We only model the environment of the program that we would like to test, and then allow the developer to connect the original system with the environmental model using \emph{dynamic dispatch}, which is provided by the \csharp runtime.
+We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_engine}
+\caption{Draft.}
+\label{fig:azurestoremodel}
+\end{figure}
+
+\subsection{Modeling the environment}
+\label{sec:method:model}
+
+Draft
+
+\subsection{Using method dispatch for modeling}
+\label{sec:method:dd}
+
+Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
+There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
+\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
+This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
+
+Using method dispatch for modeling is straightforward.
+The system under test exposes a set of APIs as \emph{virtual methods}.
+The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/models_old}
-\caption{Draft.}
-\label{fig:oldapproach}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/models_new}
-\caption{Draft.}
-\label{fig:newapproach}
-\end{figure}
-
 \PDComment{mention dependency injection pattern?}

### 2015-08-20 methodology.tex
some more text about method dispatch

@@
 
-Draft
+Draft.
 
@@
 
-We now give an example of using dynamic dispatch to model the network engine of the extent manager in the Azure Storage vNext case study.
+We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study.
+The network engine is responsible for sending to and receiving messages from the various components of the system.
+During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
+For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
+We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
+In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
+Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
+
+For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
+When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
+
+\subsection{Extensibility of the \psharp framework}
+\label{sec:method:extend}
+
+Async/Await
+
+custom schedulers, etc
+
+Logs/traces -> user can extend them?
 

### 2015-08-20 evaluation.tex
refactoring

--- 2015-08-18 paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-\documentclass{sig-alternate}
-
-\usepackage{url}
-\usepackage{xspace}
-
-\newcommand{\psharp}{P\#\xspace}
-\newcommand{\csharp}{C\#\xspace}
-
-\begin{document}
-%
-% --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
-\title{Uncovering Distributed System Bugs\\ during Testing (not Production!)}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{7} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor Pantazis Deligiannis\\
-       \affaddr{Imperial College London}\\
-       \affaddr{London, UK}\\
-       \email{p.deligiannis@imperial.ac.uk}
-% 2nd. author
-\alignauthor John Erickson\\
-       \affaddr{Microsoft}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{jerick@microsoft.com}
-% 3rd. author
-\alignauthor Cheng Huang\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{cheng.huang@microsoft.com}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Matt McCutchen\\
-       \affaddr{MIT}\\
-       \affaddr{Cambridge, MA, USA}\\
-       \email{rmccutch@mit.edu}
-% 5th. author
-\alignauthor Shaz Qadeer\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{qadeer@microsoft.com}
-% 6th. author
-\alignauthor Wolfram Schulte\\
-       \affaddr{Microsoft}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{schulte@microsoft.com}
-}
-
-\date{15 August 2015}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-Testing distributed systems is very challenging due to multiple sources of nondeterminism, such as arbitrary interleavings between event handlers and unexpected node failures.
-Stress testing, commonly used in industry today, is unable to deal with this kind of nondeterminism, which results in the most tricky bugs being missed during testing and only getting exposed in production.
-
-We present a new methodology for systematically testing unmodified distributed systems.
-Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism.
-We present two case studies of using \psharp to test production distributed systems inside Microsoft.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it.
-\psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
-
-\end{abstract}
-
-\category{D.2.4}{Software Engineering}{Software/Program Verification}
-\category{D.2.5}{Software Engineering}{Testing and Debugging}
-
-\keywords{distributed systems, testing}
-
-\section{Introduction}
-\label{sec:intro}
-
-\input{intro}
-
-\section{Overview}
-\label{sec:overview}
-
-How cheng was testing his code before and how after -- high level, pretty picture
-
-%\input{overview}
-
-\section{Background}
-\label{sec:bg}
-
-\input{background}
-
-Dependency injection pattern
-
-\section{Methodology}
-\label{sec:methodology}
-
-%\input{methodology}
-
-Very simple running example (maybe some master and some worker nodes) and describe how its environment is modeled and its tested.
-
-\section{Case studies}
-\label{sec:cases}
-
-% cheng to review
-\subsection{Distributed Extent Management System}
-
-We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system.
-This component is responsible for managing the partitioned extent metadata and works as follows.
-
-There are multiple \emph{extent managers} (EMgrs) that are in charge of managing a subset of the extents.
-Each EMgr communicates asynchronously (via remote procedure calls) with a number of \emph{extent nodes} (ENs) that store the extents.
-Each EN sends: (i) a \emph{heartbeat} every 5 seconds, to notify the EMgr that it is still available; and (ii) a \emph{sync report} every 5 minutes, to synchronize its extent with the rest of the nodes.
-The EMgr contains two data structures: an \emph{extent center} (ECtr), which is updated every time an EN syncs; and an EN map, which is updated every time it receives a heartbeat from an EN.
-The EN map runs an EN expiration loop that is responsible for removing ENs that have expired from the EN map and also delete them from the ECtr.
-Finally, the EMgr runs an extent repair loop, which examines all contents of all extents in the ECtr and schedules repairs of extents if they are out-of-date (via a repair request).
-
-\subsection{Live Azure Table Migration}
-
-We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing it.
-MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
-It moves all data from the old table to the new table in the background.
-Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
-
-% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
-The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because of the limit on traffic supported by a single Azure storage account.
-As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
-During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration.
-MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
-
-Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
-
-\subsubsection{Modeling approach}
-% TBD where this ends up in the paper ~ Matt 2015-08-17
-
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
-MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
-Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances backed by the same tables, the behavior is compliant with the specification of IChainTable2 for the combined input history.
-
-% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.
-Given a multi-page read, SpecTable can actually produce a list of all valid results.
-Our correctness property is then:
-\begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
-\end{quote}
-%
-\def\term#1{\emph{#1}}
-We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
-Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
-This makes it possible to verify the correctness property as the system executes.
-We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
-Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
-When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.
-When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.
-\psharp controls the interleaving of the backend calls.
-After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We also use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp scheduling strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
-
-To do...
-
-\section{Evaluation}
-\label{sec:eval}
-
-Paragraph that says what will be discussed in evaluation and what are the results (very short)
+We compared the bug-finding effectiveness of \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control on several benchmark bugs, described below.
+We report the length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
+\MMComment{Overview of actual results?}
 
 \subsection{Experimental Setup}
+
+All experiments were run on \MMComment{insert MSR machine, or perhaps CloudDevR1S04 if John/Matt are able to grant access to the rest of the collaborators}. \MMComment{specs}
 
@@
 Effort required to use the system
-
-\section{Related Work}
-\label{sec:rw}
-
-\input{related}
-
-\section{Conclusion}
-\label{sec:concl}
-
-Draft.
-
-%\section{Acknowledgments}
-%Draft.
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{references}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
-
-\end{document}

### 2015-08-20 paper.tex
updates in methodology
refactoring

@@
 \usepackage{color}
-\newcommand{\Comment}[1]{\textcolor{cyan}{#1}}
-\newcommand{\PDComment}[1]{\Comment{Pantazis: #1}}
+\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
+\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
+\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
 
@@
 \section{Methodology}
-\label{sec:methodology}
+\label{sec:method}
 
@@
 
-We compared the bug-finding effectiveness of \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control on several benchmark bugs, described below.
-We report the length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
-\Comment{Overview of actual results?}
-
-\subsection{Experimental Setup}
-
-All experiments were run on \Comment{insert MSR machine, or perhaps CloudDevR1S04 if John/Matt are able to grant access to the rest of the collaborators}. \Comment{specs}
-
-\subsection{Benchmarks}
-
-Benchmarks that are not production code (e.g. multipaxos) that can be used to evaluate testing
-
-Cheng's case study with the actual bug
-
-After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs:
-\begin{enumerate}
-\item Filtering on a user-defined property did not retrieve non-matching rows from the new table that might shadow a matching row from the old table.
-\item ...
-\end{enumerate}
-
-\subsection{Stressing testing vs Systematic testing}
-
-Comparison of traditional stress testing VS what we do with \psharp
-
-How long you had to run the test to find the bug (if you can find it) and how long are the traces
-
-How easy it is to pinpoint the error using the \psharp traces, comparing to traditional stress testing
-
-Effort required to use the system
+\input{evaluation}
 

### 2015-08-20 paper.tex
refactoring

@@
 
-\section{Methodology}
+\section{Testing production systems}
 \label{sec:method}
@@
 
-% cheng to review
-\subsection{Distributed Extent Management System}
-\label{sec:cases:azurestore}
-
-We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system.
-This component is responsible for managing the partitioned extent metadata and works as follows.
-
-There are multiple \emph{extent managers} (EMgrs) that are in charge of managing a subset of the extents.
-Each EMgr communicates asynchronously (via remote procedure calls) with a number of \emph{extent nodes} (ENs) that store the extents.
-Each EN sends: (i) a \emph{heartbeat} every 5 seconds, to notify the EMgr that it is still available; and (ii) a \emph{sync report} every 5 minutes, to synchronize its extent with the rest of the nodes.
-The EMgr contains two data structures: an \emph{extent center} (ECtr), which is updated every time an EN syncs; and an EN map, which is updated every time it receives a heartbeat from an EN.
-The EN map runs an EN expiration loop that is responsible for removing ENs that have expired from the EN map and also delete them from the ECtr.
-Finally, the EMgr runs an extent repair loop, which examines all contents of all extents in the ECtr and schedules repairs of extents if they are out-of-date (via a repair request).
-
-\subsection{Live Azure Table Migration}
-
-We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
-MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
-It moves all data from the old table to the new table in the background.
-Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
-
-% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
-The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
-As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
-During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration.
-MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
-
-Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
-
-\subsubsection{Modeling approach}
-% TBD where this ends up in the paper ~ Matt 2015-08-17
-
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
-MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
-Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.
-
-% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.
-Given a multi-page read, SpecTable can actually produce a list of all valid results.
-Our correctness property is then:
-\begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
-\end{quote}
-%
-\def\term#1{\emph{#1}}
-We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
-Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
-This makes it possible to verify the correctness property as the system executes.
-We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
-Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
-When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.
-When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.
-\psharp controls the interleaving of the backend calls.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
-
-% Draft
-We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.
-Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.
-By default, all event processing occurs on the .NET thread pool.
-
-We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.
-Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.
+\input{casestudies}
 

### 2015-08-20 methodology.tex
added some code to explain how the network engine is mocked

@@
 \includegraphics[width=\linewidth]{img/mocked_engine}
-\caption{Draft.}
+\caption{The model machines that replace the real environment during testing of the Extent Manager.}
 \label{fig:azurestoremodel}
@@
 
-We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study.
+We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
 The network engine is responsible for sending to and receiving messages from the various components of the system.
@@
 When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
+
+\begin{figure}[t]
+\begin{lstlisting}
+// Public interface of the real network engine
+class NetworkEngine {
+  public virtual void SendMessage();
+  public virtual void EnqueueMessage();
+}
+
+// The mocked network engine used during testing
+class MockedNetEngine : NetworkEngine {
+  ExtentManager EM; // Handle to actual system under test
+  MachineId Env; // Handle to modeled environment
+
+  public MockedNetEngine(ExtentManager em, MachineId env) {
+    this.EM = em;
+    this.Env = env;
+  }
+
+  public override void SendMessage(Socket s, Message msg) {
+    PSharpRuntime.Send(this.Env, new MsgEvent(), s, msg);
+  }
+
+  public override void EnqueueMessage(Message msg) {
+    this.EM.ProcessMessage(msg);
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{The mocked network engine used for testing the Azure Storage vNext system.}
+\label{fig:enginecode}
+%\vspace{-2mm}
+\end{figure}
 

### 2015-08-20 paper.tex
added some code to explain how the network engine is mocked

@@
 
+%\usepackage[utf8]{inputenc}
+%\usepackage[T1]{fontenc}
+%\usepackage{microtype}
+%\usepackage{flushend}
+
 \usepackage{url}
+\usepackage{listings}
 \usepackage{xspace}
@@
 \newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
+
+\definecolor{bluekeywords}{rgb}{0.13,0.13,1}
+\definecolor{greencomments}{rgb}{0,0.5,0}
+\definecolor{redstrings}{rgb}{0.9,0,0}
+
+\lstset{language=[Sharp]C,
+showspaces=false,
+showtabs=false,
+breaklines=true,
+showstringspaces=false,
+breakatwhitespace=true,
+escapeinside={(*@}{@*)},
+commentstyle=\color{greencomments},
+keywordstyle=\color{bluekeywords}\bfseries,
+stringstyle=\color{redstrings},
+basicstyle=\ttfamily\scriptsize,
+columns=fullflexible
+}
 

### 2015-08-20 casestudies.tex
new figure and update to old ones

% cheng to review
\subsection{Distributed Extent Management System}
\label{sec:cases:azurestore}

We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system. This component is responsible for managing the partitioned extent metadata and works as follows.

There are multiple \emph{extent managers} (EMgrs) that are in charge of managing a subset of the extents. Each EMgr communicates asynchronously (via remote procedure calls) with a number of \emph{extent nodes} (ENs) that store the extents. Each EN sends: (i) a \emph{heartbeat} every 5 seconds, to notify the EMgr that it is still available; and (ii) a \emph{sync report} every 5 minutes, to synchronize its extent with the rest of the nodes. The EMgr contains two data structures: an \emph{extent center} (ECtr), which is updated every time an EN syncs; and an EN map, which is updated every time it receives a heartbeat from an EN. The EN map runs an EN expiration loop that is responsible for removing ENs that have expired from the EN map and also delete them from the ECtr. Finally, the EMgr runs an extent repair loop, which examines all contents of all extents in the ECtr and schedules repairs of extents if they are out-of-date (via a repair request).

\subsection{Live Azure Table Migration}

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.  It moves all data from the old table to the new table in the background.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/livemigration}
\caption{Draft.}
\label{fig:livemigration}
\end{figure}

% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.  As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.  During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration.  MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.

Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.

\subsubsection{Modeling approach}
% TBD where this ends up in the paper ~ Matt 2015-08-17

MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.  MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.  Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

% Draft
We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.  Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.  By default, all event processing occurs on the .NET thread pool.

We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.  Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.

### 2015-08-21 methodology.tex
update
removed mentions of unmodified as this is wrong

@@
 
-In order to solve this problem, and allow \psharp to be used for testing \emph{unmodified} legacy distributed systems, we decided to take a radically different modeling approach.
+In order to solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a radically different modeling approach.
 We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
@@
 \includegraphics[width=\linewidth]{img/mocked_engine}
-\caption{The model machines that replace the real environment during testing of the Extent Manager.}
+\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
 \label{fig:azurestoremodel}

### 2015-08-21 casestudies.tex
new figure
Migration case study: add references to figures from text.

@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_migration}
+\caption{Draft.}
+\label{fig:mockedmigration}
+\end{figure}
+
 % N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
@@
 As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
-During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration.
+During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).
 MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
@@
 This makes it possible to verify the correctness property as the system executes.
-We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.
+We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable (Figure~\ref{fig:mockedmigration}).
 Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.

### 2015-08-21 casestudies.tex
Migration case study: figure captions and related text tweaks.

@@
 MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
-It moves all data from the old table to the new table in the background.
+A background ``migrator'' job moves all data from the old table to the new table.
 Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
@@
 \includegraphics[width=\linewidth]{img/livemigration}
-\caption{Draft.}
+\caption{Resharding a data set when a third Azure storage account is added. Two key ranges are each migrated to the new account using a MigratingTable instance (abbreviated MTable).}
 \label{fig:livemigration}
@@
 \includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Draft.}
+\caption{MigratingTable \psharp correctness test environment.}
 \label{fig:mockedmigration}
@@
 \begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
 \end{quote}
@@
 This makes it possible to verify the correctness property as the system executes.
-We have a \psharp \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable (Figure~\ref{fig:mockedmigration}).
+We have a \psharp \term{tables machine} containing all three SpecTables, a collection of \term{service machines} containing identically configured MigratingTables, and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
 Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.

### 2015-08-22 paper.tex
removed mentions of unmodified as this is wrong
added akash and paul (need to somehow fix the author names)
switched to the USENIX style

@@
-% This is "sig-alternate.tex" V2.0 May 2012
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
+% TEMPLATE for Usenix papers, specifically to meet requirements of
+%  USENIX '05
+% originally a template for producing IEEE-format articles using LaTeX.
+%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
+% adapted by David Beazley for his excellent SWIG paper in Proceedings,
+%   Tcl 96
+% turned into a smartass generic template by De Clarke, with thanks to
+%   both the above pioneers
+% use at your own risk.  Complaints to /dev/null.
+% make it two column with no page numbering, default is 10 point
 
-\documentclass{sig-alternate}
+% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
+% the .sty file from the LaTeX source template, so that people can
+% more easily include the .sty file into an existing document.  Also
+% changed to more closely follow the style guidelines as represented
+% by the Word sample file.
+
+% Note that since 2010, USENIX does not require endnotes. If you want
+% foot of page notes, don't include the endnotes package in the 
+% usepackage command, below.
+
+% This version uses the latex2e styles, not the very ancient 2.09 stuff.
+\documentclass[letterpaper,twocolumn,10pt]{article}
+
+\usepackage{usenix,epsfig,endnotes}
 
@@
 \begin{document}
-%
-% --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
 
-\title{Uncovering Distributed System Bugs\\ during Testing (not Production!)}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
+%don't want date printed
+\date{}
 
-\numberofauthors{6} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
+%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
+\title{\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)}
+
+%for single author (just remove % characters)
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor Pantazis Deligiannis\\
-       \affaddr{Imperial College London}\\
-       \affaddr{London, UK}\\
-       \email{p.deligiannis@imperial.ac.uk}
-% 2nd. author
-\alignauthor John Erickson\\
-       \affaddr{Microsoft}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{jerick@microsoft.com}
-% 3rd. author
-\alignauthor Cheng Huang\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{cheng.huang@microsoft.com}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Matt McCutchen\\
-       \affaddr{MIT}\\
-       \affaddr{Cambridge, MA, USA}\\
-       \email{rmccutch@mit.edu}
-% 5th. author
-\alignauthor Shaz Qadeer\\
-       \affaddr{Microsoft Research}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{qadeer@microsoft.com}
-% 6th. author
-\alignauthor Wolfram Schulte\\
-       \affaddr{Microsoft}\\
-       \affaddr{Redmond, WA, USA}\\
-       \email{schulte@microsoft.com}
-}
-
-\date{17 August 2015}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
+{\rm Pantazis Deligiannis$^1$, John Erickson$^2$, Cheng Huang$^3$, Akash Lal$^3$}\\
+{\rm Matt McCutchen$^4$, Shaz Qadeer$^3$, Wolfram Schulte$^2$, Paul Thomson$^1$}\\\\
+$^1$Imperial College London, $^2$Microsoft, $^3$Microsoft Research, $^4$MIT
+} % end author
 
 \maketitle
-\begin{abstract}
+
+% Use the following at camera-ready time to suppress page numbers.
+% Comment it out when you first submit the paper for review.
+\thispagestyle{empty}
+
+\subsection*{Abstract}
 Testing distributed systems is very challenging due to multiple sources of nondeterminism, such as arbitrary interleavings between event handlers and unexpected node failures.
@@
 
-We present a new methodology for systematically testing unmodified distributed systems.
+We present a new methodology for systematically testing distributed systems.
 Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism.
@@
 \psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
-
-\end{abstract}
-
-\category{D.2.4}{Software Engineering}{Software/Program Verification}
-\category{D.2.5}{Software Engineering}{Testing and Debugging}
-
-\keywords{distributed systems, testing}
 
@@
 
-%\section{Acknowledgments}
-%Draft.
+\section{Acknowledgments}
 
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{references}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
+A polite author always includes acknowledgments.  Thank everyone,
+especially those who funded the work.
+
+\section{Availability}
+
+It's great when this section says that MyWonderfulApp is free software, 
+available via anonymous FTP from ... Also, it's even greater when you can
+write that information is also  available on the Wonderful homepage at
+
+{\footnotesize \bibliographystyle{acm}
+\bibliography{references}}
 

### 2015-08-22 paper.tex
fix in author list

@@
 {\rm Matt McCutchen$^4$, Shaz Qadeer$^3$, Wolfram Schulte$^2$, Paul Thomson$^1$}\\\\
-$^1$Imperial College London, $^2$Microsoft, $^3$Microsoft Research, $^4$MIT
+$^1$Imperial College London, $^2$Microsoft, $^3$Microsoft Research\\
+$^4$Massachusetts Institute of Technology\\
 } % end author

### 2015-08-22 evaluation.tex
added a draft table with stats
draft results table
added cite to Paul's paper

@@
 
-Benchmarks that are not production code (e.g. multipaxos) that can be used to evaluate testing
+\newcommand{\colspacing}{\hspace{1.8em}}
+\begin{table}[t]
+\small
+\centering
+\setlength{\tabcolsep}{0.3em}
+\label{tab:stats}
+\begin{tabular}{l rrrrr rr}
+\centering
+\input{experiments/tables/statistics.tex}
+\end{tabular}
+\caption{Statistics.}
+%\vspace{-3mm}
+\end{table}
+
+\PDComment{We probably do not need this now that we have many bugs from the case studies -- benchmarks that are not production code (e.g. multipaxos) that can be used to evaluate testing}
 
@@
 
-\subsection{Stressing testing vs Systematic testing}
+\subsection{Stress testing vs \psharp testing}
+
+\setlength{\tabcolsep}{.42em}
+\begin{table*}[t]
+\small
+\centering
+\begin{tabular}{l rr rrrrr}
+\centering
+\input{experiments/tables/results.tex}
+\end{tabular}
+\caption{Results.}
+\label{tab:testing}
+\end{table*}
 
@@
 Effort required to use the system
+
+Controlled random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.

### 2015-08-23 paper.tex
added a draft table with stats
commented out acks
update in table

@@
 
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+\usepackage{pifont}
+
 \usepackage{url}
+\usepackage{multirow}
 \usepackage{listings}
+\usepackage{booktabs}
+%\usepackage{tikz}
 \usepackage{xspace}
@@
 \newcommand{\csharp}{C\#\xspace}
+
+\newcommand{\cmark}{\ding{51}}
+\newcommand{\xmark}{\ding{55}}
 
@@
 \psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
+\PDComment{we now have many bugs in Matt's system too.}
 
@@
 
-\section{Acknowledgments}
+%\section{Acknowledgments}
 
-A polite author always includes acknowledgments.  Thank everyone,
-especially those who funded the work.
-
-\section{Availability}
-
-It's great when this section says that MyWonderfulApp is free software, 
-available via anonymous FTP from ... Also, it's even greater when you can
-write that information is also  available on the Wonderful homepage at
+%Draft.
 

### 2015-08-24 casestudies.tex
some restracturing to emphasize the case studies early on

@@
-% cheng to review
-\subsection{Distributed Extent Management System}
+\subsection{Azure Storage vNext}
 \label{sec:cases:azurestore}
 
-We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system.
-This component is responsible for managing the partitioned extent metadata and works as follows.
+Azure Storage vNext is the next generation storage system for Windows Azure.
+Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node (see Figure~\ref{fig:azurestore}).
+Each extent manager is responsible for managing a subset of the extent nodes.
+Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager.
+When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/azurestore}
+\caption{Top-level components of a distributed extent management system for Windows Azure.}
+\label{fig:azurestore}
+\end{figure}
 
@@
 
+The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug.
+We discuss later in this paper, how we are able to detect and reproduce this bug using our approach.
+
+Azure Storage vNext is very challenging to test.
+All the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
+However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions.
+The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied.
+The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
+
 \subsection{Live Azure Table Migration}
+\label{sec:cases:migration}
 
-We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
-MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.
-A background ``migrator'' job moves all data from the old table to the new table.
+The Live Azure Table Migration is a library capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
+MigratingTable provides a \emph{virtual table} with an API similar to that of an ordinary Azure table, backed by a pair of \emph{old} and \emph{new}  tables.
+A background \emph{migrator} job moves all data from the old table to the new table.
 Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{MigratingTable \psharp correctness test environment.}
-\label{fig:mockedmigration}
-\end{figure}
-
 % N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
@@
 
-Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
+\subsection{Azure Service Fabric}
+\label{sec:cases:fabric}
 
-\subsubsection{Modeling approach}
-% TBD where this ends up in the paper ~ Matt 2015-08-17
-
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
-MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
-Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.
-
-% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.
-Given a multi-page read, SpecTable can actually produce a list of all valid results.
-Our correctness property is then:
-\begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
-\end{quote}
-%
-\def\term#1{\emph{#1}}
-We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
-Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
-This makes it possible to verify the correctness property as the system executes.
-We have a \psharp \term{tables machine} containing all three SpecTables, a collection of \term{service machines} containing identically configured MigratingTables, and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
-Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
-When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.
-When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.
-\psharp controls the interleaving of the backend calls.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
-
-% Draft
-We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.
-Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.
-By default, all event processing occurs on the .NET thread pool.
-
-We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.
-Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.
+Draft.

### 2015-08-24 experience.tex
some restracturing to emphasize the case studies early on

\subsection{Distributed Extent Management System}
\label{sec:cases:azurestore}

We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system. This component is responsible for managing the partitioned extent metadata and works as follows.

\subsection{Live Azure Table Migration}

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/mocked_migration}
\caption{MigratingTable \psharp correctness test environment.}
\label{fig:mockedmigration}
\end{figure}

Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.

MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.  MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.  Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \psharp \term{tables machine} containing all three SpecTables, a collection of \term{service machines} containing identically configured MigratingTables, and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

% Draft
We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.  Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.  By default, all event processing occurs on the .NET thread pool.

We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.  Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.

### 2015-08-24 intro.tex
some restracturing to emphasize the case studies early on

@@
 
-We present two case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol.
+We present three case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol.
 Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time as they did not have an effective way to reproduce the bug and nail down the culprit.
@@
 \item Our infrastructure can test production code written in \csharp, which is a mainstream language.
-\item We present two case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
+\item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
 \item ...

### 2015-08-24 methodology.tex
some restracturing to emphasize the case studies early on

@@
+Our goal in this work is to \emph{test what is being executed}.
+To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework.
+The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.
+
+\subsection{The \psharp framework}
+\label{sec:method:psharp}
+
+\psharp~\cite{deligiannis2015psharp} provides an \emph{event-driven asynchronous programming} language and a \emph{systematic concurrency testing} engine for developing highly-reliable distributed systems.
+
+The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
+\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
+This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
+Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+
+A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
+Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
+This handler might update a field, create a new machine, or send an event to another machine.
+In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+
+Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
+Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
+
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where a embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
+
+\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+
+\PDComment{Do we want to actually mention (and compare with) P in this paper?}
+%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
+
+\subsection{Methodology}
+\label{sec:method:new}
+
 In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
@@
 
+Another important operation that we had to model in order to capture the nondeterminism in  the Azure Storage vNext system, and be able to detect the liveness violation, was to model the timer operation that sends a heartbeat every 5 seconds and a sync message every 5 minutes.
+
 \subsection{Extensibility of the \psharp framework}

### 2015-08-24 overview.tex
updates in overview
refactoring
some restracturing to emphasize the case studies early on

--- 2015-08-19 overview.tex
@@
 
-As an example of a distributed system, Figure~\ref{fig:azurestore} shows the top-level components of Azure Storage vNext, a distributed extent management system for Windows Azure that is used in production inside Microsoft.
-We will use this system as the main running example in this paper.
-The system is described in more details in Section~\ref{sec:cases:azurestore}, together with another case study from Microsoft.
-
-Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node.
-Each extent manager is responsible for managing a subset of the extent nodes.
-Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager.
-When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/azurestore}
-\caption{Top-level components of a distributed extent management system for Windows Azure.}
-\label{fig:azurestore}
-\end{figure}
-
-\subsection{Bugs in distributed systems}
-\label{sec:overview:bugs}
+\subsection{Challenges in testing}
+\label{sec:bg:challenges}
 
 In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication.
-To complicate matters further, unexpected failures are the norm in production: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system.
+To complicate matters further, unexpected failures are the norm in production systems: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system.
 Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
@@
 All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
+However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.
+
+The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
+However, this is easier said than done when testing production systems.
+
+A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
+A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
+Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
+A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
+Even if the model is verified, there is no guarantee that the code that will actually execute is free of bugs.
+
+\subsection{Types of bugs}
+\label{sec:bg:bugs}
 
@@
 A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
-As an example, in Azure Storage vNext we assert that whenever a message gets dequeued there must be an action that can handle the received message.
-Note that this safety property is not specific to Azure Storage vNext, and can be used in any system based on message passing.
+An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.
 
 Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
-The liveness property that is being checked in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
-
 Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
@@
 Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
-
-\PDComment{Quickly say how we found Cheng's bug?}
-
-\subsection{Testing production systems}
-\label{sec:overview:testing}
-
-Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
-However, these techniques are not effective for testing distributed systems, which have many sources of nondeterminism.
-
-As an example, all the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
-However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions.
-The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied.
-The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
-
-The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
-However, this is easier said than done when testing production systems.
-
-A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
-A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
-Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
-A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
-Even if the model is verified, there is no guarantee that the code that will actually execute is free of bugs.
-
-Our goal in this work instead is to \emph{test what is actually being executed}.
-To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework.
-The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.
-
-\subsection{The \psharp approach}
-\label{sec:overview:psharp}
-
-The \psharp~\cite{deligiannis2015psharp} framework provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} engine for developing highly-reliable distributed systems.
-
-The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines.
-\psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems.
-This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
-Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
-
-A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods.
-Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
-This handler might update a field, create a new machine, or send an event to another machine.
-In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
-
-Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
-Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
-
-The second core component of the \psharp project is a concurrency unit testing engine that is embedded in the \psharp runtime.
-This engine allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
-
-\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
-
-%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

### 2015-08-24 casestudies.tex
Added small description of Fabric.

@@
 
-Draft.
+\emph{Azure Service Faric} (or \emph{Fabric} for short)
+is a platform and API for creating reliable services
+that execute on a cluster of machines.
+The developer writes a service that receives requests (e.g.\ from some client program via HTTP)
+and mutates its state based on these requests.
+In order to make the service \emph{reliable},
+Fabric launches several copies (\emph{replicas}) of the service,
+where each copy runs as a separate process
+on different nodes in the cluster.
+One replica is selected to be the \emph{primary}
+which serves client requests; the rest are \emph{secondaries}.
+The primary replicates state changes to the secondaries
+so that all replicas eventually have the same state.
+If the primary fails (e.g.\ if the node on which the primary is running crashes),
+Fabric elects one of the secondaries to be the new primary
+and launches another secondary;
+the new secondary will receive a full or partial copy (depending on whether persistent storage is used) 
+of the state of the new primary in order to ``catch up'' with the other secondaries. 
+Fabric provides a name-resolution service so that clients can always find the current primary.
+Fabric exposes several different APIs that developers can use to write services.

### 2015-08-27 related.tex
added cite for Verdi and added some draft text
added a citation

@@
 A significant amount of research has been conducted on how to analyze and test distributed systems~\cite{lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but a lot of these techniques either have significant limitations, or cannot be easily applied in a production environment, due to many complexities that are outside the scope of a research project.
+
+In Verdi~\cite{wilcox2015verdi}, a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
+Verdi cannot find liveness bugs.
+\psharp is also more production-friendly (works on a mainstream language).
+\PDComment{say few more stuff}
+
+D3S~\cite{liu2008d3s} \PDComment{should read the paper, they say they are debugging running systems}

### 2015-08-28 methodology.tex
added some extra text and code

@@
 
-\PDComment{Do we want to actually mention (and compare with) P in this paper?}
 %The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
 
-\subsection{Methodology}
-\label{sec:method:new}
+\subsection{Modeling the environment}
+\label{sec:method:model}
 
@@
 \end{figure}
-
-\subsection{Modeling the environment}
-\label{sec:method:model}
-
-Draft.
 
@@
 class NetworkEngine {
-  public virtual void SendMessage();
-  public virtual void EnqueueMessage();
+  public virtual void SendMessage(Socket s, Message msg);
+  public virtual void EnqueueMessage(Message msg);
 }
@@
 
-\subsection{Extensibility of the \psharp framework}
-\label{sec:method:extend}
+\subsection{Abstracting timers}
+\label{sec:method:timers}
+
+Distributed systems are often using timers to determine when an event should be fired.
+For example, in the Azure Storage vNext system, each Extent Node is associated with a timer that fires of a synchronization message every 5 minutes and a heartbeat every 5 seconds.
+However, when testing to find bugs, these timers should be abstracted away and be triggered nondeterministically.
+
+Figure~\ref{fig:timer} shows how we modeled timers in Azure Storage vNext.
+Each ExtentNode in the harness is associated with a unique \texttt{Timer} machine.
+Each \texttt{Timer}, when transitions to the \texttt{Active} state, loops infinitely and nondeterministically sends a \texttt{TimerTickEvent} to \texttt{this.Target}.
+
+\begin{figure}[t]
+\begin{lstlisting}
+internal class Timer : Machine
+{
+  MachineId Target;
+  string Info;
+
+  [Start]
+  [OnEntry(nameof(InitOnEntryAction))]
+  [OnEventGotoState(typeof(Unit), typeof(Active))]
+  class Init : MachineState { }
+
+  void InitOnEntryAction()
+  {
+    this.Target = (MachineId)(this.Payload as object[])[0];
+    this.Info = (string)(this.Payload as object[])[1];
+    this.Raise(new Unit());
+  }
+
+  [OnEntry(nameof(ProcessTickEvent))]
+  [OnEventGotoState(typeof(Unit), typeof(Active))]
+  class Active : MachineState { }
+
+  void ProcessTickEvent()
+  {
+    if (this.Nondet())
+    {
+      this.Send(this.Target, new TimerTickEvent());
+    }
+
+    this.Raise(new Unit());
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Timers in Azure Storage vNext are modeled as a nondeterministic \psharp machines.}
+\label{fig:timer}
+%\vspace{-2mm}
+\end{figure}
+
+\subsection{Abstracting node expiration}
+\label{sec:method:timers}
+
+Node expiration
+
+\begin{figure}[t]
+\begin{lstlisting}
+public virtual bool IsNodeExpired(string node_id, DateTime expiration)
+{
+  return DateTime.Compare(expiration, DateTime.Now) <= 0;
+}
+
+public override bool IsNodeExpired(string node_id, DateTime expiration)
+{
+  return deletedNodes.Contains(node_id);
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{draft}
+\label{fig:expiration}
+%\vspace{-2mm}
+\end{figure}
+
+\subsection{Handling intra-machine concurrency}
+\label{sec:method:async}
 
@@
 
+\subsection{Other}
+\label{sec:method:other}
+
 Logs/traces -> user can extend them?

### 2015-08-29 intro.tex
added a citation
aded text

@@
-Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, maddox2015test}.
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
 This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
@@
 \psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.
+We show that bugs in production do not need a large setting to be reproduced; they can be reproduced in small settings with easy to understand traces.
 
@@
 \item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
-\item ...
+\item We show that we can reproduce bugs in production systems in a small setting with easy to understand traces.
 \end{itemize}

### 2015-08-29 methodology.tex
aded text

@@
 Our goal in this work is to \emph{test what is being executed}.
-To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework.
-The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.
+Our approach involves using \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
 
@@
 \label{sec:method:psharp}
-
-\psharp~\cite{deligiannis2015psharp} provides an \emph{event-driven asynchronous programming} language and a \emph{systematic concurrency testing} engine for developing highly-reliable distributed systems.
 
@@
 
-\subsection{Modeling the environment}
+\subsection{Overview of our approach}
 \label{sec:method:model}
@@
 
-In order to solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a radically different modeling approach.
+To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a radically different approach.
 We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
@@
 Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+
+\subsection{Modeling the environment}
+\label{sec:method:model}
+
+The environment of a distributed system might consist of other distributed systems and services, clients, operating system timers, as well as libraries for networking or other purposes.
+To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
 
@@
 
-\subsection{Using method dispatch for modeling}
-\label{sec:method:dd}
+\subsubsection{Using method dispatch for modeling}
+\label{sec:method:model:dd}
 
@@
 
-\subsection{Abstracting timers}
-\label{sec:method:timers}
+\subsubsection{Abstracting timers}
+\label{sec:method:model:timers}
 
-Distributed systems are often using timers to determine when an event should be fired.
+Abstract timers away so all timing events become message passing: why and how
+
+Distributed systems are often using timers to determine when an event should be send from one component to another.
 For example, in the Azure Storage vNext system, each Extent Node is associated with a timer that fires of a synchronization message every 5 minutes and a heartbeat every 5 seconds.
@@
 
-\subsection{Abstracting node expiration}
-\label{sec:method:timers}
+\subsubsection{Abstracting node expiration}
+\label{sec:method:model:timers}
 

### 2015-08-29 methodology.tex
update
more text about modeling timers

@@
 
-We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
-The network engine is responsible for sending to and receiving messages from the various components of the system.
-During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
-For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
-We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
-In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
-Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
-
-For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
-When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
-
 \begin{figure}[t]
@@
 
-Another important operation that we had to model in order to capture the nondeterminism in  the Azure Storage vNext system, and be able to detect the liveness violation, was to model the timer operation that sends a heartbeat every 5 seconds and a sync message every 5 minutes.
+We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
+The network engine is responsible for sending to and receiving messages from the various components of the system.
+During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
+For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
+We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
+In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
+Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
+
+For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
+When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
 
@@
 
-Abstract timers away so all timing events become message passing: why and how
-
 Distributed systems are often using timers to determine when an event should be send from one component to another.
 For example, in the Azure Storage vNext system, each Extent Node is associated with a timer that fires of a synchronization message every 5 minutes and a heartbeat every 5 seconds.
-However, when testing to find bugs, these timers should be abstracted away and be triggered nondeterministically.
+This timer is related to the liveness bug that we discovered: the synchronization message that gets fired every 5 minutes can potentially race with an Extent Node failure; if it arrives after the node failed, then the bug would manifest.
+Traditional testing techniques cannot easily find such a bug, due to the very infrequent occurrence of this race due to the timer.
 
-Figure~\ref{fig:timer} shows how we modeled timers in Azure Storage vNext.
-Each ExtentNode in the harness is associated with a unique \texttt{Timer} machine.
-Each \texttt{Timer}, when transitions to the \texttt{Active} state, loops infinitely and nondeterministically sends a \texttt{TimerTickEvent} to \texttt{this.Target}.
+Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.
+
+Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study.
+The Extend Manager, as well as each Extent Node in the harness, is associated with a unique \texttt{Timer} machine.
+When creating this machine, we pass as a payload the id of the machine that owns this timer.
+When the \texttt{Timer} machine is created, it stores this id in the \texttt{Owner} field and then transitions to the \texttt{Active} state.
+In this state, the \texttt{Timer} loops infinitely and nondeterministically sends a \texttt{TimerTickEvent} to \texttt{this.Owner}.
+When the Extent Node owner receives this event, it handles it by generating a synchronization message that is being send to the Extent Manager.
+Similarly, when the Extent Manager receives a \texttt{TimerTickEvent} from its own \texttt{Timer}, it handles it by nondeterministically invoking repair-related methods in the Extent Repair Center data structure.
 
@@
 {
-  MachineId Target;
-  string Info;
+  MachineId Owner; // Id of the owner machine
 
@@
   {
-    this.Target = (MachineId)(this.Payload as object[])[0];
-    this.Info = (string)(this.Payload as object[])[1];
+    this.Owner = (MachineId)this.Payload;
+    // triggers state transition to Active
     this.Raise(new Unit());
@@
   {
+    // Nondeterministic boolean choice controlled by P#
     if (this.Nondet())
-    {
-      this.Send(this.Target, new TimerTickEvent());
-    }
-
+      // sends a timer tick event to the owner machine
+      this.Send(this.Owner, new TimerTickEvent());
+    // triggers state transition to Active
     this.Raise(new Unit());

### 2015-08-29 methodology.tex
text about modeling node failures

@@
 
-\subsubsection{Abstracting node expiration}
-\label{sec:method:model:timers}
+\subsubsection{Modeling and injecting failures}
+\label{sec:method:model:failures}
 
-Node expiration
+In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
+Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness.
+However, the Extent Manager logic relies on time intervals to detect node failures (see Figure~\ref{fig:expiration}).
+The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
 
@@
 \begin{lstlisting}
+// Real code for detecting node expiration
 public virtual bool IsNodeExpired(string node_id, DateTime expiration)
@@
 
+// Mocked code for detecting node expiration
 public override bool IsNodeExpired(string node_id, DateTime expiration)
 {
-  return deletedNodes.Contains(node_id);
+  return this.DeletedNodes.Contains(node_id);
 }
@@
 \vspace{-2mm}
-\caption{draft}
+\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
 \label{fig:expiration}
@@
 \end{figure}
+
+Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
+Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration.
+If it contains the id, then it means that the node has failed.
+The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
 

### 2015-08-29 methodology.tex
more text about modeling

@@
 To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
+
+In the following subsections, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
+In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
+We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
 
@@
 \vspace{-2mm}
-\caption{Timers in Azure Storage vNext are modeled as a nondeterministic \psharp machines.}
+\caption{Timers in Azure Storage vNext are modeled as nondeterministic \psharp machines.}
 \label{fig:timer}

### 2015-08-29 casestudies.tex
more text, updates and restructuring

@@
 
-\subsection{Azure Service Fabric}
+\subsection{Azure Service Fabric (CScale?}
 \label{sec:cases:fabric}
+
+\PDComment{I guess we should write stuff about CScale here based on our last discussion: overview of the system (collection of Fabric services connected via TPL dataflow?), of its environment (Fabric), how they interop?, why its very challenging to test?
+how they test currently?
+I am not sure if all these should go here or be split here and the experience report, we will see ...
+same for the above case studies}
 

### 2015-08-29 methodology.tex
more text, updates and restructuring

@@
 
-To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a radically different approach.
-We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
+To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
+Our approach of testing existing distributed systems using \psharp requires three key steps:
 
-We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+\begin{enumerate}
+\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines.
+We call this modeled environment the \psharp test harness.
+
+\item The top level asynchrony due to message passing, has to be exposed as event sending using the \psharp runtime API.
+If the communication layer is not based on message passing, then additional effort must be spend into refactoring the system to use message passing.
+This step allows the \psharp runtime to capture and control the message passing interleavings during systematic testing.
+
+\item Any source of nondeterminism in the system or the environment (e.g.
+failures and timers), must be modeled using the \psharp \texttt{Nondet()} method.
+This method returns a nondeterministic boolean value that is controlled by the \psharp runtime and can be used to systematically inject ``nondeterministic'' events and for bug reproduction.
+\end{enumerate}
+
+\PDComment{Step 1 and 3 are related and thus could combine, also in the following subsections I present 1 and 3 together as the process of ``modeling of environment''}
 
@@
 
+We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
+
+We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+
 In the following subsections, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
@@
 We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
+
+\subsubsection{Test harness as \psharp state machines}
+\label{sec:method:model:harness}
+
+Draft
+
+\subsubsection{Entry point to a \psharp test harness}
+\label{sec:method:model:entrypoint}
+
+A developer can specify the entry point of a \psharp test harness similar to how unit tests are typically written.
+Figure~\ref{fig:entrypoint} shows the source code of the entry point for the Azure Storage vNext \psharp test harness.
+The programmer can declare an entry point method using the \texttt{Microsoft.PSharp.Test} attribute.
+When the \psharp systematic testing engine is invoked, it will scan the binary and attempt to find a static method declared with this attribute.
+When such a method is found (in our example the \texttt{Execute()} method), the testing engine will invoke it and start testing the program.
+The systematic engine can optionally run multiple iterations of the test harness, each one potentially exploring a different schedule.
+In each iteration, the program state will be reset (any static fields must be explicitly reset by the programmer) and the entry point will be re-invoked.
+
+\begin{figure}[t]
+\begin{lstlisting}
+[Microsoft.PSharp.Test]
+public static void Execute()
+{
+  PSharpRuntime.CreateMachine(typeof(Environment));
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Static \csharp method acting as the entry point to the Azure Storage vNext \psharp test harness.}
+\label{fig:entrypoint}
+%\vspace{-2mm}
+\end{figure}
+
+\subsubsection{Using method dispatch for modeling}
+\label{sec:method:model:dispath}
+
+Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
+There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
+\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
+This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
 
@@
 \end{figure}
-
-\subsubsection{Using method dispatch for modeling}
-\label{sec:method:model:dd}
-
-Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
-There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
-\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
-This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
 
@@
 // Real code for detecting node expiration
-public virtual bool IsNodeExpired(string node_id, DateTime expiration)
+public virtual bool IsNodeExpired(string id, DateTime time)
 {
-  return DateTime.Compare(expiration, DateTime.Now) <= 0;
+  return DateTime.Compare(time, DateTime.Now) <= 0;
 }
@@
 // Mocked code for detecting node expiration
-public override bool IsNodeExpired(string node_id, DateTime expiration)
+public override bool IsNodeExpired(string id, DateTime time)
 {
-  return this.DeletedNodes.Contains(node_id);
+  return this.DeletedNodes.Contains(id);
 }

### 2015-08-29 methodology.tex
refactoring

@@
 To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
-Our approach of testing existing distributed systems using \psharp requires three key steps:
+Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
 
@@
 
-\PDComment{Step 1 and 3 are related and thus could combine, also in the following subsections I present 1 and 3 together as the process of ``modeling of environment''}
+In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
+In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
+We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
+
+We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+
+In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.
 
@@
 We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
-
-We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
-
-In the following subsections, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
-In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
-We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
 
@@
 The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
-
-\subsection{Handling intra-machine concurrency}
-\label{sec:method:async}
-
-Async/Await
-
-custom schedulers, etc
-
-\subsection{Other}
-\label{sec:method:other}
-
-Logs/traces -> user can extend them?
-
-\PDComment{mention dependency injection pattern?}

### 2015-08-29 psharp.tex
refactoring

\subsection{Handling intra-machine concurrency}
\label{sec:psharp:async}

Async/Await

custom schedulers, etc

\subsection{Other}
\label{sec:psharp:other}

Logs/traces -> user can extend them?

\PDComment{mention dependency injection pattern?}

### 2015-08-29 methodology.tex
more text

@@
+\PDComment{Based on discussion with Shaz and Cheng, I am trying to do the following: this section will be mostly about the high level overview and how we can model the environment of existing distributed systems and capture all nondeterminism (due to message passing, failures, timers, etc) -- this is not specific to \psharp and thus we could potentially move the \psharp description in the next section and keep this more abstract, but I am not convinced about this.
+Then the next section is more specifics about \psharp: how the bug-finding runtime works (just overview as its described in our PLDI paper), how the liveness algorithm works, how we handle async/await -- just the generic approach I implemented based on Matt's original work, and maybe some generic discussion about how you can implement extensions of top of \psharp}
+
 Our goal in this work is to \emph{test what is being executed}.

### 2015-08-29 methodology.tex
update in text

@@
 We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes, which makes this method approachable to product groups.
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
+\emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.
 
@@
 To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
-
-We provide the capability to model the environment of a system using \psharp, and then allow the developer to take advantage of existing language features, such as \emph{method dispatch}, to connect the system under test with the environmental model, and finally test it using the \psharp systematic concurrency testing engine.
 

### 2015-08-29 methodology.tex
more text about modeling

@@
 
-\subsubsection{Test harness as \psharp state machines}
+\subsubsection{Creating a test harness using \psharp}
 \label{sec:method:model:harness}
 
-Draft
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_engine}
+\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
+\label{fig:azurestoremodel}
+\end{figure}
+
+Before being able to use \psharp to test an existing distributed system, the environment of the components that the developer wants to test have to be modeled using the \psharp state machine APIs.
+This involves creating mock classes that inherit from the \texttt{Machine} abstract class.
+The \texttt{Machine} class exposes methods for declaring a state machine (e.g.
+machine states and state transitions) as well as methods for sending events to other machines, declaring event handlers for received events, and accessing the payload from the latest received event.
+
+In the Azure Storage vNext case study, we created the following \psharp machines: \texttt{Environment}, \texttt{ExtentNode}, \texttt{ExtentManager}, and \texttt{Timer}.
+
+The \texttt{Environment} is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute (e.g.
+how many nodes should be created, how many failures should occur, etc).
+
+The \texttt{ExtentNode} machine is a model (less than 150 lines of code) of the real Extent Node and is responsible for generating synchronization events and sending them to the \texttt{ExtentManager} to test the extent manager's repair logic.
+
+The \texttt{ExtentManager} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
+A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
+The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
+The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
+This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+
+Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g.
+synchronization messages, heartbeats and repair logic loops).
+This is discussed in detail in Section~\ref{sec:method:model:timers}.
 
@@
 This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_engine}
-\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
-\label{fig:azurestoremodel}
-\end{figure}
 

### 2015-08-30 psharp.tex
more text
more text about modeling
added some overview regarding P# systematic testing

@@
+\subsection{Overview}
+\label{sec:psharp:overview}
+
+The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
+A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions.
+We now give an overview of how this works, while more details can be found in the original paper~\cite{deligiannis2015psharp}.
+
+The \psharp bug-finding runtime attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order.
+This approach is based on systematic concurrency testing~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} (SCT) techniques that have been previously developed for testing shared memory programs.
+In bug-finding mode, the \psharp runtime serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
+The \psharp runtime will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
+The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
+
+We have implemented multiple schedulers inside the \psharp runtime: \emph{random}, \emph{depth-first-search (DFS)}, ..., \emph{PCT}.
+It is easy to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
+The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far.
+Exposing an intuititive interface for creating new schedulers is inspired by previous work~\cite{desai2015tr}.
+
+We designed the bug-finding mode of the \psharp runtime to enable easy debugging: after a bug is found, the runtime can generate a trace that represents the buggy schedule.
+Note that this trace (in contrast to typical logs generating during production) is sequential.
+
+\subsection{Writing specifications}
+\label{sec:psharp:specs}
+
+Draft
+
+\subsection{Liveness checking}
+\label{sec:psharp:liveness}
+
+Draft
+
 \subsection{Handling intra-machine concurrency}

### 2015-08-30 psharp.tex
more text about psharp bug finding

@@
 
-\subsection{Writing specifications}
-\label{sec:psharp:specs}
+\subsection{Checking safety specifications}
+\label{sec:psharp:safety}
 
-Draft
+Safety property specifications can be encoded in \psharp by using the provided \texttt{Assert(...)} method.
+This method takes as argument a predicate, which if evaluates to false denotes a safety property violation.
+The programmer is also free to use other custom assertion APIs (as \psharp executes actual code), but to enable \psharp to recognize such APIs, the suggested approach is to override them to call the \psharp \texttt{Assert(...)} method.
 
-\subsection{Liveness checking}
+\psharp also provides a way to specify global assertions by using \emph{monitors}, special state machines that
+can only receive events, but not send.
+A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it, as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is event and an optional payload.
+This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls.
+
+\subsection{Checking liveness specifications}
 \label{sec:psharp:liveness}
 
-Draft
+Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states.
+The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states.
 

### 2015-08-30 psharp.tex
text about liveness properties

@@
 A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it, as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is event and an optional payload.
-This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls.
+This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls, as the monitor has to be called deterministically (and not be interleaved).
 
@@
 The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states.
+A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g.
+a node has failed but a new one has not come up yet).
+A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.
+
+One can imagine a liveness monitor as a \emph{thermometer}: as the program executes and the liveness property is not satisfied, the liveness monitor stays in the hot state, which infinitely raises the temperature.
+If the liveness property is ever satisfied, and thus the liveness monitor transitions to a cold state, the temperature is instantly reduced to normal levels.
+Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties.
+The liveness monitors in \psharp are based on previous work~\cite{?}.
+
+\PDComment{Discuss the liveness checking algorithm: are we going to mention the terminating-harness random walk one? or the infinite-harness lasso one with partial caching?}
 

### 2015-08-30 psharp.tex
added text for handling async/await

@@
 
-Async/Await
+Vanilla \psharp is able to capture and take control of the \emph{inter-machine concurrency} due to message passing, but is unable to systematically explore any interleavings due to \emph{intra-machine concurrency} (e.g.
+async/await or TPL).
+This is problematic as nowadays, with multicore machines being a commodity, programmers tend to write multithreaded code to exploit shared memory architectures and increase the performance of individual components of a distributed system.
 
-custom schedulers, etc
+As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp \PDComment{version} language primitives.
+Asynchronous code using async/await is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
 
-\subsection{Other}
-\label{sec:psharp:other}
+\PDComment{small example of how async/await actually works with continuations}
 
-Logs/traces -> user can extend them?
+Developing a fully automatic and universal approach to handle intra-machine concurrency in .NET is very challenging, because there are many APIs that can be used for concurrency and synchronization (e.g.
+\texttt{System.Threading}, TPL, async/await, locks, semaphores) and each one has its own complexities.
+Although developers are willing to model the top-level message passing communication using \psharp, they are resistant in modeling the low-level threading and synchronization methods as this would be a very invasive procedure and such refactoring is unlikely to scale for legacy code.
 
-\PDComment{mention dependency injection pattern?}
+To test our case studies, we decided to focus our efforts in handling async/await as providing a robust solution for a subset of the .NET threading APIs is more feasible than trying to handle arbitrary threading and synchronization.
+Our solution involves using a custom \texttt{TaskScheduler} instead of the default TPL task scheduler.
+The \texttt{TaskScheduler} class \PDComment{write some stuff}.
+
+Our approach works as follows.
+We start the task of the \emph{root} \psharp machine in our custom task scheduler.
+As long as any child tasks spawned by the root machine task do not explicitly start in another scheduler (including the default TPL scheduler), then they will be scheduled for execution in our custom task scheduler.
+Our custom scheduler intercepts the call to enqueue a task, creates a special machine called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine.
+The constructor of a \texttt{TaskMachine} takes as an argument a TPL task and stores it in an field.
+The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it will start executing the wrapped task and then immediately wait on its completion.
+This forces the lifetime of the task to become the lifetime of the machine and vice versa.
+Thus, when the \psharp systematic testing scheduler schedules the \texttt{TaskMachine}, it will also schedule the corresponding task.
+This means that when a \psharp program makes a call to an \texttt{async} method, then the task created in the backend will be automatically wrapped and scheduled.
+
+For \texttt{await}, no special treatment is required because \texttt{await} statements are compiled into a continuation.
+The task associated with the continuation will also be wrapped and scheduled accordingly.

### 2015-09-01 experience.tex
MigratingTable: more about random input call generation, other revisions

@@
 
+\def\term#1{\emph{#1}}
 We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
@@
 
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
+MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the original Azure table API with one exception: it provides \term{streaming reads} with a weaker consistency property than multi-page reads in the original API, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
 MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
@@
 
+\subsubsection{Input generation}
+
+Of course, there are many possible input histories to test.
+Since we originally hoped for a comprehensive form of verification and we didn't feel we had good intuition to prepare a set of specific test cases that would make us confident of having caught any and all concurrency bugs in the protocol, it was natural for us to sample from a distribution of input histories that we defined to exercise all the features of IChainTable2 within certain bounds.
+Furthermore, it was natural to let \psharp control the choice of input history as well as the machine interleaving so we could reproduce both using a single random seed.
+
+All of our input histories include two application processes.
+Each process performs either a single streaming read or a sequence of two atomic calls, each a read or a batch write.
+Each batch write call includes one or two operations, where the operation type is chosen from the set supported by IChainTable2 (Insert, Replace, Merge, Delete, InsertOrReplace, InsertOrMerge, DeleteIfExists) and the row key is chosen from $\{0, \ldots, 5\}$.
+If the operation requires an If-Match value, it is equally likely to be \texttt{*}, the current ETag of the row (if it exists), or some non-matching value.
+Finally, the new entity includes a user-defined property \texttt{isHappy} whose value is equally likely to be true or false.
+For both atomic and streaming reads, the filter expression is equally likely to be empty (i.e., match everything), \texttt{isHappy eq true}, or \texttt{isHappy eq false}.
+
+\subsubsection{Model structure}
+
 % N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.
-Given a multi-page read, SpecTable can actually produce a list of all valid results.
+To comprehensively verify the behavior of MigratingTable under an arbitrary input history, we needed some formulation of the IChainTable2 specification to which to compare it.
+Since the specification is deterministic under sequential calls except for the results of streaming reads, we decided the easiest approach was to write an in-memory reference implementation called SpecTable.
+Given a streaming read call, SpecTable can produce a set of all results that are compliant with the specification.
 Our correctness property is then:
+% Convert to some theorem-like environment? ~ Matt
 \begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+For every execution trace of a collection of MigratingTables backed by the same pair of \term{old} and \term{new} SpecTables (where \psharp chooses the actual result of each streaming read from the valid set) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
 \end{quote}
 %
-\def\term#1{\emph{#1}}
-We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).
-Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.
-This makes it possible to verify the correctness property as the system executes.
-We have a \psharp \term{tables machine} containing all three SpecTables, a collection of \term{service machines} containing identically configured MigratingTables, and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
+We instrumented MigratingTable to report the intended \term{linearization point} of each input call, which in our setting is always one of the corresponding \term{backend calls} to the backend tables (often the last).
+Specifically, after each backend call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
+This makes it possible to verify the correctness property as the model executes.
+The model consists of a \psharp \term{tables machine} containing all three SpecTables; a collection of \term{service machines} containing identically configured MigratingTables; and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
 Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
@@
 To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.
+We use the \psharp random scheduling strategy; we were afraid that an exhaustive strategy would only be feasible within bounds so low that we would miss some bugs.
 
-% Draft
-We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.
-Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.
-By default, all event processing occurs on the .NET thread pool.
-
-We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.
-Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.
+We wanted to implement the core MigratingTable algorithms in \csharp ``async/await'' code, like most of Artifact Services, to achieve both good readability and good performance.
+We used a method similar to that described in Section~\ref{sec:psharp:async} to bring the generated TPL tasks under the control of the \psharp scheduler.
+Then we implemented an ``async'' RPC mechanism based on the .NET RealProxy class that automates the generation of proxies for objects hosted by other \psharp machines (in our setting, the service machines use proxies for the SpecTables and various auxiliary objects hosted by the tables machine).
+When a machine calls a method on a proxy, the proxy sends a \psharp message to the host machine, causing it to execute the method call on the original object and send back the result, which the proxy then returns.
+Thus, the use of these proxies as IChainTable2 backends is transparent to the MigratingTable library, thanks to dynamic dispatch.

### 2015-09-01 evaluation.tex
update in table
some restracturing to emphasize the case studies early on
MigratingTable: description of bug 3

@@
 \centering
-\begin{tabular}{l rr rrrrr}
+\begin{tabular}{l rr rrrr}
 \centering
@@
 \end{tabular}
-\caption{Results.}
+\caption{Results from 100000 iterations of \psharp random walk scheduler.}
 \label{tab:testing}
@@
 Controlled random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
+
+\subsection{An example bug in MigratingTable}
+\MMComment{Move wherever appropriate}
+
+One of the bugs in MigratingTable that we found using the \psharp test stands out because it reflects the type of oversight that tends to occur as designs evolve and it's unclear whether we would have been able to find it by any other method (\MMComment{revise remark when we have stress test results?}).
+This bug, which we named QueryStreamedBackUpNewStream, is in the implementation of a streaming read from the virtual table, which should return a stream of all rows in the table sorted by key.
+The essential implementation idea is to start streams $s_O$, $s_N$ from the old and new backend tables and merge the sorted streams by keeping track of the next row in each stream and returning the row with the lesser key.
+In parallel, the migrator job is concurrently copying rows from the old table to the new table; we had satisfied ourselves that this concurrency would not cause any problems.
+However, then we added support to the migrator job to delete the old table when it finishes copying, which triggers the virtual stream to close $s_O$.
+Suppose the virtual stream is in a state in which the next row in $s_O$ has key $k_O$ and the next row in $s_N$ has key $k_N$, where $k_O < k_N$.
+Further suppose that before the next read from the virtual stream, the migrator job copies a row with key $k$ ($k_O < k < k_N$) from the old table to the new table and then deletes the old table.
+Since $s_O$ has not yet returned this row when it is closed and $s_N$ has already advanced to $k_N$, the row with key $k$ will be missed by the virtual stream.
+A similar problem can occur if $s_N$ does not reflect rows inserted into the new table by the migrator job after $s_N$ is started, as allowed by the IChainTable2 specification.
+Restarting $s_N$ when the old table is deleted fixes both variants of the bug.
+
+\MMComment{Add a figure based on migration-bug3-explanation.pptx (probably a hybrid of slides 4 and 5, assuming we want only one figure).}

### 2015-09-02 paper.tex
some restracturing to emphasize the case studies early on
refactoring
fix
added 2 authors

@@
 \author{
-{\rm Pantazis Deligiannis$^1$, John Erickson$^2$, Cheng Huang$^3$, Akash Lal$^3$}\\
-{\rm Matt McCutchen$^4$, Shaz Qadeer$^3$, Wolfram Schulte$^2$, Paul Thomson$^1$}\\\\
+{\rm Pantazis Deligiannis$^1$, Shuo Chen$^3$, Alastair F. Donaldson$^1$, John Erickson$^2$, Cheng Huang$^3$}\\
+{\rm Akash Lal$^3$, Matt McCutchen$^4$, Shaz Qadeer$^3$, Wolfram Schulte$^2$, Paul Thomson$^1$}\\\\
 $^1$Imperial College London, $^2$Microsoft, $^3$Microsoft Research\\
@@
 
-\section{Overview}
-\label{sec:overview}
+\section{Distributed Systems}
+\label{sec:bg}
 
@@
 
-\section{Testing production systems}
+\section{Case Studies in Microsoft}
+\label{sec:cases}
+
+\input{casestudies}
+
+\section{Testing Production Systems with \psharp}
 \label{sec:method}
@@
 
-\section{Case studies}
-\label{sec:cases}
+\section{The \psharp Bug-Finding Runtime}
+\label{sec:psharp}
 
-\input{casestudies}
+\input{psharp}
+
+\section{Experience Report}
+\label{sec:exp}
+
+\input{experience}
 

### 2015-09-03 casestudies.tex
update
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex

@@
 The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
-As the traffic continues to grow over time, we will need to reshard the data set across a greater number of Azure storage accounts without interrupting service.
+As the traffic continues to grow over time, we will need \SCComment{change to "the system needs", because the word "we" is unclear} to reshard the data set across a greater number of Azure storage accounts without interrupting service.
 During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).
@@
 
-\subsection{Azure Service Fabric (CScale?}
+\subsection{Azure Service Fabric (change to CScale?)}
 \label{sec:cases:fabric}

### 2015-09-03 intro.tex
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex

@@
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}.
+A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}.\SCComment{I suggest we add "but the consequence can be catastrophic".
+Otherwise, why does such an extremely unlikely bug matter?}
 
-Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
+Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
 Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.
+\SCComment{do we have a few real stories about the damages?
+They are helpful.}
 
-We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the unified response was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the unified response \SCComment{change "unified response" to "consensus"?} was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
 The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
@@
 The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system.
-Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.
+Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.\SCComment{we want to avoid making a general point that TLA+ has no guarantee to find all bugs, because our approach cannot either.
+Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."}
 
@@
 We achieve this using \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides two key capabilities: (i) a flexible way of modeling the environment using simple language features; and (ii) a systematic concurrency testing framework that is able to capture and take control of all the nondeterminism in a real system (together with its modeled environment) and systematically explore execution paths to discover bugs.
+\SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?}
 
@@
 \end{itemize}
+
+\SCComment{let me give my overall comments about the current version of the paper here.
+This set of case studies are really cool.
+The fact that our approach is proven to be effective clearly is the primary selling point of the paper.
+Reviewers will love the results we have.
+
+About the writing, there may be 3 areas to improve:
+(1) the word "we" is overloaded, because "we" play multiple roles in this work.
+"We" built \psharp; "we", as developers, designed the TableMigration system; "we", as testers, also take the responsibility to test the legacy Azure vNext system.
+The current writing doesn't clearly differentiate these roles, so readers may wonder how much work does it requires if they want to adopt the approach.
+(2) The intro makes me feel that we compare with existing logic reasoning approaches such as TLA+, and point out the limitation of TLA+.
+Until much later in the paper, I realize that our approach doesn't have any logic reasoning ingredient in it.
+I suggest that the intro makes it clear that our approach is complementary to TLA+, not a competing approach.
+We should avoid give a comparison with TLA+, because it is not an apple-to-apple comparison.
+(3) Section 4 is the most important section about the approach.
+I feel that we spend too much text on section 4.3, which is about building the test harness.
+My concern is that readers' minds may be dominated by how we mocked all kinds of objects and override real methods by fake methods.
+They may feel that, in the end, we tested a highly simplified system.
+This is not true, because the extent manager is a REALLY complicated system (which is shown as a small box in fig.3).
+I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}

### 2015-09-03 methodology.tex
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex

@@
 
-A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where a embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
 
@@
 \item The top level asynchrony due to message passing, has to be exposed as event sending using the \psharp runtime API.
-If the communication layer is not based on message passing, then additional effort must be spend into refactoring the system to use message passing.
+If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
 This step allows the \psharp runtime to capture and control the message passing interleavings during systematic testing.
@@
 failures and timers), must be modeled using the \psharp \texttt{Nondet()} method.
-This method returns a nondeterministic boolean value that is controlled by the \psharp runtime and can be used to systematically inject ``nondeterministic'' events and for bug reproduction.
+This method returns a nondeterministic boolean value \SCComment{why only Boolean?
+even a timer returns a Boolean value?} that is controlled by the \psharp runtime and can be used to systematically inject ``nondeterministic'' events and for bug reproduction.
 \end{enumerate}
@@
 
-The \texttt{Environment} is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute (e.g.
-how many nodes should be created, how many failures should occur, etc).
+The \texttt{Environment} is essentially a ``god'' machine: \SCComment{it is helpful to indicate this machine clearly in Fig.3.
+I think it is the dashed line box} it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute (e.g.
+how many nodes should be created, how many failures should occur \SCComment{why is the failure rate limited?
+is there a supporting evidence for the limit we set?}, etc).
 

### 2015-09-03 paper.tex
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex

@@
 \newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
+\newcommand{\SCComment}[1]{\textcolor{orange}{Shuo: #1}}
 
@@
 We present a new methodology for systematically testing distributed systems.
-Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism.
+Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism\SCComment{this paper mainly focuses on the nondet due to scheduling, right?
+There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc.
+I feel that "all sources" is perhaps an overclaim}.
 We present two case studies of using \psharp to test production distributed systems inside Microsoft.
@@
 
-\section{Distributed Systems}
+\section{Distributed Systems \SCComment{change to "Distributed System Bugs"?}
 \label{sec:bg}

### 2015-09-03 psharp.tex
small update
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex

@@
 The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far.
-Exposing an intuititive interface for creating new schedulers is inspired by previous work~\cite{desai2015tr}.
+Exposing an intuitive interface for creating new schedulers is inspired by previous work~\cite{desai2015tr}.
 
@@
 Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties.
-The liveness monitors in \psharp are based on previous work~\cite{?}.
+The liveness monitors in \psharp are based on previous work~\cite{?}.\SCComment{at this point, I realize that our work (perhaps \psharp) doesn't have any symbolic reasoning ingredient in it, right?
+so the comparison with TLA+ is not an apple-to-apple comparison.}
 
@@
 
-For \texttt{await}, no special treatment is required because \texttt{await} statements are compiled into a continuation.
+For \texttt{await} statements, no special treatment is required because \texttt{await} statements are compiled into a continuation.
 The task associated with the continuation will also be wrapped and scheduled accordingly.

### 2015-09-03 paper.tex
fixed compiler errors
put the names of other student authors first as they have significant contribution in case studies, this list is tentative anyway

@@
 \newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
-\newcommand{\SCComment}[1]{\textcolor{orange}{Shuo: #1}}
+\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
 
@@
 \author{
-{\rm Pantazis Deligiannis$^1$, Shuo Chen$^3$, Alastair F. Donaldson$^1$, John Erickson$^2$, Cheng Huang$^3$}\\
-{\rm Akash Lal$^3$, Matt McCutchen$^4$, Shaz Qadeer$^3$, Wolfram Schulte$^2$, Paul Thomson$^1$}\\\\
-$^1$Imperial College London, $^2$Microsoft, $^3$Microsoft Research\\
-$^4$Massachusetts Institute of Technology\\
+{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$, Alastair F. Donaldson$^\dagger$}\\
+{\rm John Erickson$^\ddagger$, Cheng Huang$^\star$, Akash Lal$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\ddagger$}\\\\
+$^\dagger$Imperial College London, $^\ddagger$Microsoft, $^\star$Microsoft Research\\
+$^\diamond$Massachusetts Institute of Technology\\
 } % end author
@@
 There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc.
-I feel that "all sources" is perhaps an overclaim}.
+I feel that ``all sources'' is perhaps an overclaim}.
 We present two case studies of using \psharp to test production distributed systems inside Microsoft.
@@
 
-\section{Distributed Systems \SCComment{change to "Distributed System Bugs"?}
+\section{Distributed Systems \SCComment{change to "Distributed System Bugs"?}}
 \label{sec:bg}

### 2015-09-03 background.tex
some updates
updates in overview
addressed most of Shuo's comments

--- 2015-08-16 background.tex
@@
-Easy to think you can do better until it happens to you :-)
+Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer~\cite{lamport1978time}.
+Each component has its own input message queue, and when a message arrives, the component responds by executing an appropriate \emph{message handler}.
+Such a handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
 
-\subsection{Programming and testing with \psharp}
-\label{bg:psharp}
+\subsection{Challenges in testing}
+\label{sec:bg:challenges}
 
-The \psharp~\cite{deligiannis2015psharp} project provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} framework for developing highly-reliable distributed systems.
+In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication.
+To complicate matters further, unexpected failures are the norm in production systems: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system.
+Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
 
-The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines.
-\psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems.
-This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
-Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
+However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.
 
-A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods.
-Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
-This handler might update a field, create a new machine, or send an event to another machine.
-In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
+However, this is easier said than done when testing production systems.
 
-Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
-Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
+A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
+A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
+Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
+A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
+Even if the model is verified, the gap between a real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.
 
-The second core component of the \psharp project is a concurrency unit testing framework that is embedded in the \psharp runtime.
-This framework allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.
+\subsection{Types of bugs}
+\label{sec:bg:bugs}
 
-\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
+
+\begin{description}
+\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.
+
+\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
+\end{description}
+
+\noindent
+A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
+An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.
+
+Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
+Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
+Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
+Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

### 2015-09-03 casestudies.tex
addressed most of Shuo's comments

@@
 The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
-As the traffic continues to grow over time, we will need \SCComment{change to "the system needs", because the word "we" is unclear} to reshard the data set across a greater number of Azure storage accounts without interrupting service.
+As the traffic continues to grow over time, the system needs to reshard the data set across a greater number of Azure storage accounts without interrupting service.
 During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).

### 2015-09-03 intro.tex
fixed compiler errors
addressed most of Shuo's comments

@@
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}.\SCComment{I suggest we add "but the consequence can be catastrophic".
-Otherwise, why does such an extremely unlikely bug matter?}
+A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic.
 
-Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
+Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?} \PDComment{only the ones that the user has mocked using the \psharp APIs, some are handled automatically like async/await, we should probably rephrase}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
 Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.
@@
 
-We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the unified response \SCComment{change "unified response" to "consensus"?} was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
 The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
+\PDComment{We are missing a link for Google, but would be good if we can provide such a link, or for other companies like Facebook, I will try find something, but if anyone has a good suggestion please include}
 
@@
 Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.\SCComment{we want to avoid making a general point that TLA+ has no guarantee to find all bugs, because our approach cannot either.
-Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."}
+Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."} \PDComment{We should rephrase.
+By the way the main contrast with TLA+ is that TLA+ works \emph{only} on a model and not the real code, whereas \psharp works on a real code (with a modeled environment).}
 
@@
 The fact that our approach is proven to be effective clearly is the primary selling point of the paper.
-Reviewers will love the results we have.
-
+Reviewers will love the results we have.\\
 About the writing, there may be 3 areas to improve:
@@
 This is not true, because the extent manager is a REALLY complicated system (which is shown as a small box in fig.3).
-I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}
+I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}  \PDComment{Really nice suggestions.
+I totally agree with the last point you make and we should make sure this is addressed (it is super important)!}

### 2015-09-03 methodology.tex
addressed most of Shuo's comments

@@
 
-\item The top level asynchrony due to message passing, has to be exposed as event sending using the \psharp runtime API.
+\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs.
 If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
-This step allows the \psharp runtime to capture and control the message passing interleavings during systematic testing.
+This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.
 
-\item Any source of nondeterminism in the system or the environment (e.g.
-failures and timers), must be modeled using the \psharp \texttt{Nondet()} method.
-This method returns a nondeterministic boolean value \SCComment{why only Boolean?
-even a timer returns a Boolean value?} that is controlled by the \psharp runtime and can be used to systematically inject ``nondeterministic'' events and for bug reproduction.
+\item Any other source of nondeterminism in the system or the environment (e.g.
+failures and timers), must be explicitly modeled using the available \psharp APIs.
+This would allow \psharp to capture the nondeterminism and control it during systematic testing.
 \end{enumerate}
@@
 
-The \texttt{Environment} is essentially a ``god'' machine: \SCComment{it is helpful to indicate this machine clearly in Fig.3.
-I think it is the dashed line box} it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute (e.g.
-how many nodes should be created, how many failures should occur \SCComment{why is the failure rate limited?
-is there a supporting evidence for the limit we set?}, etc).
+The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute.
+The programmer is free to create either a finite test harness (e.g.
+with a predefined number of node failures) or an infinite test harness (e.g.
+with an unbounded number of failures), as long as any nondeterminism (e.g.
+when to insert a failure) is captured using the available \psharp APIs.
 
@@
 Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.
+This nondeterminism is introduced using the \psharp \texttt{Nondet()} method, which returns a nondeterministic boolean value that is controlled by the \psharp runtime during systematic testing.
 
@@
 When the \texttt{Timer} machine is created, it stores this id in the \texttt{Owner} field and then transitions to the \texttt{Active} state.
-In this state, the \texttt{Timer} loops infinitely and nondeterministically sends a \texttt{TimerTickEvent} to \texttt{this.Owner}.
+In this state, the \texttt{Timer} loops infinitely and nondeterministically (using \texttt{Nondet()}) sends a \texttt{TimerTickEvent} to \texttt{this.Owner}.
 When the Extent Node owner receives this event, it handles it by generating a synchronization message that is being send to the Extent Manager.

### 2015-09-03 paper.tex
addressed most of Shuo's comments

@@
 We present a new methodology for systematically testing distributed systems.
-Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism\SCComment{this paper mainly focuses on the nondet due to scheduling, right?
+Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism \SCComment{this paper mainly focuses on the nondet due to scheduling, right?
 There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc.
-I feel that ``all sources'' is perhaps an overclaim}.
+I feel that ``all sources'' is perhaps an overclaim} \PDComment{We should clarify: as long as you model a source of nondet using \psharp, then you can capture it and control it (not automatically)}.
 We present two case studies of using \psharp to test production distributed systems inside Microsoft.
@@
 
-\section{Distributed Systems \SCComment{change to "Distributed System Bugs"?}}
+\section{Testing Distributed Systems}
 \label{sec:bg}
 
-\input{overview}
+\input{background}
 

### 2015-09-04 paper.tex
added rashmi

@@
 \author{
-{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$, Alastair F. Donaldson$^\dagger$}\\
-{\rm John Erickson$^\ddagger$, Cheng Huang$^\star$, Akash Lal$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\ddagger$}\\\\
+{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
+{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\ddagger$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
+{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\ddagger$}\\\\
 $^\dagger$Imperial College London, $^\ddagger$Microsoft, $^\star$Microsoft Research\\

### 2015-09-04 intro.tex
added some citations

@@
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic.
+A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
 
 Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?} \PDComment{only the ones that the user has mocked using the \psharp APIs, some are handled automatically like async/await, we should probably rephrase}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
-Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers.
-\SCComment{do we have a few real stories about the damages?
-They are helpful.}
+Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
 

### 2015-09-04 intro.tex
added comment

@@
 Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?} \PDComment{only the ones that the user has mocked using the \psharp APIs, some are handled automatically like async/await, we should probably rephrase}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
-Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
+Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} \PDComment{this citation is the famous multi-billion cost for the US economy, not sure if it should go here as we are focusing in distr systems} and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
 

### 2015-09-07 casestudies.tex
added extent manager internals figure
added a section on exemplary case, partially through methodology

@@
+\CHComment{With the newly added section on exemplary case, perhaps we could re-organize this section?
+A couple of suggestions: 1) very briefly summarize the three cases and highlight the success stories in the introduction section; 2) move the detailed description of Table Migration and Service Fabric into the experience section.}
+
 \subsection{Azure Storage vNext}
@@
 \end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\linewidth]{img/extent_manager}
+\caption{Internal components of the Extent Manager in Azure Storage vNext.}
+\label{fig:extentmanager}
+\end{figure*}
 

### 2015-09-07 example_vnext.tex
added a section on exemplary case, partially through methodology

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from 10s of Petabytes (PB) in 2010 to Exabytes (EB) in 2015, with the total number of objects stored well-exceeding 60 trillion.

Azure Storage vNext is the next generation system, where the primary design target is to increase the scalability by 100x or more. Similar to the current system, vNext also employs containers, called {\em extents}, to store data. Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em extent nodes}. Different from the current system, which employs a Paxos-based, centralized mapping between extents to ENs, vNext achieves its scalability target by employing a completely distributed mapping. In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager}.

One of the many responsibilities of Extent Manager is to ensure that every extent maintains enough replicas in the system. For that, ExtMgr receives frequent periodic heartbeat messages from every EN. Failure of EN is detected by missing heartbeats. ExtMgr also receives less frequent, but still periodic {\em synchronization reports} from every EN. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, ExtMgr identifies which ENs have failed and which extents are affected and missing replicas. ExtMgr, then, schedules tasks to repair the affected extents and distributes the tasks to ENs. ENs repair the extents from their existing replicas in the system and lazily update ExtMgr via future sync reports.

To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing for Extent Manager.
$i)$ Unit testing sends emulated heartbeats and sync reports to ExtMgr and verifies the messages processed as expected.
$ii)$ Integration testing launches Extent Manager together with multiple ENs, injects EN failure and verifies that affected extents are eventually repaired.
$iii)$ Stress testing launches Extent Manager with many ENs and many extents. It keeps repeating the following process: injecting EN failure, launching new EN and verifying affected extents eventually repaired.

Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in Extent Manager. All the unit test and integration test suites successfully pass every single time. However, the stress test suite could fail {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired. The bug appears difficult to identify, reproduce and troubleshoot. First, it takes very long executions to trigger. Second, extent not being repaired is {\em not} a property that can be easily verified. In practice, the developers rely on very large time-out period to detect the bug. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.

To uncover this bug and many other similar ones, the developers are in constant search of a generic and systematic approach for testing distributed storage systems.

%Since this is a rather typical dilemma in the development of distributed storage systems, a general and systematic approach hopefully would be very helpful to the developers and greatly increase their productivity.

### 2015-09-07 methodology.tex
added a section on exemplary case, partially through methodology

@@
 
-\subsection{Modeling the environment}
+\subsection{Testing Target}
+\label{sec:method:wrap_target}
+
+The testing target in this case is Extent Manager, from the real vNext codebase. It is wrapped inside a \psharp machine in the test setup.
+
+As shown in Figure~\ref{fig:wrap_target}, the \texttt{ExtentManagerMachine} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
+A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
+The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
+The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
+This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+
+\begin{figure}[t]
+\begin{lstlisting}
+// Wrapping the target vNext component in P#
+class ExtentManagerMachine : Machine {
+  private ExtentManager extMgr; // target vNext component
+
+  void Init()
+  {
+    extMgr = new ExtentManager();
+
+    extMgr.netEngine = new MockedNetEngine(); // mock network
+    extMgr.SkippingJournal = true; // skip persistence
+    extMgr.IsMockingTimer = true;	 // disable internal timer
+  }
+
+  [OnEventDoAction(typeof(MessageFromExtentNodeEvent), nameof(DeliverMessageAction))]
+
+  private void DeliverMessageAction()
+  {
+    var msg = (CommandMessage)this.Payload;
+    // relay messages from Extent Node to Extent Manager
+    extMgr.ProcessSingleMessage(msg);
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Wrapping vNext Component in \psharp.}
+\label{fig:wrap_target}
+%\vspace{-2mm}
+\end{figure}
+
+\subsubsection{Relaying Inbound Messages}
+\label{sec:method:model:incoming}
+
+Messages from mocked Extent Nodes are delivered to the wrapper \psharp machine as events.
+The wrapper machine processes such events and relay the corresponding messages to its internal Extent Manager (\emph{extMgr.ProcessSingleMessage}).
+From the perspective of Extent Manager, these messages are no different from these from real Extent Nodes.
+
+%\subsubsection{Using method dispatch for modeling}
+\subsubsection{Proxying Outbound Messages}
+\label{sec:method:model:dispath}
+
+Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
+There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
+\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
+This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
+
+Using method dispatch for modeling is straightforward.
+The system under test exposes a set of APIs as \emph{virtual methods}.
+The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.
+
+%\begin{figure}[t]
+%\begin{lstlisting}
+%// Public interface of the real network engine
+%class NetworkEngine {
+  %public virtual void SendMessage(Socket s, Message msg);
+  %public virtual void EnqueueMessage(Message msg);
+%}
+%
+%// The mocked network engine used during testing
+%class MockedNetEngine : NetworkEngine {
+  %ExtentManager EM; // Handle to actual system under test
+  %MachineId Env; // Handle to modeled environment
+  %
+  %public MockedNetEngine(ExtentManager em, MachineId env) {
+    %this.EM = em;
+    %this.Env = env;
+  %}
+  %
+  %public override void SendMessage(Socket s, Message msg) {
+    %PSharpRuntime.Send(this.Env, new MsgEvent(), s, msg);
+  %}
+  %
+  %public override void EnqueueMessage(Message msg) {
+    %this.EM.ProcessMessage(msg);
+  %}
+%}
+%\end{lstlisting}
+%\vspace{-2mm}
+%\caption{The mocked network engine used for testing the Azure Storage vNext system.}
+%\label{fig:enginecode}
+%%\vspace{-2mm}
+%\end{figure}
+
+\begin{figure}[t]
+\begin{lstlisting}
+// network interface in vNext codebase
+class NetworkEngine {
+  public virtual void SendMessage(Socket s, Message msg);
+}
+
+// mocked network for capturing Extent Manager messages
+class MockedNetEngine : NetworkEngine {
+  public override void SendMessage(Socket s, Message msg) {
+    // capture and relay Extent Manager messages
+    PSharpRuntime.Send(this.Environment, new MessageFromExtentManagerEvent(), s, msg);
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Mocked Network Engine.}
+\label{fig:enginecode}
+%\vspace{-2mm}
+\end{figure}
+
+We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
+The network engine is responsible for sending to and receiving messages from the various components of the system.
+During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
+For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
+We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
+In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
+Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
+
+For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
+When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
+
+\subsection{Mocking Extent Node}
+\label{sec:method:mock_en}
+
+The \texttt{ExtentNode} machine is a model (less than 150 lines of code) of the real Extent Node and is responsible for generating synchronization events and sending them to the \texttt{ExtentManager} to test the extent manager's repair logic.
+
+briefly describe mocked repair logic.
+
+briefly describe mocked sync logic
+
+%\begin{figure}[t]
+%\begin{lstlisting}
+%// Wrapping the target vNext component in P#
+%class ExtentNodeMachine : Machine {
+  %private ExtentNode.ExtentCenter extCtr; // real vNext component
+%
+	%[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
+	%[OnEventDoAction(typeof(DownloadExtentRequestEvent), nameof(ProcessDownloadExtentRequestEvent))]
+	%[OnEventDoAction(typeof(DownloadExtentResponseEvent), nameof(ProcessDownloadExtentResponseEvent))]
+	%[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSyncEvent))]
+%
+  %void ProcessRepairExtentRequest()
+  %{
+    %var source = GetRepairSourceMachine(this.Payload);
+    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
+  %}
+%
+  %void ProcessDownloadExtentRequestEvent()
+  %{
+	  %var destination = GetRepairDestinationMachine(this.Payload);
+    %var ext_id = GetExtentID(this.Payload);
+    %
+		%Extent ext;
+    %if (extCtr.TryGet(ext_id, out ext))
+      %PSharpRuntime.Send(destination, new DownloadExtentResponseEvent(), ext);
+  %}
+%
+  %void ProcessDownloadExtentResponseEvent()
+  %{
+    %var ext = (Extent)this.Payload;
+    %extCtr.AddExtent(ext);
+%
+    %this.Monitor<RepairMonitor>(new NotifyExtentRepairCompletion(), this.NodeID);
+  %}
+%
+  %void ProcessExtentNodeSyncEvent()
+  %{
+    %var sync = new ExtentNodeSync(this.NodeID);
+    %extCtr.Extents.ForEach(ext =>
+      %{
+        %var rec = new ExtentRecord(ext.ExtentID, ext.Length, ext.Sealed);
+        %sync.ExtentRecs.Add(rec);
+      %});
+    %PSharpRuntime.Send(this.NameNode, new NameNodeMessageEvent(), sync);
+  %}
+%}
+%\end{lstlisting}
+%\vspace{-2mm}
+%\caption{Mocked Extent Node.}
+%\label{fig:mocked_en}
+%%\vspace{-2mm}
+%\end{figure}
+
+\begin{figure}[t]
+\begin{lstlisting}
+// Mocking Extent Node in P#
+class ExtentNodeMachine : Machine {
+  // use real vNext component whenever appropriate
+  private ExtentNode.ExtentCenter extCtr;
+
+  [OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
+  [OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSync))]
+
+  // extent repair logic
+  void ProcessRepairExtentRequest()
+  {
+    var source = GetRepairSourceMachine(this.Payload);
+    PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
+  }
+  ...
+
+  // extent node sync logic
+  void ProcessExtentNodeSync()
+  {
+    var sync = new ExtentNodeSync(this.NodeID);
+    extCtr.Extents.ForEach(ext =>
+      {
+        sync.ExtentRecords.Add(new ExtentRecord(ext));
+      });
+    PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Mocked Extent Node.}
+\label{fig:mocked_en}
+%\vspace{-2mm}
+\end{figure}
+
+%\subsection{Modeling the environment}
+\subsection{Test Harness}
 \label{sec:method:model}
@@
 
-\subsubsection{Creating a test harness using \psharp}
+%\subsubsection{Creating a test harness using \psharp}
+\subsubsection{Orchestrating Machines}
 \label{sec:method:model:harness}
@@
 
-The \texttt{ExtentNode} machine is a model (less than 150 lines of code) of the real Extent Node and is responsible for generating synchronization events and sending them to the \texttt{ExtentManager} to test the extent manager's repair logic.
-
-The \texttt{ExtentManager} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
-A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
-The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
-The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
-This is discussed in detail in Section~\ref{sec:method:model:dispath}.
-
 Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g.
@@
 This is discussed in detail in Section~\ref{sec:method:model:timers}.
+
+\subsubsection{Modeling and injecting failures}
+\label{sec:method:model:failures}
+
+In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
+Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness.
+However, the Extent Manager logic relies on time intervals to detect node failures (see Figure~\ref{fig:expiration}).
+The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
+
+\begin{figure}[t]
+\begin{lstlisting}
+// Real code for detecting node expiration
+public virtual bool IsNodeExpired(string id, DateTime time)
+{
+  return DateTime.Compare(time, DateTime.Now) <= 0;
+}
+
+// Mocked code for detecting node expiration
+public override bool IsNodeExpired(string id, DateTime time)
+{
+  return this.DeletedNodes.Contains(id);
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
+\label{fig:expiration}
+%\vspace{-2mm}
+\end{figure}
+
+Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
+Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration.
+If it contains the id, then it means that the node has failed.
+The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
 
@@
 \end{figure}
-
-\subsubsection{Using method dispatch for modeling}
-\label{sec:method:model:dispath}
-
-Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
-There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
-\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
-This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
-
-Using method dispatch for modeling is straightforward.
-The system under test exposes a set of APIs as \emph{virtual methods}.
-The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.
-
-\begin{figure}[t]
-\begin{lstlisting}
-// Public interface of the real network engine
-class NetworkEngine {
-  public virtual void SendMessage(Socket s, Message msg);
-  public virtual void EnqueueMessage(Message msg);
-}
-
-// The mocked network engine used during testing
-class MockedNetEngine : NetworkEngine {
-  ExtentManager EM; // Handle to actual system under test
-  MachineId Env; // Handle to modeled environment
-
-  public MockedNetEngine(ExtentManager em, MachineId env) {
-    this.EM = em;
-    this.Env = env;
-  }
-
-  public override void SendMessage(Socket s, Message msg) {
-    PSharpRuntime.Send(this.Env, new MsgEvent(), s, msg);
-  }
-
-  public override void EnqueueMessage(Message msg) {
-    this.EM.ProcessMessage(msg);
-  }
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{The mocked network engine used for testing the Azure Storage vNext system.}
-\label{fig:enginecode}
-%\vspace{-2mm}
-\end{figure}
-
-We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
-The network engine is responsible for sending to and receiving messages from the various components of the system.
-During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
-For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
-We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
-In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
-Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
-
-For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
-When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
 
@@
 \end{figure}
-
-\subsubsection{Modeling and injecting failures}
-\label{sec:method:model:failures}
-
-In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
-Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness.
-However, the Extent Manager logic relies on time intervals to detect node failures (see Figure~\ref{fig:expiration}).
-The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
-
-\begin{figure}[t]
-\begin{lstlisting}
-// Real code for detecting node expiration
-public virtual bool IsNodeExpired(string id, DateTime time)
-{
-  return DateTime.Compare(time, DateTime.Now) <= 0;
-}
-
-// Mocked code for detecting node expiration
-public override bool IsNodeExpired(string id, DateTime time)
-{
-  return this.DeletedNodes.Contains(id);
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
-\label{fig:expiration}
-%\vspace{-2mm}
-\end{figure}
-
-Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
-Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration.
-If it contains the id, then it means that the node has failed.
-The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.

### 2015-09-07 example_vnext.tex
small pass over the vnext example text

@@
 
-Azure Storage vNext is the next generation system, where the primary design target is to increase the scalability by 100x or more.
-Similar to the current system, vNext also employs containers, called {\em extents}, to store data.
-Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em extent nodes}.
-Different from the current system, which employs a Paxos-based, centralized mapping between extents to ENs, vNext achieves its scalability target by employing a completely distributed mapping.
-In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager}.
+Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to increase the scalability by 100x or more.
+Similar to the current system, vNext employs containers, called {\em extents}, to store data.
+Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
+In contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely distributed mapping.
+In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager} (ExtMgr).
 
-One of the many responsibilities of Extent Manager is to ensure that every extent maintains enough replicas in the system.
-For that, ExtMgr receives frequent periodic heartbeat messages from every EN.
+One of the many responsibilities of ExtMgr is to ensure that every extent maintains enough replicas in the system.
+To achieve this, ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN.
 Failure of EN is detected by missing heartbeats.
@@
 
-To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing for Extent Manager.
-$i)$ Unit testing sends emulated heartbeats and sync reports to ExtMgr and verifies the messages processed as expected.
-$ii)$ Integration testing launches Extent Manager together with multiple ENs, injects EN failure and verifies that affected extents are eventually repaired.
-$iii)$ Stress testing launches Extent Manager with many ENs and many extents.
-It keeps repeating the following process: injecting EN failure, launching new EN and verifying affected extents eventually repaired.
+To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing for Extent Manager:
+\begin{enumerate}
+\item \emph{Unit testing}, which sends emulated heartbeats and sync reports to ExtMgr and verifies that the messages are processed as expected.
 
-Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in Extent Manager.
+\item \emph{Integration testing}, which launches an ExtMgr together with multiple ENs, subsequently injects an EN failure, and finally verifies that the affected extents are eventually repaired.
+
+\item \emph{Stress testing}, which launches an ExtMgr with multiple ENs and multiple extents.
+It keeps repeating the following process: injecting an EN failure, launching a new EN and verifying that the affected extents are eventually repaired.
+\end{enumerate}
+
+Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in ExtMgr.
 All the unit test and integration test suites successfully pass every single time.
@@
 First, it takes very long executions to trigger.
-Second, extent not being repaired is {\em not} a property that can be easily verified.
+Second, an extent not being repaired is {\em not} a property that can be easily verified.
 In practice, the developers rely on very large time-out period to detect the bug.

### 2015-09-07 methodology.tex
some updates

@@
-\PDComment{Based on discussion with Shaz and Cheng, I am trying to do the following: this section will be mostly about the high level overview and how we can model the environment of existing distributed systems and capture all nondeterminism (due to message passing, failures, timers, etc) -- this is not specific to \psharp and thus we could potentially move the \psharp description in the next section and keep this more abstract, but I am not convinced about this.
-Then the next section is more specifics about \psharp: how the bug-finding runtime works (just overview as its described in our PLDI paper), how the liveness algorithm works, how we handle async/await -- just the generic approach I implemented based on Matt's original work, and maybe some generic discussion about how you can implement extensions of top of \psharp}
-
 Our goal in this work is to \emph{test what is being executed}.

### 2015-09-07 psharp.tex
update

@@
 
-We have implemented multiple schedulers inside the \psharp runtime: \emph{random}, \emph{depth-first-search (DFS)}, ..., \emph{PCT}.
-It is easy to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
+We have implemented two schedulers inside the \psharp runtime: \emph{random}, and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}.
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
 The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far.
-Exposing an intuitive interface for creating new schedulers is inspired by previous work~\cite{desai2015tr}.
+Exposing an easy-to-use interface for creating new schedulers is inspired by previous work~\cite{desai2015tr}.
 

### 2015-09-07 psharp.tex
updates

@@
-\subsection{Overview}
-\label{sec:psharp:overview}
-
 The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
@@
 
-The \psharp bug-finding runtime attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order.
+When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic testing engine} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
+The engine attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order.
 This approach is based on systematic concurrency testing~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} (SCT) techniques that have been previously developed for testing shared memory programs.
-In bug-finding mode, the \psharp runtime serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
-The \psharp runtime will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
+
+The systematic testing engine serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
+The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
 The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
 
-We have implemented two schedulers inside the \psharp runtime: \emph{random}, and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}.
+We have implemented two schedulers inside the \psharp systematic testing engine: \emph{random} and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}.
+The random scheduler takes a completely random decision at every scheduling point, whereas the PCT scheduler uses randomization in a disciplined fashion to provide a probabilistic guarantee of finding a bug.
 It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
@@
 
-We designed the bug-finding mode of the \psharp runtime to enable easy debugging: after a bug is found, the runtime can generate a trace that represents the buggy schedule.
+We designed the systematic testing engine in a way that enables easy debugging: after a bug is found, the engine can generate a trace that represents the buggy schedule.
 Note that this trace (in contrast to typical logs generating during production) is sequential.
@@
 can only receive events, but not send.
-A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it, as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is event and an optional payload.
+To declare a monitor, the programmer has to inherit from the \psharp \texttt{Monitor} class.
+A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it; as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is an event and an optional payload.
 This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls, as the monitor has to be called deterministically (and not be interleaved).
@@
 \PDComment{Discuss the liveness checking algorithm: are we going to mention the terminating-harness random walk one? or the infinite-harness lasso one with partial caching?}
+
+The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
+
+\PDComment{Show how we specified the liveness property in vNext?}
 

### 2015-09-07 psharp.tex
updates in async/await

@@
 
-As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp \PDComment{version} language primitives.
-Asynchronous code using async/await is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
-
-\PDComment{small example of how async/await actually works with continuations}
+As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+Asynchronous code using \texttt{async}/\texttt{await} is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
+The key idea behind \texttt{async}/\texttt{await} is that a method declared as \texttt{async} can use internally the \texttt{await} keyword which allows the thread executing the method to wait on a TPL task, or any other \emph{awaitable} object, \emph{without} blocking.
+This is achieved as follows.
+When an \texttt{await} statement is executed, the code following the await is wrapped as a TPL task \emph{continuation} and the method returns to the caller.
+This continuation executes when the awaitable object has completed.
 

### 2015-09-07 psharp.tex
update in section 6

--- 2015-09-07 psharp.tex
@@
 
-As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp \PDComment{version} language primitives.
+As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
 Asynchronous code using async/await is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
-
-\PDComment{small example of how async/await actually works with continuations}
+The key idea behind async/await is that a method declared as \texttt{async} can use internally the \texttt{await} keyword which allows the thread executing the method to wait on a TPL task, or any other \emph{awaitable} object, \emph{without} blocking.
+This is achieved as follows.
+When an \texttt{await} statement is executed, the code following the await is wrapped as a TPL task \emph{continuation} and the method returns to the caller.
+This continuation executes when the awaitable object has completed.
 
@@
 To test our case studies, we decided to focus our efforts in handling async/await as providing a robust solution for a subset of the .NET threading APIs is more feasible than trying to handle arbitrary threading and synchronization.
-Our solution involves using a custom \texttt{TaskScheduler} instead of the default TPL task scheduler.
-The \texttt{TaskScheduler} class \PDComment{write some stuff}.
+Our solution involves using a custom task scheduler whenever the bug-finding mode of the \psharp runtime is enabled, instead of the default TPL task scheduler.
+The \psharp task scheduler inherits from the \texttt{TaskScheduler} class, a low-level API that is responsible for enqueueing TPL tasks into threads.
 
@@
 The task associated with the continuation will also be wrapped and scheduled accordingly.
+
+Regarding support for handling non-async/await concurrency primitives, we are currently investigate feasible approaches.

### 2015-09-09 background.tex
update draft.pdf

@@
 Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
+
+Our goal in this work is to \emph{test what is being executed}.
+Our approach involves using \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
+
+\subsection{The \psharp framework}
+\label{sec:method:psharp}
+
+The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
+\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
+This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
+Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+
+A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
+Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
+This handler might update a field, create a new machine, or send an event to another machine.
+In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+
+Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
+Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
+
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
+
+\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+
+%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
+
+\subsection{Overview of our approach}
+\label{sec:method:model}
+
+In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
+First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
+The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
+However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
+Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.
+
+To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
+Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
+
+\begin{enumerate}
+\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines.
+We call this modeled environment the \psharp test harness.
+
+\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs.
+If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
+This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.
+
+\item Any other source of nondeterminism in the system or the environment (e.g.
+failures and timers), must be explicitly modeled using the available \psharp APIs.
+This would allow \psharp to capture the nondeterminism and control it during systematic testing.
+\end{enumerate}
+
+In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
+In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
+We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
+
+We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
+\emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.
+
+In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.

### 2015-09-09 casestudies.tex
update draft.pdf

--- 2015-09-03 casestudies.tex
@@
+%\CHComment{With the newly added section on exemplary case, perhaps we could re-organize this section?
+A couple of suggestions: 1) very briefly summarize the three cases and highlight the success stories in the introduction section; 2) move the detailed description of Table Migration and Service Fabric into the experience section.}
+
 \subsection{Azure Storage vNext}
@@
 
-Azure Storage vNext is the next generation storage system for Windows Azure.
+%Azure Storage vNext is the next generation storage system for Windows Azure.
 Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node (see Figure~\ref{fig:azurestore}).
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/azurestore}
-\caption{Top-level components of a distributed extent management system for Windows Azure.}
-\label{fig:azurestore}
-\end{figure}
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=\linewidth]{img/azurestore}
+%\caption{Top-level components of a distributed extent management system for Windows Azure.}
+%\label{fig:azurestore}
+%\end{figure}
 
-There are multiple \emph{extent managers} (EMgrs) that are in charge of managing a subset of the extents.
-Each EMgr communicates asynchronously (via remote procedure calls) with a number of \emph{extent nodes} (ENs) that store the extents.
-Each EN sends: (i) a \emph{heartbeat} every 5 seconds, to notify the EMgr that it is still available; and (ii) a \emph{sync report} every 5 minutes, to synchronize its extent with the rest of the nodes.
-The EMgr contains two data structures: an \emph{extent center} (ECtr), which is updated every time an EN syncs; and an EN map, which is updated every time it receives a heartbeat from an EN.
-The EN map runs an EN expiration loop that is responsible for removing ENs that have expired from the EN map and also delete them from the ECtr.
-Finally, the EMgr runs an extent repair loop, which examines all contents of all extents in the ECtr and schedules repairs of extents if they are out-of-date (via a repair request).
-
-The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+%The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
 There was an actual bug in the system, that led this property to fail for some very rare executions.
@@
 
-Azure Storage vNext is very challenging to test.
+%Azure Storage vNext is very challenging to test.
 All the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
@@
 The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
+
+vNext bug goes here.
 

### 2015-09-09 methodology.tex
update draft.pdf

@@
-Our goal in this work is to \emph{test what is being executed}.
-Our approach involves using \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
+To uncover the elusive extent repair bug in Azure Storage vNext, a testing harness is written in \psharp.
+The developers of vNext expect the bug more likely in Extent Manager than in Extent Node.
+Hence, the testing harness focuses on the real Extent Manager with mocked Extent Nodes.
 
-\subsection{The \psharp framework}
-\label{sec:method:psharp}
+The testing harness consists of multiple \psharp machines. 
+\begin{enumerate}
+\item There is a {\em Extent Manager} machine that is a thin wrapper for the real Extent Manager and several {\em Extent Node} machines that represent simple mocked ENs.
 
-The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
-\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
-This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
-Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+\item In addition, there is a {\em Testing Driver} machine that communicates with all the machines, relays messages between Extent Manager and EN machines, and drives testing scenarios.
+In one testing scenario, the testing driver launches Extent Manager with three ENs, with a single extent on one of the ENs.
+It then waits for the extent to be repaired on all three ENs.
+In another testing scenario, the testing driver fails one of the ENs and launches a new EN.
+It then waits for the extent to be repaired on all the remaining ENs.
 
-A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
-Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
-This handler might update a field, create a new machine, or send an event to another machine.
-In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
-All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
-
-Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
-Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
-
-A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
-
-\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
-
-%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
-
-\subsection{Overview of our approach}
-\label{sec:method:model}
-
-In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
-First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
-The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
-However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
-Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.
-
-To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
-Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
-
-\begin{enumerate}
-\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines.
-We call this modeled environment the \psharp test harness.
-
-\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs.
-If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
-This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.
-
-\item Any other source of nondeterminism in the system or the environment (e.g.
-failures and timers), must be explicitly modeled using the available \psharp APIs.
-This would allow \psharp to capture the nondeterminism and control it during systematic testing.
+\item Moreover, there is a {\em Monitor} machine, which collects states from all the ENs and verifies the desired property that the extent is {\em eventually} repaired.
 \end{enumerate}
 
-In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
-In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
-We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
-
-We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{generic} since a programmer can build on top of it to test more complicated use cases (see Section~\ref{}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
-\emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.
-
-In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.
-
-\subsection{Testing Target}
+\subsection{Extent Manager Machine}
 \label{sec:method:wrap_target}
 
-The testing target in this case is Extent Manager, from the real vNext codebase. It is wrapped inside a \psharp machine in the test setup.
+The testing target is the real Extent Manager in vNext, wrapped inside a \psharp machine in the testing harness.
 
-As shown in Figure~\ref{fig:wrap_target}, the \texttt{ExtentManagerMachine} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
-A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
-The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
-The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
-This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+\subsubsection{Extent Manager Internals}
 
-\begin{figure}[t]
-\begin{lstlisting}
-// Wrapping the target vNext component in P#
-class ExtentManagerMachine : Machine {
-  private ExtentManager extMgr; // target vNext component
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\linewidth]{img/extent_manager}
+\caption{Internal components of the Extent Manager in Azure Storage vNext.}
+\label{fig:extentmanager}
+\end{figure*}
 
-  void Init()
-  {
-    extMgr = new ExtentManager();
+Inside Extent Manager, there are two structures related to extent repair: an \emph{Extent Center} and an EN map.
+The Extent Center maintains mapping records from extents to their hosting ENs.
+It is updated upon every sync report from ENs.
+The EN map records the latest heartbeat time from every EN.
+It internally runs an EN expiration loop that is responsible for removing ENs that have been missing heartbeats for extended period and also cleaning up corresponding Extent Center records.
+In addition, Extent Manager runs an extent repair loop, which examines all the Extent Center records, identifies extents with missing replica, schedules extent repair tasks and sends the repair tasks to ENs.
 
-    extMgr.netEngine = new MockedNetEngine(); // mock network
-    extMgr.SkippingJournal = true; // skip persistence
-    extMgr.IsMockingTimer = true;	 // disable internal timer
-  }
-
-  [OnEventDoAction(typeof(MessageFromExtentNodeEvent), nameof(DeliverMessageAction))]
-
-  private void DeliverMessageAction()
-  {
-    var msg = (CommandMessage)this.Payload;
-    // relay messages from Extent Node to Extent Manager
-    extMgr.ProcessSingleMessage(msg);
-  }
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{Wrapping vNext Component in \psharp.}
-\label{fig:wrap_target}
-%\vspace{-2mm}
-\end{figure}
-
-\subsubsection{Relaying Inbound Messages}
-\label{sec:method:model:incoming}
-
-Messages from mocked Extent Nodes are delivered to the wrapper \psharp machine as events.
-The wrapper machine processes such events and relay the corresponding messages to its internal Extent Manager (\emph{extMgr.ProcessSingleMessage}).
-From the perspective of Extent Manager, these messages are no different from these from real Extent Nodes.
-
-%\subsubsection{Using method dispatch for modeling}
-\subsubsection{Proxying Outbound Messages}
+\subsubsection{Stubbing Network}
 \label{sec:method:model:dispath}
 
-Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
+%Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
 There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
@@
 
-Using method dispatch for modeling is straightforward.
+%Using method dispatch for modeling is straightforward.
 The system under test exposes a set of APIs as \emph{virtual methods}.
@@
 
+Extent Manager uses a network engine to send and receive messages from ENs.
+The testing harness mocks the original network engine in vNext and overrides the networking methods.
+In this way, it intercepts all outbound messages and relays to the testing driver, which is responsible for dispatching the messages to destination ENs.
+As shown in Figure~\ref{fig:enginecode}, the mocked network engine captures outbound messages from Extent Manager and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to the testing driver.
+
 \begin{figure}[t]
@@
 
-We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
+%We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
 The network engine is responsible for sending to and receiving messages from the various components of the system.
@@
 
-For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
+%For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
 When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
+
+\subsubsection{Extent Manager Machine}
+
+Figure~\ref{fig:wrap_target} shows code snippet from Extent Manager machine, as a thin wrapper of the real Extent Manager.
+A mocked network engine replaces the real one in Extent Manager.
+It captures and relays all outbound messages from Extent Manager.
+
+%A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
+The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
+The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
+This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+
+%As shown in Figure~\ref{fig:wrap_target}, the \texttt{ExtentManagerMachine} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
+A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
+The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
+The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
+This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+
+\begin{figure}[t]
+\begin{lstlisting}
+// Wrapping the target vNext component in P#
+class ExtentManagerMachine : Machine {
+  private ExtentManager extMgr; // real vNext code
+
+  void Init()
+  {
+    extMgr = new ExtentManager();
+
+    extMgr.netEngine = new MockedNetEngine(); // mock network
+    extMgr.SkippingJournal = true; // skip persistence
+    extMgr.IsMockingTimer = true;	 // disable internal timer
+  }
+
+  [OnEventDoAction(typeof(MessageFromExtentNodeEvent), nameof(DeliverMessageAction))]
+
+  private void DeliverMessageAction()
+  {
+    var msg = (CommandMessage)this.Payload;
+    // relay messages from Extent Node to Extent Manager
+    extMgr.ProcessSingleMessage(msg);
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Wrapping vNext Component in \psharp.}
+\label{fig:wrap_target}
+%\vspace{-2mm}
+\end{figure}
+
+Incoming messages from ENs are delivered to the Extent Manager machine and trigger action.
+The specific action here relays the messages to the internal Extent Manager by invoking \texttt{extMgr.ProcessSingleMessage}.
+From the Extent Manager's perspective, it is unaware of EN mocking and behaves as if it is communicating with several real ENs.
 

### 2015-09-09 evaluation.tex
updated table results
updates
update in evaluation

@@
 
-All experiments were run on \MMComment{insert MSR machine, or perhaps CloudDevR1S04 if John/Matt are able to grant access to the rest of the collaborators}. \MMComment{specs}
+We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
 
-\subsection{Benchmarks}
+Because we fixed all bugs that we discovered using the \psharp systematic testing engine in the case studies, for the evaluation we added an option to conditionally reintroduce each one of the bugs so that we can measure how fast and how often each bug can be discovered with the \psharp random and the PCT scheduler.
+
+Cheng's case study with the actual bug
+
+After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs:
+\begin{enumerate}
+\item Filtering on a user-defined property did not retrieve non-matching rows from the new table that might shadow a matching row from the old table.
+\item ...
+\end{enumerate}
 
@@
 
-\PDComment{We probably do not need this now that we have many bugs from the case studies -- benchmarks that are not production code (e.g. multipaxos) that can be used to evaluate testing}
+\subsection{Systematic testing with \psharp}
 
-Cheng's case study with the actual bug
-
-After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs:
-\begin{enumerate}
-\item Filtering on a user-defined property did not retrieve non-matching rows from the new table that might shadow a matching row from the old table.
-\item ...
-\end{enumerate}
-
-\subsection{Stress testing vs \psharp testing}
-
-\setlength{\tabcolsep}{.42em}
+\setlength{\tabcolsep}{.52em}
 \begin{table*}[t]
@@
 \centering
-\begin{tabular}{l rr rrrr}
+\begin{tabular}{l rrrr rrrr}
 \centering
@@
 \end{tabular}
-\caption{Results from 100000 iterations of \psharp random walk scheduler.}
+\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
+We report: time in seconds to find a bug (Time to bug); number of scheduling points when a bug was found (\#SP); percentage of buggy iterations (\%Buggy); and if a bug was found with a particular scheduler (Bug found?).}
 \label{tab:testing}
@@
 
-Comparison of traditional stress testing VS what we do with \psharp
+Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with an enabled bug using the random and the PCT schedulers.
+We configured the engine to perform 100,000 iterations.
+The random seed for both schedulers was generated in each iteration using the \texttt{DateTime.Now.Millisecond} API which represents the current time in milliseconds.
+The PCT scheduler was configured with a bug depth of 2 and a max number of scheduling steps to execute of 500.
+All reported times are in seconds.
 

### 2015-09-09 evaluation.tex
some updates all around

--- 2015-08-22 evaluation.tex
@@
 \centering
-\begin{tabular}{l rr rrrrr}
+\begin{tabular}{l rrrr rrrr}
 \centering
@@
 \end{tabular}
-\caption{Results.}
+\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
+We report: time in seconds to find a bug (Time to bug); number of scheduling points when a bug was found (\#SP); percentage of buggy iterations (\%Buggy); and if a bug was found with a particular scheduler (Bug found?).}
 \label{tab:testing}

### 2015-09-09 experience.tex
some updates all around

@@
-\subsection{Distributed Extent Management System}
+\subsection{Azure Storage vNext}
 \label{sec:cases:azurestore}
+
+%Azure Storage vNext is the next generation storage system for Windows Azure.
+Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node (see Figure~\ref{fig:azurestore}).
+Each extent manager is responsible for managing a subset of the extent nodes.
+Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager.
+When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.
+
+%The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug.
+We discuss later in this paper, how we are able to detect and reproduce this bug using our approach.
+
+%Azure Storage vNext is very challenging to test.
+All the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
+However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions.
+The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied.
+The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
 
@@
 
+vNext bug goes here.
+
 \subsection{Live Azure Table Migration}
+\label{sec:cases:migration}
+
+The Live Azure Table Migration is a library capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
+MigratingTable provides a \emph{virtual table} with an API similar to that of an ordinary Azure table, backed by a pair of \emph{old} and \emph{new}  tables.
+A background \emph{migrator} job moves all data from the old table to the new table.
+Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/livemigration}
+\caption{Resharding a data set when a third Azure storage account is added. Two key ranges are each migrated to the new account using a MigratingTable instance (abbreviated MTable).}
+\label{fig:livemigration}
+\end{figure}
+
+% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
+The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
+As the traffic continues to grow over time, the system needs to reshard the data set across a greater number of Azure storage accounts without interrupting service.
+During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).
+MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
 
@@
 Thus, the use of these proxies as IChainTable2 backends is transparent to the MigratingTable library, thanks to dynamic dispatch.
+
+\subsection{An example bug in MigratingTable}
+\MMComment{Move wherever appropriate}
+
+One of the bugs in MigratingTable that we found using the \psharp test stands out because it reflects the type of oversight that tends to occur as designs evolve and it's unclear whether we would have been able to find it by any other method (\MMComment{revise remark when we have stress test results?}).
+This bug, which we named QueryStreamedBackUpNewStream, is in the implementation of a streaming read from the virtual table, which should return a stream of all rows in the table sorted by key.
+The essential implementation idea is to start streams $s_O$, $s_N$ from the old and new backend tables and merge the sorted streams by keeping track of the next row in each stream and returning the row with the lesser key.
+In parallel, the migrator job is concurrently copying rows from the old table to the new table; we had satisfied ourselves that this concurrency would not cause any problems.
+However, then we added support to the migrator job to delete the old table when it finishes copying, which triggers the virtual stream to close $s_O$.
+Suppose the virtual stream is in a state in which the next row in $s_O$ has key $k_O$ and the next row in $s_N$ has key $k_N$, where $k_O < k_N$.
+Further suppose that before the next read from the virtual stream, the migrator job copies a row with key $k$ ($k_O < k < k_N$) from the old table to the new table and then deletes the old table.
+Since $s_O$ has not yet returned this row when it is closed and $s_N$ has already advanced to $k_N$, the row with key $k$ will be missed by the virtual stream.
+A similar problem can occur if $s_N$ does not reflect rows inserted into the new table by the migrator job after $s_N$ is started, as allowed by the IChainTable2 specification.
+Restarting $s_N$ when the old table is deleted fixes both variants of the bug.
+
+\MMComment{Add a figure based on migration-bug3-explanation.pptx (probably a hybrid of slides 4 and 5, assuming we want only one figure).}
+
+\subsection{Azure Service Fabric (change to CScale?)}
+\label{sec:cases:fabric}
+
+\PDComment{I guess we should write stuff about CScale here based on our last discussion: overview of the system (collection of Fabric services connected via TPL dataflow?), of its environment (Fabric), how they interop?, why its very challenging to test?
+how they test currently?
+I am not sure if all these should go here or be split here and the experience report, we will see ...
+same for the above case studies}
+
+\emph{Azure Service Faric} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines.
+The developer writes a service that receives requests (e.g.\ from some client program via HTTP) and mutates its state based on these requests.
+In order to make the service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on different nodes in the cluster.
+One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
+The primary replicates state changes to the secondaries so that all replicas eventually have the same state.
+If the primary fails (e.g.\ if the node on which the primary is running crashes), Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive a full or partial copy (depending on whether persistent storage is used) of the state of the new primary in order to ``catch up'' with the other secondaries.
+Fabric provides a name-resolution service so that clients can always find the current primary.
+Fabric exposes several different APIs that developers can use to write services.

### 2015-09-09 motivation.tex
some updates all around
update in motivation section

--- 2015-09-07 example_vnext.tex
@@
 
-Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to increase the scalability by 100x or more.
+Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to increase the scalability by more than 100$\times$.
 Similar to the current system, vNext employs containers, called {\em extents}, to store data.
-Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
-In contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely distributed mapping.
-In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager} (ExtMgr).
+Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em extent nodes} (ENs).
+However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely \emph{distributed mapping}.
+In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em extent manager} (ExtMgr).
 
-One of the many responsibilities of ExtMgr is to ensure that every extent maintains enough replicas in the system.
-To achieve this, ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN.
-Failure of EN is detected by missing heartbeats.
-ExtMgr also receives less frequent, but still periodic {\em synchronization reports} from every EN.
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/azurestore}
+\caption{Top-level components of a distributed extent management system for Windows Azure.}
+\label{fig:vnext}
+\end{figure}
+
+One of the many responsibilities of an ExtMgr is to ensure that every extent maintains enough \emph{replicas} in the system.
+To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN.
+The failure of an EN is detected by missing heartbeats.
+An ExtMgr also receives less frequent, but still periodic, {\em synchronization reports} from every EN.
 The sync reports list all the extents (and associated metadata) stored on the EN.
-Based on these two types of messages, ExtMgr identifies which ENs have failed and which extents are affected and missing replicas.
-ExtMgr, then, schedules tasks to repair the affected extents and distributes the tasks to ENs.
-ENs repair the extents from their existing replicas in the system and lazily update ExtMgr via future sync reports.
+Based on these two types of messages, an ExtMgr identifies which ENs have failed and which extents are affected and missing replicas.
+The ExtMgr, then, schedules tasks to repair the affected extents and distributes the tasks to ENs.
+The ENs repair the extents from their existing replicas in the system and lazily update the ExtMgr via future sync reports.
+All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
 
-To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing for Extent Manager:
+To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing:
 \begin{enumerate}
-\item \emph{Unit testing}, which sends emulated heartbeats and sync reports to ExtMgr and verifies that the messages are processed as expected.
+\item \emph{Unit testing}, which sends emulated heartbeats and sync reports to an ExtMgr and verifies that the messages are processed as expected.
 
@@
 
-Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in ExtMgr.
+Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in the ExtMgr logic.
 All the unit test and integration test suites successfully pass every single time.
-However, the stress test suite could fail {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired.
+However, the stress test suite fails {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired.
 The bug appears difficult to identify, reproduce and troubleshoot.

### 2015-09-09 overview.tex
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex
various updates

@@
 Such a handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
-
-\subsection{Challenges in testing}
-\label{sec:bg:challenges}
 
@@
 However, this is easier said than done when testing production systems.
-
-A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
-A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
-Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
-A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
-Even if the model is verified, there is no guarantee that the code that will actually execute is free of bugs.
 
@@
 Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
-Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
-Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
-Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
+The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
+
+\subsection{The \psharp framework}
+\label{sec:method:psharp}
+
+Our goal in this work is to \emph{test what is being executed}.
+To achieve this we use \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
+
+The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
+\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
+This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
+Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+
+A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
+Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
+This handler might update a field, create a new machine, or send an event to another machine.
+In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+
+Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
+Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
+
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
+
+\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
+
+%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
+
+\subsection{Overview of our approach}
+\label{sec:method:model}
+
+In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
+First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
+The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
+However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
+Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.
+
+To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
+Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
+
+\begin{enumerate}
+\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines.
+We call this modeled environment the \psharp test harness.
+
+\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs.
+If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
+This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.
+
+\item Any other source of nondeterminism in the system or the environment (e.g.
+failures and timers), must be explicitly modeled using the available \psharp APIs.
+This would allow \psharp to capture the nondeterminism and control it during systematic testing.
+\end{enumerate}
+
+In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
+In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
+We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
+
+We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:exp}).
+Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
+\emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.
+
+In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.

### 2015-09-09 psharp.tex
various updates

@@
 Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states.
+Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
+Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
+Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
+
 The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states.
@@
 \PDComment{Discuss the liveness checking algorithm: are we going to mention the terminating-harness random walk one? or the infinite-harness lasso one with partial caching?}
-
-The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
-There was an actual bug in the system, that led this property to fail for some very rare executions.
-Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
 

### 2015-09-09 related.tex
various updates

@@
 A significant amount of research has been conducted on how to analyze and test distributed systems~\cite{lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but a lot of these techniques either have significant limitations, or cannot be easily applied in a production environment, due to many complexities that are outside the scope of a research project.
+
+A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
+A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
+Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
+A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
+Even if the model is verified, the gap between a real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.
 

### 2015-09-10 evaluation.tex
update in evaluation

@@
 \centering
-\begin{tabular}{l rrrr rrrr}
+\begin{tabular}{rl rrrr rrrr}
 \centering
@@
 \caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
-We report: time in seconds to find a bug (Time to bug); number of scheduling points when a bug was found (\#SP); percentage of buggy iterations (\%Buggy); and if a bug was found with a particular scheduler (Bug found?).}
+We report: time in seconds to find a bug (Time to Bug); number of scheduling points when a bug was found (\#SP); percentage of buggy iterations (\%Buggy); and if a bug was found with a particular scheduler (Bug Found?).}
 \label{tab:testing}

### 2015-09-10 experience.tex
Some additions regarding Fabric model.

@@
 \emph{Azure Service Faric} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines.
-The developer writes a service that receives requests (e.g.\ from some client program via HTTP) and mutates its state based on these requests.
-In order to make the service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on different nodes in the cluster.
+The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests.
+In order to make the service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on a different node in the cluster.
 One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
-The primary replicates state changes to the secondaries so that all replicas eventually have the same state.
+The primary replicates state changes to the secondaries by sending \emph{replication requests} so that all replicas eventually have the same state.
 If the primary fails (e.g.\ if the node on which the primary is running crashes), Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive a full or partial copy (depending on whether persistent storage is used) of the state of the new primary in order to ``catch up'' with the other secondaries.
 Fabric provides a name-resolution service so that clients can always find the current primary.
-Fabric exposes several different APIs that developers can use to write services.
+
+Fabric services have a lot of asynchrony, which make them interesting targets for systematic testing with \psharp{}.
+Our primary goal was to create a \psharp{} model of Fabric to allow
+thorough testing of services, where Fabric's asynchrony is controlled 
+by the \psharp{} runtime.
+The system that we wished to test is \emph{CScale}~\cite{X},
+a big data-stream processing system that chains multiple Fabric services
+in a directed acyclic graph.
+Prior work~\cite{deligiannis2015psharp}
+created a model of Fabric with limited functionality;
+it used a mixture of \csharp{} and \psharp{} internally,
+only supported one in-flight replication request
+(which restricts the asynchrony that can be tested),
+and only supported one Fabric service.
+Our new Fabric model was re-written to use only \psharp{}
+internally,
+support an arbitrary number of Fabric services and in-flight replication requests,
+and in general be a more complete model of Fabric.
+Note that \csharp{} code is still required to interface with the existing
+\csharp{} service code.
+We refer to our \csharp{} code
+as the \emph{translation layer}
+and
+ the user-written \csharp{} service code as \emph{user code}.
+
+\begin{figure}[thb]
+\centering
+\includegraphics[width=\linewidth, clip=true, trim=0 180px 440px 0]{img/FabricModelOverview}
+\caption{Overview of the key machines (rounded boxes) and interfaces (boxes) in our Fabric model.}
+\label{fig:fabric_model}
+\end{figure}
+
+An overview of our Fabric model is shown in Figure~\ref{fig:fabric_model}.
+The \texttt{ClusterRuntime} machine 
+handles the creation and management of 
+one or more Fabric services,
+as well as service resolution requests
+which allows for client-service and inter-service communication
+within the model.
+Each Fabric service instance is managed by a \texttt{ServiceRuntime}
+machine, which in turn manages 
+several \texttt{ReplicaRuntime} machines.
+Each \texttt{ReplicaRuntime} communicates with the user code
+via several machines and interfaces from the translation layer
+(only the translation layer for the primary is shown,
+but every \texttt{ReplicaRuntime} has its own instance of the translation layer).
+Note that communication between machines is hierachical;
+thus, communication between \texttt{ReplicaRuntime}s
+(such as the sending of replication requests)
+is via the \texttt{ServiceRuntime} machine for that service.
+This approach does not necessarily reflect how Fabric works in practice.
+Instead, we chose an architecture
+that keeps the model simple
+while still allowing (what we believe to be) realistic
+asynchrony and failure scenarios.

### 2015-09-11 evaluation.tex
updated fabric figure to same style
update in evaluation

--- 2015-09-09 evaluation.tex
@@
 
-\setlength{\tabcolsep}{.52em}
+\setlength{\tabcolsep}{.72em}
 \begin{table*}[t]
@@
 \centering
-\begin{tabular}{l rrrr rrrr}
+\begin{tabular}{rl rrr rrr}
 \centering
@@
 \caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
-We report: time in seconds to find a bug (Time to bug); number of scheduling points when a bug was found (\#SP); percentage of buggy iterations (\%Buggy); and if a bug was found with a particular scheduler (Bug found?).}
+We report: time in seconds to find a bug (Time to Bug); number of scheduling points when a bug was found (\#SP); and if a bug was found with a particular scheduler (BF?).}
 \label{tab:testing}

### 2015-09-11 evaluation.tex
update in evaluation

@@
 \caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
-We report: time in seconds to find a bug (Time to Bug); number of scheduling points when a bug was found (\#SP); and if a bug was found with a particular scheduler (BF?).}
+We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
 \label{tab:testing}

### 2015-09-12 methodology.tex
resize figure
update in figure
update section 4

@@
 
-The testing target is the real Extent Manager in vNext, wrapped inside a \psharp machine in the testing harness.
+The testing target, the real Extent Manager in vNext, is wrapped inside a \psharp machine in the testing harness.
 
@@
 \centering
-\includegraphics[width=\linewidth]{img/extent_manager}
+\includegraphics[width=.9\linewidth]{img/extent_manager}
 \caption{Internal components of the Extent Manager in Azure Storage vNext.}
@@
 
-\subsubsection{Stubbing Network}
+\subsubsection{Intercepting Extent Manager Messages}
 \label{sec:method:model:dispath}
@@
 Extent Manager uses a network engine to send and receive messages from ENs.
-The testing harness mocks the original network engine in vNext and overrides the networking methods.
+The testing harness mocks the original network engine in vNext and overrides the networking interface.
 In this way, it intercepts all outbound messages and relays to the testing driver, which is responsible for dispatching the messages to destination ENs.
 As shown in Figure~\ref{fig:enginecode}, the mocked network engine captures outbound messages from Extent Manager and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to the testing driver.
+\CHComment{Conceivably, if it is desirable to emulate message loss, the mocked network engine could leverage the non-determinism support in \psharp and drop the messages non-deterministically (only if non-determinism is already described in the previous section).
+}
 
@@
   public override void SendMessage(Socket s, Message msg) {
-    // capture and relay Extent Manager messages
-    PSharpRuntime.Send(this.Environment, new MessageFromExtentManagerEvent(), s, msg);
+    // intercept and relay Extent Manager messages
+    PSharpRuntime.Send(this.TestingDriver, new MessageFromExtentManagerEvent(), s, msg);
   }
@@
 A mocked network engine replaces the real one in Extent Manager.
-It captures and relays all outbound messages from Extent Manager.
+It intercepts all outbound messages from Extent Manager and relays them to the testing driver.
 
@@
     extMgr.netEngine = new MockedNetEngine(); // mock network
-    extMgr.SkippingJournal = true; // skip persistence
-    extMgr.IsMockingTimer = true;	 // disable internal timer
+    extMgr.skippingJournal = true; // skip persistence
+    extMgr.isMockingTimer = true;	 // disable internal timer
   }
 
-  [OnEventDoAction(typeof(MessageFromExtentNodeEvent), nameof(DeliverMessageAction))]
+  [OnEvent(MessageFromExtentNode, DeliverExtentNodeMessage)]
+  void DeliverExtentNodeMessage()
+  {
+    var msg = (ExtentNodeMessage)this.Payload;
+    // relay messages from Extent Node to Extent Manager
+    extMgr.ProcessMessage(msg);
+  }
 
-  private void DeliverMessageAction()
+  [OnEvent(TimerTick), nameof(ProcessExtentRepair))]
+  void ProcessExtentRepair()
   {
-    var msg = (CommandMessage)this.Payload;
-    // relay messages from Extent Node to Extent Manager
-    extMgr.ProcessSingleMessage(msg);
+    // extent repair loop driven by external timer
+    extMgr.ProcessEvent(new ExtentRepairEvent());
   }
@@
 
-Incoming messages from ENs are delivered to the Extent Manager machine and trigger action.
-The specific action here relays the messages to the internal Extent Manager by invoking \texttt{extMgr.ProcessSingleMessage}.
-From the Extent Manager's perspective, it is unaware of EN mocking and behaves as if it is communicating with several real ENs.
+Messages from ENs do {\em not} go through the mocked network engine.
+They are delivered to the Extent Manager machine directly and trigger an action that invokes the messages on the internal Extent Manager with \texttt{extMgr.ProcessSingleMessage}.
+From Extent Manager's perspective, it is unaware of the testing harness and behaves as if running in a real distributed environment and communicating with real ENs.
 
-\subsection{Mocking Extent Node}
+It is crucial to delegate all non-determinism to \psharp.
+For that, all internal timers of Extent Manager are disabled.
+Instead, the EN expiration loop and the extent repair loop are triggered repeatedly by dedicated \psharp timers.
+Note that system correctness should {\em not} hinge on the frequency of any individual timer.
+Hence, \psharp has the complete freedom to schedule arbitrary interleaving between the timers.
+
+\subsection{Extent Node Machine}
 \label{sec:method:mock_en}
 
-The \texttt{ExtentNode} machine is a model (less than 150 lines of code) of the real Extent Node and is responsible for generating synchronization events and sending them to the \texttt{ExtentManager} to test the extent manager's repair logic.
+The \texttt{ExtentNode} machine is reduced to the simplest.
+It omits the details of real Extent Node, mocks only those logic required for the testing scenarios, and keeps the mocked logic simplest.
+The logic here includes: repairing an extent from its replica, generating sync reports periodically, and sending heartbeat messages periodically.
 
-briefly describe mocked repair logic.
+The testing harness leverages components in vNext whenever appropriate. Here, the machine re-uses \texttt{ExtentCenter}, an internal component of real Extent Node, for extent bookkeeping.
 
-briefly describe mocked sync logic
+In the mocked extent repair logic, the machine is instrumented by Extent Manager to repair an extent.
+It sends a repair request to a source Extent Node where a replica is stored.
+After receiving a \texttt{ExtentRepairResponse} from the source, it updates its internal \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
+
+In the mocked sync logic, the machine is triggered repeatedly by a \psharp timer.
+It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to Extent Manager with \texttt{PSharpRuntime.Send(...)}.
 
@@
 
+%\begin{figure}[t]
+%\begin{lstlisting}
+%// Mocking Extent Node in P#
+%class ExtentNodeMachine : Machine {
+  %// use real vNext component whenever appropriate
+  %private ExtentNode.ExtentCenter extCtr;
+%
+  %[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
+  %[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSync))]
+%
+  %// extent repair logic
+  %void ProcessRepairExtentRequest()
+  %{
+    %var source = GetRepairSourceMachine(this.Payload);
+    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
+  %}
+  %...
+%
+  %// extent node sync logic
+  %void ProcessExtentNodeSync()
+  %{
+    %var sync = new ExtentNodeSync(this.NodeID);
+    %extCtr.Extents.ForEach(ext =>
+      %{
+        %sync.ExtentRecords.Add(new ExtentRecord(ext));
+      %});
+    %PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
+  %}
+%}
+%\end{lstlisting}
+%\vspace{-2mm}
+%\caption{Mocked Extent Node.}
+%\label{fig:mocked_en}
+%%\vspace{-2mm}
+%\end{figure}
+
 \begin{figure}[t]
@@
 class ExtentNodeMachine : Machine {
-  // use real vNext component whenever appropriate
+  // leverage real vNext component whenever appropriate
   private ExtentNode.ExtentCenter extCtr;
 
-  [OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
-  [OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSync))]
-
   // extent repair logic
-  void ProcessRepairExtentRequest()
+  ...
+  [OnEvent(ExtentRepairResponse, ProcessRepairResponse)]
+  void ProcessRepairResponse()
   {
-    var source = GetRepairSourceMachine(this.Payload);
-    PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
+    // repair response from source replica
+    if (IsRepairSucceeded(this.Payload))
+    {
+      var rec = GetExtentRecord(this.Payload);
+      extCtr.AddOrUpdate(rec); // update Extent Center
     }
-  ...
+  }
 
   // extent node sync logic
+  [OnEvent(TimerTick, ProcessExtentNodeSync)]
   void ProcessExtentNodeSync()
   {
-    var sync = new ExtentNodeSync(this.NodeID);
-    extCtr.Extents.ForEach(ext =>
-      {
-        sync.ExtentRecords.Add(new ExtentRecord(ext));
-      });
+    var sync = extCtr.GetSyncReport(); // prepare sync report
     PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
@@
 
-%\subsection{Modeling the environment}
-\subsection{Test Harness}
-\label{sec:method:model}
+%\subsection{Testing Driver Machine}
+%\label{sec:method:driver}
 
-The environment of a distributed system might consist of other distributed systems and services, clients, operating system timers, as well as libraries for networking or other purposes.
-To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
+\subsection{Monitor Machine}
+\label{sec:method:monitor}
 
-%\subsubsection{Creating a test harness using \psharp}
-\subsubsection{Orchestrating Machines}
-\label{sec:method:model:harness}
+As illustrated in Figure~\ref{fig:monitor}, the monitor machine transitions between a cold state and a hot state.
+It is desirable to maintain a target number of replicas for an extent.
+Whenever an extent node fails, the monitor is notified and transitions into the hot {\em repairing} state, as the replica number now falls below the target.
+Once the extent is repaired, the monitor is again notified and transitions into the cold {\em repaired} state, when the replica number reaches the target.
+
+In the extent repair testing scenarios, it is expected that the monitor should {\em always eventually} end up in the cold repaired state.
+Otherwise, if the monitor is stuck in the hot repairing state for {\em infinitely} long, it is an indication that the corresponding execution sequence results in the extent never being repaired -- a liveness bug!
 
 \begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_engine}
-\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
-\label{fig:azurestoremodel}
+\begin{lstlisting}
+class ExtentRepairMonitor : Machine {
+  // true: with replica, false: w/o replica
+  private Dictionary<Machine, bool> ExtentNode2Replica;
+
+  [Cold] // cold state: repaired
+  [OnEvent(NotifyEnFailure, ProcessExtentNodeFailure)]
+  state Repaired : MonitorState {
+    void ProcessExtentNodeFailure()
+    {
+        var node = GetExtentNode(this.Payload);
+        ExtentNode2Replica.Remove(node);
+        goto Repairing;
+    }
+  }
+
+  [Hot] // hot state: repairing
+  [OnEvent(NotifyExtentRepaired ProcessExtentRepaired)]
+  state Repairing : MonitorState {
+    void ProcessExtentRepaired()
+    {
+      var node = GetExtentNode(this.Payload);
+      ExtentNode2Replica[node] = true;
+
+      if (ReplicaCount == Harness.REPLICATION_FACTOR)
+        goto Repaired;
+    }
+  }
+}
+\end{lstlisting}
+\vspace{-2mm}
+\caption{Monitor Machine.}
+\label{fig:monitor}
+%\vspace{-2mm}
 \end{figure}
 
-Before being able to use \psharp to test an existing distributed system, the environment of the components that the developer wants to test have to be modeled using the \psharp state machine APIs.
+%%\subsection{Modeling the environment}
+%\subsection{Test Harness}
+%\label{sec:method:model}
+%
+%The environment of a distributed system might consist of other distributed systems and services, clients, operating system timers, as well as libraries for networking or other purposes.
+To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
+%
+%%\subsubsection{Creating a test harness using \psharp}
+%\subsubsection{Orchestrating Machines}
+%\label{sec:method:model:harness}
+%
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=\linewidth]{img/mocked_vnext}
+%\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
+%\label{fig:azurestoremodel}
+%\end{figure}
+%
+%Before being able to use \psharp to test an existing distributed system, the environment of the components that the developer wants to test have to be modeled using the \psharp state machine APIs.
 This involves creating mock classes that inherit from the \texttt{Machine} abstract class.
@@
 machine states and state transitions) as well as methods for sending events to other machines, declaring event handlers for received events, and accessing the payload from the latest received event.
-
-In the Azure Storage vNext case study, we created the following \psharp machines: \texttt{Environment}, \texttt{ExtentNode}, \texttt{ExtentManager}, and \texttt{Timer}.
-
-The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute.
+%
+%In the Azure Storage vNext case study, we created the following \psharp machines: \texttt{Environment}, \texttt{ExtentNode}, \texttt{ExtentManager}, and \texttt{Timer}.
+%
+%The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute.
 The programmer is free to create either a finite test harness (e.g.
@@
 when to insert a failure) is captured using the available \psharp APIs.
-
-Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g.
+%
+%Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g.
 synchronization messages, heartbeats and repair logic loops).
 This is discussed in detail in Section~\ref{sec:method:model:timers}.
-
-\subsubsection{Modeling and injecting failures}
-\label{sec:method:model:failures}
-
-In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
+%
+%\subsubsection{Modeling and injecting failures}
+%\label{sec:method:model:failures}
+%
+%In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
 Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness.
@@
 The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
-
-\begin{figure}[t]
-\begin{lstlisting}
-// Real code for detecting node expiration
-public virtual bool IsNodeExpired(string id, DateTime time)
-{
-  return DateTime.Compare(time, DateTime.Now) <= 0;
-}
-
-// Mocked code for detecting node expiration
-public override bool IsNodeExpired(string id, DateTime time)
-{
-  return this.DeletedNodes.Contains(id);
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
-\label{fig:expiration}
+%
+%\begin{figure}[t]
+%\begin{lstlisting}
+%// Real code for detecting node expiration
+%public virtual bool IsNodeExpired(string id, DateTime time)
+%{
+  %return DateTime.Compare(time, DateTime.Now) <= 0;
+%}
+%
+%// Mocked code for detecting node expiration
+%public override bool IsNodeExpired(string id, DateTime time)
+%{
+  %return this.DeletedNodes.Contains(id);
+%}
+%\end{lstlisting}
 %\vspace{-2mm}
-\end{figure}
-
-Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
+%\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
+%\label{fig:expiration}
+%%\vspace{-2mm}
+%\end{figure}
+%
+%Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
 Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration.
@@
 The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
-
-\subsubsection{Entry point to a \psharp test harness}
-\label{sec:method:model:entrypoint}
-
-A developer can specify the entry point of a \psharp test harness similar to how unit tests are typically written.
+%
+%\subsubsection{Entry point to a \psharp test harness}
+%\label{sec:method:model:entrypoint}
+%
+%A developer can specify the entry point of a \psharp test harness similar to how unit tests are typically written.
 Figure~\ref{fig:entrypoint} shows the source code of the entry point for the Azure Storage vNext \psharp test harness.
@@
 In each iteration, the program state will be reset (any static fields must be explicitly reset by the programmer) and the entry point will be re-invoked.
-
-\begin{figure}[t]
-\begin{lstlisting}
-[Microsoft.PSharp.Test]
-public static void Execute()
-{
-  PSharpRuntime.CreateMachine(typeof(Environment));
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{Static \csharp method acting as the entry point to the Azure Storage vNext \psharp test harness.}
-\label{fig:entrypoint}
+%
+%\begin{figure}[t]
+%\begin{lstlisting}
+%[Microsoft.PSharp.Test]
+%public static void Execute()
+%{
+  %PSharpRuntime.CreateMachine(typeof(Environment));
+%}
+%\end{lstlisting}
 %\vspace{-2mm}
-\end{figure}
-
-\subsubsection{Abstracting timers}
-\label{sec:method:model:timers}
-
-Distributed systems are often using timers to determine when an event should be send from one component to another.
+%\caption{Static \csharp method acting as the entry point to the Azure Storage vNext \psharp test harness.}
+%\label{fig:entrypoint}
+%%\vspace{-2mm}
+%\end{figure}
+%
+%\subsubsection{Abstracting timers}
+%\label{sec:method:model:timers}
+%
+%Distributed systems are often using timers to determine when an event should be send from one component to another.
 For example, in the Azure Storage vNext system, each Extent Node is associated with a timer that fires of a synchronization message every 5 minutes and a heartbeat every 5 seconds.
@@
 Traditional testing techniques cannot easily find such a bug, due to the very infrequent occurrence of this race due to the timer.
-
-Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.
+%
+%Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.
 This nondeterminism is introduced using the \psharp \texttt{Nondet()} method, which returns a nondeterministic boolean value that is controlled by the \psharp runtime during systematic testing.
-
-Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study.
+%
+%Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study.
 The Extend Manager, as well as each Extent Node in the harness, is associated with a unique \texttt{Timer} machine.
@@
 Similarly, when the Extent Manager receives a \texttt{TimerTickEvent} from its own \texttt{Timer}, it handles it by nondeterministically invoking repair-related methods in the Extent Repair Center data structure.
-
-\begin{figure}[t]
-\begin{lstlisting}
-internal class Timer : Machine
-{
-  MachineId Owner; // Id of the owner machine
-
-  [Start]
-  [OnEntry(nameof(InitOnEntryAction))]
-  [OnEventGotoState(typeof(Unit), typeof(Active))]
-  class Init : MachineState { }
-
-  void InitOnEntryAction()
-  {
-    this.Owner = (MachineId)this.Payload;
-    // triggers state transition to Active
-    this.Raise(new Unit());
-  }
-
-  [OnEntry(nameof(ProcessTickEvent))]
-  [OnEventGotoState(typeof(Unit), typeof(Active))]
-  class Active : MachineState { }
-
-  void ProcessTickEvent()
-  {
-    // Nondeterministic boolean choice controlled by P#
-    if (this.Nondet())
-      // sends a timer tick event to the owner machine
-      this.Send(this.Owner, new TimerTickEvent());
-    // triggers state transition to Active
-    this.Raise(new Unit());
-  }
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{Timers in Azure Storage vNext are modeled as nondeterministic \psharp machines.}
-\label{fig:timer}
+%
+%\begin{figure}[t]
+%\begin{lstlisting}
+%internal class Timer : Machine
+%{
+  %MachineId Owner; // Id of the owner machine
+%
+  %[Start]
+  %[OnEntry(nameof(InitOnEntryAction))]
+  %[OnEventGotoState(typeof(Unit), typeof(Active))]
+  %class Init : MachineState { }
+%
+  %void InitOnEntryAction()
+  %{
+    %this.Owner = (MachineId)this.Payload;
+    %// triggers state transition to Active
+    %this.Raise(new Unit());
+  %}
+%
+  %[OnEntry(nameof(ProcessTickEvent))]
+  %[OnEventGotoState(typeof(Unit), typeof(Active))]
+  %class Active : MachineState { }
+%
+  %void ProcessTickEvent()
+  %{
+    %// Nondeterministic boolean choice controlled by P#
+    %if (this.Nondet())
+      %// sends a timer tick event to the owner machine
+      %this.Send(this.Owner, new TimerTickEvent());
+    %// triggers state transition to Active
+    %this.Raise(new Unit());
+  %}
+%}
+%\end{lstlisting}
 %\vspace{-2mm}
-\end{figure}
+%\caption{Timers in Azure Storage vNext are modeled as nondeterministic \psharp machines.}
+%\label{fig:timer}
+%%\vspace{-2mm}
+%\end{figure}

### 2015-09-12 experience.tex
Small change.
updated fabric figure to same style
Explain the use of generic and bug-specific test cases for MigratingTable in the evaluation section.
Also, better address the merits of generic and specific test cases in
the experience section, and associated clarifications.

@@
 \subsubsection{Input generation}
+\label{sec:mtable:input}
 
-Of course, there are many possible input histories to test.
-Since we originally hoped for a comprehensive form of verification and we didn't feel we had good intuition to prepare a set of specific test cases that would make us confident of having caught any and all concurrency bugs in the protocol, it was natural for us to sample from a distribution of input histories that we defined to exercise all the features of IChainTable2 within certain bounds.
-Furthermore, it was natural to let \psharp control the choice of input history as well as the machine interleaving so we could reproduce both using a single random seed.
+% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
+Of course, there are many possible input histories for MigratingTable.
+We could have tested specific input histories, but we were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for interactions between different parts of the code that would be difficult for us to foresee.
+If we had no formalization of the specification and had to rely on expected outputs worked out by hand, this might be the best we could do.
+However, since the IChainTable2 specification is relatively simple and is almost deterministic under sequential calls, it was straightforward to write an in-memory reference implementation called SpecTable to which we can compare the output of MigratingTable on an arbitrary input history.
+This gave us the attractive option to sample from a distribution we defined over all possible input histories within certain bounds.
+It was convenient to let \psharp control the choice of input history as well as the schedule so we could reproduce both using a single random seed.
+Then, under the ``small scope hypothesis'' that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
 
@@
 
-% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-To comprehensively verify the behavior of MigratingTable under an arbitrary input history, we needed some formulation of the IChainTable2 specification to which to compare it.
-Since the specification is deterministic under sequential calls except for the results of streaming reads, we decided the easiest approach was to write an in-memory reference implementation called SpecTable.
-Given a streaming read call, SpecTable can produce a set of all results that are compliant with the specification.
+As mentioned above, the IChainTable2 specification is almost deterministic under sequential calls; the only nondeterminism is in the results of streaming reads.
+Given a streaming read, SpecTable can compute the set of all results that are compliant with the specification, so we can simply check if the result of MigratingTable is in this set.
+To test MigratingTable, we must supply it with backing tables.
+We use SpecTable for this purpose as well, with \psharp choosing the actual result of each streaming read from the valid set.
 Our correctness property is then:
@@
 \begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of \term{old} and \term{new} SpecTables (where \psharp chooses the actual result of each streaming read from the valid set) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+For every execution trace of a collection of MigratingTables backed by the same pair of \term{old} and \term{new} SpecTables in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
 \end{quote}
@@
 \centering
-\includegraphics[width=\linewidth, clip=true, trim=0 180px 440px 0]{img/FabricModelOverview}
+\includegraphics[width=\linewidth]{img/fabricmodel}
 \caption{Overview of the key machines (rounded boxes) and interfaces (boxes) in our Fabric model.}

### 2015-09-12 methodology.tex
update section 4

@@
 
-The testing harness consists of multiple \psharp machines. 
-\begin{enumerate}
-\item There is a {\em Extent Manager} machine that is a thin wrapper for the real Extent Manager and several {\em Extent Node} machines that represent simple mocked ENs.
+The testing harness consists of multiple \psharp machines:
+%\begin{enumerate}
+\begin{description}
+\item[Extent Manager machine] is a thin wrapper for the real Extent Manager in vNext.
 
-\item In addition, there is a {\em Testing Driver} machine that communicates with all the machines, relays messages between Extent Manager and EN machines, and drives testing scenarios.
+\item[Extent Node machines] represent simple mocked ENs.
+
+\item[Testing Driver machine] communicates with all other machines, relays messages between the Extent Manager machine and the EN machines, and drives testing scenarios.
+In one scenario, the testing driver launches one Extent Manager and three ENs, with a single extent on one of the ENs.
+It then waits for the extent to be replicated at the other ENs.
+In another scenario, the testing driver fails one of the ENs and launches a new one.
+It then waits for the extent to be repaired on the new EN.
+
+\item[Monitor machine] collects states from all the ENs and verifies desired properties, such as the extent is replicated or repaired at the expected ENs.
+\end{description}
+%\item {\em Extent Manager machine} is a thin wrapper for the real Extent Manager in vNext.
+%
+%\item {\em Extent Node machine} represents simple mocked EN.
+%
+%\item {\em Testing Driver machine} communicates with all other machines, relays messages between the Extent Manager machine and multiple EN machines, and drives testing scenarios.
 In one testing scenario, the testing driver launches Extent Manager with three ENs, with a single extent on one of the ENs.
@@
 It then waits for the extent to be repaired on all the remaining ENs.
-
-\item Moreover, there is a {\em Monitor} machine, which collects states from all the ENs and verifies the desired property that the extent is {\em eventually} repaired.
-\end{enumerate}
+%
+%\item Moreover, there is a {\em Monitor} machine, which collects states from all the ENs and verifies the desired property that the extent is {\em eventually} repaired.
+%\end{enumerate}
 
@@
 
-Inside Extent Manager, there are two structures related to extent repair: an \emph{Extent Center} and an EN map.
-The Extent Center maintains mapping records from extents to their hosting ENs.
-It is updated upon every sync report from ENs.
+Inside Extent Manager, there are two data structures related to extent replication and repair: an \texttt{ExtentCenter} and an \texttt{ExtentNode Map}.
+The extent center maintains the mapping records from extents to their hosting ENs.
+It is updated upon the periodic sync reports from the ENs.
+Again recall that the sync report from a particular EN lists all the extents stored at the EN.
+Its purpose is to update Extent Manager's possible out-of-date view of the EN with the ground truth.
 The EN map records the latest heartbeat time from every EN.
-It internally runs an EN expiration loop that is responsible for removing ENs that have been missing heartbeats for extended period and also cleaning up corresponding Extent Center records.
-In addition, Extent Manager runs an extent repair loop, which examines all the Extent Center records, identifies extents with missing replica, schedules extent repair tasks and sends the repair tasks to ENs.
+
+Extent Manager internally runs a periodic {\em EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period, as well as cleaning up the corresponding records in the extent center.
+In addition, it runs a periodic {\em extent repair loop}, which examines all the extent center records, identifies extents with missing replica, schedules extent repair tasks and sends them to the ENs.
 
@@
 
-Extent Manager uses a network engine to send and receive messages from ENs.
-The testing harness mocks the original network engine in vNext and overrides the networking interface.
-In this way, it intercepts all outbound messages and relays to the testing driver, which is responsible for dispatching the messages to destination ENs.
-As shown in Figure~\ref{fig:enginecode}, the mocked network engine captures outbound messages from Extent Manager and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to the testing driver.
-\CHComment{Conceivably, if it is desirable to emulate message loss, the mocked network engine could leverage the non-determinism support in \psharp and drop the messages non-deterministically (only if non-determinism is already described in the previous section).
-}
+Extent Manager uses a network engine to send messages to the ENs.
+The testing harness mocks the original network engine in vNext and overrides its interface.
+In this way, the mocked network engine intercepts all outbound messages and relays them to the testing driver, which is responsible for dispatching the messages to the destination ENs.
+As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from Extent Manager and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to the testing driver.
+Conceivably, the mocked network engine could leverage the non-determinism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.
+\CHComment{only if non-determinism is already covered in the previous section.}
 
@@
 \begin{lstlisting}
-// network interface in vNext codebase
+// network interface in vNext
 class NetworkEngine {
@@
 
-// mocked network for capturing Extent Manager messages
+// mocked engine for intercepting Extent Manager messages
 class MockedNetEngine : NetworkEngine {
@@
 
-Figure~\ref{fig:wrap_target} shows code snippet from Extent Manager machine, as a thin wrapper of the real Extent Manager.
-A mocked network engine replaces the real one in Extent Manager.
-It intercepts all outbound messages from Extent Manager and relays them to the testing driver.
+Figure~\ref{fig:wrap_target} shows some code snippets from the Extent Manager machine.
+It is a thin wrapper of the real Extent Manager.
+The mocked network engine replaces the real one in Extent Manager, which intercepts all the outbound messages from Extent Manager and relays them to the testing driver.
 
@@
     extMgr.netEngine = new MockedNetEngine(); // mock network
-    extMgr.skippingJournal = true; // skip persistence
     extMgr.isMockingTimer = true;	 // disable internal timer
@@
 \vspace{-2mm}
-\caption{Wrapping vNext Component in \psharp.}
+\caption{Extent Manager Machine: the Testing Target.}
 \label{fig:wrap_target}
@@
 
-Messages from ENs do {\em not} go through the mocked network engine.
+Messages coming from the ENs do {\em not} go through the mocked network engine.
 They are delivered to the Extent Manager machine directly and trigger an action that invokes the messages on the internal Extent Manager with \texttt{extMgr.ProcessSingleMessage}.
@@
 
-It is crucial to delegate all non-determinism to \psharp.
-For that, all internal timers of Extent Manager are disabled.
-Instead, the EN expiration loop and the extent repair loop are triggered repeatedly by dedicated \psharp timers.
+All internal timers of Extent Manager are disabled, because it is crucial to delegate all non-determinism to \psharp.
+Instead, the EN expiration loop and the extent repair loop are driven by external timers provided by \psharp.
 Note that system correctness should {\em not} hinge on the frequency of any individual timer.
@@
 
-The \texttt{ExtentNode} machine is reduced to the simplest.
-It omits the details of real Extent Node, mocks only those logic required for the testing scenarios, and keeps the mocked logic simplest.
+The Extent Node machine is reduced to the simplest.
+It omits the details of a real Extent Node, mocks only those logic necessary for the testing scenarios, and keeps the mocked logic simplest.
 The logic here includes: repairing an extent from its replica, generating sync reports periodically, and sending heartbeat messages periodically.
 
-The testing harness leverages components in vNext whenever appropriate. Here, the machine re-uses \texttt{ExtentCenter}, an internal component of real Extent Node, for extent bookkeeping.
+The testing harness leverages components in vNext whenever appropriate. Here, the Extent Node machine re-uses \texttt{ExtentCenter}, an internal component of real Extent Node, for extent bookkeeping.
 
-In the mocked extent repair logic, the machine is instrumented by Extent Manager to repair an extent.
-It sends a repair request to a source Extent Node where a replica is stored.
-After receiving a \texttt{ExtentRepairResponse} from the source, it updates its internal \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
+In the mocked extent repair logic, the machine takes action upon receiving an extent repair request from Extent Manager.
+It sends a copy request to a source Extent Node where a replica is stored.
+After receiving a \texttt{ExtentCopyResponse} from the source, it updates the internal extent center, as illustrated in Figure~\ref{fig:mocked_en}.
 
-In the mocked sync logic, the machine is triggered repeatedly by a \psharp timer.
+In the mocked extent node sync logic, the machine is again driven by an external timer provided by \psharp.
 It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to Extent Manager with \texttt{PSharpRuntime.Send(...)}.
@@
   ...
-  [OnEvent(ExtentRepairResponse, ProcessRepairResponse)]
-  void ProcessRepairResponse()
+  [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
+  void ProcessExtentCopyResponse()
   {
-    // repair response from source replica
-    if (IsRepairSucceeded(this.Payload))
+    // extent copy response from source replica
+    if (IsCopySucceeded(this.Payload))
     {
@@
 
-As illustrated in Figure~\ref{fig:monitor}, the monitor machine transitions between a cold state and a hot state.
-It is desirable to maintain a target number of replicas for an extent.
-Whenever an extent node fails, the monitor is notified and transitions into the hot {\em repairing} state, as the replica number now falls below the target.
-Once the extent is repaired, the monitor is again notified and transitions into the cold {\em repaired} state, when the replica number reaches the target.
+The monitor machine transitions between a cold state and a hot state.
+Whenever an extent node fails, the monitor is notified.
+As soon as the number of extent replica falls below a specified target, the monitor transitions into the hot {\em repairing} state, where the missing replica is being repaired.
+Whenever a replica is repaired, the monitor is also notified.
+It transitions into the cold {\em repaired} state, when the replica number reaches the target again, as illustrated in Figure~\ref{fig:monitor}.
 
-In the extent repair testing scenarios, it is expected that the monitor should {\em always eventually} end up in the cold repaired state.
-Otherwise, if the monitor is stuck in the hot repairing state for {\em infinitely} long, it is an indication that the corresponding execution sequence results in the extent never being repaired -- a liveness bug!
+%It verifies the property that the number of extent replica {\em eventually} reaches the target.
+
+In the extent repair testing scenarios, the monitor verifies that it should {\em always eventually} end up in the cold repaired state.
+Otherwise, the monitor is stuck in the hot repairing state for {\em infinitely} long.
+This indicates that the corresponding execution sequence results in extent replica never being repaired -- a liveness bug!
 
@@
 class ExtentRepairMonitor : Machine {
-  // true: with replica, false: w/o replica
+  // true: EN has replica, false: EN has no replica
   private Dictionary<Machine, bool> ExtentNode2Replica;
@@
   [Hot] // hot state: repairing
-  [OnEvent(NotifyExtentRepaired ProcessExtentRepaired)]
+  [OnEvent(NotifyExtentRepaired, ProcessExtentRepaired)]
   state Repairing : MonitorState {
@@
 
-      if (ReplicaCount == Harness.REPLICATION_FACTOR)
+      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
         goto Repaired;

### 2015-09-12 evaluation.tex

@@
-We compared the bug-finding effectiveness of \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control on several benchmark bugs, described below.
-We report the length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
+After we fixed all the bugs we could find in the programs under test, we added options to the code to introduce various benchmark bugs (some that were actually in the original programs and some we made up) one at a time so we could evaluate different methods to detect and diagnose them.
+We compare \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control; we report the average length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
 \MMComment{Overview of actual results?}
@@
 We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-
-Because we fixed all bugs that we discovered using the \psharp systematic testing engine in the case studies, for the evaluation we added an option to conditionally reintroduce each one of the bugs so that we can measure how fast and how often each bug can be discovered with the \psharp random and the PCT scheduler.
 
@@
 
+For MigratingTable, the upper section of the table uses the random input generator described in Section~\ref{sec:mtable:input}.
+For each bug that led to at least one test failure, we manually reviewed one of the failure traces to confirm it reflected the intended bug.
+For each of the remaining bugs, we repeated the test using a test case custom written to trigger that particular bug in order to confirm that the failure to detect the bug in the original test was due to unlucky random choices of inputs and schedules and not some other problem with the experimental setup.
+These results are in the lower section of the table; they can serve as additional cases in which to compare the random and PCT schedulers but do not represent a testing method one could use to find unknown bugs in software.
+It may have been interesting to report results for specific test cases written before we knew the bugs as mentioned in \ref{sec:mtable:input}, but we did not do so in this study.
+
 How long you had to run the test to find the bug (if you can find it) and how long are the traces

### 2015-09-12 experience.tex

@@
 
+% It might be nice to include excerpts of a trace like in migration-bug3-explanation.pptx.
+Unfortunately, the trace in migration-bug3-explanation.pptx doesn't match my recollection of the original diagnosis, which I want to write about truthfully (the former looks like it involves a stale streaming read, while I'm fairly sure the latter involved only the $k_O < k < k_N$ case).
+If it's important, I could try to get a new trace consistent with the original diagnosis.
+~ Matt
+This bug took us only about 10 minutes to diagnose from the trace; granted, this is after we had days of experience analyzing MigratingTable traces and had added our own trace output to the test harness, since \psharp's built-in trace output is too low-level and does not include event payloads and other diagnostic data that is not passed between machines.
+We started from the failure symptom: the virtual stream returned end-of-stream when according to the reference SpecTable, it should have returned an additional row with key 4.
+We filtered the trace for actions by the same service machine and saw that $s_O$ was closed before the virtual stream had returned the row with key 4, but $s_N$ had already advanced past key 4 before the migrator inserted the row in the new table.
+At first we found this phenomenon hard to believe, but soon we were convinced it reflected a gap in our design.
+
 \MMComment{Add a figure based on migration-bug3-explanation.pptx (probably a hybrid of slides 4 and 5, assuming we want only one figure).}

### 2015-09-12 methodology.tex
update section 4

@@
 
-The testing harness consists of multiple \psharp machines:
+The testing harness consists of multiple \psharp machines, as shown in Figure~\ref{fig:azurestoremodel}:
 %\begin{enumerate}
@@
 \end{description}
-%\item {\em Extent Manager machine} is a thin wrapper for the real Extent Manager in vNext.
-%
-%\item {\em Extent Node machine} represents simple mocked EN.
-%
-%\item {\em Testing Driver machine} communicates with all other machines, relays messages between the Extent Manager machine and multiple EN machines, and drives testing scenarios.
-In one testing scenario, the testing driver launches Extent Manager with three ENs, with a single extent on one of the ENs.
-It then waits for the extent to be repaired on all three ENs.
-In another testing scenario, the testing driver fails one of the ENs and launches a new EN.
-It then waits for the extent to be repaired on all the remaining ENs.
-%
-%\item Moreover, there is a {\em Monitor} machine, which collects states from all the ENs and verifies the desired property that the extent is {\em eventually} repaired.
-%\end{enumerate}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_vnext}
+\caption{Real Extent Manager with a Mocked Environment (each box presents one \psharp machine).}
+\label{fig:azurestoremodel}
+\end{figure}
 
@@
 
-%\subsection{Testing Driver Machine}
-%\label{sec:method:driver}
+\subsection{Testing Driver Machine}
+\label{sec:method:driver}
+
+The testing driver machine drives two testing scenarios.
+In one scenario, it launches one Extent Manager and three ENs, with a single extent on one of the ENs.
+It waits for the extent to be replicated at the other ENs.
+In another scenario, it fails one of the ENs and launches a new one.
+It waits for the extent to be repaired on the new EN.
+Code snippets are skipped due to space constraint.
 

### 2015-09-13 methodology.tex
updates in Section 4

@@
-To uncover the elusive extent repair bug in Azure Storage vNext, a testing harness is written in \psharp.
-The developers of vNext expect the bug more likely in Extent Manager than in Extent Node.
-Hence, the testing harness focuses on the real Extent Manager with mocked Extent Nodes.
+To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a testing harness in \psharp.
+The developers of vNext expected that it was more likely for the bug to occur in the ExtMgr logic, rather than the EN logic.
+Hence, the testing harness focuses on the real ExtMgr using mocked ENs.
 
-The testing harness consists of multiple \psharp machines, as shown in Figure~\ref{fig:azurestoremodel}:
+The testing harness consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
 %\begin{enumerate}
 \begin{description}
-\item[Extent Manager machine] is a thin wrapper for the real Extent Manager in vNext.
+\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext.
 
-\item[Extent Node machines] represent simple mocked ENs.
+\item[ExtentNode] represents a simplified mocked EN.
 
-\item[Testing Driver machine] communicates with all other machines, relays messages between the Extent Manager machine and the EN machines, and drives testing scenarios.
-In one scenario, the testing driver launches one Extent Manager and three ENs, with a single extent on one of the ENs.
-It then waits for the extent to be replicated at the other ENs.
-In another scenario, the testing driver fails one of the ENs and launches a new one.
-It then waits for the extent to be repaired on the new EN.
+\item[TestingDriver] communicates with all other machines, relays messages between the \texttt{ExtentManager} and the \texttt{ExtentNode} machines, and is responsible for driving testing scenarios.
 
-\item[Monitor machine] collects states from all the ENs and verifies desired properties, such as the extent is replicated or repaired at the expected ENs.
+\item[RepairMonitor] collects states from all \texttt{ExtentNode} machines in the testing harness and verifies desired properties, such as that an extent is replicated or repaired at the expected ENs.
 \end{description}
@@
 \includegraphics[width=\linewidth]{img/mocked_vnext}
-\caption{Real Extent Manager with a Mocked Environment (each box presents one \psharp machine).}
+\caption{Real Extent Manager with a mocked environment (each box represents one \psharp machine).}
 \label{fig:azurestoremodel}
@@
 
-\subsection{Extent Manager Machine}
+\subsection{The ExtentManager machine}
 \label{sec:method:wrap_target}
 
-The testing target, the real Extent Manager in vNext, is wrapped inside a \psharp machine in the testing harness.
+The component under test, the real ExtMgr in vNext, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
 
-\subsubsection{Extent Manager Internals}
+\subsubsection{Internals of the real Extent Manager}
 
@@
 \includegraphics[width=.9\linewidth]{img/extent_manager}
-\caption{Internal components of the Extent Manager in Azure Storage vNext.}
+\caption{Internal components of the real Extent Manager in Azure Storage vNext.}
 \label{fig:extentmanager}
@@
 
-Inside Extent Manager, there are two data structures related to extent replication and repair: an \texttt{ExtentCenter} and an \texttt{ExtentNode Map}.
-The extent center maintains the mapping records from extents to their hosting ENs.
+Inside the real ExtMgr, there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
+The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs.
 It is updated upon the periodic sync reports from the ENs.
-Again recall that the sync report from a particular EN lists all the extents stored at the EN.
-Its purpose is to update Extent Manager's possible out-of-date view of the EN with the ground truth.
+Recall that the sync report from a particular EN lists all the extents stored at the EN.
+Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth.
 The EN map records the latest heartbeat time from every EN.
 
-Extent Manager internally runs a periodic {\em EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period, as well as cleaning up the corresponding records in the extent center.
-In addition, it runs a periodic {\em extent repair loop}, which examines all the extent center records, identifies extents with missing replica, schedules extent repair tasks and sends them to the ENs.
+The ExtMgr runs internally a periodic {\em EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period, as well as cleaning up the corresponding records in the \texttt{ExtentCenter}.
+In addition, the ExtMgr runs a periodic {\em extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.
 
-\subsubsection{Intercepting Extent Manager Messages}
+\subsubsection{Intercepting messages between machines}
 \label{sec:method:model:dispath}
@@
 
-Extent Manager uses a network engine to send messages to the ENs.
+The real ExtMgr uses a network engine to send messages to the ENs.
 The testing harness mocks the original network engine in vNext and overrides its interface.
-In this way, the mocked network engine intercepts all outbound messages and relays them to the testing driver, which is responsible for dispatching the messages to the destination ENs.
-As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from Extent Manager and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to the testing driver.
+In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
+As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
 Conceivably, the mocked network engine could leverage the non-determinism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.
@@
 \vspace{-2mm}
-\caption{Mocked Network Engine.}
+\caption{Mocked network engine in vNext.}
 \label{fig:enginecode}
@@
 
-\subsubsection{Extent Manager Machine}
+\subsubsection{Specifics of the ExtentManager machine}
 
-Figure~\ref{fig:wrap_target} shows some code snippets from the Extent Manager machine.
-It is a thin wrapper of the real Extent Manager.
-The mocked network engine replaces the real one in Extent Manager, which intercepts all the outbound messages from Extent Manager and relays them to the testing driver.
+Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine.
+\texttt{ExtentManager} is a thin wrapper of the real ExtMgr.
+The mocked network engine replaces the real one in ExtMgr, intercepts all the outbound messages from ExtMgr and relays them to \texttt{TestingDriver}.
 
@@
 \begin{lstlisting}
-// Wrapping the target vNext component in P#
+// Wrapping the target vNext component in a P# machine
 class ExtentManagerMachine : Machine {
@@
 
-  void Init()
-  {
+  void Init() {
     extMgr = new ExtentManager();
-
     extMgr.netEngine = new MockedNetEngine(); // mock network
@@
   [OnEvent(MessageFromExtentNode, DeliverExtentNodeMessage)]
-  void DeliverExtentNodeMessage()
-  {
+  void DeliverExtentNodeMessage() {
     var msg = (ExtentNodeMessage)this.Payload;
@@
   [OnEvent(TimerTick), nameof(ProcessExtentRepair))]
-  void ProcessExtentRepair()
-  {
+  void ProcessExtentRepair() {
     // extent repair loop driven by external timer
@@
 \vspace{-2mm}
-\caption{Extent Manager Machine: the Testing Target.}
+\caption{Component under test: the real Extent Manager wrapped inside an \texttt{ExtentManager} \psharp machine.}
 \label{fig:wrap_target}
@@
 Messages coming from the ENs do {\em not} go through the mocked network engine.
-They are delivered to the Extent Manager machine directly and trigger an action that invokes the messages on the internal Extent Manager with \texttt{extMgr.ProcessSingleMessage}.
-From Extent Manager's perspective, it is unaware of the testing harness and behaves as if running in a real distributed environment and communicating with real ENs.
+They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}.
+The benefit of this approach is that ExtMgr is unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.
 
-All internal timers of Extent Manager are disabled, because it is crucial to delegate all non-determinism to \psharp.
+All internal timers of ExtMgr are disabled, because it is crucial to delegate all non-determinism to \psharp.
 Instead, the EN expiration loop and the extent repair loop are driven by external timers provided by \psharp.
 Note that system correctness should {\em not} hinge on the frequency of any individual timer.
-Hence, \psharp has the complete freedom to schedule arbitrary interleaving between the timers.
+Hence, \psharp has the complete freedom to schedule arbitrary interleavings between the timers.
 
-\subsection{Extent Node Machine}
+\subsection{The ExtentNode machine}
 \label{sec:method:mock_en}
 
-The Extent Node machine is reduced to the simplest.
-It omits the details of a real Extent Node, mocks only those logic necessary for the testing scenarios, and keeps the mocked logic simplest.
+The \texttt{ExtentNode} machine is a much simplified version of the original EN.
+It omits the details of a real EN, mocks only logic necessary for the testing scenarios, and keeps the mocked logic as simple as possible.
 The logic here includes: repairing an extent from its replica, generating sync reports periodically, and sending heartbeat messages periodically.
 
-The testing harness leverages components in vNext whenever appropriate. Here, the Extent Node machine re-uses \texttt{ExtentCenter}, an internal component of real Extent Node, for extent bookkeeping.
+The testing harness leverages components in vNext whenever appropriate.
+Here, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, an internal component of a real EN used for extent bookkeeping.
 
-In the mocked extent repair logic, the machine takes action upon receiving an extent repair request from Extent Manager.
-It sends a copy request to a source Extent Node where a replica is stored.
-After receiving a \texttt{ExtentCopyResponse} from the source, it updates the internal extent center, as illustrated in Figure~\ref{fig:mocked_en}.
+In the mocked extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine.
+It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored.
+After receiving an \texttt{ExtentCopyResponse} event from the source, it updates the internal \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
 
-In the mocked extent node sync logic, the machine is again driven by an external timer provided by \psharp.
-It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to Extent Manager with \texttt{PSharpRuntime.Send(...)}.
+In the mocked EN sync logic, the machine is again driven by an external timer provided by \psharp.
+It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to \texttt{ExtentManager} with \texttt{PSharpRuntime.Send(...)}.
 
@@
   [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
-  void ProcessExtentCopyResponse()
-  {
+  void ProcessExtentCopyResponse() {
     // extent copy response from source replica
-    if (IsCopySucceeded(this.Payload))
-    {
+    if (IsCopySucceeded(this.Payload)) {
       var rec = GetExtentRecord(this.Payload);
@@
   [OnEvent(TimerTick, ProcessExtentNodeSync)]
-  void ProcessExtentNodeSync()
-  {
+  void ProcessExtentNodeSync() {
     var sync = extCtr.GetSyncReport(); // prepare sync report
@@
 \vspace{-2mm}
-\caption{Mocked Extent Node.}
+\caption{The mocked EN in vNext.}
 \label{fig:mocked_en}
@@
 
-\subsection{Testing Driver Machine}
+\subsection{The TestingDriver machine}
 \label{sec:method:driver}
 
-The testing driver machine drives two testing scenarios.
-In one scenario, it launches one Extent Manager and three ENs, with a single extent on one of the ENs.
+The \texttt{TestingDriver} machine drives two testing scenarios.
+In the first scenario, it launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs.
 It waits for the extent to be replicated at the other ENs.
-In another scenario, it fails one of the ENs and launches a new one.
+In the second scenario, it fails one of the \texttt{ExtentNode} machines and launches a new one.
 It waits for the extent to be repaired on the new EN.
@@
 
-\subsection{Monitor Machine}
+\subsection{The RepairMonitor machine}
 \label{sec:method:monitor}
 
-The monitor machine transitions between a cold state and a hot state.
-Whenever an extent node fails, the monitor is notified.
-As soon as the number of extent replica falls below a specified target, the monitor transitions into the hot {\em repairing} state, where the missing replica is being repaired.
-Whenever a replica is repaired, the monitor is also notified.
-It transitions into the cold {\em repaired} state, when the replica number reaches the target again, as illustrated in Figure~\ref{fig:monitor}.
+The \texttt{RepairMonitor} machine transitions between a cold state and a hot state.
+Whenever an extent node fails, \texttt{RepairMonitor} is notified.
+As soon as the number of extent replicas falls below a specified target, \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
+Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified.
+It transitions into the cold \emph{repaired} state, when the replica number reaches the target again, as illustrated in Figure~\ref{fig:monitor}.
 
@@
 
-In the extent repair testing scenarios, the monitor verifies that it should {\em always eventually} end up in the cold repaired state.
-Otherwise, the monitor is stuck in the hot repairing state for {\em infinitely} long.
-This indicates that the corresponding execution sequence results in extent replica never being repaired -- a liveness bug!
+In the extent repair testing scenarios, \texttt{RepairMonitor} verifies that it should {\em always eventually} end up in the cold repaired state.
+Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long.
+This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug!
 
@@
 \begin{lstlisting}
-class ExtentRepairMonitor : Machine {
+class RepairMonitor : Machine {
   // true: EN has replica, false: EN has no replica
@@
 \vspace{-2mm}
-\caption{Monitor Machine.}
+\caption{The RepairMonitor machine.}
 \label{fig:monitor}

### 2015-09-13 motivation.tex
updates in Section 4

@@
 Similar to the current system, vNext employs containers, called {\em extents}, to store data.
-Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em extent nodes} (ENs).
+Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
 However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely \emph{distributed mapping}.
-In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em extent manager} (ExtMgr).
+In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager} (ExtMgr).
 

### 2015-09-13 methodology.tex
updates to section 4

@@
     // intercept and relay Extent Manager messages
-    PSharpRuntime.Send(this.TestingDriver, new MessageFromExtentManagerEvent(), s, msg);
+    PSharpRuntime.Send(this.TestingDriver,
+      new MessageFromExtentManagerEvent(), s, msg);
   }
@@
     var sync = extCtr.GetSyncReport(); // prepare sync report
-    PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
+    PSharpRuntime.Send(this.ExtentManagerMachine,
+      new MessageFromExtentNodeEvent(), sync);
   }
@@
 
-  [Cold] // cold state: repaired
+  // cold state: repaired
   [OnEvent(NotifyEnFailure, ProcessExtentNodeFailure)]
-  state Repaired : MonitorState {
-    void ProcessExtentNodeFailure()
-    {
+  cold state Repaired : MonitorState {
+    void ProcessExtentNodeFailure() {
         var node = GetExtentNode(this.Payload);
@@
 
-  [Hot] // hot state: repairing
+  // hot state: repairing
   [OnEvent(NotifyExtentRepaired, ProcessExtentRepaired)]
-  state Repairing : MonitorState {
-    void ProcessExtentRepaired()
-    {
+  hot state Repairing : MonitorState {
+    void ProcessExtentRepaired() {
       var node = GetExtentNode(this.Payload);
       ExtentNode2Replica[node] = true;
-
       if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)

### 2015-09-13 paper.tex
added a section on exemplary case, partially through methodology
updates
update draft.pdf
some updates all around
various updates
Small change.
updates to section 4

@@
 \usepackage{color}
+\definecolor{light-gray}{gray}{0.40}
 \newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
@@
 \newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
+\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
+\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}
 
@@
 escapeinside={(*@}{@*)},
+morekeywords={OnEvent, nameof, var, state, cold, hot},
 commentstyle=\color{greencomments},
@@
 
+\section{Motivating Example}
+\label{sec:motivation}
+
+\input{motivation}
+
 \section{Testing Distributed Systems}
-\label{sec:bg}
+\label{sec:overview}
 
-\input{background}
+\input{overview}
 
-\section{Case Studies in Microsoft}
-\label{sec:cases}
-
-\input{casestudies}
-
-\section{Testing Production Systems with \psharp}
+\section{Testing Azure Storage vNext with \psharp}
 \label{sec:method}
@@
 
-\section{The \psharp Bug-Finding Runtime}
+\section{The \psharp Systematic Testing Engine}
 \label{sec:psharp}

### 2015-09-13 methodology.tex
updates to section 4

@@
   [OnEvent(NotifyEnFailure, ProcessExtentNodeFailure)]
-  cold state Repaired : MonitorState {
+  cold state Repaired {
     void ProcessExtentNodeFailure() {
@@
   [OnEvent(NotifyExtentRepaired, ProcessExtentRepaired)]
-  hot state Repairing : MonitorState {
+  hot state Repairing {
     void ProcessExtentRepaired() {

### 2015-09-14 paper.tex
updated abstract

@@
 
-We present a new methodology for systematically testing distributed systems.
-Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism \SCComment{this paper mainly focuses on the nondet due to scheduling, right?
-There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc.
-I feel that ``all sources'' is perhaps an overclaim} \PDComment{We should clarify: as long as you model a source of nondet using \psharp, then you can capture it and control it (not automatically)}.
-We present two case studies of using \psharp to test production distributed systems inside Microsoft.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it.
+We discuss our experience of using \psharp inside Microsoft to test three distributed storage systems for the Windows Azure cloud computing platform.
+\psharp is an extension of the \csharp language that provides environmental modeling capabilities and a systematic concurrency testing engine.
+As long as the developer has expressed all sources of nondeterminism in a system using \psharp, then the \psharp testing engine can systematically explore all these sources of nondeterminism to find bugs.
+
+The power of \psharp is that it enables the developer to \emph{test what is being executed} and is engineered to work on production-level code.
+Using \psharp we managed to find and reproduce more than 10 hard-to-find with traditional testing techniques bugs in our case studies.
+As an example, \psharp found a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it.
 \psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
-\PDComment{we now have many bugs in Matt's system too.}
 

### 2015-09-14 experience.tex
Updated Fabric model discussion.

@@
 \includegraphics[width=\linewidth]{img/fabricmodel}
-\caption{Overview of the key machines (rounded boxes) and interfaces (boxes) in our Fabric model.}
+\caption{Overview of the key machines and interfaces in our Fabric model.}
 \label{fig:fabric_model}
@@
 asynchrony and failure scenarios.
+
+\textbf{Replication example:} In order to replicate a state-mutating operation,
+user code at the primary replica 
+calls \texttt{IStateReplicator.ReplicateAsync},
+passing the serialized operation
+object.
+The operation is sent to the \texttt{ServiceRuntime},
+where it is assigned a \emph{logical sequence number} (LSN);
+each operation is assigned a consecutive LSN
+to track the total-order in which operations should be applied.
+The LSN is sent back to the primary replica
+where it is returned from the \texttt{ReplicateAsync} call,
+along with a \texttt{Task} object that will ``complete'' once
+the operation has been replicated to a majority of
+secondaries;
+thus, the user code can wait on the \texttt{Task}
+before confirming to any clients that the request has been applied reliably.
+The \texttt{ServiceRuntime} adds the operation to its list of in-flight
+replication requests and sends $n$ events to itself to signal that the request
+must be sent to a replica, where $n$ is the number of secondary replicas.
+The reason for sending $n$ events to itself instead of simply sending events
+directly to each secondary is so that the \texttt{ServiceRuntime}
+can process a simulated failure event inbetween the sending of replica requests
+to each secondary.
+This is an example of where we carefully considered
+the granularity of actions so that we could 
+model failures appropriately.
+The user code at a secondary receives the operation,
+applies it and then calls \texttt{Acknowledge} on the operation object;
+we implement this to send an event to the \texttt{ReplicaRuntime}
+which forwards the acknowledgement to the \texttt{ServiceRuntime}.
+Once a majority of secondaries have acknowledged, the
+\texttt{ServiceRuntime} removes the replication request from its list
+of replication requests
+and sends an acknowledgement to the primary,
+where the previously returned \texttt{Task} completes.
+
+\subsubsection{Fabric model correctness}
+Our model does not attempt to
+simulate the internals of Fabric accurately,
+as its purpose is to find bugs in the user code
+and not in Fabric itself (which we assume to be correct).   
+However,
+due to lack of documentation,
+it is not always clear how Fabric should behave
+in certain scenarios.
+Thus,
+we ran several variants of a simple Fabric service
+that logs calls into the user code in order to
+reverse-engineer the actual behaviour of Fabric;
+we ensured that our model has the same behaviour,
+although this is an ongoing process as we encounter
+additional scenarios.
+A further problem is that our model may contain bugs.
+In order to find bugs in our model effectively,
+we wrote a \psharp{} service made up of a single machine
+which takes the place
+of the user code and translation layer in Figure \ref{fig:fabric_model},
+for each \texttt{ReplicaRuntime}.
+Thus, we were able to run this pure \psharp{} system
+under \psharp{}'s systematic testing mode
+and uncover many assertion failures within our model.
+We tested a scenario where the primary fails at some non-deterministic point
+within the execution.
+
+\textbf{Example Fabric model assertion failure:}
+In the buggy trace,
+the \texttt{ServiceRuntime}
+sends an \texttt{EEpochInfo} event to the second
+\texttt{ReplicaRuntime}
+indicating that this is the first epoch and the
+replica will be a secondary.
+An \emph{epoch} represents a configuration of primary and
+secondary replicas; when a different replica becomes the primary,
+this indicates the start of a new epoch.
+The \texttt{ReplicaRuntime} acknowledges that it has become
+a secondary by responding with the same event type.
+The \psharp{} service sends an \texttt{ESecondaryCopyContextOp};
+this indicates what state the secondary has
+and, thus, what the primary should send to this secondary so that it can catch
+up.
+The \texttt{ESecondaryCopyContextOp} event is forwarded to the
+\texttt{ServiceRuntime}. 
+The \texttt{ServiceRuntime} then receives and handles an \texttt{EKillPrimary}
+event, which causes the second replica to become the new primary.
+Thus,
+the \texttt{ServiceRuntime} sends another \texttt{EEpochInfo}
+to the second \texttt{ReplicaRuntime}
+indicating that this is the second epoch
+and the replica will be a primary.
+As part of this change,
+the \texttt{ReplicaRuntime}
+sends an event to the \psharp{} service
+indicating that it should stop waiting for
+the state to be copied from the old primary to this replica,
+which is acknowledged by sending an event to the \texttt{ReplicaRuntime}.
+This event causes the \texttt{ReplicaRuntime}
+to send an \texttt{ESecondaryCopyStateDone}
+event to the \texttt{ServiceRuntime},
+which unfortunately responds with an event indicating that the
+\texttt{ReplicaRuntime} is now an \emph{active} secondary
+(i.e.\ a secondary that has caught up with the primary).
+However, this causes an assertion failure because
+the \texttt{ReplicaRuntime} is becoming a primary and, thus,
+cannot be a secondary.
+Our fix to this bug was to ensure that the 
+\texttt{ESecondaryCopyStateDone} event was marked as part of the first
+epoch (as the \texttt{ReplicaRuntime} had not yet acknowledged the
+change to primary); thus, 
+the \texttt{ServiceRuntime} ignores the event and does not try to make the
+replica an active secondary.
+
+\subsubsection{CScale}

### 2015-09-15 experience.tex

@@
-\subsection{Azure Storage vNext}
+\section{Liveness Bug in Azure Storage vNext}
 \label{sec:cases:azurestore}
 
-%Azure Storage vNext is the next generation storage system for Windows Azure.
-Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node (see Figure~\ref{fig:azurestore}).
-Each extent manager is responsible for managing a subset of the extent nodes.
-Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager.
-When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.
+It took only tens of seconds before the testing harness reported the first occurrence of a liveness bug. Upon examining the debug trace, the vNext developers were able to confirm the bug.
 
-%The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
-There was an actual bug in the system, that led this property to fail for some very rare executions.
-Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug.
-We discuss later in this paper, how we are able to detect and reproduce this bug using our approach.
+However, the trace didn't include enough details, which prevented the developers from identifying a root cause.
+Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration.
+After several iterations, the developers were able to pinpoint the exact culprit and immediately proposed a solution to fix the bug.
+Once the proposed solution was implemented, the developers again ran the testing harness, which reported no more bug for $100,000$ iterations in tens of minutes.
 
-%Azure Storage vNext is very challenging to test.
-All the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time.
-However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions.
-The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied.
-The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.
+The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} fails one of the \texttt{ExtentNode} and launches a new one.
+The \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.
 
-We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system.
-This component is responsible for managing the partitioned extent metadata and works as follows.
+Here is one particular execution sequence resulting in the bug.
+$i)$ EN$_0$ fails and is detected by the EN expiration loop; $ii)$ EN$_0$ is removed from the EN map; $iii)$ the extent center is updated and the replica count drops from 3, the target, to 2; $iv)$ \texttt{ExtMgr} receives a sync report from EN$_0$; $v)$ the extent center is updated and the replica count increases from 2 to 3 again.
+This is problematic because, on one hand, the replica count is equal to the target, so the extent repair loop never schedules any repair task.
+On the other hand, there are only two true replicas in the system, one fewer than the target.
+This execution sequence leads to one replica missing.
+Conceivably, repeating this another two times would result in all replicas missing, while \texttt{ExtMgr} still thinks all replicas are healthy.
+If deployed in production, such bug would have caused a very serious incident of customer data loss.
 
-vNext bug goes here.
+The culprit is in $iv)$, where \texttt{ExtMgr} receives a sync report from EN$_0$ after deleting the EN.
+This may occur in \psharp because of arbitrary message interleaving.
+It may also occur, albeit much less frequently, in the stress testing due to messages being delayed in the network.
+This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger.
+In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution verified effectively, all of which should vastly increase the productivity of distributed storage system development.
 
-\subsection{Live Azure Table Migration}
+\section{Additional Example I: Live Azure Table Migration}
 \label{sec:cases:migration}
@@
 
-\subsection{Azure Service Fabric (change to CScale?)}
+\section{Additional Example II: Azure Service Fabric (change to CScale?)}
 \label{sec:cases:fabric}

### 2015-09-15 evaluation.tex
update in the evaluation

@@
-After we fixed all the bugs we could find in the programs under test, we added options to the code to introduce various benchmark bugs (some that were actually in the original programs and some we made up) one at a time so we could evaluate different methods to detect and diagnose them.
-We compare \psharp systematic concurrency testing to traditional ``stress testing'' without systematic concurrency control; we report the average length of time it took for the test to fail and remark on the effort needed to diagnose the bug from the test output.
-\MMComment{Overview of actual results?}
+We report our experience of applying \psharp on the three case studies discussed in this paper: Azure Storage vNext, Live Migration Table and Azure Service Fabric.
+We first discuss how much human effort was spent in modeling the environment of these case studies so they become amenable to testing with \psharp.
+We then present runtime results of using \psharp with two different systematic testing schedulers to find bugs in the case studies.
 
-\subsection{Experimental Setup}
-
-We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-
-Cheng's case study with the actual bug
-
-After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs:
-\begin{enumerate}
-\item Filtering on a user-defined property did not retrieve non-matching rows from the new table that might shadow a matching row from the old table.
-\item ...
-\end{enumerate}
+\subsection{Cost of modeling}
 
@@
 \end{tabular}
-\caption{Statistics.}
+\caption{Statistics from environmental modeling in the three case studies.}
 %\vspace{-3mm}
@@
 
-\subsection{Systematic testing with \psharp}
+\subsection{Cost of systematic testing}
+
+After we fixed all the bugs we could find in the programs under test, we added options to the code to introduce various benchmark bugs (some that were actually in the original programs and some we made up) one at a time so we could evaluate different methods to detect and diagnose them.
+
+After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs.
+
+We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
 
@@
 
-How long you had to run the test to find the bug (if you can find it) and how long are the traces
-
-How easy it is to pinpoint the error using the \psharp traces, comparing to traditional stress testing
-
-Effort required to use the system
-
 Controlled random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.

### 2015-09-15 evaluation.tex
update in the evaluation

@@
-We report our experience of applying \psharp on the three case studies discussed in this paper: Azure Storage vNext, Live Migration Table and Azure Service Fabric.
-We first discuss how much human effort was spent in modeling the environment of these case studies so they become amenable to testing with \psharp.
-We then present runtime results of using \psharp with two different systematic testing schedulers to find bugs in the case studies.
+We report our experience of applying \psharp on the three case studies discussed in this paper. We aim to answer the two following questions:
 
-\subsection{Cost of modeling}
+\begin{enumerate}
+\item How much human effort was spent in modeling the environment of a distributed system using \psharp?
+
+\item How much computational time was spent in systematically testing a distributed system using \psharp?
+\end{enumerate}
+
+\subsection{Cost of environmental modeling}
+\label{sec:eval:human_cost}
 
@@
 \setlength{\tabcolsep}{0.3em}
-\label{tab:stats}
 \begin{tabular}{l rrrrr rr}
@@
 \end{tabular}
-\caption{Statistics from environmental modeling in the three case studies.}
+\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test.}
+\label{tab:stats}
 %\vspace{-3mm}
@@
 
+Environmental modeling is a core activity of using \psharp.
+It is required for \emph{closing the environment} of a system under test and making it amenable to systematic testing.
+Table~\ref{tab:stats} presents program statistics for our three case studies.
+We report: lines of code for the system under test (\#LoC); number of bugs found in the system (\#B); lines of \psharp code for the environmental model (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
+
+Modeling the environment of the vNext
+
 \subsection{Cost of systematic testing}
+\label{sec:eval:machine_cost}
+
+We then present runtime results of using \psharp with two different systematic testing schedulers to find bugs in the case studies.
 

### 2015-09-15 evaluation.tex
update in the evaluation

@@
 
-Modeling the environment of the vNext
+Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately 2 weeks of part-time developing.
+The \psharp model for testing this system is the smallest (in lines of code) from all three case studies.
+This was because the modeling effort was targeting the particular liveness bug that was haunting the developers of vNext.
+We are currently in the process of modeling other components of vNext, such as a ChainReplication and a Paxos system.
+
+Modeling the Live Migration Table required \PDComment{waiting confirmation from Matt}.
+This case study is interesting because the development of the actual system and its \psharp environmental model occurred side-by-side.
+This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and at a later stage of the development process.
+
+Modeling Fabric required approximately 4-5 months.
+Although this is a significant amount of time, it is a one time effort activity.
+Our plan is to reuse the developed Fabric model for testing arbitrary user services built for the Azure Service Fabric system.
 

### 2015-09-15 evaluation.tex
update in the evaluation

@@
 
-We then present runtime results of using \psharp with two different systematic testing schedulers to find bugs in the case studies.
-
-After we fixed all the bugs we could find in the programs under test, we added options to the code to introduce various benchmark bugs (some that were actually in the original programs and some we made up) one at a time so we could evaluate different methods to detect and diagnose them.
-
-After fixing all the bugs we found in MigratingTable, we added an option to conditionally reintroduce each of the following bugs.
-
-We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-
 \setlength{\tabcolsep}{.72em}
@@
 
-Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with an enabled bug using the random and the PCT schedulers.
+Using \psharp we managed to uncover more than 10 serious bugs in our case studies.
+As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting.
+According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner.
+After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one so that we can evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.
+
+Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using the random and the PCT schedulers.
+We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
 We configured the engine to perform 100,000 iterations.
-The random seed for both schedulers was generated in each iteration using the \texttt{DateTime.Now.Millisecond} API which represents the current time in milliseconds.
-The PCT scheduler was configured with a bug depth of 2 and a max number of scheduling steps to execute of 500.
+The random seed for both schedulers was generated in each iteration using the \texttt{DateTime.Now.Millisecond} API which returns the current time in milliseconds.
+The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps to execute of 500.
 All reported times are in seconds.

### 2015-09-15 evaluation.tex
update in the evaluation

@@
 Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using the random and the PCT schedulers.
+The CS column shows which case study corresponds to each bug: 1 is for the Azure Storage vNext; and 2 is for the Live Migration Table.
+\PDComment{update with 3 for Fabric}
+
 We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
@@
 All reported times are in seconds.
+
+For the vNext case study, the random scheduler was able to reproduce the bug in less than 10 seconds.
+The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in Section~\ref{} we leave the program to run for a long time before checking if the liveness property holds.
+The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.
 

### 2015-09-15 experience.tex
CScale text.

@@
 by the \psharp{} runtime.
-The system that we wished to test is \emph{CScale}~\cite{X},
-a big data-stream processing system that chains multiple Fabric services
-in a directed acyclic graph.
 Prior work~\cite{deligiannis2015psharp}
@@
 \subsubsection{CScale}
+The system that we wished to test is \emph{CScale}~\cite{X},
+a big data-stream processing system that chains multiple Fabric services
+in a directed acyclic graph.
+A key challenge is that
+CScale contains inter-service communication
+that resolves services
+via Fabric but communicates
+using a non-Fabric remote procedure call protocol.
+In order to close the system,
+we first implemented service resolution in our Fabric model,
+which resolves Fabric service names to IP addresses.
+We then replaced the inter-service communication
+classes in CScale with our own;
+our replacement includes the ability to map
+URLs to \psharp{} machines
+so that remote procedure calls are now implemented
+by sending and receiving \psharp{} events.
+Thus,
+we converted a multi-process system
+that runs on the Fabric platform
+and 
+uses network communication
+into 
+a single process system that uses multiple threads,
+contains
+the Fabric model
+and the CScale services,
+and does not use network communication.
+Additional changes were needed to
+remove certain static (per-process) fields,
+as these were inadvertently
+being shared between services
+after making the system a single process.
+
+A key challenge in our work
+was to test CScale despite the fact that it
+uses various synchronous and asynchronous APIs
+other than \psharp{}.
+This work is still in-progress.
+However,
+we were able to find a \texttt{NullReferenceException}
+bug
+in CScale
+by running it on our model.

### 2015-09-15 paper.tex
split and update section 6
update
update in the evaluation
updated abstract and author affiliations

@@
 {\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
-{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\ddagger$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
-{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\ddagger$}\\\\
-$^\dagger$Imperial College London, $^\ddagger$Microsoft, $^\star$Microsoft Research\\
+{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
+{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
+$^\dagger$Imperial College London, $^\star$Microsoft\\
 $^\diamond$Massachusetts Institute of Technology\\
@@
 \subsection*{Abstract}
-Testing distributed systems is very challenging due to multiple sources of nondeterminism, such as arbitrary interleavings between event handlers and unexpected node failures.
-Stress testing, commonly used in industry today, is unable to deal with this kind of nondeterminism, which results in the most tricky bugs being missed during testing and only getting exposed in production.
+Testing distributed systems is challenging due to multiple sources of nondeterminism in the execution, e.g., arbitrary interleavings between event handlers and unexpected node failures.
+State-of-the-art testing methods commonly used in industry today ---unit, integration, and stress testing--- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
+In this paper, we present a new approach to testing distributed storage systems.
+Our approach significantly improves test coverage and allows us to discover during testing errors that are often caught after deployment.
 
-We discuss our experience of using \psharp inside Microsoft to test three distributed storage systems for the Windows Azure cloud computing platform.
-\psharp is an extension of the \csharp language that provides environmental modeling capabilities and a systematic concurrency testing engine.
-As long as the developer has expressed all sources of nondeterminism in a system using \psharp, then the \psharp testing engine can systematically explore all these sources of nondeterminism to find bugs.
+Our method for testing distributed storage systems is based on \psharp, an extension to the \csharp language,
+that provides support for modeling, specification, and systematic testing of asynchronous systems.
+We have used \psharp to test three distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing platform.
+To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts---a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
+The \psharp testing engine then systematically explores behaviors of the test harness and validates them against the provided specifications.
 
-The power of \psharp is that it enables the developer to \emph{test what is being executed} and is engineered to work on production-level code.
-Using \psharp we managed to find and reproduce more than 10 hard-to-find with traditional testing techniques bugs in our case studies.
-As an example, \psharp found a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it.
-\psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem.
+Our approach enables the developer to \emph{test what is being executed} and is engineered to work on production code.
+We have used our approach to find, reproduce, and fix numerous bugs in production systems;
+these bugs are exhibited by executions with subtle combination of concurrency and failures, making them extremely difficult
+to find with traditional testing techniques.
+An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace,
+which dramatically increases the productivity of debugging.
 
@@
 
-\section{Experience Report}
-\label{sec:exp}
+%\section{Experience Report}
+%\label{sec:exp}
 
@@
 
-\section{Evaluation}
+\section{Quantifying the Cost of Using \psharp}
 \label{sec:eval}

### 2015-09-15 experience.tex
save one line space

@@
 \label{sec:cases:migration}
+
+- \psharp generalized beyond vNext. \psharp is not only applicable to vNext but in many systems.
+- Advanced safety properties
+- Handles intra-machine concurrency
+- Writing the protocol while testing it (co-development)
 
@@
 \label{sec:cases:fabric}
+
+- \psharp generalized beyond vNext. \psharp is not only applicable to vNext but in many systems.
+- (Advanced) safety properties?
+- What we developed is a model that other people can reuse. Plugin the model and test the user application.
+- Example of advanced modeling capability of \psharp. One component talks to another sophisticated API. We demonstrate environmental modeling expressivity in system engineering.
 

### 2015-09-15 paper.tex
save one line space

@@
 {\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
-$^\dagger$Imperial College London, $^\star$Microsoft\\
-$^\diamond$Massachusetts Institute of Technology\\
+$^\dagger$Imperial College London, $^\star$Microsoft, $^\diamond$Massachusetts Institute of Technology\\
 } % end author

### 2015-09-15 paper.tex
quick fix in abstract

@@
 Testing distributed systems is challenging due to multiple sources of nondeterminism in the execution, e.g., arbitrary interleavings between event handlers and unexpected node failures.
-State-of-the-art testing methods commonly used in industry today ---unit, integration, and stress testing--- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
+State-of-the-art testing methods commonly used in industry today --- unit, integration, and stress testing --- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
 In this paper, we present a new approach to testing distributed storage systems.
-Our approach significantly improves test coverage and allows us to discover during testing errors that are often caught after deployment.
+Our approach significantly improves test coverage and allows us to discover errors during testing that are often caught after deployment.
 

### 2015-09-15 experience.tex
updates in case studies

--- 2015-09-15 experience.tex
@@
 
-\section{Additional Example I: Live Azure Table Migration}
+\section{Additional Case Studies}
+\label{sec:cases:other}
+
+The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system.
+\psharp is a generic framework, capable of handling arbitrary distributed systems.
+We showcase this capability by presenting developer experience in using \psharp to model and test two other distributed storage systems used in production in Microsoft: the Live Azure Table Migration library; and the Azure Service Fabric system.
+
+\subsection{Live Azure Table Migration}
 \label{sec:cases:migration}
 
-The Live Azure Table Migration is a library capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
+- Advanced safety properties\\
+- Handles intra-machine concurrency\\
+- Writing the protocol while testing it (co-development)
+
+The Live Azure Table Migration (MigratingTable) is a library for transparently migrating a data set between tables in an Azure Storage service, while an application is accessing this data set.
+
 MigratingTable provides a \emph{virtual table} with an API similar to that of an ordinary Azure table, backed by a pair of \emph{old} and \emph{new} tables.
 A background \emph{migrator} job moves all data from the old table to the new table.
-Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.
+Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol that guarantees linearizability of operations on the virtual table across multiple application processes, assuming that the backend tables respect their own linearizability guarantees.
 
@@
 % N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
-The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
+%The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.
 As the traffic continues to grow over time, the system needs to reshard the data set across a greater number of Azure storage accounts without interrupting service.
@@
 MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.
-
-\def\term#1{\emph{#1}}
-We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.
 
@@
 
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the original Azure table API with one exception: it provides \term{streaming reads} with a weaker consistency property than multi-page reads in the original API, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
+MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the original Azure table API with one exception: it provides \emph{streaming reads} with a weaker consistency property than multi-page reads in the original API, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
 MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
@@
 \begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of \term{old} and \term{new} SpecTables in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+For every execution trace of a collection of MigratingTables backed by the same pair of \emph{old} and \emph{new} SpecTables in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
 \end{quote}
 %
-We instrumented MigratingTable to report the intended \term{linearization point} of each input call, which in our setting is always one of the corresponding \term{backend calls} to the backend tables (often the last).
+We instrumented MigratingTable to report the intended \emph{linearization point} of each input call, which in our setting is always one of the corresponding \emph{backend calls} to the backend tables (often the last).
 Specifically, after each backend call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
 This makes it possible to verify the correctness property as the model executes.
-The model consists of a \psharp \term{tables machine} containing all three SpecTables; a collection of \term{service machines} containing identically configured MigratingTables; and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
+The model consists of a \psharp \emph{tables machine} containing all three SpecTables; a collection of \emph{service machines} containing identically configured MigratingTables; and a \emph{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
 Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
@@
 
-\subsection{An example bug in MigratingTable}
-\MMComment{Move wherever appropriate}
+\subsubsection{An example bug in MigratingTable}
 
@@
 
-\MMComment{Add a figure based on migration-bug3-explanation.pptx (probably a hybrid of slides 4 and 5, assuming we want only one figure).}
-
-\section{Additional Example II: Azure Service Fabric (change to CScale?)}
+\subsection{Azure Service Fabric}
 \label{sec:cases:fabric}
 
-\PDComment{I guess we should write stuff about CScale here based on our last discussion: overview of the system (collection of Fabric services connected via TPL dataflow?), of its environment (Fabric), how they interop?, why its very challenging to test?
-how they test currently?
-I am not sure if all these should go here or be split here and the experience report, we will see ...
-same for the above case studies}
+- \psharp generalized beyond vNext. \psharp is not only applicable to vNext but in many systems.\\
+- (Advanced) safety properties?\\
+- What we developed is a model that other people can reuse. Plugin the model and test the user application.\\
+- Example of advanced modeling capability of \psharp. One component talks to another sophisticated API. We demonstrate environmental modeling expressivity in system engineering.
 

### 2015-09-15 experience.tex
updates in case studies

@@
 
-- Advanced safety properties\\
-- Handles intra-machine concurrency\\
-- Writing the protocol while testing it (co-development)
-
 The Live Azure Table Migration (MigratingTable) is a library for transparently migrating a data set between tables in an Azure Storage service, while an application is accessing this data set.
+This case study is interesting because its developers built the \psharp model of the system, while developing the library, to increase confidence in its implementation; indeed they discovered many bugs during this \emph{co-development} process (see Section~\ref{}).
+Another interesting aspect of this system is that it heavily uses the \texttt{async}/\texttt{await} primitives, which were not supported in the original \psharp runtime, and thus required us to extend \psharp with support for intra-machine concurrency (see Section~\ref{}).
+Finally, a lot of complex safety properties were written to test MigratingTable with \psharp, whereas the vNext model focused on a single liveness property.
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/livemigration}
-\caption{Resharding a data set when a third Azure storage account is added. Two key ranges are each migrated to the new account using a MigratingTable instance (abbreviated MTable).}
-\label{fig:livemigration}
-\end{figure}
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=\linewidth]{img/livemigration}
+%\caption{Resharding a data set when a third Azure storage account is added. Two key ranges are each migrated to the new account using a MigratingTable instance (abbreviated MTable).}
+%\label{fig:livemigration}
+%\end{figure}
 
@@
 \includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{MigratingTable \psharp correctness test environment.}
+\caption{Environmental model of MigratingTable (each box with a dotted line represents one \psharp machine).}
 \label{fig:mockedmigration}
@@
 
-Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
+%Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
 
@@
 
-All of our input histories include two application processes.
+%All of our input histories include two application processes.
 Each process performs either a single streaming read or a sequence of two atomic calls, each a read or a batch write.

### 2015-09-15 experience.tex
Made cuts in the Fabric section.

@@
 
-The Live Azure Table Migration (MigratingTable) is a library for transparently migrating a data set between tables in an Azure Storage service, while an application is accessing this data set.
-This case study is interesting because its developers built the \psharp model of the system, while developing the library, to increase confidence in its implementation; indeed they discovered many bugs during this \emph{co-development} process (see Section~\ref{}).
+The Live Azure Table Migration (MigratingTable) is a library for \emph{transparently migrating} a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
+This case study is interesting because its developers built the \psharp model of the system while developing the library to increase confidence in its implementation; indeed they discovered many bugs during this \emph{co-development} process (see Section~\ref{}).
 Another interesting aspect of this system is that it heavily uses the \texttt{async}/\texttt{await} primitives, which were not supported in the original \psharp runtime, and thus required us to extend \psharp with support for intra-machine concurrency (see Section~\ref{}).
@@
 
-MigratingTable provides a \emph{virtual table} with an API similar to that of an ordinary Azure table, backed by a pair of \emph{old} and \emph{new} tables.
-A background \emph{migrator} job moves all data from the old table to the new table.
-Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol that guarantees linearizability of operations on the virtual table across multiple application processes, assuming that the backend tables respect their own linearizability guarantees.
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_migration}
+\caption{Environmental model of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\label{fig:mockedmigration}
+\end{figure}
+
+MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary Azure table.
+This interface is named \texttt{IChainTable}.
+The virtual table is backed by a pair of \emph{old} and \emph{new} tables.
+A background \emph{migrator} job is responsible for moving all data from the old table to the new table.
+Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol that guarantees linearizability of operations on the virtual table across multiple application processes, assuming that the backend tables respect their own linearizability guarantees.
+
+The \psharp testing goal was to ensure that when multiple application processes, implementing the \texttt{IChainTable} interface, issue read and write operations to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of \texttt{IChainTable} for the combined \emph{input history}.
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Environmental model of MigratingTable (each box with a dotted line represents one \psharp machine).}
-\label{fig:mockedmigration}
-\end{figure}
-
 %Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.
 
-MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the original Azure table API with one exception: it provides \emph{streaming reads} with a weaker consistency property than multi-page reads in the original API, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
-MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.
-Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.
+%MigratingTable implements an interface called \texttt{IChainTable}, which provides the core read and write functionality of the original Azure table API with one exception: it provides \emph{streaming reads} with a weaker consistency property than multi-page reads in the original API, since the original property would have been difficult to achieve for no benefit to applications we could foresee.
+MigratingTable requires that its backend tables also implement \texttt{IChainTable}, and we wrote a simple adapter to expose physical Azure tables as \texttt{IChainTable}.
 
-\subsubsection{Input generation}
-\label{sec:mtable:input}
+% N.B. \texttt{SpecTable} = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
 
-% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-Of course, there are many possible input histories for MigratingTable.
-We could have tested specific input histories, but we were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for interactions between different parts of the code that would be difficult for us to foresee.
-If we had no formalization of the specification and had to rely on expected outputs worked out by hand, this might be the best we could do.
-However, since the IChainTable2 specification is relatively simple and is almost deterministic under sequential calls, it was straightforward to write an in-memory reference implementation called SpecTable to which we can compare the output of MigratingTable on an arbitrary input history.
+The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
+The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult to foresee interactions between different parts of the code.
+
+\subsubsection{Testing the MigratingTable library}
+
+Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history.
+This enabled sampling from a distribution, which was defined over all possible input histories within certain bounds.
+\psharp takes control of the choice of input history, so that it can systematically explore different input scenarios.
+
+%\psharp takes control of the choice of input history, as well as the schedule, so both can be reproduced using a single random seed.
+Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
+
+%If we had no formalization of the specification and had to rely on expected outputs worked out by hand, this might be the best we could do.
+However, since the \texttt{IChainTable} specification is relatively simple and is almost deterministic under sequential calls, it was straightforward to write an in-memory reference implementation called \texttt{SpecTable} to which we can compare the output of MigratingTable on an arbitrary input history.
 This gave us the attractive option to sample from a distribution we defined over all possible input histories within certain bounds.
-It was convenient to let \psharp control the choice of input history as well as the schedule so we could reproduce both using a single random seed.
-Then, under the ``small scope hypothesis'' that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
+
+%It was convenient to let \psharp control the choice of input history as well as the schedule so we could reproduce both using a single random seed.
+Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
 
@@
 Each process performs either a single streaming read or a sequence of two atomic calls, each a read or a batch write.
-Each batch write call includes one or two operations, where the operation type is chosen from the set supported by IChainTable2 (Insert, Replace, Merge, Delete, InsertOrReplace, InsertOrMerge, DeleteIfExists) and the row key is chosen from $\{0, \ldots, 5\}$.
+Each batch write call includes one or two operations, where the operation type is chosen from the set supported by \texttt{IChainTable} (Insert, Replace, Merge, Delete, InsertOrReplace, InsertOrMerge, DeleteIfExists) and the row key is chosen from $\{0, \ldots, 5\}$.
 If the operation requires an If-Match value, it is equally likely to be \texttt{*}, the current ETag of the row (if it exists), or some non-matching value.
@@
 
-\subsubsection{Model structure}
+%As mentioned above, the \texttt{IChainTable} specification is almost deterministic under sequential calls; the only nondeterminism is in the results of streaming reads.
+Given a streaming read, \texttt{SpecTable} can compute the set of all results that are compliant with the specification, so we can simply check if the result of MigratingTable is in this set.
 
-As mentioned above, the IChainTable2 specification is almost deterministic under sequential calls; the only nondeterminism is in the results of streaming reads.
-Given a streaming read, SpecTable can compute the set of all results that are compliant with the specification, so we can simply check if the result of MigratingTable is in this set.
 To test MigratingTable, we must supply it with backing tables.
-We use SpecTable for this purpose as well, with \psharp choosing the actual result of each streaming read from the valid set.
+We use \texttt{SpecTable} for this purpose as well, with \psharp choosing the actual result of each streaming read from the valid set.
 Our correctness property is then:
@@
 \begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of \emph{old} and \emph{new} SpecTables in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
+For every execution trace of a collection of MigratingTables backed by the same pair of \emph{old} and \emph{new} \texttt{SpecTable}s in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' \texttt{SpecTable} on the linearized input.
 \end{quote}
@@
 This makes it possible to verify the correctness property as the model executes.
-The model consists of a \psharp \emph{tables machine} containing all three SpecTables; a collection of \emph{service machines} containing identically configured MigratingTables; and a \emph{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).
-Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.
-When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.
-When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.
-\psharp controls the interleaving of the backend calls.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
-We use the \psharp random scheduling strategy; we were afraid that an exhaustive strategy would only be feasible within bounds so low that we would miss some bugs.
 
-We wanted to implement the core MigratingTable algorithms in \csharp ``async/await'' code, like most of Artifact Services, to achieve both good readability and good performance.
+The \psharp environmental model of MigratingTable consists of a \psharp \texttt{Tables} machine containing all three \texttt{SpecTable} implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
+
+Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine.
+When MigratingTable reports the linearization point of an input call, the \texttt{Service} machine sends that input call to the reference table.
+When an input call completes, the \texttt{Service} machine checks that the results from the MigratingTable and the reference table agree.
+
+\psharp captures and controls the interleaving of the backend calls.
+To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
+
+%We use the \psharp random scheduling strategy; we were afraid that an exhaustive strategy would only be feasible within bounds so low that we would miss some bugs.
+
+%We wanted to implement the core MigratingTable algorithms in \csharp ``async/await'' code, like most of Artifact Services, to achieve both good readability and good performance.
 We used a method similar to that described in Section~\ref{sec:psharp:async} to bring the generated TPL tasks under the control of the \psharp scheduler.
-Then we implemented an ``async'' RPC mechanism based on the .NET RealProxy class that automates the generation of proxies for objects hosted by other \psharp machines (in our setting, the service machines use proxies for the SpecTables and various auxiliary objects hosted by the tables machine).
+Then we implemented an ``async'' RPC mechanism based on the .NET RealProxy class that automates the generation of proxies for objects hosted by other \psharp machines (in our setting, the service machines use proxies for the \texttt{SpecTable}s and various auxiliary objects hosted by the tables machine).
 When a machine calls a method on a proxy, the proxy sends a \psharp message to the host machine, causing it to execute the method call on the original object and send back the result, which the proxy then returns.
-Thus, the use of these proxies as IChainTable2 backends is transparent to the MigratingTable library, thanks to dynamic dispatch.
-
-\subsubsection{An example bug in MigratingTable}
-
-One of the bugs in MigratingTable that we found using the \psharp test stands out because it reflects the type of oversight that tends to occur as designs evolve and it's unclear whether we would have been able to find it by any other method (\MMComment{revise remark when we have stress test results?}).
-This bug, which we named QueryStreamedBackUpNewStream, is in the implementation of a streaming read from the virtual table, which should return a stream of all rows in the table sorted by key.
-The essential implementation idea is to start streams $s_O$, $s_N$ from the old and new backend tables and merge the sorted streams by keeping track of the next row in each stream and returning the row with the lesser key.
-In parallel, the migrator job is concurrently copying rows from the old table to the new table; we had satisfied ourselves that this concurrency would not cause any problems.
-However, then we added support to the migrator job to delete the old table when it finishes copying, which triggers the virtual stream to close $s_O$.
-Suppose the virtual stream is in a state in which the next row in $s_O$ has key $k_O$ and the next row in $s_N$ has key $k_N$, where $k_O < k_N$.
-Further suppose that before the next read from the virtual stream, the migrator job copies a row with key $k$ ($k_O < k < k_N$) from the old table to the new table and then deletes the old table.
-Since $s_O$ has not yet returned this row when it is closed and $s_N$ has already advanced to $k_N$, the row with key $k$ will be missed by the virtual stream.
-A similar problem can occur if $s_N$ does not reflect rows inserted into the new table by the migrator job after $s_N$ is started, as allowed by the IChainTable2 specification.
-Restarting $s_N$ when the old table is deleted fixes both variants of the bug.
-
-% It might be nice to include excerpts of a trace like in migration-bug3-explanation.pptx.
-Unfortunately, the trace in migration-bug3-explanation.pptx doesn't match my recollection of the original diagnosis, which I want to write about truthfully (the former looks like it involves a stale streaming read, while I'm fairly sure the latter involved only the $k_O < k < k_N$ case).
-If it's important, I could try to get a new trace consistent with the original diagnosis.
-~ Matt
-This bug took us only about 10 minutes to diagnose from the trace; granted, this is after we had days of experience analyzing MigratingTable traces and had added our own trace output to the test harness, since \psharp's built-in trace output is too low-level and does not include event payloads and other diagnostic data that is not passed between machines.
-We started from the failure symptom: the virtual stream returned end-of-stream when according to the reference SpecTable, it should have returned an additional row with key 4.
-We filtered the trace for actions by the same service machine and saw that $s_O$ was closed before the virtual stream had returned the row with key 4, but $s_N$ had already advanced past key 4 before the migrator inserted the row in the new table.
-At first we found this phenomenon hard to believe, but soon we were convinced it reflected a gap in our design.
+Thus, the use of these proxies as \texttt{IChainTable} backends is transparent to the MigratingTable library, thanks to dynamic dispatch.
 
@@
 \emph{Azure Service Faric} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
-The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests.
-In order to make the service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on a different node in the cluster.
-One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
+%The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
+In order to make the user-written service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on a different node in the cluster.
+\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
+%One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
 The primary replicates state changes to the secondaries by sending \emph{replication requests} so that all replicas eventually have the same state.
@@
 Fabric provides a name-resolution service so that clients can always find the current primary.
+The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
+Since user-written Fabric services are complex asynchronous and distributed applications, 
+they are challenging to test.
+% They are interesting targets for systematic testing with \psharp{}.
 
-Fabric services have a lot of asynchrony, which make them interesting targets for systematic testing with \psharp{}.
+\PTComment{Could be improved:}
 Our primary goal was to create a \psharp{} model of Fabric to allow
@@
 by the \psharp{} runtime.
-Prior work~\cite{deligiannis2015psharp}
-created a model of Fabric with limited functionality;
-it used a mixture of \csharp{} and \psharp{} internally,
-only supported one in-flight replication request
-(which restricts the asynchrony that can be tested),
-and only supported one Fabric service.
-Our new Fabric model was re-written to use only \psharp{}
-internally,
-support an arbitrary number of Fabric services and in-flight replication requests,
-and in general be a more complete model of Fabric.
-Note that \csharp{} code is still required to interface with the existing
-\csharp{} service code.
-We refer to our \csharp{} code
-as the \emph{translation layer}
-and
- the user-written \csharp{} service code as \emph{user code}.
+Unlike the other case studies,
+this model is written once
+to include all behaviours of Fabric
+so that it can be used again and again to test arbitrary Fabric services.
 
-\begin{figure}[thb]
-\centering
-\includegraphics[width=\linewidth]{img/fabricmodel}
-\caption{Overview of the key machines and interfaces in our Fabric model.}
-\label{fig:fabric_model}
-\end{figure}
+\PTComment{Cut: prior work created a Fabric model\ldots}
+% Prior work~\cite{deligiannis2015psharp}
+% created a model of Fabric with limited functionality;
+% it used a mixture of \csharp{} and \psharp{} internally,
+% only supported one in-flight replication request
+% (which restricts the asynchrony that can be tested),
+% and only supported one Fabric service.
+% Our new Fabric model was re-written to use only \psharp{}
+% internally,
+% support an arbitrary number of Fabric services and in-flight replication requests,
+% and in general be a more complete model of Fabric.
+% Note that \csharp{} code is still required to interface with the existing
+% \csharp{} service code.
+% We refer to our \csharp{} code
+% as the \emph{translation layer}
+% and
+%  the user-written \csharp{} service code as \emph{user code}.
 
-An overview of our Fabric model is shown in Figure~\ref{fig:fabric_model}.
-The \texttt{ClusterRuntime} machine 
-handles the creation and management of 
-one or more Fabric services,
-as well as service resolution requests
-which allows for client-service and inter-service communication
-within the model.
-Each Fabric service instance is managed by a \texttt{ServiceRuntime}
-machine, which in turn manages 
-several \texttt{ReplicaRuntime} machines.
-Each \texttt{ReplicaRuntime} communicates with the user code
-via several machines and interfaces from the translation layer
-(only the translation layer for the primary is shown,
-but every \texttt{ReplicaRuntime} has its own instance of the translation layer).
-Note that communication between machines is hierachical;
-thus, communication between \texttt{ReplicaRuntime}s
-(such as the sending of replication requests)
-is via the \texttt{ServiceRuntime} machine for that service.
-This approach does not necessarily reflect how Fabric works in practice.
-Instead, we chose an architecture
-that keeps the model simple
-while still allowing (what we believe to be) realistic
-asynchrony and failure scenarios.
+\PTComment{Cut: Figure of model.}
+% \begin{figure}[thb]
+% \centering
+% \includegraphics[width=\linewidth]{img/fabricmodel}
+% \caption{Overview of the key machines and interfaces in our Fabric model.}
+% \label{fig:fabric_model}
+% \end{figure}
 
-\textbf{Replication example:} In order to replicate a state-mutating operation,
-user code at the primary replica 
-calls \texttt{IStateReplicator.ReplicateAsync},
-passing the serialized operation
-object.
-The operation is sent to the \texttt{ServiceRuntime},
-where it is assigned a \emph{logical sequence number} (LSN);
-each operation is assigned a consecutive LSN
-to track the total-order in which operations should be applied.
-The LSN is sent back to the primary replica
-where it is returned from the \texttt{ReplicateAsync} call,
-along with a \texttt{Task} object that will ``complete'' once
-the operation has been replicated to a majority of
-secondaries;
-thus, the user code can wait on the \texttt{Task}
-before confirming to any clients that the request has been applied reliably.
-The \texttt{ServiceRuntime} adds the operation to its list of in-flight
-replication requests and sends $n$ events to itself to signal that the request
-must be sent to a replica, where $n$ is the number of secondary replicas.
-The reason for sending $n$ events to itself instead of simply sending events
-directly to each secondary is so that the \texttt{ServiceRuntime}
-can process a simulated failure event inbetween the sending of replica requests
-to each secondary.
-This is an example of where we carefully considered
-the granularity of actions so that we could 
-model failures appropriately.
-The user code at a secondary receives the operation,
-applies it and then calls \texttt{Acknowledge} on the operation object;
-we implement this to send an event to the \texttt{ReplicaRuntime}
-which forwards the acknowledgement to the \texttt{ServiceRuntime}.
-Once a majority of secondaries have acknowledged, the
-\texttt{ServiceRuntime} removes the replication request from its list
-of replication requests
-and sends an acknowledgement to the primary,
-where the previously returned \texttt{Task} completes.
+\PTComment{Cut: overview of model.}
+% An overview of our Fabric model is shown in Figure~\ref{fig:fabric_model}.
+% The \texttt{ClusterRuntime} machine 
+% handles the creation and management of 
+% one or more Fabric services,
+% as well as service resolution requests
+% which allows for client-service and inter-service communication
+% within the model.
+% Each Fabric service instance is managed by a \texttt{ServiceRuntime}
+% machine, which in turn manages 
+% several \texttt{ReplicaRuntime} machines.
+% Each \texttt{ReplicaRuntime} communicates with the user code
+% via several machines and interfaces from the translation layer
+% (only the translation layer for the primary is shown,
+% but every \texttt{ReplicaRuntime} has its own instance of the translation layer).
+% Note that communication between machines is hierachical;
+% thus, communication between \texttt{ReplicaRuntime}s
+% (such as the sending of replication requests)
+% is via the \texttt{ServiceRuntime} machine for that service.
+% This approach does not necessarily reflect how Fabric works in practice.
+% Instead, we chose an architecture
+% that keeps the model simple
+% while still allowing (what we believe to be) realistic
+% asynchrony and failure scenarios.
 
-\subsubsection{Fabric model correctness}
-Our model does not attempt to
-simulate the internals of Fabric accurately,
-as its purpose is to find bugs in the user code
-and not in Fabric itself (which we assume to be correct).   
-However,
-due to lack of documentation,
-it is not always clear how Fabric should behave
-in certain scenarios.
-Thus,
-we ran several variants of a simple Fabric service
-that logs calls into the user code in order to
-reverse-engineer the actual behaviour of Fabric;
-we ensured that our model has the same behaviour,
-although this is an ongoing process as we encounter
-additional scenarios.
-A further problem is that our model may contain bugs.
-In order to find bugs in our model effectively,
-we wrote a \psharp{} service made up of a single machine
-which takes the place
-of the user code and translation layer in Figure \ref{fig:fabric_model},
-for each \texttt{ReplicaRuntime}.
-Thus, we were able to run this pure \psharp{} system
-under \psharp{}'s systematic testing mode
-and uncover many assertion failures within our model.
-We tested a scenario where the primary fails at some non-deterministic point
-within the execution.
+\PTComment{Cut: Description of how replication works in the model, including how we modeled at a fine granularity.}
+% \textbf{Replication example:} In order to replicate a state-mutating operation,
+% user code at the primary replica 
+% calls \texttt{IStateReplicator.ReplicateAsync},
+% passing the serialized operation
+% object.
+% The operation is sent to the \texttt{ServiceRuntime},
+% where it is assigned a \emph{logical sequence number} (LSN);
+% each operation is assigned a consecutive LSN
+% to track the total-order in which operations should be applied.
+% The LSN is sent back to the primary replica
+% where it is returned from the \texttt{ReplicateAsync} call,
+% along with a \texttt{Task} object that will ``complete'' once
+% the operation has been replicated to a majority of
+% secondaries;
+% thus, the user code can wait on the \texttt{Task}
+% before confirming to any clients that the request has been applied reliably.
+% The \texttt{ServiceRuntime} adds the operation to its list of in-flight
+% replication requests and sends $n$ events to itself to signal that the request
+% must be sent to a replica, where $n$ is the number of secondary replicas.
+% The reason for sending $n$ events to itself instead of simply sending events
+% directly to each secondary is so that the \texttt{ServiceRuntime}
+% can process a simulated failure event inbetween the sending of replica requests
+% to each secondary.
+% This is an example of where we carefully considered
+% the granularity of actions so that we could 
+% model failures appropriately.
+% The user code at a secondary receives the operation,
+% applies it and then calls \texttt{Acknowledge} on the operation object;
+% we implement this to send an event to the \texttt{ReplicaRuntime}
+% which forwards the acknowledgement to the \texttt{ServiceRuntime}.
+% Once a majority of secondaries have acknowledged, the
+% \texttt{ServiceRuntime} removes the replication request from its list
+% of replication requests
+% and sends an acknowledgement to the primary,
+% where the previously returned \texttt{Task} completes.
 
-\textbf{Example Fabric model assertion failure:}
-In the buggy trace,
-the \texttt{ServiceRuntime}
-sends an \texttt{EEpochInfo} event to the second
-\texttt{ReplicaRuntime}
-indicating that this is the first epoch and the
-replica will be a secondary.
-An \emph{epoch} represents a configuration of primary and
-secondary replicas; when a different replica becomes the primary,
-this indicates the start of a new epoch.
-The \texttt{ReplicaRuntime} acknowledges that it has become
-a secondary by responding with the same event type.
-The \psharp{} service sends an \texttt{ESecondaryCopyContextOp};
-this indicates what state the secondary has
-and, thus, what the primary should send to this secondary so that it can catch
-up.
-The \texttt{ESecondaryCopyContextOp} event is forwarded to the
-\texttt{ServiceRuntime}. 
-The \texttt{ServiceRuntime} then receives and handles an \texttt{EKillPrimary}
-event, which causes the second replica to become the new primary.
-Thus,
-the \texttt{ServiceRuntime} sends another \texttt{EEpochInfo}
-to the second \texttt{ReplicaRuntime}
-indicating that this is the second epoch
-and the replica will be a primary.
-As part of this change,
-the \texttt{ReplicaRuntime}
-sends an event to the \psharp{} service
-indicating that it should stop waiting for
-the state to be copied from the old primary to this replica,
-which is acknowledged by sending an event to the \texttt{ReplicaRuntime}.
-This event causes the \texttt{ReplicaRuntime}
-to send an \texttt{ESecondaryCopyStateDone}
-event to the \texttt{ServiceRuntime},
-which unfortunately responds with an event indicating that the
-\texttt{ReplicaRuntime} is now an \emph{active} secondary
-(i.e.\ a secondary that has caught up with the primary).
-However, this causes an assertion failure because
-the \texttt{ReplicaRuntime} is becoming a primary and, thus,
-cannot be a secondary.
-Our fix to this bug was to ensure that the 
-\texttt{ESecondaryCopyStateDone} event was marked as part of the first
-epoch (as the \texttt{ReplicaRuntime} had not yet acknowledged the
-change to primary); thus, 
-the \texttt{ServiceRuntime} ignores the event and does not try to make the
-replica an active secondary.
+\PTComment{Cut: We reverse-engineered Fabric as needed.}
+% \subsubsection{Fabric model correctness}
+% Our model does not attempt to
+% simulate the internals of Fabric accurately,
+% as its purpose is to find bugs in user code
+% and not in Fabric itself (which we assume to be correct).   
+% However,
+% due to lack of documentation,
+% it is not always clear how Fabric should behave
+% in certain scenarios.
+% Thus,
+% we ran several variants of a simple Fabric service
+% that logs calls into the user code in order to
+% reverse-engineer the actual behaviour of Fabric;
+% we ensured that our model has the same behaviour,
+% although this is an ongoing process as we encounter
+% additional scenarios.
 
-\subsubsection{CScale}
-The system that we wished to test is \emph{CScale}~\cite{X},
+\PTComment{Cut: We used systematic testing to find bugs in our model!}
+% A further problem is that our model may contain bugs.
+% In order to find bugs in our model effectively,
+% we wrote a \psharp{} service made up of a single machine
+% which takes the place
+% of the user code and translation layer in Figure \ref{fig:fabric_model},
+% for each \texttt{ReplicaRuntime}.
+% Thus, we were able to run this pure \psharp{} system
+% under \psharp{}'s systematic testing mode
+% and uncover many assertion failures within our model.
+% We tested a scenario where the primary fails at some non-deterministic point
+% within the execution.
+
+\PTComment{Cut: Example bug found in our model plus the fix.}
+% \textbf{Example Fabric model assertion failure:}
+% In the buggy trace,
+% the \texttt{ServiceRuntime}
+% sends an \texttt{EEpochInfo} event to the second
+% \texttt{ReplicaRuntime}
+% indicating that this is the first epoch and the
+% replica will be a secondary.
+% An \emph{epoch} represents a configuration of primary and
+% secondary replicas; when a different replica becomes the primary,
+% this indicates the start of a new epoch.
+% The \texttt{ReplicaRuntime} acknowledges that it has become
+% a secondary by responding with the same event type.
+% The \psharp{} service sends an \texttt{ESecondaryCopyContextOp};
+% this indicates what state the secondary has
+% and, thus, what the primary should send to this secondary so that it can catch
+% up.
+% The \texttt{ESecondaryCopyContextOp} event is forwarded to the
+% \texttt{ServiceRuntime}. 
+% The \texttt{ServiceRuntime} then receives and handles an \texttt{EKillPrimary}
+% event, which causes the second replica to become the new primary.
+% Thus,
+% the \texttt{ServiceRuntime} sends another \texttt{EEpochInfo}
+% to the second \texttt{ReplicaRuntime}
+% indicating that this is the second epoch
+% and the replica will be a primary.
+% As part of this change,
+% the \texttt{ReplicaRuntime}
+% sends an event to the \psharp{} service
+% indicating that it should stop waiting for
+% the state to be copied from the old primary to this replica,
+% which is acknowledged by sending an event to the \texttt{ReplicaRuntime}.
+% This event causes the \texttt{ReplicaRuntime}
+% to send an \texttt{ESecondaryCopyStateDone}
+% event to the \texttt{ServiceRuntime},
+% which unfortunately responds with an event indicating that the
+% \texttt{ReplicaRuntime} is now an \emph{active} secondary
+% (i.e.\ a secondary that has caught up with the primary).
+% However, this causes an assertion failure because
+% the \texttt{ReplicaRuntime} is becoming a primary and, thus,
+% cannot be a secondary.
+% Our fix to this bug was to ensure that the 
+% \texttt{ESecondaryCopyStateDone} event was marked as part of the first
+% epoch (as the \texttt{ReplicaRuntime} had not yet acknowledged the
+% change to primary); thus, 
+% the \texttt{ServiceRuntime} ignores the event and does not try to make the
+% replica an active secondary.
+
+% \subsubsection{CScale}
+
+The main system that we tested is \emph{CScale}~\cite{X},
 a big data-stream processing system that chains multiple Fabric services
 in a directed acyclic graph.
-A key challenge is that
-CScale contains inter-service communication
-that resolves services
-via Fabric but communicates
-using a non-Fabric remote procedure call protocol.
+A key challenge was that
+CScale contains inter-service communication.
+\PTComment{Cut: detail.}
+% that resolves services
+% via Fabric but communicates
+% using a non-Fabric remote procedure call protocol.
 In order to close the system,
-we first implemented service resolution in our Fabric model,
-which resolves Fabric service names to IP addresses.
-We then replaced the inter-service communication
+\PTComment{Cut: we implemented service resolution}
+% we first implemented service resolution in our Fabric model,
+% which resolves Fabric service names to IP addresses.
+%We then replaced 
+we replaced
+the inter-service communication
 classes in CScale with our own;
@@
 Thus,
-we converted a multi-process system
-that runs on the Fabric platform
-and 
-uses network communication
+we converted a distributed system
+that uses both Fabric and its own
+network communication protocol
+% that runs on the Fabric platform
+% and 
+% uses network communication
 into 
-a single process system that uses multiple threads,
-contains
-the Fabric model
-and the CScale services,
-and does not use network communication.
-Additional changes were needed to
-remove certain static (per-process) fields,
-as these were inadvertently
-being shared between services
-after making the system a single process.
-
+a closed single process system. 
+% that uses multiple threads,
+% contains
+% the Fabric model
+% and the CScale services,
+% and does not use network communication.
+\PTComment{Cut: we had to remove static fields.}
+% Additional changes were needed to
+% remove certain static (per-process) fields,
+% as these were inadvertently
+% being shared between services
+% after making the system a single process.
 A key challenge in our work

### 2015-09-15 experience.tex
updates in case studies

@@
 
-To test MigratingTable, we must supply it with backing tables.
+%To test MigratingTable, we must supply it with backing tables.
 We use \texttt{SpecTable} for this purpose as well, with \psharp choosing the actual result of each streaming read from the valid set.
@@
 % Convert to some theorem-like environment? ~ Matt
-\begin{quote}
-For every execution trace of a collection of MigratingTables backed by the same pair of \emph{old} and \emph{new} \texttt{SpecTable}s in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' \texttt{SpecTable} on the linearized input.
-\end{quote}
+%\begin{quote}
+%For every execution trace of a collection of MigratingTables backed by the same pair of \emph{old} and \emph{new} \texttt{SpecTable}s in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' \texttt{SpecTable} on the linearized input.
+%\end{quote}
 %
-We instrumented MigratingTable to report the intended \emph{linearization point} of each input call, which in our setting is always one of the corresponding \emph{backend calls} to the backend tables (often the last).
+
+%The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call, which in our setting is always one of the corresponding \emph{backend calls} to the backend tables (often the last).
 Specifically, after each backend call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
-This makes it possible to verify the correctness property as the model executes.
+This makes it possible to check the correctness property as the model executes.
 
-The \psharp environmental model of MigratingTable consists of a \psharp \texttt{Tables} machine containing all three \texttt{SpecTable} implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
+The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call.
+Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
+This makes it possible to check the correctness property as the model executes.
+
+The \psharp environmental model of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
 

### 2015-09-15 intro.tex
update

--- 2015-09-04 intro.tex
@@
-Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
-This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
-%
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}. T
+his is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
@@
 
-Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?} \PDComment{only the ones that the user has mocked using the \psharp APIs, some are handled automatically like async/await, we should probably rephrase}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production.
-Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
+Implementors of production distributed systems use many techniques to test their systems,
+such as unit testing, integration testing, stress testing, and fault injection.
+In spite of extensive use of these testing methods,
+many bugs that arise from subtle combinations of concurrency and failure events
+are missed during testing and get exposed only after a system has been put in production.
+Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} 
+and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
 
-We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
-The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
-\PDComment{We are missing a link for Google, but would be good if we can provide such a link, or for other companies like Facebook, I will try find something, but if anyone has a good suggestion please include}
-
-Amazon recently published an article~\cite{newcombe2015aws} that describes their use of TLA+~\cite{lamport1994temporal} to detect distributed system bugs and prevent them from reaching production.
-TLA+ is a powerful specification language for verifying distributed protocols, but it is unable to verify the code that is actually being executed.
-The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system.
-Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.\SCComment{we want to avoid making a general point that TLA+ has no guarantee to find all bugs, because our approach cannot either.
-Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."} \PDComment{We should rephrase.
-By the way the main contrast with TLA+ is that TLA+ works \emph{only} on a model and not the real code, whereas \psharp works on a real code (with a modeled environment).}
+We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development.
+The consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+The need for better testing techniques is not specific to Microsoft;
+other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
 
 In this work, our goal is to \emph{test what is being executed}.
-We present a new methodology for testing legacy distributed systems and uncovering bugs before these systems are released in the wild.
-We achieve this using \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides two key capabilities: (i) a flexible way of modeling the environment using simple language features; and (ii) a systematic concurrency testing framework that is able to capture and take control of all the nondeterminism in a real system (together with its modeled environment) and systematically explore execution paths to discover bugs.
-\SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?}
+We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
+We have implemented our testing method in \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides
+support for modeling, specification, and systematic testing of asynchronous systems.
+To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts---a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
+The \psharp testing engine then systematically explores behaviors of the test harness and validates them against the provided specifications.
 
-We present three case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol.
-Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time as they did not have an effective way to reproduce the bug and nail down the culprit.
-\psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.
-We show that bugs in production do not need a large setting to be reproduced; they can be reproduced in small settings with easy to understand traces.
+We applied our approach to test three production distributed storage systems implemented inside Microsoft for Windows Azure cloud computing
+platform---a distributed storage management system, a system for migrating data across partitions, and a big-data computing system.
+We uncovered numerous bugs in these systems using our approach.
+In particular, we uncovered a subtle liveness bug that was intermittently manifested during stress testing for months without being debugged.
+Our testing approach uncovered this bug in a very small setting, which made it easy to examine traces, identify, and eventually fix the problem.
+Our experience demonstrates that bugs in production do not need a large setting to be reproduced;
+they can be reproduced in small settings with easy to understand traces.
 
 To summarize, our contributions are as follows:
-
 \begin{itemize}
@@
 \end{itemize}
-
-\SCComment{let me give my overall comments about the current version of the paper here.
-This set of case studies are really cool.
-The fact that our approach is proven to be effective clearly is the primary selling point of the paper.
-Reviewers will love the results we have.\\
-About the writing, there may be 3 areas to improve:
-(1) the word "we" is overloaded, because "we" play multiple roles in this work.
-"We" built \psharp; "we", as developers, designed the TableMigration system; "we", as testers, also take the responsibility to test the legacy Azure vNext system.
-The current writing doesn't clearly differentiate these roles, so readers may wonder how much work does it requires if they want to adopt the approach.
-(2) The intro makes me feel that we compare with existing logic reasoning approaches such as TLA+, and point out the limitation of TLA+.
-Until much later in the paper, I realize that our approach doesn't have any logic reasoning ingredient in it.
-I suggest that the intro makes it clear that our approach is complementary to TLA+, not a competing approach.
-We should avoid give a comparison with TLA+, because it is not an apple-to-apple comparison.
-(3) Section 4 is the most important section about the approach.
-I feel that we spend too much text on section 4.3, which is about building the test harness.
-My concern is that readers' minds may be dominated by how we mocked all kinds of objects and override real methods by fake methods.
-They may feel that, in the end, we tested a highly simplified system.
-This is not true, because the extent manager is a REALLY complicated system (which is shown as a small box in fig.3).
-I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}  \PDComment{Really nice suggestions.
-I totally agree with the last point you make and we should make sure this is addressed (it is super important)!}

### 2015-09-15 paper.tex
update

--- 2015-09-15 paper.tex
@@
 In this paper, we present a new approach to testing distributed storage systems.
-Our approach significantly improves test coverage and allows us to discover during testing errors that are often caught after deployment.
+Our approach enables the developer to \emph{test what is being executed} and is engineered to work on production code.
+It significantly improves coverage of important system behaviors and allows us to discover during testing errors that are often caught after deployment.
 
-Our method for testing distributed storage systems is based on \psharp, an extension to the \csharp language,
-that provides support for modeling, specification, and systematic testing of asynchronous systems.
-We have used \psharp to test three distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing platform.
-To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts---a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
-The \psharp testing engine then systematically explores behaviors of the test harness and validates them against the provided specifications.
-
-Our approach enables the developer to \emph{test what is being executed} and is engineered to work on production code.
-We have used our approach to find, reproduce, and fix numerous bugs in production systems;
-these bugs are exhibited by executions with subtle combination of concurrency and failures, making them extremely difficult
-to find with traditional testing techniques.
+We have used our approach to test three distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing platform.
+In the process, we found, reproduced, and fixed numerous bugs in these systems.
+The bugs discovered by our method are exhibited by executions with subtle combination of concurrency and failures,
+making them extremely difficult to find with traditional testing techniques.
 An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace,

### 2015-09-15 evaluation.tex
update in the evaluation

@@
-We report our experience of applying \psharp on the three case studies discussed in this paper. We aim to answer the two following questions:
+We report our experience of applying \psharp on the three case studies discussed in this paper. We aim to answer the following two questions:
 
@@
 It may have been interesting to report results for specific test cases written before we knew the bugs as mentioned in \ref{sec:mtable:input}, but we did not do so in this study.
+\PDComment{this paragraph needs to be trimmed down probably}
+
+One of the bugs, QueryStreamedBackUpNewStream, that we found using \psharp in MigratingTable stands out because it reflects the type of oversight that tends to occur as designs evolve.
+Although this bug could be potentially found with other approaches, \psharp discovered it in a matter of seconds.
+The developers spend about 10 minutes analyzing the trace to diagnose what was happening; granted, this was after days of experience analyzing MigratingTable traces.
+The developers had also added their own trace output to the \psharp test harness, since \psharp's built-in trace output is too low-level and does not include event payloads and other diagnostic data that is not passed between machines.
+
+%We started from the failure symptom: the virtual stream returned end-of-stream when according to the reference \texttt{SpecTable}, it should have returned an additional row with key 4.
+We filtered the trace for actions by the same service machine and saw that $s_O$ was closed before the virtual stream had returned the row with key 4, but $s_N$ had already advanced past key 4 before the migrator inserted the row in the new table.
+At first we found this phenomenon hard to believe, but soon we were convinced it reflected a gap in our design.
+
+%This bug, which we named QueryStreamedBackUpNewStream, is in the implementation of a streaming read from the virtual table, which should return a stream of all rows in the table sorted by key.
+%The essential implementation idea is to start streams $s_O$, $s_N$ from the old and new backend tables and merge the sorted streams by keeping track of the next row in each stream and returning the row with the lesser key.
+In parallel, the migrator job is concurrently copying rows from the old table to the new table; we had satisfied ourselves that this concurrency would not cause any problems.
+However, then we added support to the migrator job to delete the old table when it finishes copying, which triggers the virtual stream to close $s_O$.
+Suppose the virtual stream is in a state in which the next row in $s_O$ has key $k_O$ and the next row in $s_N$ has key $k_N$, where $k_O < k_N$.
+Further suppose that before the next read from the virtual stream, the migrator job copies a row with key $k$ ($k_O < k < k_N$) from the old table to the new table and then deletes the old table.
+Since $s_O$ has not yet returned this row when it is closed and $s_N$ has already advanced to $k_N$, the row with key $k$ will be missed by the virtual stream.
+A similar problem can occur if $s_N$ does not reflect rows inserted into the new table by the migrator job after $s_N$ is started, as allowed by the \texttt{IChainTable} specification.
+Restarting $s_N$ when the old table is deleted fixes both variants of the bug.
+
+% It might be nice to include excerpts of a trace like in migration-bug3-explanation.pptx.
+Unfortunately, the trace in migration-bug3-explanation.pptx doesn't match my recollection of the original diagnosis, which I want to write about truthfully (the former looks like it involves a stale streaming read, while I'm fairly sure the latter involved only the $k_O < k < k_N$ case).
+If it's important, I could try to get a new trace consistent with the original diagnosis.
+~ Matt
+
+%This bug took us only about 10 minutes to diagnose from the trace; granted, this is after we had days of experience analyzing MigratingTable traces and had added our own trace output to the test harness, since \psharp's built-in trace output is too low-level and does not include event payloads and other diagnostic data that is not passed between machines.
+
+%We started from the failure symptom: the virtual stream returned end-of-stream when according to the reference \texttt{SpecTable}, it should have returned an additional row with key 4.
+We filtered the trace for actions by the same service machine and saw that $s_O$ was closed before the virtual stream had returned the row with key 4, but $s_N$ had already advanced past key 4 before the migrator inserted the row in the new table.
+At first we found this phenomenon hard to believe, but soon we were convinced it reflected a gap in our design.
 

### 2015-09-15 paper.tex

@@
 Testing distributed systems is challenging due to multiple sources of nondeterminism in the execution, e.g., arbitrary interleavings between event handlers and unexpected node failures.
-State-of-the-art testing methods commonly used in industry today ---unit, integration, and stress testing--- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
+State-of-the-art testing methods commonly used in industry today --- unit, integration, and stress testing --- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
 In this paper, we present a new approach to testing distributed storage systems.

### 2015-09-15 overview.tex
merged sections 3 and 5

@@
 However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.
-
-The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths.
-However, this is easier said than done when testing production systems.
-
-\subsection{Types of bugs}
-\label{sec:bg:bugs}
-
-We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
-
-\begin{description}
-\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.
-
-\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
-\end{description}
-
-\noindent
-A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
-An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.
-
-Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
-The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
-There was an actual bug in the system, that led this property to fail for some very rare executions.
-Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
-
-\subsection{The \psharp framework}
-\label{sec:method:psharp}
 
@@
 
-Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code.
-Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.
-
-A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
-
-\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.
-
-%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.
-
-\subsection{Overview of our approach}
-\label{sec:method:model}
-
-In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows.
-First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine.
-The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp.
-However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext.
-Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.
-
-To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach.
 Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
@@
 
-In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
+In Section~\ref{sec:method}, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
 In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
@@
 
-In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.
+\subsection{Specifications}
+\label{sec:bg:bugs}
+
+We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
+
+\begin{description}
+\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.
+A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
+An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.
+
+\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
+Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
+The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
+There was an actual bug in the system, that led this property to fail for some very rare executions.
+Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
+\end{description}
+
+We now describe how safety and liveness specifications are provided in our system.
+
+\subsubsection{Safety}
+\label{sec:psharp:safety}
+
+Safety property specifications can be encoded in \psharp by using the provided \texttt{Assert(...)} method.
+This method takes as argument a predicate, which if evaluates to false denotes a safety property violation.
+The programmer is also free to use other custom assertion APIs (as \psharp executes actual code), but to enable \psharp to recognize such APIs, the suggested approach is to override them to call the \psharp \texttt{Assert(...)} method.
+
+\psharp also provides a way to specify global assertions by using \emph{monitors}, special state machines that
+can only receive events, but not send.
+To declare a monitor, the programmer has to inherit from the \psharp \texttt{Monitor} class.
+A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it; as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is an event and an optional payload.
+This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls, as the monitor has to be called deterministically (and not be interleaved).
+
+\subsubsection{Liveness}
+\label{sec:psharp:liveness}
+
+Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states.
+Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
+Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
+Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
+
+The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states.
+A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g.
+a node has failed but a new one has not come up yet).
+A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.
+
+One can imagine a liveness monitor as a \emph{thermometer}: as the program executes and the liveness property is not satisfied, the liveness monitor stays in the hot state, which infinitely raises the temperature.
+If the liveness property is ever satisfied, and thus the liveness monitor transitions to a cold state, the temperature is instantly reduced to normal levels.
+Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties.
+The liveness monitors in \psharp are based on previous work~\cite{?}.
+
+\subsection{Systematic testing}
+In this section, we present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.
+
+A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
+
+The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
+A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions.
+We now give an overview of how this works, while more details can be found in the original paper~\cite{deligiannis2015psharp}.
+
+When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic testing engine} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
+The engine attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order.
+This approach is based on systematic concurrency testing~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} (SCT) techniques that have been previously developed for testing shared memory programs.
+
+The systematic testing engine serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
+The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
+The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
+
+We have implemented two schedulers inside the \psharp systematic testing engine: \emph{random} and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}.
+The random scheduler takes a completely random decision at every scheduling point, whereas the PCT scheduler uses randomization in a disciplined fashion to provide a probabilistic guarantee of finding a bug.
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
+The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far.
+Exposing an easy-to-use interface for creating new schedulers is inspired by previous work~\cite{DesaiQS15}.
+
+We designed the systematic testing engine in a way that enables easy debugging: after a bug is found, the engine can generate a trace that represents the buggy schedule.
+Note that this trace (in contrast to typical logs generating during production) is sequential.
+
+\subsection{Handling intra-machine concurrency}
+\label{sec:psharp:async}
+
+Vanilla \psharp is able to capture and take control of the \emph{inter-machine concurrency} due to message passing, but is unable to systematically explore any interleavings due to \emph{intra-machine concurrency} (e.g.
+async/await or TPL).
+This is problematic as nowadays, with multicore machines being a commodity, programmers tend to write multithreaded code to exploit shared memory architectures and increase the performance of individual components of a distributed system.
+
+As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+Asynchronous code using async/await is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
+The key idea behind async/await is that a method declared as \texttt{async} can use internally the \texttt{await} keyword which allows the thread executing the method to wait on a TPL task, or any other \emph{awaitable} object, \emph{without} blocking.
+This is achieved as follows.
+When an \texttt{await} statement is executed, the code following the await is wrapped as a TPL task \emph{continuation} and the method returns to the caller.
+This continuation executes when the awaitable object has completed.
+
+Developing a fully automatic and universal approach to handle intra-machine concurrency in .NET is very challenging, because there are many APIs that can be used for concurrency and synchronization (e.g.
+\texttt{System.Threading}, TPL, async/await, locks, semaphores) and each one has its own complexities.
+Although developers are willing to model the top-level message passing communication using \psharp, they are resistant in modeling the low-level threading and synchronization methods as this would be a very invasive procedure and such refactoring is unlikely to scale for legacy code.
+
+To test our case studies, we decided to focus our efforts in handling async/await as providing a robust solution for a subset of the .NET threading APIs is more feasible than trying to handle arbitrary threading and synchronization.
+Our solution involves using a custom task scheduler whenever the bug-finding mode of the \psharp runtime is enabled, instead of the default TPL task scheduler.
+The \psharp task scheduler inherits from the \texttt{TaskScheduler} class, a low-level API that is responsible for enqueueing TPL tasks into threads.
+
+Our approach works as follows.
+We start the task of the \emph{root} \psharp machine in our custom task scheduler.
+As long as any child tasks spawned by the root machine task do not explicitly start in another scheduler (including the default TPL scheduler), then they will be scheduled for execution in our custom task scheduler.
+Our custom scheduler intercepts the call to enqueue a task, creates a special machine called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine.
+The constructor of a \texttt{TaskMachine} takes as an argument a TPL task and stores it in an field.
+The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it will start executing the wrapped task and then immediately wait on its completion.
+This forces the lifetime of the task to become the lifetime of the machine and vice versa.
+Thus, when the \psharp systematic testing scheduler schedules the \texttt{TaskMachine}, it will also schedule the corresponding task.
+This means that when a \psharp program makes a call to an \texttt{async} method, then the task created in the backend will be automatically wrapped and scheduled.
+
+For \texttt{await} statements, no special treatment is required because \texttt{await} statements are compiled into a continuation.
+The task associated with the continuation will also be wrapped and scheduled accordingly.
+
+Regarding support for handling non-async/await concurrency primitives, we are currently investigate feasible approaches.

### 2015-09-15 paper.tex
merged sections 3 and 5
changing the order of institutions to alphabetical

@@
 {\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
-$^\dagger$Imperial College London, $^\star$Microsoft, $^\diamond$Massachusetts Institute of Technology\\
+$^\dagger$Imperial College London, $^\diamond$Massachusetts Institute of Technology, $^\star$Microsoft\\
 } % end author
@@
 
-\section{Testing Distributed Systems}
+\section{Our Approach to Testing Distributed Systems}
 \label{sec:overview}
@@
 
-\section{The \psharp Systematic Testing Engine}
-\label{sec:psharp}
-
-\input{psharp}
+%% \section{The \psharp Systematic Testing Engine}
+%% \label{sec:psharp}
+%% \input{psharp}
 

### 2015-09-15 intro.tex
minor fixes

@@
-Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}. T
-his is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
+This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
@@
 
-We applied our approach to test three production distributed storage systems implemented inside Microsoft for Windows Azure cloud computing
-platform---a distributed storage management system, a system for migrating data across partitions, and a big-data computing system.
+We applied our approach to test three production distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing
+platform: a distributed storage management system; a system for migrating data across partitions; and a big-data computing system.
 We uncovered numerous bugs in these systems using our approach.

### 2015-09-15 paper.tex
minor fixes

@@
 Testing distributed systems is challenging due to multiple sources of nondeterminism in the execution, e.g., arbitrary interleavings between event handlers and unexpected node failures.
-State-of-the-art testing methods commonly used in industry today --- unit, integration, and stress testing --- are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
+State-of-the-art testing methods commonly used in industry today---unit, integration, and stress testing---are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
 In this paper, we present a new approach to testing distributed storage systems.

### 2015-09-15 experience.tex
merged 5 to 4

@@
-\section{Liveness Bug in Azure Storage vNext}
-\label{sec:cases:azurestore}
-
-It took only tens of seconds before the testing harness reported the first occurrence of a liveness bug. Upon examining the debug trace, the vNext developers were able to confirm the bug.
-
-However, the trace didn't include enough details, which prevented the developers from identifying a root cause.
-Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration.
-After several iterations, the developers were able to pinpoint the exact culprit and immediately proposed a solution to fix the bug.
-Once the proposed solution was implemented, the developers again ran the testing harness, which reported no more bug for $100,000$ iterations in tens of minutes.
-
-The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} fails one of the \texttt{ExtentNode} and launches a new one.
-The \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.
-
-Here is one particular execution sequence resulting in the bug.
-$i)$ EN$_0$ fails and is detected by the EN expiration loop; $ii)$ EN$_0$ is removed from the EN map; $iii)$ the extent center is updated and the replica count drops from 3, the target, to 2; $iv)$ \texttt{ExtMgr} receives a sync report from EN$_0$; $v)$ the extent center is updated and the replica count increases from 2 to 3 again.
-This is problematic because, on one hand, the replica count is equal to the target, so the extent repair loop never schedules any repair task.
-On the other hand, there are only two true replicas in the system, one fewer than the target.
-This execution sequence leads to one replica missing.
-Conceivably, repeating this another two times would result in all replicas missing, while \texttt{ExtMgr} still thinks all replicas are healthy.
-If deployed in production, such bug would have caused a very serious incident of customer data loss.
-
-The culprit is in $iv)$, where \texttt{ExtMgr} receives a sync report from EN$_0$ after deleting the EN.
-This may occur in \psharp because of arbitrary message interleaving.
-It may also occur, albeit much less frequently, in the stress testing due to messages being delayed in the network.
-This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger.
-In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution verified effectively, all of which should vastly increase the productivity of distributed storage system development.
-
-\section{Additional Case Studies}
-\label{sec:cases:other}
-
 The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system.

### 2015-09-15 methodology.tex
merged 5 to 4

@@
 
+\subsection{Liveness Bug in Azure Storage vNext}
+\label{sec:method:azurestore}
+
+It took only tens of seconds before the testing harness reported the first occurrence of a liveness bug. Upon examining the debug trace, the vNext developers were able to confirm the bug.
+
+However, the trace didn't include enough details, which prevented the developers from identifying a root cause.
+Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration.
+After several iterations, the developers were able to pinpoint the exact culprit and immediately proposed a solution to fix the bug.
+Once the proposed solution was implemented, the developers again ran the testing harness, which reported no more bug for $100,000$ iterations in tens of minutes.
+
+The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} fails one of the \texttt{ExtentNode} and launches a new one.
+The \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.
+
+Here is one particular execution sequence resulting in the bug.
+$i)$ EN$_0$ fails and is detected by the EN expiration loop; $ii)$ EN$_0$ is removed from the EN map; $iii)$ the extent center is updated and the replica count drops from 3, the target, to 2; $iv)$ \texttt{ExtMgr} receives a sync report from EN$_0$; $v)$ the extent center is updated and the replica count increases from 2 to 3 again.
+This is problematic because, on one hand, the replica count is equal to the target, so the extent repair loop never schedules any repair task.
+On the other hand, there are only two true replicas in the system, one fewer than the target.
+This execution sequence leads to one replica missing.
+Conceivably, repeating this another two times would result in all replicas missing, while \texttt{ExtMgr} still thinks all replicas are healthy.
+If deployed in production, such bug would have caused a very serious incident of customer data loss.
+
+The culprit is in $iv)$, where \texttt{ExtMgr} receives a sync report from EN$_0$ after deleting the EN.
+This may occur in \psharp because of arbitrary message interleaving.
+It may also occur, albeit much less frequently, in the stress testing due to messages being delayed in the network.
+This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger.
+In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution verified effectively, all of which should vastly increase the productivity of distributed storage system development.
+
 %%\subsection{Modeling the environment}

### 2015-09-15 evaluation.tex
update in evaluation

@@
 
-Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using the random and the PCT schedulers.
+Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using the random and PCT schedulers.
+We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
 The CS column shows which case study corresponds to each bug: 1 is for the Azure Storage vNext; and 2 is for the Live Migration Table.
@@
 At first we found this phenomenon hard to believe, but soon we were convinced it reflected a gap in our design.
-
-Controlled random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.

### 2015-09-16 evaluation.tex
typos

@@
 
-Modeling the Live Migration Table required \PDComment{waiting confirmation from Matt}.
-This case study is interesting because the development of the actual system and its \psharp environmental model occurred side-by-side.
-This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and at a later stage of the development process.
+Modeling the Live Migration Table required \PDComment{waiting confirmation from
+Matt}. The modeling effort was done in parallel with the development of the actual system. 
+This is in contrast with the other two case studies discussed in this paper, where the modeling 
+activity occurred independently and at a later stage of the development process.
 
-Modeling Fabric required approximately 4-5 months.
-Although this is a significant amount of time, it is a one time effort activity.
-Our plan is to reuse the developed Fabric model for testing arbitrary user services built for the Azure Service Fabric system.
+Modeling Fabric required approximately 5 person months. Although this is a
+significant amount of time, it is a one time effort (once per release). Our plan is to reuse the developed Fabric model 
+for testing arbitrary user services built for the Azure Service Fabric system.
 
@@
 The random seed for both schedulers was generated in each iteration using the \texttt{DateTime.Now.Millisecond} API which returns the current time in milliseconds.
-The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps to execute of 500.
+The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps of 500.
 All reported times are in seconds.

### 2015-09-16 experience.tex
typos

@@
 
-The Live Azure Table Migration (MigratingTable) is a library for \emph{transparently migrating} a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
-This case study is interesting because its developers built the \psharp model of the system while developing the library to increase confidence in its implementation; indeed they discovered many bugs during this \emph{co-development} process (see Section~\ref{}).
-Another interesting aspect of this system is that it heavily uses the \texttt{async}/\texttt{await} primitives, which were not supported in the original \psharp runtime, and thus required us to extend \psharp with support for intra-machine concurrency (see Section~\ref{}).
-Finally, a lot of complex safety properties were written to test MigratingTable with \psharp, whereas the vNext model focused on a single liveness property.
+The Live Azure Table Migration (MigratingTable) is a library for
+\emph{transparently migrating} a data set between tables in an Azure Storage
+service \emph{while} an application is accessing this data set. This case study
+is interesting because its developers built the \psharp model of the system
+while developing the library to increase confidence in its implementation;
+indeed they discovered many bugs during this \emph{co-development} process (see
+Section~\ref{}). Another interesting aspect of this system is that it heavily
+uses the \texttt{async}/\texttt{await} primitives, which were not supported in
+the original \psharp runtime, and thus required us to extend \psharp with
+support for intra-machine concurrency (see Section~\ref{}). Finally, multiple 
+complex safety properties were written to test MigratingTable with \psharp, 
+whereas the vNext model focused on a single liveness property.
 
@@
 
-The \psharp testing goal was to ensure that when multiple application processes, implementing the \texttt{IChainTable} interface, issue read and write operations to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of \texttt{IChainTable} for the combined \emph{input history}.
+The testing goal was to ensure that when multiple application processes, implementing the \texttt{IChainTable} interface, issue read and write operations to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of \texttt{IChainTable} for the combined \emph{input history}.
 

### 2015-09-16 experience.tex
Changes to Fabric section.

@@
 
-\emph{Azure Service Faric} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
+\emph{Azure Service Fabric}\footnote{\url{http://azure.microsoft.com/en-gb/campaigns/service-fabric/}} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines.
 %The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
 In order to make the user-written service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on a different node in the cluster.
-\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
+%\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
 %One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
@@
 
-\PTComment{Could be improved:}
 Our primary goal was to create a \psharp{} model of Fabric to allow
@@
 by the \psharp{} runtime.
-Unlike the other case studies,
-this model is written once
+The model is written once
 to include all behaviours of Fabric
 so that it can be used again and again to test arbitrary Fabric services.
+This was the largest among the case studies
+and required multiple rounds of debugging.
+The availability of systematic testing
+allowed us to debug the model to
+a point where reported assertion violations indicate a bug in the
+user service.
+Without systematic testing,
+even bugs in the model would have been hard to find.
+Note that we model the lowest Fabric API layer (\texttt{Fabric.dll})
+which is not documented for use externally.
+Eventually, we will lift the model
+to higher layers
+but, for this paper, we study internally-developed services
+that directly target the the lowest layer.
 
-\PTComment{Cut: prior work created a Fabric model\ldots}
+%\PTComment{Cut: prior work created a Fabric model\ldots}
 % Prior work~\cite{deligiannis2015psharp}
@@
 
-\PTComment{Cut: Figure of model.}
+%\PTComment{Cut: Figure of model.}
 % \begin{figure}[thb]
@@
 
-\PTComment{Cut: overview of model.}
+%\PTComment{Cut: overview of model.}
 % An overview of our Fabric model is shown in Figure~\ref{fig:fabric_model}.
@@
 
-\PTComment{Cut: Description of how replication works in the model, including how we modeled at a fine granularity.}
+%\PTComment{Cut: Description of how replication works in the model, including how we modeled at a fine granularity.}
 % \textbf{Replication example:} In order to replicate a state-mutating operation,
@@
 
-\PTComment{Cut: We reverse-engineered Fabric as needed.}
+%\PTComment{Cut: We reverse-engineered Fabric as needed.}
 % \subsubsection{Fabric model correctness}
@@
 
-\PTComment{Cut: We used systematic testing to find bugs in our model!}
+%\PTComment{Cut: We used systematic testing to find bugs in our model!}
 % A further problem is that our model may contain bugs.
@@
 
-\PTComment{Cut: Example bug found in our model plus the fix.}
+%\PTComment{Cut: Example bug found in our model plus the fix.}
 % \textbf{Example Fabric model assertion failure:}
@@
 CScale contains inter-service communication.
-\PTComment{Cut: detail.}
+%\PTComment{Cut: detail.}
 % that resolves services
@@
 In order to close the system,
-\PTComment{Cut: we implemented service resolution}
+%\PTComment{Cut: we implemented service resolution}
 % we first implemented service resolution in our Fabric model,
@@
 % and does not use network communication.
-\PTComment{Cut: we had to remove static fields.}
+%\PTComment{Cut: we had to remove static fields.}
 % Additional changes were needed to

### 2015-09-16 experience.tex
Small change.

@@
 %The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
-In order to make the user-written service \emph{reliable}, Fabric launches several copies (\emph{replicas}) of the service, where each copy runs as a separate process on a different node in the cluster.
+In order to make the user-written service \emph{reliable}, Fabric launches several \emph{replicas} (copies) of the service, where each replica runs as a separate process on a different node in the cluster.
 %\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
@@
 to higher layers
-but, for this paper, we study internally-developed services
+but for this paper we study internally-developed services
 that directly target the the lowest layer.

### 2015-09-16 related.tex
update in related work

@@
-A significant amount of research has been conducted on how to analyze and test distributed systems~\cite{lamport1994temporal, schuppan2004efficient, killian2007life, gupta2008diecast, yang2009modist}, but a lot of these techniques either have significant limitations, or cannot be easily applied in a production environment, due to many complexities that are outside the scope of a research project.
+\psharp is most closely related to \textsc{MoDist}~\cite{yang2009modist}, a model checker that can be applied on unmodified distributed systems.
+To achieve this, the tool uses binary instrumentation (reuses the instrumentation engine of the D$^3$S testing tool) to expose all actions of the system under test, and then uses a model checking engine to systematically explore these actions.
+\textsc{MoDist} also provides a virtual clock manager that can simulate timeouts and accelerate the passage of time.
+A major limitation of \textsc{MoDist} is that it must be applied on a whole system, which does not scale for production systems.
+In contrast, \psharp has built-in support for modeling the environment of individual components of a system, which allows the tool to achieve scalability.
+Further, \psharp is an extension of the \csharp language and has built-in systematic testing support; it does not need to perform binary instrumentation which can be computationally expensive.
+
+\textsc{MaceMc}~\cite{killian2007life} is a model checker for distributed systems implemented in the \textsc{Mace} language.
+The focus of \textsc{MaceMc} is to find liveness property violations using an algorithm based on bounded random walk and the use of heuristics.
+\psharp differs from \textsc{MaceMc} in that it can be applied on legacy code written in a mainstream language, whereas a system to be tested with \textsc{MaceMc} has to be written in \textsc{Mace}, which makes \textsc{MaceMc} harder to be applied in an industrial setting.
+Further, \psharp uses a simpler, but effective, algorithm for detecting liveness bugs.
+
+\textsc{Fate} and \textsc{Destini} is a framework for systematically injecting (combination of) failures in distributed systems~\cite{gunawi2011fate}.
+
+DieCast~\cite{gupta2008diecast}
 
@@
 
-In Verdi~\cite{wilcox2015verdi}, a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
+Another approach is Verdi~\cite{wilcox2015verdi}, where a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
 Verdi cannot find liveness bugs.
-\psharp is also more production-friendly (works on a mainstream language).
-\PDComment{say few more stuff}
-
-D3S~\cite{liu2008d3s} \PDComment{should read the paper, they say they are debugging running systems}
+\psharp is also more production-friendly: it works on \csharp, which is a mainstream language.

### 2015-09-16 experience.tex

@@
 but for this paper we study internally-developed services
-that directly target the the lowest layer.
+that target the lowest layer.
 
@@
 % using a non-Fabric remote procedure call protocol.
-In order to close the system,
+To close the system,
 %\PTComment{Cut: we implemented service resolution}
@@
 %We then replaced 
-we replaced
-the inter-service communication
-classes in CScale with our own;
-our replacement includes the ability to map
-URLs to \psharp{} machines
-so that remote procedure calls are now implemented
+we 
+replaced the relevant classes
+% replaced the inter-service communication classes with our own,
+% ;
+% our replacement includes the ability to map
+% URLs to \psharp{} machines
+so that remote procedure calls are implemented
 by sending and receiving \psharp{} events.

### 2015-09-16 experience.tex
update in experience

@@
 \label{sec:cases:fabric}
-
-- \psharp generalized beyond vNext. \psharp is not only applicable to vNext but in many systems.\\
-- (Advanced) safety properties?\\
-- What we developed is a model that other people can reuse. Plugin the model and test the user application.\\
-- Example of advanced modeling capability of \psharp. One component talks to another sophisticated API. We demonstrate environmental modeling expressivity in system engineering.
 

### 2015-09-16 related.tex
update in related work

@@
 \textsc{Fate} and \textsc{Destini} is a framework for systematically injecting (combination of) failures in distributed systems~\cite{gunawi2011fate}.
-
-DieCast~\cite{gupta2008diecast}
+This framework focuses in exercising failure scenarios, whereas \psharp can be used for testing generic safety and liveness properties of distributed systems.
+\psharp also provides language extensions and modeling capabilities that further differentiate it from prior work.
 
@@
 
-Another approach is Verdi~\cite{wilcox2015verdi}, where a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
+Another relevant approach is Verdi~\cite{wilcox2015verdi}, where a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
 Verdi cannot find liveness bugs.

### 2015-09-16 experience.tex
update

@@
 while developing the library to increase confidence in its implementation;
-indeed they discovered many bugs during this \emph{co-development} process (see
-Section~\ref{}). Another interesting aspect of this system is that it heavily
+indeed they discovered many bugs during this \emph{co-development} process.
+Another interesting aspect of this system is that it heavily
 uses the \texttt{async}/\texttt{await} primitives, which were not supported in
 the original \psharp runtime, and thus required us to extend \psharp with
-support for intra-machine concurrency (see Section~\ref{}). Finally, multiple 
+support for intra-machine concurrency (see Section~\ref{sec:psharp:async}). Finally, multiple 
 complex safety properties were written to test MigratingTable with \psharp, 

### 2015-09-16 intro.tex
update intro

@@
 
-Implementors of production distributed systems use many techniques to test their systems,
+Developers of production distributed systems use many techniques to test their systems,
 such as unit testing, integration testing, stress testing, and fault injection.
@@
 
-We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development.
+We interviewed technical leaders and senior managers in the Microsoft Azure team regarding the top problems in distributed system development.
 The consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
@@
 
-In this work, our goal is to \emph{test what is being executed}.
+Arguably, the most advanced practice in industry so far appeals in Amazon Web Services where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''.
+The gist of this approach is to extract high level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
+While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\ref{}.
+
+In our opinion, high level specification is necessary but not sufficient, due to the gap between the specification and the executable code that ultimately powers the distributed systems.
+In this work, our goal is to bridge this gap with a new framework that allows verification techniques for high level specification directly applicable in the executable code.
+Instead of requiring developers to switch to unfamiliar special programming language~\ref{}, or manually annotate and instrument their code~\ref{}, we allow the developers to {\em test unmodified production code} by writing testing harness in our framework with familiar programming language.
+Based on our experiences, this significantly lowered acceptance barrier and was the key to the adoption of the framework in the Microsoft Azure team.
+
 We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.

### 2015-09-16 evaluation.tex

@@
 For the vNext case study, the random scheduler was able to reproduce the bug in less than 10 seconds.
-The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in Section~\ref{} we leave the program to run for a long time before checking if the liveness property holds.
+The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in Section~\ref{sec:psharp:testing} we leave the program to run for a long time before checking if the liveness property holds.
 The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.

### 2015-09-16 experience.tex

@@
 
-The main system that we tested is \emph{CScale}~\cite{X},
+The main system that we tested is 
+\emph{CScale}~\cite{faleiro2012},
 a big data-stream processing system that chains multiple Fabric services

### 2015-09-16 overview.tex

@@
 We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:exp}).
+We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}).
 Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
@@
 Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties.
-The liveness monitors in \psharp are based on previous work~\cite{?}.
+The liveness monitors in \psharp are based on previous work~\cite{Desai:EECS-2015-198}.
 
 \subsection{Systematic testing}
+\label{sec:psharp:testing}
 In this section, we present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.

### 2015-09-17 intro.tex
update abstract/intro

--- 2015-09-15 intro.tex
@@
 
-We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development.
-The consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+We interviewed technical leaders and senior managers in the Microsoft Azure team regarding the top problems in distributed system development.
+The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
 The need for better testing techniques is not specific to Microsoft;
@@
 
-In this work, our goal is to \emph{test what is being executed}.
-We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
-We have implemented our testing method in \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides
-support for modeling, specification, and systematic testing of asynchronous systems.
-To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts---a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
+Arguably, the most advanced practice in industry so far appears in Amazon Web Services where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\ref{}.
+The gist of their approach is to extract high level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
+While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\ref{}.
+
+We have found high level specification is necessary but not sufficient, due to the gap between the specification and the executable code.
+%that ultimately powers the distributed systems.
+In this work, our goal is to bridge this gap with a new framework that makes verification techniques for high level specification directly applicable to the executable code.
+Our framework is different from prior approaches where the developers are required to either switch to unfamiliar special programming language~\ref{}, or manually annotate and instrument production code~\ref{}.
+Instead, the developers can test {\em unmodified production code} by writing test harness in our framework using familiar programming language.
+Based on our experiences, this significantly lowered acceptance barrier and was the key to adoption in the Microsoft Azure team.
+
+%We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
+We have implemented our framework in \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides support for modeling, specification, and systematic testing of asynchronous systems.
+To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
 The \psharp testing engine then systematically explores behaviors of the test harness and validates them against the provided specifications.
 
-We applied our approach to test three production distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing
-platform: a distributed storage management system; a system for migrating data across partitions; and a big-data computing system.
-We uncovered numerous bugs in these systems using our approach.
-In particular, we uncovered a subtle liveness bug that was intermittently manifested during stress testing for months without being debugged.
+So far, three distributed storage systems in the Microsoft Azure team have adopted our framework: Azure Storage vNext, Azure Table Live Migration, and Azure Service Fabric.
+%a distributed storage system; a system for migrating data across partitions; and a big-data computing system.
+Numerous bugs were uncovered in these systems using our approach,
+including, most notably, a subtle liveness bug that was intermittently manifested during stress testing for months without being fixed.
 Our testing approach uncovered this bug in a very small setting, which made it easy to examine traces, identify, and eventually fix the problem.

### 2015-09-17 paper.tex
merged 5 to 4
added a draft conclusion
update abstract/intro

@@
 \subsection*{Abstract}
-Testing distributed systems is challenging due to multiple sources of nondeterminism in the execution, e.g., arbitrary interleavings between event handlers and unexpected node failures.
-State-of-the-art testing methods commonly used in industry today---unit, integration, and stress testing---are unable to cover many behaviors, which results in tricky bugs being missed during testing and getting exposed in production.
-In this paper, we present a new approach to testing distributed storage systems.
-Our approach enables the developer to \emph{test what is being executed} and is engineered to work on production code.
-It significantly improves coverage of important system behaviors and allows us to discover during testing errors that are often caught after deployment.
+Testing distributed systems is challenging due to multiple sources of nondeterminism.
+Conventional testing approaches are ineffective in preventing serious but subtle bugs from reaching production.
+Advanced industry practice, such as AWS using formal methods~\ref{}, only verifies high level specification and falls short of checking executable code.
+In this paper, we present a new framework for testing distributed storage systems.
+Our approach makes verification techniques for high level specification directly applicable to the executable code.
+It therefore significantly improves coverage of important system behaviors and uncovers during testing subtle bugs that are often caught after deployment.
 
-We have used our approach to test three distributed storage systems implemented inside Microsoft for the Windows Azure cloud computing platform.
-In the process, we found, reproduced, and fixed numerous bugs in these systems.
+Our framework has been applied to three distributed storage systems in the Microsoft Azure platform.
+In the process, numerous bugs were identified, reproduced, fixed and confirmed.
 The bugs discovered by our method are exhibited by executions with subtle combination of concurrency and failures,
-making them extremely difficult to find with traditional testing techniques.
+making them extremely difficult to find with conventional testing techniques.
 An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace,
@@
 
-%% \section{The \psharp Systematic Testing Engine}
-%% \label{sec:psharp}
-%% \input{psharp}
-
-%\section{Experience Report}
-%\label{sec:exp}
+\section{Additional Case Studies}
+\label{sec:cases}
 
@@
 \section{Conclusion}
-\label{sec:concl}
+\label{sec:conclusion}
 
-Draft.
+\input{conclusion}
 

### 2015-09-17 overview.tex

@@
 
-All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.
+All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
 Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
-However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.
+However, these techniques are not effective for testing distributed systems, as they are not able to control the many sources of nondeterminism.
 
-Our goal in this work is to \emph{test what is being executed}.
-To achieve this we use \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
-
-The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines.
-\psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems.
-This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g.
-Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).
+The goal of our work is to develop testing techniques to detect and debug Heisenbugs in distributed storage systems prior to deployment.
+To achieve our goal, we use \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
 
@@
 
-Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:
-
+Our approach of testing distributed systems using \psharp requires the developer to perform three key modeling tasks;
+these tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a case study.
 \begin{enumerate}
-\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines.
+\item
+The computation model underlying \psharp is communicating state machines.
+The environment of the system-under-test must be modeled using \psharp machines,
+while the real components of the system must be wrapped inside \psharp machines.
 We call this modeled environment the \psharp test harness.
 
-\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs.
-If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing.
-This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.
+\item
+The asynchrony due to message passing must be exposed using the \psharp APIs for sending events.
+Similarly, other sources of nondeterministically-generated events, such as failures and timer expiration,
+must also be explicitly modeled using the available \psharp APIs;
+This step allows the \psharp runtime to explore the nondeterministic choices during testing.
 
-\item Any other source of nondeterminism in the system or the environment (e.g.
-failures and timers), must be explicitly modeled using the available \psharp APIs.
-This would allow \psharp to capture the nondeterminism and control it during systematic testing.
+\item
+The correctness of each execution generated during testing must be specified.
+Specifications can be either \emph{safety} and \emph{liveness}~\cite{lamport1977proving}.
+A safety specification is a generalization of assertions;
+a safety violation is a finite trace leading to an erroneous state.
+A liveness specification is a generalization of nontermination;
+a liveness violation is an infinite trace that exhibits lack of progress.
 \end{enumerate}
-
-In Section~\ref{sec:method}, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp.
-In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool.
-We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.
+In principle, this modeling methodology is not specific to \psharp but can be used in combination with any other programming framework
+with equivalent capabilities.
 
@@
 We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g.
-\emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.
+Furthermore, the language features that are required to be used to connect the real code with the modeled code (e.g., \emph{virtual method dispatch}), are already being heavily used in production for testing purposes , which significantly lowers the bar for practitioners to embrace \psharp for testing.
 
@@
 \label{sec:bg:bugs}
+In  this section, we describe how safety and liveness specifications are expressed in our system.
 
-We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.
-
-\begin{description}
-\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.
-A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state.
-An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.
-
-\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
-Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states.
-The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.
-There was an actual bug in the system, that led this property to fail for some very rare executions.
-Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.
-\end{description}
-
-We now describe how safety and liveness specifications are provided in our system.
-
-\subsubsection{Safety}
-\label{sec:psharp:safety}
-
+{\bf Safety}.
 Safety property specifications can be encoded in \psharp by using the provided \texttt{Assert(...)} method.
@@
 
-\subsubsection{Liveness}
-\label{sec:psharp:liveness}
-
+{\bf Liveness}.
 Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states.

### 2015-09-17 intro.tex
made a pass over abstract/intro

@@
 We interviewed technical leaders and senior managers in the Microsoft Azure team regarding the top problems in distributed system development.
-The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production.
+The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes to production.
 The need for better testing techniques is not specific to Microsoft;
@@
 
-Arguably, the most advanced practice in industry so far appears in Amazon Web Services where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\ref{}.
-The gist of their approach is to extract high level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
+Recently, Amazon Web Services used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\ref{}.
+The gist of their approach is to extract high-level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
 While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\ref{}.
 
-We have found high level specification is necessary but not sufficient, due to the gap between the specification and the executable code.
+We have found that high-level specifications are necessary but not sufficient, due to the gap between the specification and the executable code.
 %that ultimately powers the distributed systems.
-In this work, our goal is to bridge this gap with a new framework that makes verification techniques for high level specification directly applicable to the executable code.
-Our framework is different from prior approaches where the developers are required to either switch to unfamiliar special programming language~\ref{}, or manually annotate and instrument production code~\ref{}.
-Instead, the developers can test {\em unmodified production code} by writing test harness in our framework using familiar programming language.
-Based on our experiences, this significantly lowered acceptance barrier and was the key to adoption in the Microsoft Azure team.
+Our goal is to bridge this gap. We propose a new framework that
+applies verification techniques for high-level specifications directly
+to the executable code. Our framework is different from prior
+approaches where the developers were required to either switch to an
+unfamiliar special programming language~\ref{}, or manually annotate
+and instrument code~\ref{}. Instead, we allow developers to test
+{\em unmodified production code} by writing test harnesses in a
+mainstream programming language. This
+significantly lowered the acceptance barrier for adoption by the
+Microsoft Azure team.
 
 %We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
-We have implemented our framework in \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides support for modeling, specification, and systematic testing of asynchronous systems.
-To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications for each behavior of the system under the test harness.
-The \psharp testing engine then systematically explores behaviors of the test harness and validates them against the provided specifications.
+Our framework is based on \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides support for modeling, specification, and systematic testing of asynchronous systems.
+To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
+The \psharp testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
 
-So far, three distributed storage systems in the Microsoft Azure team have adopted our framework: Azure Storage vNext, Azure Table Live Migration, and Azure Service Fabric.
+We have applied our framework to three distributed storage systems in Microsoft Azure: Azure Storage vNext, Azure Table Live Migration, 
+and Azure Service Fabric.
 %a distributed storage system; a system for migrating data across partitions; and a big-data computing system.
-Numerous bugs were uncovered in these systems using our approach,
-including, most notably, a subtle liveness bug that was intermittently manifested during stress testing for months without being fixed.
+We uncovered numerous bugs in these systems, including, most notably, a subtle liveness bug that only intermittently 
+manifested during stress testing for months without being fixed.
 Our testing approach uncovered this bug in a very small setting, which made it easy to examine traces, identify, and eventually fix the problem.
@@
 \item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms.
-\item Our infrastructure can test production code written in \csharp, which is a mainstream language.
+\item Our infrastructure can test production code written in \csharp.
 \item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.

### 2015-09-17 paper.tex
made a pass over abstract/intro

@@
 Conventional testing approaches are ineffective in preventing serious but subtle bugs from reaching production.
-Advanced industry practice, such as AWS using formal methods~\ref{}, only verifies high level specification and falls short of checking executable code.
+Formal techniques, such as AWS's use of TLA+~\ref{}, only verifies high level specification and falls short of checking executable code.
 In this paper, we present a new framework for testing distributed storage systems.
-Our approach makes verification techniques for high level specification directly applicable to the executable code.
-It therefore significantly improves coverage of important system behaviors and uncovers during testing subtle bugs that are often caught after deployment.
+Our approach applies verification techniques for high-level specifications directly on the executable code and
+significantly improves coverage of important system behaviors.
 
@@
 In the process, numerous bugs were identified, reproduced, fixed and confirmed.
-The bugs discovered by our method are exhibited by executions with subtle combination of concurrency and failures,
+These bugs required subtle combination of concurrency and failures,
 making them extremely difficult to find with conventional testing techniques.

### 2015-09-17 intro.tex
fix

@@
 
-Recently, Amazon Web Services used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\ref{}.
+Recently, Amazon Web Services used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
 The gist of their approach is to extract high-level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
-While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\ref{}.
+While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
 

### 2015-09-17 evaluation.tex
fixes

@@
 Environmental modeling is a core activity of using \psharp.
-It is required for \emph{closing the environment} of a system under test and making it amenable to systematic testing.
+It is required for \emph{closing the environment} of a system-under-test and making it amenable to systematic testing.
 Table~\ref{tab:stats} presents program statistics for our three case studies.
-We report: lines of code for the system under test (\#LoC); number of bugs found in the system (\#B); lines of \psharp code for the environmental model (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
+We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system (\#B); lines of \psharp code for the environmental model (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
 

### 2015-09-17 related.tex
fixes

@@
 \psharp is most closely related to \textsc{MoDist}~\cite{yang2009modist}, a model checker that can be applied on unmodified distributed systems.
-To achieve this, the tool uses binary instrumentation (reuses the instrumentation engine of the D$^3$S testing tool) to expose all actions of the system under test, and then uses a model checking engine to systematically explore these actions.
+To achieve this, the tool uses binary instrumentation (reuses the instrumentation engine of the D$^3$S testing tool) to expose all actions of the system-under-test, and then uses a model checking engine to systematically explore these actions.
 \textsc{MoDist} also provides a virtual clock manager that can simulate timeouts and accelerate the passage of time.

### 2015-09-17 experience.tex
fixes

@@
 
-\emph{Azure Service Fabric}\footnote{\url{http://azure.microsoft.com/en-gb/campaigns/service-fabric/}} (or \emph{Fabric} for short) is a platform and API for creating reliable services that execute on a cluster of machines.
+Azure Service Fabric\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}} (or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
 %The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 

### 2015-09-17 paper.tex
fix
update in url packages

@@
 
-\usepackage{url}
+\usepackage{hyperref}
 \usepackage{multirow}
@@
 Conventional testing approaches are ineffective in preventing serious but subtle bugs from reaching production.
-Formal techniques, such as AWS's use of TLA+~\ref{}, only verifies high level specification and falls short of checking executable code.
+Formal techniques, such as the use of TLA+ in AWS~\cite{newcombe2015aws}, only verifies high level specification and falls short of checking executable code.
 In this paper, we present a new framework for testing distributed storage systems.
@@
 Our framework has been applied to three distributed storage systems in the Microsoft Azure platform.
-In the process, numerous bugs were identified, reproduced, fixed and confirmed.
+In the process, numerous bugs were identified, reproduced, confirmed and fixed.
 These bugs required subtle combination of concurrency and failures,

### 2015-09-17 experience.tex
update to address some comments by Akash

@@
 Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history.
-This enabled sampling from a distribution, which was defined over all possible input histories within certain bounds.
-\psharp takes control of the choice of input history, so that it can systematically explore different input scenarios.
+This enabled sampling from a distribution that was defined over all possible input histories within certain bounds.
+The \psharp systematic testing engine automatically takes care of enumerating different input histories.
 

### 2015-09-17 methodology.tex
update to address some comments by Akash

@@
 All internal timers of ExtMgr are disabled, because it is crucial to delegate all non-determinism to \psharp.
-Instead, the EN expiration loop and the extent repair loop are driven by external timers provided by \psharp.
+Instead, the EN expiration loop and the extent repair loop are driven by external timers modeled in \psharp.
 Note that system correctness should {\em not} hinge on the frequency of any individual timer.
-Hence, \psharp has the complete freedom to schedule arbitrary interleavings between the timers.
+Hence, the \psharp systematic testing engine has the complete freedom to schedule arbitrary interleavings between the timers.
 

### 2015-09-17 evaluation.tex
update to evaluation

@@
 Modeling Fabric required approximately 5 person months. Although this is a
-significant amount of time, it is a one time effort (once per release). Our plan is to reuse the developed Fabric model 
+significant amount of time, it is a one time effort (once per release).
+\PDComment{should say that 5 months is required for release 1.0, subsequent releases should be much faster} Our plan is to reuse the developed Fabric model
 for testing arbitrary user services built for the Azure Service Fabric system.
@@
 
-For MigratingTable, the upper section of the table uses the random input generator described in Section~\ref{sec:mtable:input}.
-For each bug that led to at least one test failure, we manually reviewed one of the failure traces to confirm it reflected the intended bug.
-For each of the remaining bugs, we repeated the test using a test case custom written to trigger that particular bug in order to confirm that the failure to detect the bug in the original test was due to unlucky random choices of inputs and schedules and not some other problem with the experimental setup.
-These results are in the lower section of the table; they can serve as additional cases in which to compare the random and PCT schedulers but do not represent a testing method one could use to find unknown bugs in software.
-It may have been interesting to report results for specific test cases written before we knew the bugs as mentioned in \ref{sec:mtable:input}, but we did not do so in this study.
-\PDComment{this paragraph needs to be trimmed down probably}
+For the MigratingTable case study, the first 7 bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history.
+For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug.
+The remaining 4 bugs (denoted with $\diamond$) are known bugs that we were unable to find with the fully-randomized test harness in the 100,000 iterations.
+We believe this is due to unlucky random choices of inputs and schedules by the \psharp systematic testing engine.
+We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them.
+Note that using a custom test harness does not represent a testing method one could use to find unknown bugs in software.
+However, in this case the objective was to simply reproduce these known bugs.
 
-One of the bugs, QueryStreamedBackUpNewStream, that we found using \psharp in MigratingTable stands out because it reflects the type of oversight that tends to occur as designs evolve.
-Although this bug could be potentially found with other approaches, \psharp discovered it in a matter of seconds.
-The developers spend about 10 minutes analyzing the trace to diagnose what was happening; granted, this was after days of experience analyzing MigratingTable traces.
-The developers had also added their own trace output to the \psharp test harness, since \psharp's built-in trace output is too low-level and does not include event payloads and other diagnostic data that is not passed between machines.
+The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects the type of oversight that tends to occur as designs evolve.
+We will not discuss the details of the bug due to space constraints, but \psharp managed to discover this bug in a matter of seconds.
+The MigratingTable developers spent about 10 minutes analyzing the trace to diagnose what was happening; granted, this was after days of experience analyzing traces.
+The developers had to extend the logging of \psharp with additional trace information to understand the bug.
+This is expected since \psharp only outputs trace information related to its communicating state machines.
+The trace information can be easily extended though; this was done in all our case studies.
 

### 2015-09-17 overview.tex
update

@@
 We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}).
+We also argue that our approach is \emph{general} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}).
 Furthermore, the language features that are required to be used to connect the real code with the modeled code (e.g., \emph{virtual method dispatch}), are already being heavily used in production for testing purposes , which significantly lowers the bar for practitioners to embrace \psharp for testing.
@@
 {\bf Safety}.
-Safety property specifications can be encoded in \psharp by using the provided \texttt{Assert(...)} method.
-This method takes as argument a predicate, which if evaluates to false denotes a safety property violation.
-The programmer is also free to use other custom assertion APIs (as \psharp executes actual code), but to enable \psharp to recognize such APIs, the suggested approach is to override them to call the \psharp \texttt{Assert(...)} method.
+In  addition to the usual assertions useful for stating local safety properties of a machine, 
+\psharp also provides a way to specify global assertions by using a \emph{safety monitor},
+which is a special state machine that can receive but not send events.
+A safety monitor maintain local state that is modified in response to messages received from ordinary machines.
+This local state is used to maintain a history of the computation that is relevant to the property being specified.
+An erroneous global behavior is flagged via an assertion on the private state of the monitor.
 
-\psharp also provides a way to specify global assertions by using \emph{monitors}, special state machines that
-can only receive events, but not send.
-To declare a monitor, the programmer has to inherit from the \psharp \texttt{Monitor} class.
-A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it; as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is an event and an optional payload.
-This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls, as the monitor has to be called deterministically (and not be interleaved).
+To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class.
+A machine does not need a reference to a \texttt{Monitor} object to send an event to it.
+If an instance of \texttt{M}, a class that inherits from \texttt{Monitor}, has been created,
+a machine can invoke it by calling \texttt{Monitor<M>(e)},
+where the parameter \texttt{e} is an event together with an optional payload.
 
 {\bf Liveness}.
-Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states.
+Liveness property specifications are more difficult to encode since they are intended to capture program progress
+over infinite executions.
+Violation of a liveness property can only be demonstrated by an infinite execution.
+Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
+We take a different approach and allow the programmer to write a \emph{liveness monitor}.
+Similar to a safety monitor, a liveness monitor receives events but does not send them.
+Unlike a safety monitor, a liveness monitor~\cite{Desai:EECS-2015-198}
+contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
+A hot state denotes a point in the execution where progress is required but has not happened yet,
+e.g., a node has failed but a new one has not come up yet.
+A liveness monitor enters a hot state upon receiving notification of an event that requires the system
+to make progress and leaves the hot state upon receiving another event notifying it of progress.
+An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
+Our liveness monitors can encode arbitrary temporal logic properties.
+
+A thermometer analogy is useful for understanding a liveness monitor.
+A hot state increases the temperature by a small value.
+A cold state resets the temperature to zero.
+A liveness error happens if the temperature becomes infinite.
+%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events,
+%such as a periodic request or a periodic timer expiration.
+
+\subsection{Systematic testing}
+\label{sec:psharp:testing}
+
+The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
+A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions.
+We now give an overview of how this works; more details can be found in a previous paper~\cite{deligiannis2015psharp}.
+
+When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic concurrency testing engine}
+~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources
+of non-determinism that are \emph{known} to the \psharp runtime.
+In particular, the runtime is aware of nondeterminism due to interleaving of event handlers in different machines.
+Each send operation to an ordinary (non-monitor) machine and each create opetion of an ordinary (non-monitor) machine creates an
+interleaving point where the runtime makes a scheduling decision.
+The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings,
+until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+The testing is fully automatic, has no false-positives (assuming an accurate environmental model),
+and can reproduce found bugs by replaying buggy schedules.
+After a bug is found, the engine generates a trace that represents the buggy schedule.
+In contrast to typical logs generating during production, this log provides a global order of all communication events
+and is easy to debug.
+
+We have implemented two schedulers inside the \psharp systematic testing engine:
+a \emph{random} scheduler that makes each nondeterministic choice randomly
+and \emph{PCT} scheduler~\cite{burckhardt2010pct}.
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{DesaiQS15}
+exposed by the \psharp libraries.
+The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next,
+and can be used for developing both generic and application-specific schedulers.
+
 Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
@@
 Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
-
-The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states.
-A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g.
-a node has failed but a new one has not come up yet).
-A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.
-
-One can imagine a liveness monitor as a \emph{thermometer}: as the program executes and the liveness property is not satisfied, the liveness monitor stays in the hot state, which infinitely raises the temperature.
-If the liveness property is ever satisfied, and thus the liveness monitor transitions to a cold state, the temperature is instantly reduced to normal levels.
-Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties.
-The liveness monitors in \psharp are based on previous work~\cite{Desai:EECS-2015-198}.
-
-\subsection{Systematic testing}
-\label{sec:psharp:testing}
-In this section, we present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.
-
-A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.
-
-The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions.
-We now give an overview of how this works, while more details can be found in the original paper~\cite{deligiannis2015psharp}.
-
-When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic testing engine} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
-The engine attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order.
-This approach is based on systematic concurrency testing~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} (SCT) techniques that have been previously developed for testing shared memory programs.
-
-The systematic testing engine serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
-The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
-The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
-
-We have implemented two schedulers inside the \psharp systematic testing engine: \emph{random} and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}.
-The random scheduler takes a completely random decision at every scheduling point, whereas the PCT scheduler uses randomization in a disciplined fashion to provide a probabilistic guarantee of finding a bug.
-It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries.
-The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far.
-Exposing an easy-to-use interface for creating new schedulers is inspired by previous work~\cite{DesaiQS15}.
-
-We designed the systematic testing engine in a way that enables easy debugging: after a bug is found, the engine can generate a trace that represents the buggy schedule.
-Note that this trace (in contrast to typical logs generating during production) is sequential.
 

### 2015-09-18 paper.tex
updated abstract

@@
 Conventional testing approaches are ineffective in preventing serious but subtle bugs from reaching production.
-Formal techniques, such as the use of TLA+ in AWS~\cite{newcombe2015aws}, only verifies high level specification and falls short of checking executable code.
+Formal techniques, such as the use of TLA+ in Amazon Web Services, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking executable code.
 In this paper, we present a new framework for testing distributed storage systems.
-Our approach applies verification techniques for high-level specifications directly on the executable code and
-significantly improves coverage of important system behaviors.
+Our approach applies verification techniques for high-level specifications directly on the executable code and significantly improves coverage of important system behaviors.
 
@@
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.
-These bugs required subtle combination of concurrency and failures,
-making them extremely difficult to find with conventional testing techniques.
-An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace,
-which dramatically increases the productivity of debugging.
+These bugs required subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques.
+An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace, which dramatically increases the productivity of debugging.
 

### 2015-09-18 paper.tex
updated abstract

--- 2015-09-17 paper.tex
@@
 Testing distributed systems is challenging due to multiple sources of nondeterminism.
-Conventional testing approaches are ineffective in preventing serious but subtle bugs from reaching production.
-Formal techniques, such as the use of TLA+ in AWS~\cite{newcombe2015aws}, only verifies high level specification and falls short of checking executable code.
+Conventional testing techniques are ineffective in preventing serious but subtle bugs from reaching production.
+Formal techniques, such as the use of TLA+ in Amazon Web Services, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
 In this paper, we present a new framework for testing distributed storage systems.
-Our approach applies verification techniques for high-level specifications directly on the executable code and
-significantly improves coverage of important system behaviors.
+Our approach applies systematic testing to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
 

### 2015-09-18 intro.tex
updated intro

@@
 Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
-This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
+This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside components, unexpected node failures, data losses due to unreliable communication channels, and interaction with (human) clients.
 All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
+A bug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
 
-Developers of production distributed systems use many techniques to test their systems,
+Developers of production distributed systems use many testing techniques,
 such as unit testing, integration testing, stress testing, and fault injection.
@@
 many bugs that arise from subtle combinations of concurrency and failure events
-are missed during testing and get exposed only after a system has been put in production.
-Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money~\cite{tassey2002economic} 
-and many dissatisfied customers~\cite{amazon2012aws, google2014outage}.
+are missed during testing and get exposed only after a system has been put into production.
+However, allowing bugs to reach production, before they are found and fixed, can cost organizations a lot of money~\cite{tassey2002economic} and lead to customer dissatisfaction~\cite{amazon2012aws, google2014outage}.
 
@@
 
-Recently, Amazon Web Services used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
+Recently, Amazon Web Services (AWS) used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
 The gist of their approach is to extract high-level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
@@
 
-We have found that high-level specifications are necessary but not sufficient, due to the gap between the specification and the executable code.
+We have found that checking high-level specifications is necessary but not sufficient, due to the gap between the specification and the executable code.
 %that ultimately powers the distributed systems.
@@
 To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
-The \psharp testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
+The \psharp systematic testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
 
@@
 \begin{itemize}
-\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms.
-\item Our infrastructure can test production code written in \csharp.
+\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms in \psharp.
+\item Because \psharp is an extension of \csharp, our infrastructure can test production code written in \csharp.
 \item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.

### 2015-09-18 motivation.tex
small change
fix
updated motivation

@@
 Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from 10s of Petabytes (PB) in 2010 to Exabytes (EB) in 2015, with the total number of objects stored well-exceeding 60 trillion.
+It has grown from 10s of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
 
@@
 Similar to the current system, vNext employs containers, called {\em extents}, to store data.
-Extents are typically several Gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
+Extents are typically several gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
 However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely \emph{distributed mapping}.
 In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager} (ExtMgr).
+This partitioning is illustrated in Figure~\ref{fig:vnext}.
 
@@
 \includegraphics[width=\linewidth]{img/azurestore}
-\caption{Top-level components of a distributed extent management system for Windows Azure.}
+\caption{Top-level components of vNext, a distributed extent management system for Microsoft Azure.}
 \label{fig:vnext}
@@
 The sync reports list all the extents (and associated metadata) stored on the EN.
-Based on these two types of messages, an ExtMgr identifies which ENs have failed and which extents are affected and missing replicas.
-The ExtMgr, then, schedules tasks to repair the affected extents and distributes the tasks to ENs.
-The ENs repair the extents from their existing replicas in the system and lazily update the ExtMgr via future sync reports.
+Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result.
+The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs.
+The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via future sync reports.
 All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
@@
 \begin{enumerate}
-\item \emph{Unit testing}, which sends emulated heartbeats and sync reports to an ExtMgr and verifies that the messages are processed as expected.
+\item \emph{Unit testing}, in which emulated heartbeats and sync reports are sent to an ExtMgr. These tests check that the messages are processed as expected.
 
-\item \emph{Integration testing}, which launches an ExtMgr together with multiple ENs, subsequently injects an EN failure, and finally verifies that the affected extents are eventually repaired.
+\item \emph{Integration testing}, in which an ExtMgr in launched together with multiple ENs.
+An EN failure is subsequently injected.
+These tests check that the affected extents are eventually repaired.
 
-\item \emph{Stress testing}, which launches an ExtMgr with multiple ENs and multiple extents.
-It keeps repeating the following process: injecting an EN failure, launching a new EN and verifying that the affected extents are eventually repaired.
+\item \emph{Stress testing}, in which an ExtMgr in launched together with multiple ENs and multiple extents.
+The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
 \end{enumerate}
@@
 However, the stress test suite fails {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired.
-The bug appears difficult to identify, reproduce and troubleshoot.
+The bug has proven to be difficult to identify, reproduce and troubleshoot.
 First, it takes very long executions to trigger.

### 2015-09-18 methodology.tex
some updates in section 4

@@
 
-\item[RepairMonitor] collects states from all \texttt{ExtentNode} machines in the testing harness and verifies desired properties, such as that an extent is replicated or repaired at the expected ENs.
+\item[RepairMonitor] collects states from all \texttt{ExtentNode} machines in the testing harness and checks desired properties, such as that an extent is replicated or repaired at the expected ENs.
 \end{description}
@@
 
-The \texttt{RepairMonitor} machine transitions between a cold state and a hot state.
+\texttt{RepairMonitor} is a monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold state and a hot state.
 Whenever an extent node fails, \texttt{RepairMonitor} is notified.
@@
 
-%It verifies the property that the number of extent replica {\em eventually} reaches the target.
+%It checks the property that the number of extent replica {\em eventually} reaches the target.
 
-In the extent repair testing scenarios, \texttt{RepairMonitor} verifies that it should {\em always eventually} end up in the cold repaired state.
+In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should {\em always eventually} end up in the cold repaired state.
 Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long.
@@
 \begin{lstlisting}
-class RepairMonitor : Machine {
+class RepairMonitor : Monitor {
   // true: EN has replica, false: EN has no replica
@@
 This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger.
-In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution verified effectively, all of which should vastly increase the productivity of distributed storage system development.
+In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution tested effectively and thoroughly, all of which should vastly increase the productivity of distributed storage system development.
 

### 2015-09-18 methodology.tex
some update in section 4

@@
 
-The component under test, the real ExtMgr in vNext, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
+The system-under-test, the real ExtMgr in vNext, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
 
@@
 %Using method dispatch for modeling is straightforward.
-The system under test exposes a set of APIs as \emph{virtual methods}.
+The system-under-test exposes a set of APIs as \emph{virtual methods}.
 The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.
@@
 %class MockedNetEngine : NetworkEngine {
-  %ExtentManager EM; // Handle to actual system under test
+  %ExtentManager EM; // Handle to actual system-under-test
   %MachineId Env; // Handle to modeled environment
@@
 \vspace{-2mm}
-\caption{Component under test: the real Extent Manager wrapped inside an \texttt{ExtentManager} \psharp machine.}
+\caption{System-under-test: instance of the real Extent Manager wrapped inside the \texttt{ExtentManager} machine.}
 \label{fig:wrap_target}

### 2015-09-18 evaluation.tex
Fabric stuff

@@
 \end{tabular}
-\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test.}
+\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
 \label{tab:stats}
@@
 
-Modeling Fabric required approximately 5 person months. Although this is a
-significant amount of time, it is a one time effort (once per release).
-\PDComment{should say that 5 months is required for release 1.0, subsequent releases should be much faster} Our plan is to reuse the developed Fabric model
-for testing arbitrary user services built for the Azure Service Fabric system.
+Modeling Fabric required approximately $5$ person months. Although this is a
+significant amount of time, it is a one time effort, which only needs incremental refinement with each release.
 

### 2015-09-18 experience.tex
Fabric stuff

@@
 The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
-Since user-written Fabric services are complex asynchronous and distributed applications, 
-they are challenging to test.
+User-written Fabric services are complex asynchronous and distributed applications and are, thus,
+challenging to test.
 % They are interesting targets for systematic testing with \psharp{}.
@@
 The model is written once
-to include all behaviours of Fabric
-so that it can be used again and again to test arbitrary Fabric services.
-This was the largest among the case studies
-and required multiple rounds of debugging.
-The availability of systematic testing
-allowed us to debug the model to
-a point where reported assertion violations indicate a bug in the
-user service.
-Without systematic testing,
-even bugs in the model would have been hard to find.
+to include all behaviours of Fabric, including simulating failures and recovery,
+so that it can be reused repeatedly to test multiple Fabric services.
+This was the largest modeling exercise among the case studies but the cost
+was amortized across testing multiple servies. This is analogous to the work
+on driver verification \cite{BallLR11} where the cost of developing a model
+of the Windows kernel was amortized across testing multiple device drivers.
 Note that we model the lowest Fabric API layer (\texttt{Fabric.dll})
-which is not documented for use externally.
-Eventually, we will lift the model
-to higher layers
-but for this paper we study internally-developed services
-that target the lowest layer.
+which is currently not documented for use externally.
+We target internally-developed services that invoke the lowest layer.
+
+\psharp's systematic testing was very helpful in debugging the model itself.
+Without systematic testing, we feel even catching bugs in the model 
+would have been difficult.
 
@@
 
-The main system that we tested is 
-\emph{CScale}~\cite{faleiro2012},
-a big data-stream processing system that chains multiple Fabric services
-in a directed acyclic graph.
-A key challenge was that
-CScale contains inter-service communication.
-%\PTComment{Cut: detail.}
-% that resolves services
-% via Fabric but communicates
-% using a non-Fabric remote procedure call protocol.
-To close the system,
-%\PTComment{Cut: we implemented service resolution}
-% we first implemented service resolution in our Fabric model,
-% which resolves Fabric service names to IP addresses.
-%We then replaced 
-we 
-replaced the relevant classes
-% replaced the inter-service communication classes with our own,
-% ;
-% our replacement includes the ability to map
-% URLs to \psharp{} machines
-so that remote procedure calls are implemented
+The main system that we tested is \emph{CScale}~\cite{faleiro2012},
+a big data-stream processing system. In prior work~\cite{deligiannis2015psharp}, 
+we used an earlier version of the Fabric model and reported bugs in sample serives.
+Supporting CScale required significant additions to the model, making it much more
+feature complete.
+CScale chains multiple Fabric services, which communicate via remote procedure calls (RPCs).
+To close the system, we modeled RPCs 
 by sending and receiving \psharp{} events.
-Thus,
-we converted a distributed system
+This was enough to converted a distributed system
 that uses both Fabric and its own
@@
 This work is still in-progress.
-However,
-we were able to find a \texttt{NullReferenceException}
-bug
-in CScale
-by running it on our model.
+However, we were able to find a \texttt{NullReferenceException}
+bug in CScale by running it on our model. The bug has been
+communicated to the developers of CScale, but we are still awaiting a
+confirmation.

### 2015-09-18 experience.tex
various fixes

@@
 was amortized across testing multiple servies. This is analogous to the work
-on driver verification \cite{BallLR11} where the cost of developing a model
+on driver verification \cite{ball2011slam} where the cost of developing a model
 of the Windows kernel was amortized across testing multiple device drivers.

### 2015-09-18 motivation.tex
various fixes

@@
 Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from 10s of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
+It has grown from 10s of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
 
-Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to increase the scalability by more than 100$\times$.
+Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
 Similar to the current system, vNext employs containers, called {\em extents}, to store data.
@@
 The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs.
-The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via future sync reports.
+The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via their next periodic sync reports.
 All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

### 2015-09-18 overview.tex
many updates in section 3 to address Ally's comments

@@
 
-A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods.
-Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler.
-This handler might update a field, create a new machine, or send an event to another machine.
-In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+A \psharp program consists of multiple \emph{state machines} that communicate with each other \emph{asynchronously} by sending and receiving \emph{events}, which may contain an optional \emph{payload}.
+A \psharp machine is similar to a class in \csharp: it can contain any number of fields and methods, but it also contains an input event queue, and one or more states.
+Each state can register \emph{actions} to handle incoming events.
+\psharp machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking the registered action.
+This action might access a field, call a method, transition the machine to a new state, create a new machine, or send an event to another machine.
+In \psharp, a send operation is non-blocking; the event is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
 All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
 
-Our approach of testing distributed systems using \psharp requires the developer to perform three key modeling tasks;
-these tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a case study.
+Our approach of using \psharp to test distributed systems, implemented in the .NET framework, requires the developer to perform the following three key modeling tasks.
+These tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a running example.
 \begin{enumerate}
@@
 The computation model underlying \psharp is communicating state machines.
-The environment of the system-under-test must be modeled using \psharp machines,
-while the real components of the system must be wrapped inside \psharp machines.
+The environment of the system-under-test must be modeled using the \psharp APIs, while the real components of the system must be wrapped inside \psharp machines.
 We call this modeled environment the \psharp test harness.
@@
 \item
-The asynchrony due to message passing must be exposed using the \psharp APIs for sending events.
-Similarly, other sources of nondeterministically-generated events, such as failures and timer expiration,
-must also be explicitly modeled using the available \psharp APIs;
-This step allows the \psharp runtime to explore the nondeterministic choices during testing.
+All the asynchrony due to message passing between system components must become explicit and be modeled using the \psharp event-sending APIs.
+Similarly, all other sources of nondeterminism (e.g.
+failures and timer expiration) must be modeled using appropriate \psharp APIs.
+This step allows the \psharp runtime to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing.
 
 \item
-The correctness of each execution generated during testing must be specified.
-Specifications can be either \emph{safety} and \emph{liveness}~\cite{lamport1977proving}.
-A safety specification is a generalization of assertions;
-a safety violation is a finite trace leading to an erroneous state.
-A liveness specification is a generalization of nontermination;
-a liveness violation is an infinite trace that exhibits lack of progress.
+The criteria for correctness of an execution must be specified.
+Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties.
+Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state.
+Liveness specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.
 \end{enumerate}
-In principle, this modeling methodology is not specific to \psharp but can be used in combination with any other programming framework
-with equivalent capabilities.
 
-We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-We also argue that our approach is \emph{general} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}).
-Furthermore, the language features that are required to be used to connect the real code with the modeled code (e.g., \emph{virtual method dispatch}), are already being heavily used in production for testing purposes , which significantly lowers the bar for practitioners to embrace \psharp for testing.
+\psharp uses modern languages features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
+Programmers in industry are used to work with such features, and heavily use them in production for testing purposes.
+In our experience, this significantly lowers the bar for product teams inside Microsoft to embrace \psharp for testing.
+
+In principle, our modeling methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
+We also argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
 
@@
 \label{sec:bg:bugs}
-In  this section, we describe how safety and liveness specifications are expressed in our system.
+In  this section, we describe how safety and liveness specifications are expressed in \psharp.
 
 {\bf Safety}.
-In  addition to the usual assertions useful for stating local safety properties of a machine, 
-\psharp also provides a way to specify global assertions by using a \emph{safety monitor},
-which is a special state machine that can receive but not send events.
-A safety monitor maintain local state that is modified in response to messages received from ordinary machines.
+In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special \psharp machine that can receive, but not send, events.
+A safety monitor maintain local state that is modified in response to events received from regular (i.e.
+non-monitor) machines.
 This local state is used to maintain a history of the computation that is relevant to the property being specified.
-An erroneous global behavior is flagged via an assertion on the private state of the monitor.
+An erroneous global behavior is flagged via an assertion on the private state of the safety monitor.
 
 To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class.
-A machine does not need a reference to a \texttt{Monitor} object to send an event to it.
-If an instance of \texttt{M}, a class that inherits from \texttt{Monitor}, has been created,
-a machine can invoke it by calling \texttt{Monitor<M>(e)},
-where the parameter \texttt{e} is an event together with an optional payload.
+Monitors in \psharp are singleton instances and, thus, a regular machine does not need a reference to a monitor to send an event to it.
+A regular machine can send an event to a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send together with an optional payload \texttt{p}.
 
@@
 Similar to a safety monitor, a liveness monitor receives events but does not send them.
-Unlike a safety monitor, a liveness monitor~\cite{Desai:EECS-2015-198}
+Unlike a safety monitor, a liveness monitor~\cite{desai2015building}
 contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
@@
 In particular, the runtime is aware of nondeterminism due to interleaving of event handlers in different machines.
-Each send operation to an ordinary (non-monitor) machine and each create opetion of an ordinary (non-monitor) machine creates an
+Each send operation to an ordinary (non-monitor) machine and each create operation of an ordinary (non-monitor) machine creates an
 interleaving point where the runtime makes a scheduling decision.
@@
 and \emph{PCT} scheduler~\cite{burckhardt2010pct}.
-It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{DesaiQS15}
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic}
 exposed by the \psharp libraries.

### 2015-09-18 overview.tex
many updates in section 3 to address Ally's comments

@@
 \label{sec:bg:bugs}
+
 In  this section, we describe how safety and liveness specifications are expressed in \psharp.
 
-{\bf Safety}.
-In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special \psharp machine that can receive, but not send, events.
-A safety monitor maintain local state that is modified in response to events received from regular (i.e.
+\textbf{Safety}.
+In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.
+A safety monitor maintain local state that is modified in response to events received from ordinary (i.e.
 non-monitor) machines.
@@
 To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class.
-Monitors in \psharp are singleton instances and, thus, a regular machine does not need a reference to a monitor to send an event to it.
-A regular machine can send an event to a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send together with an optional payload \texttt{p}.
+Monitors in \psharp are singleton instances, which means that an ordinary machine does not need a reference to a monitor to send it an event.
+An ordinary machine can send an event to a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send together with an optional payload \texttt{p}.
 
-{\bf Liveness}.
+\textbf{Liveness}.
 Liveness property specifications are more difficult to encode since they are intended to capture program progress
@@
 The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions.
-We now give an overview of how this works; more details can be found in a previous paper~\cite{deligiannis2015psharp}.
+A key capability of the \psharp runtime is that it can execute in bug-finding mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.
 
-When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic concurrency testing engine}
-~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources
-of non-determinism that are \emph{known} to the \psharp runtime.
-In particular, the runtime is aware of nondeterminism due to interleaving of event handlers in different machines.
-Each send operation to an ordinary (non-monitor) machine and each create operation of an ordinary (non-monitor) machine creates an
-interleaving point where the runtime makes a scheduling decision.
-The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings,
-until it either reaches a bound (in number of iterations or time), or it hits a property violation.
-The testing is fully automatic, has no false-positives (assuming an accurate environmental model),
-and can reproduce found bugs by replaying buggy schedules.
+When the \psharp runtime executes in bug-finding mode, an embedded \emph{systematic concurrency testing engine}~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
+In particular, the runtime is aware of nondeterminism due to the interleaving of event handlers in different machines.
+Each send operation to an ordinary (non-monitor) machine, and each create operation of an ordinary machine, creates an interleaving point where the runtime makes a scheduling decision.
+The runtime is also aware of all nondeterministic events (e.g.
+failures) that were captured during the modeling process.
+
+The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+This testing process is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
 After a bug is found, the engine generates a trace that represents the buggy schedule.
-In contrast to typical logs generating during production, this log provides a global order of all communication events
-and is easy to debug.
+In contrast to logs typically generating during production, this log provides a global order of all communication events and, thus, is easier to debug.
 
-We have implemented two schedulers inside the \psharp systematic testing engine:
-a \emph{random} scheduler that makes each nondeterministic choice randomly
-and \emph{PCT} scheduler~\cite{burckhardt2010pct}.
-It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic}
-exposed by the \psharp libraries.
-The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next,
-and can be used for developing both generic and application-specific schedulers.
+We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.
+The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.
 
@@
 
-Vanilla \psharp is able to capture and take control of the \emph{inter-machine concurrency} due to message passing, but is unable to systematically explore any interleavings due to \emph{intra-machine concurrency} (e.g.
-async/await or TPL).
-This is problematic as nowadays, with multicore machines being a commodity, programmers tend to write multithreaded code to exploit shared memory architectures and increase the performance of individual components of a distributed system.
+The original implementation of \psharp is able to capture and take control of the \emph{inter-machine concurrency} due to message passing, but is unable to systematically explore any interleavings due to \emph{intra-machine concurrency} (e.g.
+async/await, which we discuss below).
+This is problematic since programmers often write multithreaded code that exploits shared memory multicore architectures, to increase the performance of individual components of a distributed system.
 
 As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
-Asynchronous code using async/await is more readable because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
+Asynchronous code using async/await is more readable than older asynchronous programming libraries in .NET, because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
 The key idea behind async/await is that a method declared as \texttt{async} can use internally the \texttt{await} keyword which allows the thread executing the method to wait on a TPL task, or any other \emph{awaitable} object, \emph{without} blocking.
@@
 \texttt{System.Threading}, TPL, async/await, locks, semaphores) and each one has its own complexities.
-Although developers are willing to model the top-level message passing communication using \psharp, they are resistant in modeling the low-level threading and synchronization methods as this would be a very invasive procedure and such refactoring is unlikely to scale for legacy code.
+Although developers are willing to model the top-level message passing communication using \psharp, in our experience they are resistant to modeling the low-level threading and synchronization methods as this would be a very invasive procedure and such refactoring is unlikely to scale for legacy code.
 
 To test our case studies, we decided to focus our efforts in handling async/await as providing a robust solution for a subset of the .NET threading APIs is more feasible than trying to handle arbitrary threading and synchronization.
+By choosing to focus on async/await, we cannot currently systematically test programs that use other mechanisms for intra-machine concurrency.
+We can still apply \psharp on them, but we lose the benefit of systematic testing.
+In future work, we plan to investigate support for non-async/await concurrency primitives.
+
 Our solution involves using a custom task scheduler whenever the bug-finding mode of the \psharp runtime is enabled, instead of the default TPL task scheduler.
 The \psharp task scheduler inherits from the \texttt{TaskScheduler} class, a low-level API that is responsible for enqueueing TPL tasks into threads.
-
 Our approach works as follows.
@@
 As long as any child tasks spawned by the root machine task do not explicitly start in another scheduler (including the default TPL scheduler), then they will be scheduled for execution in our custom task scheduler.
-Our custom scheduler intercepts the call to enqueue a task, creates a special machine called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine.
+Our custom scheduler intercepts the call to enqueue a task, creates a special machine (only internally available to \psharp) called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine.
 The constructor of a \texttt{TaskMachine} takes as an argument a TPL task and stores it in an field.
-The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it will start executing the wrapped task and then immediately wait on its completion.
+The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it starts executing the wrapped task and then immediately waits on its completion.
 This forces the lifetime of the task to become the lifetime of the machine and vice versa.
@@
 The task associated with the continuation will also be wrapped and scheduled accordingly.
-
-Regarding support for handling non-async/await concurrency primitives, we are currently investigate feasible approaches.

### 2015-09-18 evaluation.tex
State the time taken to develop MigratingTable.

@@
 
-Modeling the Live Migration Table required \PDComment{waiting confirmation from
-Matt}. The modeling effort was done in parallel with the development of the actual system. 
+Developing the Live Migration Table and its \psharp test harness took about 5 weeks. The test harness was developed in parallel with the actual system.
 This is in contrast with the other two case studies discussed in this paper, where the modeling 

### 2015-09-19 evaluation.tex
minor changes

@@
 We configured the engine to perform 100,000 iterations.
-The random seed for both schedulers was generated in each iteration using the \texttt{DateTime.Now.Millisecond} API which returns the current time in milliseconds.
+The random seed for both schedulers was generated using the current time.
 The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps of 500.
@@
 The remaining 4 bugs (denoted with $\diamond$) are known bugs that we were unable to find with the fully-randomized test harness in the 100,000 iterations.
-We believe this is due to unlucky random choices of inputs and schedules by the \psharp systematic testing engine.
+We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine.
 We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them.
@@
 The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects the type of oversight that tends to occur as designs evolve.
-We will not discuss the details of the bug due to space constraints, but \psharp managed to discover this bug in a matter of seconds.
-The MigratingTable developers spent about 10 minutes analyzing the trace to diagnose what was happening; granted, this was after days of experience analyzing traces.
-The developers had to extend the logging of \psharp with additional trace information to understand the bug.
-This is expected since \psharp only outputs trace information related to its communicating state machines.
-The trace information can be easily extended though; this was done in all our case studies.
+We do not discuss the details of the bug due to space constraints, but \psharp managed to discover this bug in a matter of seconds.
+The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces.
+The developers extended the logging of \psharp with additional trace information to understand the bug.
+\psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend and was done in all our case studies.
 

### 2015-09-19 evaluation.tex
updates in evaluation

@@
 For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug.
-The remaining 4 bugs (denoted with $\diamond$) are known bugs that we were unable to find with the fully-randomized test harness in the 100,000 iterations.
+The remaining 4 bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations.
 We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine.
 We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them.
-Note that using a custom test harness does not represent a testing method one could use to find unknown bugs in software.
-However, in this case the objective was to simply reproduce these known bugs.
+The design of these custom harnesses was influenced by the knowledge of these bugs; our objective here was to simply see if \psharp can reproduce these known bugs.
 

### 2015-09-19 intro.tex
update

@@
 Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
-This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside components, unexpected node failures, data losses due to unreliable communication channels, and interaction with (human) clients.
-All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute.
-A bug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
+This challenge is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside components, unexpected node failures, data losses due to unreliable communication channels, and interaction with (human) clients.
+All these sources of nondeterminism can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix. 
+A Heisenbug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
 

### 2015-09-19 overview.tex
update to remove inbalanced subsubsections
update

@@
-Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer~\cite{lamport1978time}.
-Each component has its own input message queue, and when a message arrives, the component responds by executing an appropriate \emph{message handler}.
-Such a handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.
-
-In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication.
-To complicate matters further, unexpected failures are the norm in production systems: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system.
-Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.
-
-All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix.
-Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code.
-However, these techniques are not effective for testing distributed systems, as they are not able to control the many sources of nondeterminism.
-
 The goal of our work is to develop testing techniques to detect and debug Heisenbugs in distributed storage systems prior to deployment.
@@
 These tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a running example.
+
 \begin{enumerate}
@@
 
+\noindent
 \psharp uses modern languages features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
@@
 An erroneous global behavior is flagged via an assertion on the private state of the safety monitor.
+Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).
 
@@
 
+A liveness violation is witnessed by an infinite execution in which all concurrently executing \psharp machines are scheduled fairly.
+Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time.
+Therefore, our implementation of liveness checking approximates an infinite execution using several heuristics.
+One heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
+Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution
+and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state
+graph of the program.
+
 \subsection{Systematic testing}
@@
 
-Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}.
-Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property.
-Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
-
-\subsection{Handling intra-machine concurrency}
+\subsubsection{Handling \csharp 5.0 asynchrony}
 \label{sec:psharp:async}
 
-The original implementation of \psharp is able to capture and take control of the \emph{inter-machine concurrency} due to message passing, but is unable to systematically explore any interleavings due to \emph{intra-machine concurrency} (e.g.
-async/await, which we discuss below).
-This is problematic since programmers often write multithreaded code that exploits shared memory multicore architectures, to increase the performance of individual components of a distributed system.
-
-As an example, the Live Azure Migration system uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
-Asynchronous code using async/await is more readable than older asynchronous programming libraries in .NET, because it looks like traditional procedural code, but it is translated by the compiler to an event-driven state machine that is built on top of TPL to achieve performance.
-The key idea behind async/await is that a method declared as \texttt{async} can use internally the \texttt{await} keyword which allows the thread executing the method to wait on a TPL task, or any other \emph{awaitable} object, \emph{without} blocking.
-This is achieved as follows.
-When an \texttt{await} statement is executed, the code following the await is wrapped as a TPL task \emph{continuation} and the method returns to the caller.
-This continuation executes when the awaitable object has completed.
-
-Developing a fully automatic and universal approach to handle intra-machine concurrency in .NET is very challenging, because there are many APIs that can be used for concurrency and synchronization (e.g.
-\texttt{System.Threading}, TPL, async/await, locks, semaphores) and each one has its own complexities.
-Although developers are willing to model the top-level message passing communication using \psharp, in our experience they are resistant to modeling the low-level threading and synchronization methods as this would be a very invasive procedure and such refactoring is unlikely to scale for legacy code.
-
-To test our case studies, we decided to focus our efforts in handling async/await as providing a robust solution for a subset of the .NET threading APIs is more feasible than trying to handle arbitrary threading and synchronization.
-By choosing to focus on async/await, we cannot currently systematically test programs that use other mechanisms for intra-machine concurrency.
-We can still apply \psharp on them, but we lose the benefit of systematic testing.
-In future work, we plan to investigate support for non-async/await concurrency primitives.
+The Live Azure Migration system (Section~\ref{sec:cases:migration}) uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+An \texttt{async} procedure is allowed to perform \texttt{await} on a TPL task without blocking the current thread.
+The \csharp compiler achieves this non-blocking behavior by wrapping the \emph{continuation} of the \texttt{await} statement
+in a TPL task and returning it to the caller.
+The TPL scheduler executes this task when the task being awaited has completed.
+In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing.
+However, in our experience, refactoring code to model in \psharp the asynchrony in async/await is difficult and unlikely
+to scale for large code bases.
+Therefore, our implementation automatically captures asychronously created tasks in .NET runtime as \psharp machines.
 

### 2015-09-19 experience.tex
minor edits

@@
 The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system.
-\psharp is a generic framework, capable of handling arbitrary distributed systems.
-We showcase this capability by presenting developer experience in using \psharp to model and test two other distributed storage systems used in production in Microsoft: the Live Azure Table Migration library; and the Azure Service Fabric system.
+\psharp is a generic framework.
+We showcase this capability by presenting developer experience of using \psharp to model and test two other distributed storage systems used in production in Microsoft.
 

### 2015-09-19 evaluation.tex
lots of updates to address Ally's comments

@@
 Environmental modeling is a core activity of using \psharp.
-It is required for \emph{closing the environment} of a system-under-test and making it amenable to systematic testing.
-Table~\ref{tab:stats} presents program statistics for our three case studies.
-We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system (\#B); lines of \psharp code for the environmental model (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
+It is required for \emph{closing the environment} of a system to make it amenable to systematic testing.
+Table~\ref{tab:stats} presents program statistics for the three case studies.
+The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp model'' refer to the \psharp test harness.
+We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system-under-test (\#B); lines of \psharp code for the test harness (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
 
-Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately 2 weeks of part-time developing.
-The \psharp model for testing this system is the smallest (in lines of code) from all three case studies.
-This was because the modeling effort was targeting the particular liveness bug that was haunting the developers of vNext.
-We are currently in the process of modeling other components of vNext, such as a ChainReplication and a Paxos system.
+Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two weeks of part-time developing.
+The \psharp model for testing this system is the smallest (in lines of code) from the three case studies.
+This was because this modeling exercise aimed to reproduce the particular liveness bug that was haunting the developers of vNext.
 
-Developing the Live Migration Table and its \psharp test harness took about 5 weeks. The test harness was developed in parallel with the actual system.
-This is in contrast with the other two case studies discussed in this paper, where the modeling 
-activity occurred independently and at a later stage of the development process.
+Developing the Live Migration Table and its \psharp test harness took about five weeks.
+The test harness was developed in parallel with the actual system.
+This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and at a later stage of the development process.
 
-Modeling Fabric required approximately $5$ person months. Although this is a
-significant amount of time, it is a one time effort, which only needs incremental refinement with each release.
+Modeling Fabric required approximately five person months. Although this is a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.
 
@@
 
-Using \psharp we managed to uncover more than 10 serious bugs in our case studies.
+Using \psharp we managed to uncover 12 serious bugs in our case studies \PDComment{update for Fabric}.
 As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting.
 According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner.
-After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one so that we can evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.
+After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one to evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.
 
@@
 
-For the MigratingTable case study, the first 7 bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history.
+For the MigratingTable case study, the first seven bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history.
 For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug.
-The remaining 4 bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations.
+The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations.
 We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine.

### 2015-09-19 experience.tex
lots of updates to address Ally's comments

@@
 The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system.
-\psharp is a generic framework.
-We showcase this capability by presenting developer experience of using \psharp to model and test two other distributed storage systems used in production in Microsoft.
+\psharp is a generic framework for .NET distributed systems.
+We showcase this capability by presenting developer experience of using \psharp to model and test two other systems used in production inside Microsoft.
 
@@
 The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
-The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult to foresee interactions between different parts of the code.
+The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
 
-\subsubsection{Testing the MigratingTable library}
-
+\textbf{Modeling and testing}.
 Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history.
@@
 The \psharp systematic testing engine automatically takes care of enumerating different input histories.
+The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call.
+Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
+This makes it possible to check the correctness property as the model executes.
 
@@
 
-The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call.
-Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
-This makes it possible to check the correctness property as the model executes.
-
-The \psharp environmental model of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
-
+The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
 Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine.
@@
 The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
-User-written Fabric services are complex asynchronous and distributed applications and are, thus,
-challenging to test.
+User-written Fabric services are complex asynchronous and distributed applications, and are thus challenging to test.
 % They are interesting targets for systematic testing with \psharp{}.
@@
 would have been difficult.
+\PDComment{See Ally's feedback and maybe try to address this as we now have some extra space in the paper: ``\psharp systematic testing was very helpful in debugging the model itself.'' this short paragraph is very intriguing, and begs a little more elaboration about what sorts of bugs you typically found in the model.}
 
@@
 Supporting CScale required significant additions to the model, making it much more
-feature complete.
+feature-complete.
 CScale chains multiple Fabric services, which communicate via remote procedure calls (RPCs).

### 2015-09-19 methodology.tex
update to remove inbalanced subsubsections
lots of updates to address Ally's comments

@@
 
-The system-under-test, the real ExtMgr in vNext, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
-
-\subsubsection{Internals of the real Extent Manager}
+The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
 
@@
 
-Inside the real ExtMgr, there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
+\textbf{Internals of the real Extent Manager}.
+Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
 The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs.
@@
 
-\subsubsection{Intercepting messages between machines}
-\label{sec:method:model:dispath}
+\textbf{Intercepting messages between machines}.
+The real ExtMgr uses a network engine to send messages to the ENs.
+The testing harness mocks the original network engine in vNext and overrides its interface.
+In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
+As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
+Conceivably, the mocked network engine could leverage the nondeterminism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.
 
@@
 
-The real ExtMgr uses a network engine to send messages to the ENs.
-The testing harness mocks the original network engine in vNext and overrides its interface.
-In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
-As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
-Conceivably, the mocked network engine could leverage the non-determinism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.
-\CHComment{only if non-determinism is already covered in the previous section.}
-
 \begin{figure}[t]
@@
 
-\subsubsection{Specifics of the ExtentManager machine}
-
+\textbf{Specifics of the ExtentManager machine}.
 Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine.
@@
 
-Messages coming from the ENs do {\em not} go through the mocked network engine.
+Messages coming from the ENs do \emph{not} go through the mocked network engine.
 They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}.
-The benefit of this approach is that ExtMgr is unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.
+The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.
 
-All internal timers of ExtMgr are disabled, because it is crucial to delegate all non-determinism to \psharp.
-Instead, the EN expiration loop and the extent repair loop are driven by external timers modeled in \psharp.
-Note that system correctness should {\em not} hinge on the frequency of any individual timer.
-Hence, the \psharp systematic testing engine has the complete freedom to schedule arbitrary interleavings between the timers.
+System correctness should \emph{not} hinge on the frequency of any individual timer.
+Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
+To achieve this, all timers inside ExtMgr are disabled, and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
+The \psharp timers send nondeterministic timeout events that are controlled by the \psharp runtime.
+Hence, the \psharp testing engine has the freedom to schedule arbitrary interleavings between timeout events and all other regular system events.
 
@@
 
-The \texttt{ExtentNode} machine is a much simplified version of the original EN.
-It omits the details of a real EN, mocks only logic necessary for the testing scenarios, and keeps the mocked logic as simple as possible.
-The logic here includes: repairing an extent from its replica, generating sync reports periodically, and sending heartbeat messages periodically.
+The \texttt{ExtentNode} machine is a simplified version of the original EN.
+The machine omits much of the complex details of a real EN, and only mocks the logic necessary for the testing scenarios.
+This mocked logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to the \texttt{ExtentManager} machine.
 
-The testing harness leverages components in vNext whenever appropriate.
-Here, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, an internal component of a real EN used for extent bookkeeping.
+The testing harness leverages components of Azure Storage vNext whenever it is appropriate.
+For example, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, a component that is used inside a real EN for extent bookkeeping.
 
@@
 It waits for the extent to be repaired on the new EN.
-Code snippets are skipped due to space constraint.
+We omit illustrative code snippets due to space constraints.
 
@@
 
-\texttt{RepairMonitor} is a monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold state and a hot state.
+\texttt{RepairMonitor} is a liveness monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
 Whenever an extent node fails, \texttt{RepairMonitor} is notified.
-As soon as the number of extent replicas falls below a specified target, \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
+As soon as the number of extent replicas falls below a specified target (three replicas in the current test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
 Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified.
-It transitions into the cold \emph{repaired} state, when the replica number reaches the target again, as illustrated in Figure~\ref{fig:monitor}.
-
-%It checks the property that the number of extent replica {\em eventually} reaches the target.
+It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.
 
@@
 Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long.
-This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug!
+This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug.
 
@@
 
-It took only tens of seconds before the testing harness reported the first occurrence of a liveness bug. Upon examining the debug trace, the vNext developers were able to confirm the bug.
+It took less than ten seconds for the \psharp testing engine to report the first occurrence of a liveness bug in vNext (see Section~\ref{sec:eval}).
+Upon examining the debug trace, the vNext developers were able to confirm the bug.
 
-However, the trace didn't include enough details, which prevented the developers from identifying a root cause.
+However, the original \psharp trace did not include sufficient detail to allow the developers to identify the root cause of the problem.
 Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration.
-After several iterations, the developers were able to pinpoint the exact culprit and immediately proposed a solution to fix the bug.
-Once the proposed solution was implemented, the developers again ran the testing harness, which reported no more bug for $100,000$ iterations in tens of minutes.
+After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug.
+Once the proposed solution was implemented, the developers ran again the test harness.
+No bugs were found during 100,000 iterations, a process that required only tens of minutes.
 
@@
 
-Here is one particular execution sequence resulting in the bug.
-$i)$ EN$_0$ fails and is detected by the EN expiration loop; $ii)$ EN$_0$ is removed from the EN map; $iii)$ the extent center is updated and the replica count drops from 3, the target, to 2; $iv)$ \texttt{ExtMgr} receives a sync report from EN$_0$; $v)$ the extent center is updated and the replica count increases from 2 to 3 again.
-This is problematic because, on one hand, the replica count is equal to the target, so the extent repair loop never schedules any repair task.
-On the other hand, there are only two true replicas in the system, one fewer than the target.
-This execution sequence leads to one replica missing.
-Conceivably, repeating this another two times would result in all replicas missing, while \texttt{ExtMgr} still thinks all replicas are healthy.
-If deployed in production, such bug would have caused a very serious incident of customer data loss.
+The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from the EN map; (iii) the extent center is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3.
+This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task.
+At the same time, there are only two \emph{true} replicas in the system, which is one less than the target.
+This execution sequence leads to one replica missing; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy.
+In production, such a bug can cause a very serious incident of customer data loss.
 
-The culprit is in $iv)$, where \texttt{ExtMgr} receives a sync report from EN$_0$ after deleting the EN.
-This may occur in \psharp because of arbitrary message interleaving.
-It may also occur, albeit much less frequently, in the stress testing due to messages being delayed in the network.
-This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger.
-In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution tested effectively and thoroughly, all of which should vastly increase the productivity of distributed storage system development.
+The culprit is in (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN.
+This may occur in \psharp due to arbitrary interleaving of events.
+It may also occur, albeit much less frequently, during stress testing due to messages being delayed in the network.
+This explains why the bug only occurs from time to time during stress testing and requires long executions to manifest.
+In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit to be identified promptly, and the fix to be tested effectively and thoroughly, all of which have the potential to vastly increase the productivity of distributed storage system development.
 

### 2015-09-19 motivation.tex
update to remove inbalanced subsubsections
lots of updates to address Ally's comments

@@
 
-\item \emph{Integration testing}, in which an ExtMgr in launched together with multiple ENs.
+\item \emph{Integration testing}, in which an ExtMgr is launched together with multiple ENs.
 An EN failure is subsequently injected.
@@
 
-\item \emph{Stress testing}, in which an ExtMgr in launched together with multiple ENs and multiple extents.
+\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents.
 The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
@@
 
+\noindent
 Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in the ExtMgr logic.

### 2015-09-19 overview.tex
lots of updates to address Ally's comments

@@
 
-\textbf{Safety}.
+\textbf{Safety specifications}.
 In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.
-A safety monitor maintain local state that is modified in response to events received from ordinary (i.e.
-non-monitor) machines.
+A safety monitor maintains local state that is modified in response to events received from ordinary (non-monitor) machines.
 This local state is used to maintain a history of the computation that is relevant to the property being specified.
@@
 
-\textbf{Liveness}.
+\textbf{Liveness specifications}.
 Liveness property specifications are more difficult to encode since they are intended to capture program progress
@@
 
-The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
 This testing process is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.

### 2015-09-19 intro.tex
minor edits
update in intro

@@
 We have found that checking high-level specifications is necessary but not sufficient, due to the gap between the specification and the executable code.
-%that ultimately powers the distributed systems.
-Our goal is to bridge this gap. We propose a new framework that
-applies verification techniques for high-level specifications directly
-to the executable code. Our framework is different from prior
-approaches where the developers were required to either switch to an
-unfamiliar special programming language~\ref{}, or manually annotate
-and instrument code~\ref{}. Instead, we allow developers to test
-{\em unmodified production code} by writing test harnesses in a
-mainstream programming language. This
-significantly lowered the acceptance barrier for adoption by the
-Microsoft Azure team.
+Our goal is to bridge this gap.
+We propose a new framework that validates high-level specifications directly on the executable code.
+Our framework is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{}.
+Instead, we allow developers to test \emph{unmodified production code} by writing test harnesses in \csharp, a mainstream programming language.
+This significantly lowered the acceptance barrier for adoption by the Microsoft Azure team.
 
 %We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
-Our framework is based on \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream language \csharp that provides support for modeling, specification, and systematic testing of asynchronous systems.
-To test a distributed system with the \psharp framework, the programmer augments the system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
-The \psharp systematic testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
+Our framework is based on \psharp~\cite{deligiannis2015psharp}, an extension of \csharp that provides support for modeling, specification, and systematic testing of distributed systems written in Microsoft's .NET framework.
+To test a distributed system with \psharp, the programmer augments the original system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
+The \psharp testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
 
-We have applied our framework to three distributed storage systems in Microsoft Azure: Azure Storage vNext, Azure Table Live Migration, 
-and Azure Service Fabric.
+We have applied \psharp to three distributed storage systems in Microsoft Azure: Azure Storage vNext, Azure Table Live Migration, and Azure Service Fabric.
 %a distributed storage system; a system for migrating data across partitions; and a big-data computing system.

### 2015-09-19 overview.tex
Small typos fixed.

@@
 In \psharp, a send operation is non-blocking; the event is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
-All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
 
@@
 \psharp uses modern languages features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
-Programmers in industry are used to work with such features, and heavily use them in production for testing purposes.
+Programmers in industry are used to working with such features, and heavily use them in production for testing purposes.
 In our experience, this significantly lowers the bar for product teams inside Microsoft to embrace \psharp for testing.

### 2015-09-19 experience.tex
Added some detail to the Fabric section.

@@
 %\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
-%One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}.
-The primary replicates state changes to the secondaries by sending \emph{replication requests} so that all replicas eventually have the same state.
-If the primary fails (e.g.\ if the node on which the primary is running crashes), Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive a full or partial copy (depending on whether persistent storage is used) of the state of the new primary in order to ``catch up'' with the other secondaries.
-Fabric provides a name-resolution service so that clients can always find the current primary.
-The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
+\PTComment{Re-added the detail here so we can discuss the bug below:}
+One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}. The primary replicates state changes to the secondaries 
+%by sending them \emph{replication requests}.
+so that all replicas eventually have the same state. 
+If the primary fails,
+% (e.g.\ if the node on which the primary is running crashes), 
+Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive 
+%a full or partial copy (depending on whether persistent storage is used) 
+a copy
+of the state of the new primary in order to ``catch up'' with the other secondaries. 
+%Fabric provides a name-resolution service so that clients can always find the current primary.
+%The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
 User-written Fabric services are complex asynchronous and distributed applications, and are thus challenging to test.
@@
 
+\PTComment{Changed this para; added detail about the bug: }
 \psharp's systematic testing was very helpful in debugging the model itself.
-Without systematic testing, we feel even catching bugs in the model 
-would have been difficult.
-\PDComment{See Ally's feedback and maybe try to address this as we now have some extra space in the paper: ``\psharp systematic testing was very helpful in debugging the model itself.'' this short paragraph is very intriguing, and begs a little more elaboration about what sorts of bugs you typically found in the model.}
+% Without systematic testing, we feel even catching bugs in the model 
+% would have been difficult.
+In order to perform systematic testing on our model,
+we wrote a simple service in \psharp{}
+that runs on our Fabric model.
+%made up of a single machine
+% and executed it on the Fabric model,
+% thus yielding a pure \psharp{} system.
+We tested a scenario where the primary replica
+fails at some non-deterministic point
+within the execution.
+One such bug that we found
+involved
+the primary replica
+failing
+when a secondary replica
+was about to receive a copy of the state;
+the secondary was then elected to be the new primary
+and yet, because the secondary stopped waiting for
+a copy of the state, it was
+also promoted to be an \emph{active} secondary
+(one that has caught up with the other secondaries).
+This caused an assertion failure in our model
+allowing us to detect and fix this incorrect behaviour.
 

### 2015-09-19 experience.tex
removed two comments towards getting a final draft

@@
 %\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
-\PTComment{Re-added the detail here so we can discuss the bug below:}
 One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}. The primary replicates state changes to the secondaries 
@@
 
-\PTComment{Changed this para; added detail about the bug: }
 \psharp's systematic testing was very helpful in debugging the model itself.

### 2015-09-20 intro.tex
update in intro
updates to intro

@@
 This challenge is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside components, unexpected node failures, data losses due to unreliable communication channels, and interaction with (human) clients.
-All these sources of nondeterminism can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix. 
+All these sources of nondeterminism can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, corner-case bugs that are difficult to detect, diagnose and fix.
 A Heisenbug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
 
-Developers of production distributed systems use many testing techniques,
-such as unit testing, integration testing, stress testing, and fault injection.
-In spite of extensive use of these testing methods,
-many bugs that arise from subtle combinations of concurrency and failure events
-are missed during testing and get exposed only after a system has been put into production.
+Developers of production distributed systems use many testing techniques, such as unit testing, integration testing, stress testing, and fault injection.
+In spite of extensive use of these testing methods, many bugs that arise from subtle combinations of concurrency and failure events are missed during testing and get exposed only after a system has been put into production.
 However, allowing bugs to reach production, before they are found and fixed, can cost organizations a lot of money~\cite{tassey2002economic} and lead to customer dissatisfaction~\cite{amazon2012aws, google2014outage}.
@@
 The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes to production.
-The need for better testing techniques is not specific to Microsoft;
-other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.
+The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems that are used in production.
 
-Recently, Amazon Web Services (AWS) used formal methods, where ``engineers use TLA+ to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
-The gist of their approach is to extract high-level logic from production systems, represent the logic as specification in an expressive language (such as TLA+), and verify the specification using a model checker.
-While highly effective, as demonstrated by the numerous successes in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'' and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
+Recently, engineers in Amazon Web Services (AWS) described their use of formal methods ``to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
+The gist of their approach is to extract the high-level logic from a production system, represent this logic as specifications in the expressive TLA+~\cite{lamport1994temporal} language, and finally verify the specifications using a model checker.
+While highly effective, as demonstrated by its use in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'', and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
 
@@
 Our goal is to bridge this gap.
-We propose a new framework that validates high-level specifications directly on the executable code.
-Our framework is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{}.
-Instead, we allow developers to test \emph{unmodified production code} by writing test harnesses in \csharp, a mainstream programming language.
+We propose a new methodology that validates high-level specifications directly on the executable code.
+Our methodology is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{}.
+Instead, we allow developers to test \emph{production code} by writing test harnesses in \csharp, a mainstream programming language.
 This significantly lowered the acceptance barrier for adoption by the Microsoft Azure team.
@@
 %We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
-Our framework is based on \psharp~\cite{deligiannis2015psharp}, an extension of \csharp that provides support for modeling, specification, and systematic testing of distributed systems written in Microsoft's .NET framework.
+Our methodology is based on \psharp~\cite{deligiannis2015psharp}, an extension of \csharp that provides support for modeling, specification, and systematic testing of distributed systems written in Microsoft's .NET framework.
 To test a distributed system with \psharp, the programmer augments the original system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
@@
 
-We have applied \psharp to three distributed storage systems in Microsoft Azure: Azure Storage vNext, Azure Table Live Migration, and Azure Service Fabric.
+We have applied \psharp to three distributed storage systems inside Microsoft: Azure Storage vNext; Azure Table Live Migration; and Azure Service Fabric.
 %a distributed storage system; a system for migrating data across partitions; and a big-data computing system.
-We uncovered numerous bugs in these systems, including, most notably, a subtle liveness bug that only intermittently 
-manifested during stress testing for months without being fixed.
-Our testing approach uncovered this bug in a very small setting, which made it easy to examine traces, identify, and eventually fix the problem.
-Our experience demonstrates that bugs in production do not need a large setting to be reproduced;
-they can be reproduced in small settings with easy to understand traces.
+We uncovered numerous bugs in these systems, including, most notably, a subtle liveness bug that only intermittently manifested during stress testing for months without being fixed.
+Our testing approach uncovered this bug in a very small setting, which made it easy for developers to examine traces, identify, and fix the problem.
+%Our experience demonstrates that bugs in production do not need a large setting to be reproduced; they can be reproduced in small settings with easy to understand traces.
 
@@
 \begin{itemize}
-\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms in \psharp.
+\item We present a new methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms in \psharp.
 \item Because \psharp is an extension of \csharp, our infrastructure can test production code written in \csharp.
 \item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
-\item We show that we can reproduce bugs in production systems in a small setting with easy to understand traces.
+\item We show that \psharp can reproduce bugs in production-scale distributed systems in a small setting with easy to understand traces.
 \end{itemize}

### 2015-09-20 paper.tex
balance citations in the last page
Small typos fixed.
updates to abstract

@@
 \usepackage{xspace}
+
+\usepackage{balance}
 
@@
 Testing distributed systems is challenging due to multiple sources of nondeterminism.
-Conventional testing techniques are ineffective in preventing serious but subtle bugs from reaching production.
-Formal techniques, such as the use of TLA+ in Amazon Web Services, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
-In this paper, we present a new framework for testing distributed storage systems.
+Conventional testing techniques, such as unit and stress testing, are ineffective in preventing serious but subtle bugs from reaching production.
+Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
+In this paper, we present a new methodology for testing production distributed storage systems.
 Our approach applies systematic testing to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
 
-Our framework has been applied to three distributed storage systems in the Microsoft Azure platform.
+Our technique has been applied to three production distributed storage systems in the Microsoft Azure platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.
-These bugs required subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques.
+These bugs required a subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques.
 An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace, which dramatically increases the productivity of debugging.
@@
 
+% Balances the bibliography in the last page
+\balance
+
 {\footnotesize \bibliographystyle{acm}

### 2015-09-20 paper.tex
updates to abstract

@@
 \subsection*{Abstract}
-Testing distributed systems is challenging due to multiple sources of nondeterminism.
+Testing distributed storage systems is challenging due to multiple sources of nondeterminism.
 Conventional testing techniques, such as unit and stress testing, are ineffective in preventing serious but subtle bugs from reaching production.
 Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
-In this paper, we present a new methodology for testing production distributed storage systems.
-Our approach applies systematic testing to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
+In this paper, we present a new methodology for testing production distributed systems.
+Our approach applies systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
 
-Our technique has been applied to three production distributed storage systems in the Microsoft Azure platform.
+Our methodology has been applied to three production distributed storage systems in the Microsoft Azure platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.

### 2015-09-20 motivation.tex
updates to motivation

@@
 
-To uncover this bug and many other similar ones, the developers are in constant search of a generic and systematic approach for testing distributed storage systems.
+To uncover this bug, and other tricky ones, the developers are in constant search of a generic and systematic approach for testing distributed storage systems.
 

### 2015-09-20 motivation.tex
updates to motivation

@@
 
-To uncover this bug, and other tricky ones, the developers are in constant search of a generic and systematic approach for testing distributed storage systems.
+To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing production-scale distributed storage systems.
 

### 2015-09-20 overview.tex
updates in section 3

@@
-The goal of our work is to develop testing techniques to detect and debug Heisenbugs in distributed storage systems prior to deployment.
-To achieve our goal, we use \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
+The goal of our work is to develop testing techniques for detecting and debugging Heisenbugs in distributed storage systems prior to deployment.
+To achieve our goal, we use \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream \csharp language that provides: (i) language support for \emph{modeling the environment} of distributed systems developed using Microsoft's .NET framework; and (ii) a \emph{testing engine} that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
 
@@
 
-Our approach of using \psharp to test distributed systems, implemented in the .NET framework, requires the developer to perform the following three key modeling tasks.
+Our approach of using \psharp to test distributed systems, implemented in .NET, requires the developer to perform the following three key modeling tasks.
 These tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a running example.
@@
 \noindent
-\psharp uses modern languages features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
+\psharp uses modern object-oriented language features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
 Programmers in industry are used to working with such features, and heavily use them in production for testing purposes.

### 2015-09-20 overview.tex
updates in section 3

@@
 \textbf{Safety specifications}.
-In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.
+In addition to the usual assertions for stating safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.
+
 A safety monitor maintains local state that is modified in response to events received from ordinary (non-monitor) machines.
@@
 An erroneous global behavior is flagged via an assertion on the private state of the safety monitor.
-Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).
-
-To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class.
-Monitors in \psharp are singleton instances, which means that an ordinary machine does not need a reference to a monitor to send it an event.
-An ordinary machine can send an event to a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send together with an optional payload \texttt{p}.
+Thus, a \psharp monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).
 
 \textbf{Liveness specifications}.
-Liveness property specifications are more difficult to encode since they are intended to capture program progress
-over infinite executions.
+Liveness property specifications are more difficult to encode since they are intended to capture program progress over infinite executions.
 Violation of a liveness property can only be demonstrated by an infinite execution.
 Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
-We take a different approach and allow the programmer to write a \emph{liveness monitor}.
-Similar to a safety monitor, a liveness monitor receives events but does not send them.
-Unlike a safety monitor, a liveness monitor~\cite{desai2015building}
-contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
-A hot state denotes a point in the execution where progress is required but has not happened yet,
-e.g., a node has failed but a new one has not come up yet.
-A liveness monitor enters a hot state upon receiving notification of an event that requires the system
-to make progress and leaves the hot state upon receiving another event notifying it of progress.
+We take a different approach and allow the programmer to write a \emph{liveness monitor}~\cite{desai2015building}.
+Similar to a safety monitor, a liveness monitor can receive, but not send, events.
+Unlike a safety monitor, a liveness monitor contains two types of states: \emph{hot} and \emph{cold}.
+
+A hot state denotes a point in the execution where progress is required but has not happened yet, for example a node has failed but a new one has not come up yet.
+A liveness monitor enters a hot state upon receiving notification of an event that requires the system to make progress.
+The liveness monitor leaves the hot state and enters a cold state upon receiving another event notifying it of progress.
 An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
 Our liveness monitors can encode arbitrary temporal logic properties.
+A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
+%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events, such as a periodic request or a periodic timer expiration.
 
-A thermometer analogy is useful for understanding a liveness monitor.
-A hot state increases the temperature by a small value.
-A cold state resets the temperature to zero.
-A liveness error happens if the temperature becomes infinite.
-%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events,
-%such as a periodic request or a periodic timer expiration.
+A liveness violation is witnessed by an \emph{infinite} execution in which all concurrently executing \psharp machines are \emph{fairly} scheduled.
+Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time.
+Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics.
+One such heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
+Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution, and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state graph of the program.
 
-A liveness violation is witnessed by an infinite execution in which all concurrently executing \psharp machines are scheduled fairly.
-Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time.
-Therefore, our implementation of liveness checking approximates an infinite execution using several heuristics.
-One heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
-Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution
-and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state
-graph of the program.
+\textbf{Using monitors}.
+To declare a safety or a liveness monitor, the programmer must inherit from the \psharp \texttt{Monitor} class.
+\psharp monitors are singleton instances and, thus, an ordinary machine does not need a reference to a monitor to send it an event.
+Instead, an ordinary machine can invoke a monitor by calling \texttt{Monitor<M>(e, p)}, where the parameter \texttt{e} is the event being send, and the parameter \texttt{p} is an optional payload.
 

### 2015-09-20 overview.tex
updates in section 3

@@
 The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in bug-finding mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.
+A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.
 
@@
 
-\subsubsection{Handling \csharp 5.0 asynchrony}
-\label{sec:psharp:async}
+\textbf{Handling \csharp 5.0 asynchrony}.
+The Live Azure Table Migration system (see Section~\ref{sec:cases:migration}) heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread.
+The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in an internal TPL \emph{task continuation}, and returning execution to the caller of the \texttt{async} method.
+The TPL scheduler executes this task continuation when the task being awaited has completed.
+In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing.
+However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus unlikely to scale for large code bases.
+To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.
 
-The Live Azure Migration system (Section~\ref{sec:cases:migration}) uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
-An \texttt{async} procedure is allowed to perform \texttt{await} on a TPL task without blocking the current thread.
-The \csharp compiler achieves this non-blocking behavior by wrapping the \emph{continuation} of the \texttt{await} statement
-in a TPL task and returning it to the caller.
-The TPL scheduler executes this task when the task being awaited has completed.
-In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing.
-However, in our experience, refactoring code to model in \psharp the asynchrony in async/await is difficult and unlikely
-to scale for large code bases.
-Therefore, our implementation automatically captures asychronously created tasks in .NET runtime as \psharp machines.
-
-Our solution involves using a custom task scheduler whenever the bug-finding mode of the \psharp runtime is enabled, instead of the default TPL task scheduler.
-The \psharp task scheduler inherits from the \texttt{TaskScheduler} class, a low-level API that is responsible for enqueueing TPL tasks into threads.
-Our approach works as follows.
-We start the task of the \emph{root} \psharp machine in our custom task scheduler.
-As long as any child tasks spawned by the root machine task do not explicitly start in another scheduler (including the default TPL scheduler), then they will be scheduled for execution in our custom task scheduler.
-Our custom scheduler intercepts the call to enqueue a task, creates a special machine (only internally available to \psharp) called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine.
-The constructor of a \texttt{TaskMachine} takes as an argument a TPL task and stores it in an field.
-The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it starts executing the wrapped task and then immediately waits on its completion.
-This forces the lifetime of the task to become the lifetime of the machine and vice versa.
-Thus, when the \psharp systematic testing scheduler schedules the \texttt{TaskMachine}, it will also schedule the corresponding task.
-This means that when a \psharp program makes a call to an \texttt{async} method, then the task created in the backend will be automatically wrapped and scheduled.
-
-For \texttt{await} statements, no special treatment is required because \texttt{await} statements are compiled into a continuation.
-The task associated with the continuation will also be wrapped and scheduled accordingly.
+Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
+Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon), any TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler.
+These tasks will be intercepted by our custom task scheduler and wrapped inside a special, available only internally, task-machine.
+When the \psharp testing engine schedules a task-machine for execution, the corresponding task is unwrapped and executed.
+This approach works nicely with the \texttt{async} and \texttt{await} \csharp 5.0 primitives: when an \texttt{async} method is called, the asynchronous TPL task will be automatically wrapped in a task-machine and scheduled by \psharp; for \texttt{await}, no special treatment is required since the task continuation will also be automatically wrapped and scheduled accordingly.

### 2015-09-20 evaluation.tex
updates in case studies

@@
 For the vNext case study, the random scheduler was able to reproduce the bug in less than 10 seconds.
-The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in Section~\ref{sec:psharp:testing} we leave the program to run for a long time before checking if the liveness property holds.
+The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:psharp:testing} we leave the program to run for a long time before checking if the liveness property holds.
 The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.

### 2015-09-20 experience.tex
updates in case studies

@@
 
-The Live Azure Table Migration (MigratingTable) is a library for
-\emph{transparently migrating} a data set between tables in an Azure Storage
-service \emph{while} an application is accessing this data set. This case study
-is interesting because its developers built the \psharp model of the system
-while developing the library to increase confidence in its implementation;
-indeed they discovered many bugs during this \emph{co-development} process.
-Another interesting aspect of this system is that it heavily
-uses the \texttt{async}/\texttt{await} primitives, which were not supported in
-the original \psharp runtime, and thus required us to extend \psharp with
-support for intra-machine concurrency (see Section~\ref{sec:psharp:async}). Finally, multiple 
-complex safety properties were written to test MigratingTable with \psharp, 
-whereas the vNext model focused on a single liveness property.
+The Live Azure Table Migration (MigratingTable) is a library for \emph{transparently migrating} a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
+This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp model to increase confidence in the former's implementation.
+Indeed, this co-development process resulted in many tricky bugs discovered before the system went into production (see \S\ref{sec:eval}).
+The MigratingTable modeling effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property; whereas the MigratingTable developers had to write many complex safety properties.
+MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp.
+To be able to test MigratingTable, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).
 
@@
 %We wanted to implement the core MigratingTable algorithms in \csharp ``async/await'' code, like most of Artifact Services, to achieve both good readability and good performance.
-We used a method similar to that described in Section~\ref{sec:psharp:async} to bring the generated TPL tasks under the control of the \psharp scheduler.
+We used a method similar to that described in \S\ref{sec:psharp:async} to bring the generated TPL tasks under the control of the \psharp scheduler.
 Then we implemented an ``async'' RPC mechanism based on the .NET RealProxy class that automates the generation of proxies for objects hosted by other \psharp machines (in our setting, the service machines use proxies for the \texttt{SpecTable}s and various auxiliary objects hosted by the tables machine).

### 2015-09-20 methodology.tex
updates in case studies

@@
 The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
-This is discussed in detail in Section~\ref{sec:method:model:dispath}.
+This is discussed in detail in \S\ref{sec:method:model:dispath}.
 
@@
 
-\texttt{RepairMonitor} is a liveness monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
+\texttt{RepairMonitor} is a \psharp liveness monitor machine (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
 Whenever an extent node fails, \texttt{RepairMonitor} is notified.
@@
 
-It took less than ten seconds for the \psharp testing engine to report the first occurrence of a liveness bug in vNext (see Section~\ref{sec:eval}).
-Upon examining the debug trace, the vNext developers were able to confirm the bug.
+It took less than ten seconds for the \psharp testing engine to report the first occurrence of a liveness bug in vNext (see \S\ref{sec:eval}).
+Upon examining the debug trace, the developers of vNext were able to quickly confirm the bug.
 
@@
 synchronization messages, heartbeats and repair logic loops).
-This is discussed in detail in Section~\ref{sec:method:model:timers}.
+This is discussed in detail in \S\ref{sec:method:model:timers}.
 %

### 2015-09-20 overview.tex
updates in case studies

@@
 Our approach of using \psharp to test distributed systems, implemented in .NET, requires the developer to perform the following three key modeling tasks.
-These tasks are illustrated in Section~\ref{sec:method} using the Azure Storage vNext as a running example.
+These tasks are illustrated in \S\ref{sec:method} using the Azure Storage vNext as a running example.
 
@@
 The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.
+A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:bg:bugs}.
 
@@
 \textbf{Handling \csharp 5.0 asynchrony}.
-The Live Azure Table Migration system (see Section~\ref{sec:cases:migration}) heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+The Live Azure Table Migration system (see \S\ref{sec:cases:migration}) heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
 An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread.
-The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in an internal TPL \emph{task continuation}, and returning execution to the caller of the \texttt{async} method.
+The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in a TPL \emph{task continuation}, and then returning execution to the caller of the \texttt{async} method.
 The TPL scheduler executes this task continuation when the task being awaited has completed.

### 2015-09-20 intro.tex
updates in related work

@@
 We propose a new methodology that validates high-level specifications directly on the executable code.
-Our methodology is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{}.
+Our methodology is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{simsa2011dbug}.
 Instead, we allow developers to test \emph{production code} by writing test harnesses in \csharp, a mainstream programming language.

### 2015-09-20 related.tex
updates in related work

@@
-\psharp is most closely related to \textsc{MoDist}~\cite{yang2009modist}, a model checker that can be applied on unmodified distributed systems.
-To achieve this, the tool uses binary instrumentation (reuses the instrumentation engine of the D$^3$S testing tool) to expose all actions of the system-under-test, and then uses a model checking engine to systematically explore these actions.
-\textsc{MoDist} also provides a virtual clock manager that can simulate timeouts and accelerate the passage of time.
-A major limitation of \textsc{MoDist} is that it must be applied on a whole system, which does not scale for production systems.
-In contrast, \psharp has built-in support for modeling the environment of individual components of a system, which allows the tool to achieve scalability.
-Further, \psharp is an extension of the \csharp language and has built-in systematic testing support; it does not need to perform binary instrumentation which can be computationally expensive.
+Model checking~\cite{godefroid1997verisoft} is a powerful technique that has been widely used in the past for finding tricky bugs in the actual implementation of distributed systems~\cite{killian2007life, yang2009modist, yabandeh2009crystalball, guerraoui2011model, guo2011practical, simsa2011dbug, leesatapornwongsa2014samc}.
+
+State-of-the-art model checkers, such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug}, typically focus on testing entire, often unmodified, distributed systems, an approach that easily leads to state-space explosion.
+In contrast, we try to offer a more pragmatic approach for handling state-space explosion: \psharp provides the capability for creating models of individual components of a large system, which helps towards component isolation and enables compositional testing.
+Our approach aims to enhance unit and integration testing, techniques widely used in production, where only individual (or a small number of) components are tested at each time.
+
+Tools such as CrystalBall~\cite{yabandeh2009crystalball} and \textsc{DeMeter}~\cite{guo2011practical} employ techniques based on dynamic partial order reduction~\cite{lauterburg2010evaluating} to reduce the exploration state-space.
+The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing, but the focus is still on reducing the set of interleavings.
+Although the SAMC authors claim that their tool is orders of magnitude faster than previous testing techniques, they mention that their ``extensive evaluation exercised more than 100,000 executions and used approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
+\psharp is able to explore a similar number of iterations in a matter of hours, and find serious bugs in seconds (see \S\ref{sec:eval}).
+However, we do \emph{not} claim that \psharp is faster; as aforementioned, we focus on compositional testing, whereas prior work typically focuses on testing an entire system.
+We are just trying to solve a similar problem at a very different level.
 
@@
 \psharp differs from \textsc{MaceMc} in that it can be applied on legacy code written in a mainstream language, whereas a system to be tested with \textsc{MaceMc} has to be written in \textsc{Mace}, which makes \textsc{MaceMc} harder to be applied in an industrial setting.
-Further, \psharp uses a simpler, but effective, algorithm for detecting liveness bugs.
 
@@
 
-A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.
-A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking.
-Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}.
+Formal methods have been successfully used in industry to verify the correctness of distributed protocols.
+A recent notable example is the use of TLA+~\cite{lamport1994temporal} by the Amazon Web Services team~\cite{newcombe2015aws}.
+TLA+ is an expressive formal specification language that can be used to design and verify concurrent programs via model checking.
 A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system.
-Even if the model is verified, the gap between a real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.
+Even if the model is verified, the gap between the real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.
 
-Another relevant approach is Verdi~\cite{wilcox2015verdi}, where a distributed system is written and verified in Coq, and then OCaml code is produced for execution.
-Verdi cannot find liveness bugs.
-\psharp is also more production-friendly: it works on \csharp, which is a mainstream language.
+Another formal approach is Verdi~\cite{wilcox2015verdi}, a framework for writing and verifying distributed systems in Coq~\cite{barras1997coq}.
+Verdi generates OCaml code from the verifying system, which can be used for execution.
+In contrast, \psharp performs bounded testing on a system already written in \csharp, which in our experience lowers the bar for adoption by product groups.
+Furthermore, Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed storage systems.

### 2015-09-20 paper.tex
update in abstract

@@
 Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
-In this paper, we present a new methodology for testing production distributed systems.
-Our approach applies systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
+In this paper, we present a new methodology for testing distributed storage systems.
+Our approach applies advanced systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
 

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableFactory/ATableAdapter.cs
Add Code coverage snapshot from stress test iteration

@@ -1,0 +1,82 @@
+﻿using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableFactory
+{
+    class ATableAdapter : IChainTable
+    {
+        // This Adapter adapts a normal Azure Table (ATable) to the IComposableTable interface.
+
+        public ATableAdapter(CloudTable table)
+        {
+            this.table = table;
+        }
+
+        public Table.TableResult Execute(Table.TableOperation operation, Table.TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            var res = table.Execute(operation, requestOptions, operationContext);
+            return res;
+        }
+
+        public IList<Table.TableResult> ExecuteBatch(Table.TableBatchOperation batch, Table.TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            try
+            {
+                return table.ExecuteBatch(batch, requestOptions, operationContext);
+            }
+            catch (StorageException e)
+            {
+                // find the error index
+                // WARNING: this is not very reliable
+                // c.f. http://stackoverflow.com/questions/14282385/azure-cloudtable-executebatchtablebatchoperation-throws-a-storageexception-ho/14290910#14290910
+                var msg = e.RequestInformation.ExtendedErrorInformation.ErrorMessage;
+                var parts = msg.Split(':');
+                var errorIndex = int.Parse(parts[0]);
+
+                throw new ChainTableInterface.ChainTableBatchException(errorIndex, e);
+            }
+        }
+
+        public Task<Table.TableResult> ExecuteAsync(Table.TableOperation operation, Table.TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            return table.ExecuteAsync(operation, requestOptions, operationContext);
+        }
+
+        public bool IsReadOnly(OperationContext operationContext)
+        {
+            return false;
+        }
+
+
+
+        private CloudTable table;
+
+
+        public string GetTableID()
+        {
+            return table.Name;
+        }
+
+
+        public bool CreateIfNotExists()
+        {
+            return table.CreateIfNotExists();
+        }
+
+        public bool DeleteIfExists()
+        {
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableFactory/DefaultChainTableService.cs

@@ -1,0 +1,76 @@
+﻿using Microsoft.Azure.Toolkit.Replication;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using Microsoft.WindowsAzure.Storage.RTable;
+using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.IO;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableFactory
+{
+    public class DefaultChainTableService : IChainTableService
+    {
+        public DefaultChainTableService(string configFile)
+        {
+            this.configFileName = configFile;
+            StreamReader fs = System.IO.File.OpenText(configFileName);
+            string line = fs.ReadLine();
+
+            rTabConfLocs = new List<ConfigurationStoreLocationInfo>();
+            rTabDataChain = new List<ReplicaInfo>();
+
+            bool first = true;
+            while (!fs.EndOfStream)
+            {
+                line = fs.ReadLine();
+                string[] tokens = line.Split();
+                string accountName = tokens[0];
+                string accountKey = tokens[1];
+
+                string connStr = String.Format("DefaultEndpointsProtocol=https;AccountName={0};AccountKey={1}", accountName, accountKey);
+                if (first)
+                {
+                    first = false;
+                    csa = CloudStorageAccount.Parse(connStr);
+                    client = csa.CreateCloudTableClient();
+                    rTabConfLocs.Add(new ConfigurationStoreLocationInfo() {
+                        StorageAccountName = accountName,
+                        StorageAccountKey = accountKey,
+                        BlobPath = Constants.RTableConfigurationBlobLocationContainerName + "/myRTableConfig1"
+                    });
+                }
+
+                rTabDataChain.Add(new ReplicaInfo() {
+                    StorageAccountName = accountName,
+                    StorageAccountKey = accountKey
+                });
+            }
+
+            ReplicatedTableConfigurationService rtableConfig = new ReplicatedTableConfigurationService(rTabConfLocs, true);
+            rtableConfig.UpdateConfiguration(rTabDataChain, 0);
+
+            fs.Close();
+        }
+
+        public IChainTable GetChainTable(string tableId)
+        {
+            if (tableId.StartsWith("__RTable_"))
+            {
+                var name = tableId.Substring(9);
+                ReplicatedTableConfigurationService rtableConfig = new ReplicatedTableConfigurationService(rTabConfLocs, true);
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableFactory/RTableAdapter.cs

@@ -1,0 +1,366 @@
+﻿//#define RPERF
+
+using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Reflection;
+using System.Text;
+using System.Threading.Tasks;
+using System.Net;
+using System.Diagnostics;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using Microsoft.Azure.Toolkit.Replication;
+
+namespace Microsoft.WindowsAzure.Storage.RTable
+{
+    public class RTableAdapter : IChainTable  // public for demo
+    {
+        public RTableAdapter(ReplicatedTable rtable)
+        {
+            this.rtable = rtable;
+        }
+
+        public TableResult Execute(TableOperation operation,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            var oldOp = operation;
+
+            // Encapsulate
+            operation = EncapsulateOp(oldOp, operationContext);
+
+            // Execute
+
+            var res = rtable.Execute(operation, requestOptions, operationContext);
+
+
+            // Currently RTable do not throw exception on error, so we throw it...
+            var opType = GetOpType(oldOp);
+
+            // Chunzhi: the published rtable returns 204 for insert :(
+            if (opType == TableOperationType.Insert && res.HttpStatusCode == (int)HttpStatusCode.NoContent)
+                res.HttpStatusCode = (int)HttpStatusCode.Created;
+            // Chunzhi: the published rtable returns 409 for precondition failure on merge / replace / delete
+            if ((opType == TableOperationType.Merge || opType == TableOperationType.Replace || opType == TableOperationType.Delete)
+                && res.HttpStatusCode == (int)HttpStatusCode.Conflict)
+                res.HttpStatusCode = (int)HttpStatusCode.PreconditionFailed;
+
+            if (
+                (opType == TableOperationType.Insert && res.HttpStatusCode != (int)HttpStatusCode.Created) ||
+                (opType == TableOperationType.Replace && res.HttpStatusCode != (int)HttpStatusCode.NoContent) ||
+                (opType == TableOperationType.Merge && res.HttpStatusCode != (int)HttpStatusCode.NoContent) ||
+                (opType == TableOperationType.InsertOrReplace && res.HttpStatusCode != (int)HttpStatusCode.NoContent) ||
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/AzureTableAccessors.cs

@@ -1,0 +1,62 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Reflection;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    public static class AzureTableAccessors
+    {
+        // C# doesn't support extension properties.  Defining an extension method named
+        // "get_OperationType" does not work either because it doesn't have the "specialname"
+        // MSIL attribute, and even so, I don't know if the method lookup done by a property
+        // lookup supports extension methods.
+
+        private static readonly PropertyInfo TableOperation_OperationType =
+            typeof(TableOperation).GetProperty("OperationType",
+                BindingFlags.GetProperty | BindingFlags.Instance | BindingFlags.NonPublic);
+
+        public static TableOperationType GetOperationType(this TableOperation operation)
+        {
+            return (TableOperationType)TableOperation_OperationType.GetValue(operation, null);
+        }
+
+        private static readonly PropertyInfo TableOperation_Entity =
+            typeof(TableOperation).GetProperty("Entity",
+                BindingFlags.GetProperty | BindingFlags.Instance | BindingFlags.NonPublic);
+
+        public static ITableEntity GetEntity(this TableOperation operation)
+        {
+            return (ITableEntity)TableOperation_Entity.GetValue(operation, null);
+        }
+
+        private static readonly PropertyInfo TableOperation_RetrievePartitionKey =
+            typeof(TableOperation).GetProperty("RetrievePartitionKey",
+                BindingFlags.GetProperty | BindingFlags.Instance | BindingFlags.NonPublic);
+        public static string GetRetrievePartitionKey(this TableOperation operation)
+        {
+            return (string)TableOperation_RetrievePartitionKey.GetValue(operation, null);
+        }
+
+        private static readonly PropertyInfo TableOperation_RetrieveRowKey =
+            typeof(TableOperation).GetProperty("RetrieveRowKey",
+                BindingFlags.GetProperty | BindingFlags.Instance | BindingFlags.NonPublic);
+        public static string GetRetrieveRowKey(this TableOperation operation)
+        {
+            return (string)TableOperation_RetrieveRowKey.GetValue(operation, null);
+        }
+
+        private static readonly PropertyInfo TableOperation_RetrieveResolver =
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/ChainTableBatchException.cs

@@ -1,0 +1,31 @@
+﻿using System;
+using System.Runtime.Serialization;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    [Serializable]
+    public class ChainTableBatchException : StorageException
+    {
+        public int FailedOpIndex { get; private set; }
+
+        public ChainTableBatchException(int failedOpIndex, StorageException storageEx)
+            : base(storageEx.RequestInformation,
+                  string.Format("{0}: {1}", failedOpIndex, storageEx.Message),
+                  storageEx.InnerException)
+        {
+            this.FailedOpIndex = failedOpIndex;
+        }
+
+        protected ChainTableBatchException(SerializationInfo info, StreamingContext context)
+            : base(info, context)
+        {
+            FailedOpIndex = info.GetInt32("FailedOpIndex");
+        }
+
+        public override void GetObjectData(SerializationInfo info, StreamingContext context)
+        {
+            base.GetObjectData(info, context);
+            info.AddValue("FailedOpIndex", FailedOpIndex);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/ChainTableUtils.cs

@@ -1,0 +1,325 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using Microsoft.WindowsAzure.Storage.Table.Protocol;
+using System;
+using System.Collections.Generic;
+using System.Collections.Immutable;
+using System.Diagnostics;
+using System.Linq;
+using System.Net;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    public class PrimaryKey : IEquatable<PrimaryKey>, IComparable<PrimaryKey>
+    {
+        public string PartitionKey { get; private set; }
+        public string RowKey { get; private set; }
+        public PrimaryKey(string PartitionKey, string RowKey)
+        {
+            if (PartitionKey == null) throw new ArgumentNullException("PartitionKey");
+            if (RowKey == null) throw new ArgumentNullException("RowKey");
+            this.PartitionKey = PartitionKey;
+            this.RowKey = RowKey;
+        }
+        internal static PrimaryKey Of(ITableEntity entity)
+        {
+            return new PrimaryKey(entity.PartitionKey, entity.RowKey);
+        }
+
+        public int CompareTo(PrimaryKey other)
+        {
+            // XXX Framework please
+            int cmp = string.CompareOrdinal(PartitionKey, other.PartitionKey);
+            if (cmp != 0) return cmp;
+            return string.CompareOrdinal(RowKey, other.RowKey);
+        }
+
+        public bool Equals(PrimaryKey that)
+        {
+            return Equals(PartitionKey, that.PartitionKey) && Equals(RowKey, that.RowKey);
+        }
+        public override bool Equals(object o)
+        {
+            return Equals(o as PrimaryKey);
+        }
+
+        public override int GetHashCode()
+        {
+            // Avoid a dependency on Migration.Utils for now.
+            unchecked
+            {
+                return 31 * PartitionKey.GetHashCode() + RowKey.GetHashCode();
+            }
+        }
+
+        public override string ToString()
+        {
+            return string.Format("PrimaryKey({0}, {1})", PartitionKey, RowKey);
+        }
+    }
+
+    public static partial class ChainTableUtils
+    {
+        // No extension properties. :(
+        // It would be even more awesome if extension properties could be used
+        // in the property constructor syntax.
+        public static PrimaryKey GetPrimaryKey(this ITableEntity entity)
+        {
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/FilterExpression.cs

@@ -1,0 +1,204 @@
+﻿using Irony.Parsing;
+using Microsoft.WindowsAzure.Storage.Table;
+using Microsoft.WindowsAzure.Storage.Table.Protocol;
+using System;
+using System.Collections.Generic;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    // All the parsing stuff is thrown together, but it appears to work...
+
+    class FilterStringGrammar : Grammar
+    {
+        internal readonly NonTerminal comparison, booleanOpExpr;
+        FilterStringGrammar()
+        {
+            var propertyName = new IdentifierTerminal("propertyName");
+            var compareOp = new NonTerminal("compareOp");
+            compareOp.Rule = ToTerm(QueryComparisons.Equal) | ToTerm(QueryComparisons.NotEqual)
+                | ToTerm(QueryComparisons.LessThan) | ToTerm(QueryComparisons.LessThanOrEqual)
+                | ToTerm(QueryComparisons.GreaterThan) | ToTerm(QueryComparisons.GreaterThanOrEqual);
+            var stringLiteral = new StringLiteral("stringLiteral", "'", StringOptions.AllowsDoubledQuote);
+            var booleanLiteral = new ConstantTerminal("booleanLiteral");
+            booleanLiteral.Add("true", true);
+            booleanLiteral.Add("false", false);
+            var literal = new NonTerminal("literal");
+            literal.Rule = stringLiteral | booleanLiteral;
+            comparison = new NonTerminal("comparison");
+            comparison.Rule = propertyName + compareOp + literal;
+            var booleanOp = new NonTerminal("booleanOp");
+            booleanOp.Rule = ToTerm(TableOperators.And) | ToTerm(TableOperators.Or);
+            var booleanExpr = new NonTerminal("booleanExpr");
+            booleanOpExpr = new NonTerminal("booleanOpExpr");
+            booleanOpExpr.Rule = "(" + booleanExpr + ")" + booleanOp + "(" + booleanExpr + ")";
+            booleanExpr.Rule = comparison | booleanOpExpr;
+            Root = booleanExpr;
+        }
+
+        internal static readonly FilterStringGrammar Grammar = new FilterStringGrammar();
+        internal static readonly Parser Parser = new Parser(Grammar);
+    }
+
+    public abstract class FilterExpression
+    {
+        // ITableEntity does not provide random access to properties. :(
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/IChainTable.cs

@@ -1,0 +1,69 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    // Any storage implementation that implements IChainTable can be used in TxTable
+    public interface IChainTable
+    {
+        /*
+         * Execute API:
+         *   (1) Retrieve
+         *          On success, return TableResult, code = 200 (OK);
+         *          If the row does not exist, return TableResult, code = 404 (NotFound);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         *   (2) Insert
+         *          On success, return TableResult, code = 201 (Created);
+         *          If the row already exists, throw storage exception, code = 409 (Conflict);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         *   (3) Replace / Merge / Delete
+         *          On success, return TableResult, code = 204 (NoContent);
+         *          If the row does not exist, throw storage exception, code = 404 (NotFound);
+         *          If ETag mismatches, throw storage exception, code = 412 (Precondition);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         *   (4) InsertOrReplace / InsertOrMerge
+         *          On success, return TableResult, code = 204 (NoContent);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         */         
+        TableResult Execute(TableOperation operation,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+
+        /*
+         * Batch Execute API:
+         *   Caller should not put retrieve operations inside a batch, so any error will lead to
+         *   an Exception.
+         *   
+         *   On success, return a list of TableResult. Each TableResult should follow the API
+         *   defined in Execute operation.
+         *   
+         *   Otherwise, throw a ChainTableBatchException, with failedOpIndex marks to the first
+         *   operation that causes the failure, and storageEx being the corresponding exception
+         *   defined in the Execute API.
+         */
+        IList<TableResult> ExecuteBatch(TableBatchOperation batch,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+
+        Task<TableResult> ExecuteAsync(TableOperation operation,
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/IChainTable2.cs

@@ -1,0 +1,160 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Collections.Immutable;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    public static class ChainTable2Constants
+    {
+        public const string ETAG_ANY = "*";  // Azure client library doesn't use a constant for this.
+        public const string ETAG_DELETE_IF_EXISTS = "**";
+    }
+
+    // Interface used by the migration library, intended to converge with IChainTable eventually.
+    public interface IChainTable2
+    {
+        /*
+         * WARNING: Like IChainTable, this interface uses optional parameters.
+         * Strange behavior that I don't fully understand can occur if a method
+         * overrides a superclass method or implements an interface method and
+         * doesn't specify the same optional parameters, or in certain separate
+         * compilation scenarios.  See
+         * http://stackoverflow.com/questions/8909811/c-sharp-optional-parameters-on-overridden-methods
+         * for some information (not verified to be up to date).  Duplicating
+         * the default values in subclasses is still less work than defining
+         * different overloads.
+         */
+
+        // XXX: Existing practice (at least in STable) seems to be to pass the
+        // same requestOptions and operationContext down to all the outgoing
+        // calls and not worry about them for ReadEntity/WriteEntity.  Is this
+        // actually sensible?  Should we consider stashing them in the
+        // ExecutionContext to reduce boilerplate?
+
+        // All methods can also throw StorageException if they get an unexpected
+        // result or exception from a backend table.
+
+        /*
+         * Execute API:
+         *   (1) Retrieve
+         *          On success, return TableResult, code = 200 (OK);
+         *          If the row does not exist, return TableResult, code = 404 (NotFound);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         *   (2) Insert
+         *          On success, return TableResult, code = 201 (Created);
+         *          If the row already exists, throw storage exception, code = 409 (Conflict);
+         *          Otherwise, throw storage exception with corresponding http error code.
+         *   (3) Replace / Merge / Delete
+         *          On success, return TableResult, code = 204 (NoContent);
+         *          If the row does not exist, throw storage exception, code = 404 (NotFound);
+         *          If ETag mismatches, throw storage exception, code = 412 (Precondition);
+         *          Otherwise, throw storage exception with corresponding http error code.
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/ChainTableInterface/IChainTableService.cs

@@ -1,0 +1,14 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Microsoft.WindowsAzure.Storage.ChainTableInterface
+{
+    public interface IChainTableService
+    {
+        // Translate a Kiwi table identity to a Kiwi table
+        IChainTable GetChainTable(string tableId);
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/ConfigurationService.cs

@@ -1,0 +1,105 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Migration
+{
+    // Configuration service with the ability to block on subscribers to
+    // acknowledge a new configuration.
+    public interface IReadOnlyConfigurationService<TConfig>
+    {
+        // Warning!  Currently the configuration service keeps a strong
+        // reference to the subscriber, so the object owning the subscriber must
+        // know its own intended lifetime (e.g., by implementing IDisposable
+        // itself) and dispose the subscription accordingly to avoid leaking
+        // subscriptions.  I looked into making subscribers weakly referenced,
+        // but it quickly became complicated.  XXX: Reconsider as a low
+        // priority.
+        IDisposable Subscribe(IConfigurationSubscriber<TConfig> subscriber, out TConfig currentConfig);
+    }
+
+    public interface IConfigurationService<TConfig> : IReadOnlyConfigurationService<TConfig>
+    {
+        // No concurrent calls please.  (For now; we could add support later.)
+        Task PushConfigurationAsync(TConfig newConfig);
+    }
+
+    // We could use IObserver if the data item itself included a way for the
+    // recipient to acknowledge it, but I don't see the benefit outweighing the
+    // awkwardness.  We could use a delegate, which would save clients some
+    // boilerplate at the expense of clarity.
+    public interface IConfigurationSubscriber<TConfig>
+    {
+        // It's OK if the subscriber unsubscribes first, in which case it
+        // doesn't matter if the task returned by this method ever completes.
+        Task ApplyConfigurationAsync(TConfig newConfig);
+    }
+
+    // Subscriber that never accepts updates, so updates can only be unblocked
+    // by its unsubscription.
+    public class FixedSubscriber<TConfig> : IConfigurationSubscriber<TConfig>
+    {
+        private FixedSubscriber() { }
+        public static readonly FixedSubscriber<TConfig> Instance = new FixedSubscriber<TConfig>();
+        private static readonly Task eternalTask = new TaskCompletionSource<object>().Task;
+
+        public Task ApplyConfigurationAsync(TConfig newConfig)
+        {
+            return eternalTask;
+        }
+    }
+
+    // For now, this is for modeling only.
+    // XXX: Make this extensible so the configuration can be persisted?
+    public class InMemoryConfigurationService<TConfig> : IConfigurationService<TConfig>
+    {
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/IChainTableMonitor.cs

@@ -1,0 +1,66 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Migration
+{
+    public class SpuriousETagChange
+    {
+        public readonly string partitionKey;
+        public readonly string rowKey;
+        public readonly string newETag;
+        public SpuriousETagChange(string partitionKey, string rowKey, string newETag)
+        {
+            this.partitionKey = partitionKey;
+            this.rowKey = rowKey;
+            this.newETag = newETag;
+        }
+    }
+
+    public interface IChainTableMonitor
+    {
+        /*
+         * The backend and monitor calls by verifiable table layers should follow this sequence
+         * (this does not preclude other local work from being done in parallel):
+         *
+         *  (
+         *    await backend call
+         *    await AnnotateLastOutgoingCallAsync
+         *  )*
+         *
+         * This applies even if the backend call throws an exception that is
+         * part of the semantics.  The expected handling of unexpected
+         * exceptions is not clearly defined.
+         *
+         * XXX: Have the verification framework enforce this automatically?
+         */
+
+        /*
+         * If the current incoming call is an ExecuteBatchAsync that is going to
+         * succeed, then successfulBatchResult must be its result, including the
+         * new ETags.  This allows the verification harness to mirror the batch on
+         * the reference table using the same new ETags before unlocking the
+         * TablesMachine for other ServiceMachines.  (The verification does not
+         * use ExecuteAsync.)
+         *
+         * In all other cases, successfulBatchResult must be null.
+         *
+         * For now, the verification does not deal with ExecuteQueryStreamed at
+         * all.  ExecuteQueryStreamed backend calls should not be annotated.  A
+         * verifiable table layer's implementation of ExecuteQueryStreamed
+         * should not report a linearization point, but it should annotate
+         * backend calls (other than ExecuteQueryStreamed) as normal.
+         *
+         * Passing one or more spurious ETag changes in combination with
+         * wasLinearizationPoint = true is disallowed until we have a use case
+         * to decide what its semantics should be.
+         */
+        Task AnnotateLastBackendCallAsync(
+            bool wasLinearizationPoint = false,
+            IList<TableResult> successfulBatchResult = null,
+            IList<SpuriousETagChange> spuriousETagChanges = null);
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/InMemoryTable.cs

@@ -1,0 +1,229 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Threading.Tasks;
+using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.Table;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using System.Linq;
+using System.Net;
+
+namespace Migration
+{
+    class InMemoryTable : AbstractMirrorChainTable2
+    {
+
+        int nextEtag = 1;
+        SortedDictionary<PrimaryKey, DynamicTableEntity> table = new SortedDictionary<PrimaryKey, DynamicTableEntity>();
+
+        public SortedDictionary<PrimaryKey, DynamicTableEntity> Dump()
+        {
+            var dump = new SortedDictionary<PrimaryKey, DynamicTableEntity>();
+            foreach (KeyValuePair<PrimaryKey, DynamicTableEntity> kvp in table)
+                dump.Add(kvp.Key, ChainTableUtils.CopyEntity<DynamicTableEntity>(kvp.Value));
+            return dump;
+        }
+
+        private void Merge(DynamicTableEntity to, ITableEntity from)
+        {
+            foreach (KeyValuePair<string,EntityProperty> kvp in from.WriteEntity(null))
+            {
+                to.Properties[kvp.Key] = ChainTableUtils.CopyProperty(kvp.Value);
+            }
+        }
+
+        public override Task<IList<TableResult>> ExecuteMirrorBatchAsync(
+            TableBatchOperation originalBatch, IList<TableResult> originalResponse,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            ChainTableUtils.GetBatchPartitionKey(originalBatch);  // For validation; result ignored
+
+            // Copy the table.  Entities are aliased to the original table, so don't mutate them.
+            var tmpTable = new SortedDictionary<PrimaryKey, DynamicTableEntity>(table);
+            int tmpNextEtag = nextEtag;
+            var results = new List<TableResult>();
+            for (int i = 0; i < originalBatch.Count; i++)
+            {
+                TableOperation op = originalBatch[i];
+                TableOperationType opType = op.GetOperationType();
+                ITableEntity passedEntity = op.GetEntity();
+                PrimaryKey key = passedEntity.GetPrimaryKey();
+                DynamicTableEntity oldEntity = tmpTable.GetValueOrDefault(key);
+
+                DynamicTableEntity newEntity = null;
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/MTableEntity.cs

@@ -1,0 +1,73 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System.Collections.Generic;
+using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+
+namespace Migration
+{
+    enum MTablePartitionState
+    {
+        //POPULATING, // Used by Kstart optimization, not yet implemented.
+        POPULATED,
+        SWITCHED
+    }
+
+    class MTableEntity : TableEntity
+    {
+        // Extra fields for the partition meta row (only)
+        internal const string KEY_PARTITION_STATE = "_mtable_partition_state";
+        internal MTablePartitionState? partitionState = null;
+
+        // Extra fields for ordinary rows
+        internal const string KEY_DELETED = "_mtable_deleted";
+        internal bool deleted = false;
+
+        internal IDictionary<string, EntityProperty> userProperties = new Dictionary<string, EntityProperty>();
+
+        public override void ReadEntity(IDictionary<string, EntityProperty> properties, OperationContext operationContext)
+        {
+            EntityProperty propDeleted;
+            // N.B. The cast will throw an exception if BooleanValue is null,
+            // but I think such properties are not supposed to be passed to ReadEntity.
+            deleted = properties.TryGetValue(KEY_DELETED, out propDeleted) ? (bool)propDeleted.BooleanValue : false;
+
+            EntityProperty propPartitionState;
+            // XXX Validate or use string?  (See BlobMetaDataRowTableEntity in Artifact Services for an example.)
+            partitionState = properties.TryGetValue(KEY_PARTITION_STATE, out propPartitionState)
+                ? (MTablePartitionState?)propPartitionState.Int32Value : null;
+
+            // If the rest of the code is prepared for DynamicTableEntity.ReadEntity to
+            // mutate the passed dictionary, we may as well do it too.
+            properties.Remove(KEY_DELETED);
+            properties.Remove(KEY_PARTITION_STATE);
+            userProperties = properties;
+        }
+
+        public override IDictionary<string, EntityProperty> WriteEntity(OperationContext operationContext)
+        {
+            // Partial aliasing, no worse than DynamicTableEntity.WriteEntity.
+            var properties = new Dictionary<string, EntityProperty>(userProperties);
+            // TODO: Optimization: don't output "deleted: false" except in InsertOrMerge case
+            // (to reduce the number of rows that have to be modified in post-migration cleanup).
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/MasterMigratingTable.cs

@@ -1,0 +1,289 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Threading.Tasks;
+using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using System.Net;
+using Microsoft.WindowsAzure.Storage.Table.Protocol;
+
+namespace Migration
+{
+    public class MasterMigratingTable : MigratingTable
+    {
+        public static MTableConfiguration INITIAL_CONFIGURATION =
+            new MTableConfiguration(TableClientState.USE_OLD_HIDE_METADATA);
+
+        new IConfigurationService<MTableConfiguration> configService;
+
+        public MasterMigratingTable(IConfigurationService<MTableConfiguration> configService,
+            IChainTable2 oldTable, IChainTable2 newTable,
+            MTableOptionalBug? bugToEnable)
+            : base(configService, oldTable, newTable, bugToEnable)
+        {
+            this.configService = configService;
+        }
+
+        public MasterMigratingTable(IConfigurationService<MTableConfiguration> configService,
+            IChainTable2 oldTable, IChainTable2 newTable, IChainTableMonitor monitor, MTableOptionalBug? bugToEnable)
+            : base(configService, oldTable, newTable, monitor, bugToEnable)
+        {
+            this.configService = configService;
+        }
+
+        Task CopyAsync(TableRequestOptions requestOptions, OperationContext operationContext)
+        {
+            string previousPartitionKey = null;
+            return WalkTableInParallel(oldTable, requestOptions, operationContext, async (oldEntity) =>
+            {
+                if (RowKeyIsInternal(oldEntity.RowKey))
+                    return null;
+
+                if (oldEntity.PartitionKey != previousPartitionKey)
+                {
+                    previousPartitionKey = oldEntity.PartitionKey;
+                    await EnsurePartitionSwitchedAsync(oldEntity.PartitionKey, requestOptions, operationContext);
+                    if (IsBugEnabled(MTableOptionalBug.InsertBehindMigrator))
+                    {
+                        Console.WriteLine("Copied Partition: {0}", previousPartitionKey);
+                        // More reasonable formulation of this bug.  If we're going
+                        // to allow CopyAsync while some clients are still writing
+                        // to the old table, give the developer credit for realizing
+                        // that the stream might be stale by the time the partition
+                        // is switched.
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/MigratingTable.cs

@@ -1,0 +1,1054 @@
+﻿using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections.Generic;
+using System.Threading.Tasks;
+using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using System.Linq;
+using Microsoft.WindowsAzure.Storage.Table.Protocol;
+using System.Net;
+using System.Diagnostics;
+using System.Runtime.Serialization;
+
+namespace Migration
+{
+    public enum MTableOptionalBug
+    {
+        // These were all made at some point during actual development.
+        // (I'm not 100% sure they're 100% equivalent to the original versions, but close enough.)
+        QueryAtomicFilterShadowing,
+        QueryStreamedFilterShadowing,
+        QueryStreamedLock,
+        QueryStreamedBackUpNewStream,
+        QueryStreamedSaveNewConfig,
+        DeleteNoLeaveTombstonesETag,
+        DeletePrimaryKey,
+        EnsurePartitionSwitchedFromPopulated,
+        TombstoneOutputETag,
+
+        // Some other "bugs", which a developer might not realistically make in this
+        // form but represent some ways we can break the protocol guarantees.
+        MigrateSkipPreferOld,
+        MigrateSkipUseNewWithTombstones,
+        InsertBehindMigrator,
+
+        NumBugs,
+    }
+
+    public enum TableClientState
+    {
+        // In order to push configuration without synchronizing all clients,
+        // clients in each pair of successive states must be able to coexist.
+        // Similarly, to transition from direct access to the old table to
+        // MigratingTable and ultimately to direct access to the new table,
+        // clients in the first (last) state must be able to coexist with
+        // clients directly accessing the old (respectively, new) table.
+
+        // Use oldTable only but strip MTable metadata from query results.
+        USE_OLD_HIDE_METADATA,
+
+        // Read the overlaid view.  Writes mark the partition as populated
+        // and then go to the old table unless the partition is always
+        // switched.  These clients can coexist with both USE_OLD_EXPORT
+        // and PREFER_NEW.
+        PREFER_OLD,
+
+        // Main state: read the composite view, and always ensure a
+        // partition is switched before writing so as not to insert data
+        // behind the migrator.  (Migrator is copying rows.)
+        PREFER_NEW,
+
+        // Use newTable only (now that all copying is done), but continue
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/PSharpUtils/MachineSynchronizationContext.cs

@@ -1,0 +1,50 @@
+﻿using System;
+using Microsoft.PSharp;
+using System.Threading;
+
+namespace Migration
+{
+    // Common interface for payloads so they can be handled by a single machine action.
+    // We could use a delegate, but it may be nice to introspect the fields.
+    interface IDispatchable
+    {
+        void Dispatch();
+    }
+
+    class SynchronizationContextWorkUnit : IDispatchable
+    {
+        readonly SendOrPostCallback callback;
+        readonly object arg;
+        internal SynchronizationContextWorkUnit(SendOrPostCallback callback, object arg)
+        {
+            this.callback = callback;
+            this.arg = arg;
+        }
+        public void Dispatch()
+        {
+            callback(arg);
+        }
+    }
+    // Hint: [OnEventDoAction(typeof(GenericDispatchableEvent), nameof(DispatchPayload))]
+    // "Generic" in the common sense, not C# "generic".
+    class GenericDispatchableEvent : Event { }
+    // Hint: Create only one per machine so that the TPL runs work units inline
+    // if the same synchronization context is already active.
+    class MachineSynchronizationContext : SynchronizationContext
+    {
+        readonly MachineId machineId;
+        public MachineSynchronizationContext(MachineId machineId)
+        {
+            this.machineId = machineId;
+        }
+        public override void Send(SendOrPostCallback d, object state)
+        {
+            // Not supported.  Apparently Send is only used for cancellation callbacks.
+            throw new NotSupportedException();
+        }
+        public override void Post(SendOrPostCallback d, object state)
+        {
+            PSharpRuntime.SendEvent(machineId, new GenericDispatchableEvent(), new SynchronizationContextWorkUnit(d, state));
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/PSharpUtils/PSharpNondeterminism.cs

@@ -1,0 +1,25 @@
+﻿using Microsoft.PSharp;
+using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using System.Threading.Tasks;
+
+namespace Migration
+{
+    public static class PSharpNondeterminism
+    {
+        public static int Choice(int numChoices)
+        {
+            int num = 0;
+            int denom = 1;
+            while (denom < numChoices)
+            {
+                denom *= 2;
+                num = 2 * num + (PSharpRuntime.Nondeterministic() ? 1 : 0);
+            }
+            // XXX unbalanced
+            return (num >= numChoices) ? num - numChoices : num;
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/PSharpUtils/PSharpProxy.cs

@@ -1,0 +1,160 @@
+﻿using System;
+using System.Threading.Tasks;
+using Microsoft.PSharp;
+using System.Runtime.Remoting.Proxies;
+using System.Runtime.Remoting.Messaging;
+using System.Reflection;
+using System.Diagnostics;
+using System.Linq;
+
+namespace Migration
+{
+    public delegate Event EventFactory();
+
+    /*
+     * PSharpProxy proxies all async methods of an interface so that they
+     * execute on the event loop of the host machine.  Replies are sent back to
+     * the caller machine.
+     *
+     * Interfaces that take or return further shared objects (as opposed to
+     * conceptually transferring ownership) will need custom proxies that wrap
+     * those objects in proxies.  It seems to be easiest to wrap the PSharpProxy
+     * transparent proxy in another object and write the methods that don't need
+     * special handling as boilerplate passthroughs.  It looks like trying to
+     * avoid that boilerplate by hooking Invoke to implement the special
+     * handling in terms of the reflected call, or generating a subclass of the
+     * transparent proxy class with some methods overridden, is going to be more
+     * trouble than it's worth.
+     */
+    // http://stackoverflow.com/questions/15733900/dynamically-creating-a-proxy-class
+    // Or (more authoritative?): https://msdn.microsoft.com/en-us/library/System.Runtime.Remoting.Messaging.IMethodMessage(v=vs.110).aspx
+    class PSharpRealProxy : RealProxy
+    {
+        internal readonly MachineId callerMachineId;
+        internal readonly MachineId hostMachineId;
+        internal readonly object target;
+        internal readonly string targetDebugName;
+        // The anticipated use case is to control the event type (for selective
+        // deferral, etc.) since we'll always overwrite the payload.
+        readonly EventFactory eventFactory;
+
+        static readonly Type typeofTaskGeneric = typeof(Task<object>).GetGenericTypeDefinition();
+
+        // Important: TIface must be an interface (or subclass of MarshalByRefObject).
+        // This is easy to mess up if you allow it to be inferred.
+        public static TIface MakeTransparentProxy<TIface>(MachineId callerMachineId, MachineId hostMachineId,
+            TIface target, string targetDebugName, EventFactory eventFactory)
+        {
+            // XXX: Verify up front that TIface is an interface with all methods acceptable?
+            var realProxy = new PSharpRealProxy(typeof(TIface), callerMachineId, hostMachineId, target, targetDebugName, eventFactory);
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/PSharpUtils/Replies.cs

@@ -1,0 +1,43 @@
+﻿using System;
+using System.Threading.Tasks;
+using Microsoft.PSharp;
+
+namespace Migration
+{
+    class ReplyTarget<TResult>
+    {
+        readonly object replyId;
+        readonly MachineId machineId;
+        internal readonly TaskCompletionSource<TResult> tcs;
+        bool replied = false;
+        public ReplyTarget(object replyId, MachineId machineId)
+        {
+            this.replyId = replyId;
+            this.machineId = machineId;
+            tcs = new TaskCompletionSource<TResult>(TaskCreationOptions.RunContinuationsAsynchronously);
+        }
+        public void SetOutcome(Outcome<TResult, Exception> outcome)
+        {
+            if (replied) throw new InvalidOperationException();
+            PSharpRuntime.SendEvent(machineId, new GenericDispatchableEvent(), new ReplyPayload<TResult>(tcs, outcome));
+            replied = true;
+        }
+    }
+
+    class ReplyPayload<TResult> : IDispatchable
+    {
+        readonly TaskCompletionSource<TResult> tcs;
+        readonly Outcome<TResult, Exception> outcome;
+
+        internal ReplyPayload(TaskCompletionSource<TResult> tcs, Outcome<TResult, Exception> outcome)
+        {
+            this.tcs = tcs;
+            this.outcome = outcome;
+        }
+
+        public void Dispatch()
+        {
+            tcs.SetOutcome(outcome);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Utils/Collections.cs

@@ -1,0 +1,60 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Linq;
+
+namespace Migration
+{
+    public static class CollectionStatics
+    {
+
+        /* Learning a few C# concepts!  I'd like an analogue of Java's Map.get.  This only works if TValue
+           is nullable (i.e., TValue : class or T? itself as an edge case).  If TValue is not nullable,
+           (i.e., TValue : struct), I want the return type to be TValue? instead.  These two forms can't be
+           overloaded without a hack:
+           http://stackoverflow.com/questions/7764911/generic-contraints-on-method-overloads
+           If we use different names, then from one POV, we may as well generalize the return-TValue case
+           to GetValueOrDefault with no constraint.  I personally don't think the default value of a
+           non-nullable type is useful, but others might.
+           
+           Note, "TValue : class" _can_ overload with "TValue = T?".  It's reasonable for these to be
+           different overloads. */
+        public static TValue GetValueOrDefault<TKey, TValue>(this IReadOnlyDictionary<TKey, TValue> dict, TKey key)
+        {
+            TValue value;
+            dict.TryGetValue(key, out value);
+            return value;
+        }
+        public static TValue? GetValueNullable<TKey, TValue>(this IReadOnlyDictionary<TKey, TValue> dict, TKey key) where TValue : struct
+        {
+            TValue value;
+            if (dict.TryGetValue(key, out value))
+                return value;
+            else
+                return null;
+        }
+        public static TValue GetValueOr<TKey, TValue>(this IReadOnlyDictionary<TKey, TValue> dict, TKey key, TValue defaultValue)
+        {
+            TValue value;
+            if (dict.TryGetValue(key, out value))
+                return value;
+            else
+                return defaultValue;
+        }
+        // IDictionary does not extend IReadOnlyDictionary, so if we want to
+        // support IDictionary (e.g., as used by ITableEntity) without making
+        // Dictionary ambiguous, we would need two more overloads.  Ridiculous?
+
+        // XXX: This algorithm is slow.
+        public static IEnumerable<T> IntersectAll<T>(this IEnumerable<IEnumerable<T>> sets)
+        {
+            IEnumerable<T> partialIntersection = null;
+            foreach (IEnumerable<T> set in sets)
+            {
+                partialIntersection = (partialIntersection == null) ? set : partialIntersection.Intersect(set);
+            }
+            if (partialIntersection == null)
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Utils/Comparer.cs

@@ -1,0 +1,243 @@
+﻿using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using Microsoft.WindowsAzure.Storage.Table;
+using System;
+using System.Collections;
+using System.Collections.Generic;
+using System.Collections.Immutable;
+using System.Diagnostics.Contracts;
+using System.Linq;
+using System.Reflection;
+using System.Threading;
+using System.Threading.Tasks;
+
+namespace Migration
+{
+    /* Equality comparer that compares the content of some objects we care
+     * about that don't override Equals (and GetHashCode) themselves.
+     *
+     * This is only guaranteed to be right for the objects we use.  While we
+     * try to be targeted in the additional comparisons we define, it's hard
+     * to rule out the existence of subclasses for which the comparisons here
+     * might be inappropriate.
+     *
+     * In general, we compare collections without regard to their type
+     * arguments, as in Java.  Ideally, one would use the fastest
+     * EqualityComparer for the actual type arguments, which would require
+     * some fancy type manipulation, but this isn't worth worrying about for
+     * our purposes. */
+    public class BetterComparer : EqualityComparer<object>
+    {
+        private BetterComparer() { }
+        public static readonly BetterComparer Instance = new BetterComparer();
+
+        public override bool Equals(object x, object y)
+        {
+            // If we had a lot of these cases, we could make some framework
+            // to fixpoint a combination of EqualityComparers.
+            IOutcome<object, Exception> xOutcome, yOutcome;
+            if ((xOutcome = x as IOutcome<object, Exception>) != null
+                && (yOutcome = y as IOutcome<object, Exception>) != null)
+            {
+                return Equals(xOutcome.Result, yOutcome.Result)
+                    && Equals(xOutcome.Exception, yOutcome.Exception);
+            }
+            StorageException xSE, ySE;
+            if ((xSE = x as StorageException) != null
+                && (ySE = y as StorageException) != null)
+            {
+                // This is probably the first decision in BetterComparer that is
+                // definitely not general-purpose.  This will have the effect of
+                // comparing status codes and FailedOpIndex (if applicable) for
+                // StorageExceptions that are part of the semantics.
+                return Equals(xSE.Message, ySE.Message);
+            }
+            ITableEntity xEntity, yEntity;
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Utils/Hasher.cs

@@ -1,0 +1,31 @@
+﻿namespace Migration
+{
+    // We want to make this a struct, so we have to live with people being
+    // able to create default instances.
+    public struct Hasher
+    {
+        readonly int x;
+        public Hasher(int x)
+        {
+            this.x = x;
+        }
+        public Hasher With(int y)
+        {
+            unchecked
+            {
+                return new Hasher(31 * x + y);
+            }
+        }
+        // With(object o)?  But the caller might want to use a custom EqualityComparer.
+        // I'd rather pay the boilerplate in the caller than try to deal with that here.
+        public static implicit operator int (Hasher h)
+        {
+            // If we cared about good distribution (e.g., different hashes
+            // for dictionaries that differ by a permutation of the values),
+            // we'd apply some function to h.x here.
+            return h.x;
+        }
+
+        public static readonly Hasher Start = new Hasher(17);
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Utils/Outcome.cs

@@ -1,0 +1,80 @@
+﻿using System;
+using System.Threading.Tasks;
+
+namespace Migration
+{
+    // Default Outcome represents default TResult: reasonable.
+    // I could define good Equals, GetHashCode, ToString here, but they
+    // would have to call the "better" ones on the fields (or accept
+    // implementations for the fields as parameters), which would be weird.
+    public struct Outcome<TResult, TException> : IOutcome<TResult, TException>
+        where TException : Exception
+    {
+        private readonly TResult result;
+        private readonly TException exception;
+
+        public Outcome(TResult result)
+        {
+            this.result = result;
+            this.exception = null;
+        }
+        public Outcome(TException exception)
+        {
+            this.result = default(TResult);
+            this.exception = exception;
+        }
+
+        public TResult Result { get { return result; } }
+        public TException Exception { get { return exception; } }
+    }
+
+    // For introspection by BetterComparator, etc. without wildcard capture.
+    public interface IOutcome<out TResult, out TException> where TException : Exception
+    {
+        TResult Result { get; }
+        TException Exception { get; }
+    }
+
+    public static class Catching<TException> where TException : Exception
+    {
+        public static Outcome<TResult, TException> Run<TResult>(Func<TResult> func)
+        {
+            TResult result;
+            try
+            {
+                result = func();
+            }
+            catch (TException ex)
+            {
+                return new Outcome<TResult, TException>(ex);
+            }
+            return new Outcome<TResult, TException>(result);
+        }
+        public static async Task<Outcome<TResult, TException>> Task<TResult>(Task<TResult> task)
+        {
+            TResult result;
+            try
+            {
+                result = await task;
+            }
+            catch (TException ex)
+            {
+                return new Outcome<TResult, TException>(ex);
+            }
+            return new Outcome<TResult, TException>(result);
+        }
+    }
+
+    public static class OutcomeStatics
+    {
+        public static void SetOutcome<TResult, TException>(
+            this TaskCompletionSource<TResult> tcs, Outcome<TResult, TException> outcome)
+            where TException : Exception
+        {
+            if (outcome.Exception != null)
+                tcs.SetException(outcome.Exception);
+            else
+                tcs.SetResult(outcome.Result);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Utils/SynchronizationContextUse.cs

@@ -1,0 +1,37 @@
+﻿using System;
+using System.Threading;
+
+namespace Migration
+{
+    class SynchronizationContextUse : IDisposable
+    {
+        SynchronizationContext sc, previous;
+        internal SynchronizationContextUse(SynchronizationContext sc)
+        {
+            if (sc == null) throw new ArgumentNullException("sc");
+            this.sc = sc;
+            previous = SynchronizationContext.Current;
+            SynchronizationContext.SetSynchronizationContext(sc);
+        }
+
+        public void Dispose()
+        {
+            if (sc != null)  // detect redundant calls
+            {
+                if (SynchronizationContext.Current != sc)
+                    throw new InvalidOperationException("The SynchronizationContext we are trying to deactivate is no longer current.");
+                SynchronizationContext.SetSynchronizationContext(previous);
+                sc = previous = null;
+            }
+        }
+    }
+
+    public static class SynchronizationContextUseStatics
+    {
+        // Intended for use in a using statement.
+        public static IDisposable AsCurrent(this SynchronizationContext sc)
+        {
+            return new SynchronizationContextUse(sc);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/Migration/Verification.cs

@@ -1,0 +1,950 @@
+﻿using System;
+using System.Threading.Tasks;
+using Microsoft.PSharp;
+using System.Collections.Generic;
+using Microsoft.WindowsAzure.Storage.Table;
+using Microsoft.WindowsAzure.Storage.Table.Protocol;
+using Microsoft.WindowsAzure.Storage;
+using Microsoft.WindowsAzure.Storage.ChainTableInterface;
+using System.Collections.Immutable;
+using System.Diagnostics;
+using System.Linq;
+
+namespace Migration
+{
+    interface IMirrorChainTable2 : IChainTable2
+    {
+        /*
+         * Like ExecuteBatch but sets the new ETags the same as in
+         * originalResponse (assuming the mirror batch succeeds).  If
+         * originalResponse is null, equivalent to ExecuteBatchAsync.
+         *
+         * originalBatch must be a copy of the original batch with the
+         * _original_ ETag (i.e., If-Match) fields for correct processing.
+         */
+        Task<IList<TableResult>> ExecuteMirrorBatchAsync(
+            TableBatchOperation originalBatch, IList<TableResult> originalResponse,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+    }
+    abstract class AbstractMirrorChainTable2 : AbstractChainTable2, IMirrorChainTable2
+    {
+        public abstract Task<IList<TableResult>> ExecuteMirrorBatchAsync(
+            TableBatchOperation originalBatch, IList<TableResult> originalResponse,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+
+        public override Task<IList<TableResult>> ExecuteBatchAsync(TableBatchOperation batch,
+            TableRequestOptions requestOptions = null, OperationContext operationContext = null)
+        {
+            return ExecuteMirrorBatchAsync(batch, null, requestOptions, operationContext);
+        }
+    }
+
+    // This is expected to be a single call to one of the ITable methods.
+    delegate Task<object> TableCall(IChainTable2 table);
+    delegate Task<object> MirrorTableCall(IMirrorChainTable2 table);
+
+    static class MigrationModel
+    {
+        // Start with a single partition per John's suggestion.
+        internal static readonly string SINGLE_PARTITION_KEY = "";
+
+        internal static readonly int NUM_SERVICE_MACHINES = 2;
+        internal static readonly int NUM_CALLS_PER_MACHINE = 2;
+
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/CloudBlobHelpers.cs

@@ -1,0 +1,131 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Table;
+    using Microsoft.WindowsAzure.Storage.Blob;
+
+    public class CloudBlobHelpers  // Used by ShardingAzureCloudTableClientFactory for demo
+    {
+        public static CloudBlockBlob GetBlockBlob(string configurationStorageConnectionString, string configurationLocation)
+        {
+            CloudStorageAccount blobStorageAccount = CloudStorageAccount.Parse(configurationStorageConnectionString);
+            CloudBlobClient blobClient = blobStorageAccount.CreateCloudBlobClient();
+
+            CloudBlobContainer container = blobClient.GetContainerReference(GetContainerName(configurationLocation));
+            if (container.CreateIfNotExists())
+            {
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ConfigurationStore.cs

@@ -1,0 +1,46 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Runtime.Serialization;
+
+    [DataContract(Namespace = "http://schemas.microsoft.com/windowsazure")]
+    public class ConfigurationStore
+    {
+        public ConfigurationStore()
+        {
+            this.LeaseDuration = Constants.LeaseDurationInSec;
+            this.Timestamp = DateTime.UtcNow;
+            this.ViewId = 1; // minimum ViewId is 1.
+        }
+
+        [DataMember(IsRequired = true)]
+        public long ViewId { get; set; }
+
+        [DataMember(IsRequired = true)]
+        public int LeaseDuration { get; set; }
+
+        [DataMember(IsRequired = true)]
+        public DateTime Timestamp { get; set; }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ConfigurationStoreLocationInfo.cs

@@ -1,0 +1,62 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Runtime.Serialization;
+
+    [DataContract(Namespace = "http://schemas.microsoft.com/windowsazure")]
+    public class ConfigurationStoreLocationInfo
+    {
+        [DataMember(IsRequired = true)]
+        public string StorageAccountName { get; set; }
+
+        //Eventually, need a way to NOT pass this in here
+        [DataMember(IsRequired = true)]
+        public string StorageAccountKey { get; set; }
+
+        [DataMember(IsRequired = true)]
+        public string BlobPath { get; set; }
+
+        public ConfigurationStoreLocationInfo()
+        {
+            this.StorageAccountName = string.Empty;
+            this.StorageAccountKey = string.Empty;
+        }
+
+        /// <summary>
+        /// 
+        /// </summary>
+        /// <param name="storageAccountConnectionString"></param>
+        /// <exception cref="IndexOutOfRangeException"></exception>
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/Constants.cs

@@ -1,0 +1,56 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    public class Constants
+    {
+        /// <summary>
+        /// The duration for which the lease on the current configuration is valid. 
+        /// </summary>
+        public static int LeaseDurationInSec = 60;
+
+        /// <summary>
+        /// The clock factor takes into account that two machine's clock might count up or down at different speeds. This is also
+        /// known as time dilation. This is an upper bound on time delay introduced due to the clock factor
+        /// </summary>
+        public static int ClockFactorInSec = 5;
+
+        public static int LeaseRenewalExpirationWatermark = 5;
+
+        /// <summary>
+        /// When LeaseRenewalIntervalInSec has expired since we last renewed the lease, a new renewal is needed
+        /// </summary>
+        public static int LeaseRenewalIntervalInSec = LeaseDurationInSec - LeaseRenewalExpirationWatermark;
+
+        public static string ConfigurationStoreUpdatingText =
+            "The configuration store is in the process of being updated...";
+
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/DynamicReplicatedTableEntity.cs

@@ -1,0 +1,225 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Collections.Generic;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    /// <summary>
+    /// InitDynamicReplicatedTableEntity is the same as DynamicReplicatedTableEntity except KeyNotFoundException is caught in ReadEntity().
+    /// InitDynamicReplicatedTableEntity is used instead of DynamicReplicatedTableEntity when ConvertXStoreTableMode = true. This 
+    /// is used to aid in transitioning an existing Azure Table to ReplicatedTable
+    /// </summary>
+    public class InitDynamicReplicatedTableEntity : DynamicReplicatedTableEntity
+    {
+        public InitDynamicReplicatedTableEntity()
+            : base()
+        {
+        }
+
+        public InitDynamicReplicatedTableEntity(string partitionKey, string rowKey)
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/IReplicatedTable.cs

@@ -1,0 +1,49 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System.Collections.Generic;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    public interface IReplicatedTable
+    {
+        string TableName { get; }
+        bool CreateIfNotExists(TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+        bool Exists();
+        bool DeleteIfExists();
+        TableResult Execute(TableOperation operation, TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+        IList<TableResult> CheckRetrieveInBatch(TableBatchOperation batch, TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+        IList<TableResult> ExecuteBatch(TableBatchOperation batch, TableRequestOptions requestOptions = null, OperationContext operationContext = null);
+        TableResult Retrieve(TableOperation operation, TableRequestOptions requestOptions = null, OperationContext operationContext = null);
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/IReplicatedTableEntity.cs

@@ -1,0 +1,55 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    //this is the class that users extend to store their own data in a row
+    public interface IReplicatedTableEntity : ITableEntity
+    {
+        // lock bit: used to detect that replication is in progress
+        bool _rtable_RowLock { get; set; }
+
+        // version: used as (virtual) etag for replicated table with different (physical) etags  
+        long _rtable_Version { get; set; }
+
+        // tomb stone: to support deletes
+        bool _rtable_Tombstone { get; set; }
+
+        // view number: to support changes in replica configurations
+        long _rtable_ViewId { get; set; }
+
+        // _rtable_Operation type: last operation type that modified this data
+        string _rtable_Operation { get; set; }
+
+        // Batch id: to support batch operations
+        Guid _rtable_BatchId { get; set; }
+
+        /// <summary>
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/JsonStore.cs

@@ -1,0 +1,120 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Globalization;
+    using System.IO;
+    using System.Runtime.Serialization;
+    using System.Runtime.Serialization.Json;
+    using System.Text;
+
+    public class JsonStore<T> where T : class  // public for demo
+    {
+        /// <summary>
+        /// JSON Serialization
+        /// </summary>
+        public static string Serialize(T t)
+        {
+            string jsonString = string.Empty;
+            MemoryStream memoryStream = new MemoryStream();
+
+            try
+            {
+                DataContractJsonSerializer serializer = new DataContractJsonSerializer(typeof(T));
+                serializer.WriteObject(memoryStream, t);
+                jsonString = Encoding.UTF8.GetString(memoryStream.ToArray());
+            }
+            catch (Exception e)
+            {
+                string message = String.Format(CultureInfo.InvariantCulture, "Could not perform Json Serialization on {0}", t);
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/PeriodicTimer.cs

@@ -1,0 +1,88 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Threading;
+    
+    /// <summary>
+    /// http://msdn.microsoft.com/en-us/library/ah1h85ch.aspx for instantiating Timer Class
+    /// </summary>
+    class PeriodicTimer
+    {
+        private Timer timer;
+
+        public TimerCallback Callback { get; private set; }
+        public Object State { get; private set; }
+        public TimeSpan DueTime { get; private set; }
+        public TimeSpan Period { get; private set; }
+
+        /// <summary>
+        /// Initializes a new instance of the <see cref="PeriodicTimer" /> class.
+        /// </summary>
+        /// <param name="callback">The callback method.</param>
+        /// <param name="period">Periodicity of the task.</param>
+        public PeriodicTimer(TimerCallback callback, TimeSpan period) : this(callback, null, TimeSpan.Zero, period) { }
+
+        /// <summary>
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReconfigurationStatus.cs

@@ -1,0 +1,35 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    // Enum for status codes returned during recovery/reconfiguration
+    public enum ReconfigurationStatus
+    {
+        SUCCESS = 0,
+        PARTIAL_FAILURE = 1,
+        LOCK_FAILURE = 2,
+        UNLOCK_FAILURE = 4,
+        FAULTY_WRITE_VIEW = 8,
+        TABLE_NOT_FOUND = 16,
+        FAILURE = 32
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicaInfo.cs

@@ -1,0 +1,52 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System.Runtime.Serialization;
+    
+    [DataContract(Namespace = "http://schemas.microsoft.com/windowsazure")]
+    public class ReplicaInfo
+    {
+        [DataMember(IsRequired = true)]
+        public string StorageAccountName { get; set; }
+
+        //Eventually, need a way to NOT pass this in here
+        [DataMember(IsRequired = true)]
+        public string StorageAccountKey { get; set; }
+
+        [DataMember(IsRequired = true)]
+        public long ViewInWhichAddedToChain { get; set; }
+
+        public ReplicaInfo()
+        {
+            this.ViewInWhichAddedToChain = 1;
+        }
+
+        public override string ToString()
+        {
+            return string.Format("Account Name: {0}, AccountKey: {1}, ViewInWhichAddedToChain: {2}", 
+                this.StorageAccountName, 
+                "***********", 
+                this.ViewInWhichAddedToChain);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTable.cs

@@ -1,0 +1,2274 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Collections.Generic;
+    using System.Data.Services.Client;
+    using System.Linq;
+    using System.Net;
+    using System.Reflection;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    public class ReplicatedTable : IReplicatedTable
+    {
+        private ReplicatedTableConfigurationService _replicatedTableConfigurationService;
+        private string myName;
+
+        public string TableName
+        {
+            get
+            {
+                return myName;
+            }
+            private set
+            {
+                myName = value;
+            }
+        }
+
+        // Current view
+        private View CurrentView
+        {
+            get { return _replicatedTableConfigurationService.GetWriteView(); }
+        }
+
+        // Used for read optimization to read from random replica
+        private static Random random = new Random();
+
+        // 2PC protocol constants
+        private const int PREPARE_PHASE = 1;
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableConfigurationService.cs

@@ -1,0 +1,340 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Collections.Generic;
+    using System.Threading;
+    using System.Threading.Tasks;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Blob;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    public class ReplicatedTableConfigurationService : IDisposable
+    {
+        private List<ConfigurationStoreLocationInfo> blobLocations;
+        private Dictionary<string, CloudBlockBlob> blobs = new Dictionary<string, CloudBlockBlob>();
+
+        private bool useHttps;
+
+        private int quorumSize = 0;
+        private string cloudStorageAccountTemplate = "DefaultEndpointsProtocol={0};AccountName={1};AccountKey={2}";
+
+        private DateTime lastViewRefreshTime;
+        private View lastRenewedReadView;
+        private View lastRenewedWriteView;
+
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableConfigurationStore.cs

@@ -1,0 +1,46 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System.Collections.Generic;
+    using System.Runtime.Serialization;
+
+    [DataContract(Namespace = "http://schemas.microsoft.com/windowsazure")]
+    public class ReplicatedTableConfigurationStore : ConfigurationStore
+    {
+        public ReplicatedTableConfigurationStore()
+            : base()
+        {
+            this.ReplicaChain = new List<ReplicaInfo>();
+        }
+
+        [DataMember(IsRequired = true)]
+        public List<ReplicaInfo> ReplicaChain { get; set; }
+
+        [DataMember(IsRequired = true)]
+        public int ReadViewHeadIndex { get; set; }
+
+        [DataMember(IsRequired = false)]
+        public bool ConvertXStoreTableMode { get; set; }
+
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableConflictException.cs

@@ -1,0 +1,43 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using Microsoft.WindowsAzure.Storage;
+
+    public class ReplicatedTableConflictException : StorageException
+    {
+        public ReplicatedTableConflictException()
+        {
+        }
+
+        public ReplicatedTableConflictException(string message)
+            : base(message)
+        {
+        }
+
+        public ReplicatedTableConflictException(string message, Exception inner)
+            : base(message, inner)
+        {
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableEntity.cs

@@ -1,0 +1,146 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Collections.Generic;
+    using Microsoft.WindowsAzure.Storage;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    //this is the class that users extend to store their own data in a row
+    public class ReplicatedTableEntity : IReplicatedTableEntity
+    {
+
+        /// Gets or sets the properties in the table entity, indexed by property name.
+        /// </summary>
+        /// <value>The entity properties.</value>
+        public IDictionary<string, EntityProperty> Properties { get; set; }
+
+        /// <summary>
+        /// Gets or sets the entity's partition key.
+        /// </summary>
+        /// <value>The entity partition key.</value>
+        public string PartitionKey { get; set; }
+
+        /// <summary>
+        /// Gets or sets the entity's row key.
+        /// </summary>
+        /// <value>The entity row key.</value>
+        public string RowKey { get; set; }
+
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableEventSource.cs

@@ -1,0 +1,57 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System.Diagnostics.Tracing;
+
+    [EventSource(Name = Constants.LogSourceName)]
+    internal class ReplicatedTableEventSource : EventSource
+    {
+        internal ReplicatedTableEventSource()
+        {
+        }
+
+        [Event(1, Level = EventLevel.Error)]
+        internal void Error(string message)
+        {
+            this.WriteEvent(1, message);
+        }
+
+        [Event(2, Level = EventLevel.Warning)]
+        internal void Warning(string message)
+        {
+            this.WriteEvent(2, message);
+        }
+
+        [Event(3, Level = EventLevel.Informational)]
+        internal void Informational(string message)
+        {
+            this.WriteEvent(3, message);
+        }
+
+        [Event(4, Level = EventLevel.Verbose)]
+        internal void Verbose(string message)
+        {
+            this.WriteEvent(4, message);
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableLogger.cs

@@ -1,0 +1,60 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    public static class ReplicatedTableLogger
+    {
+        private static ReplicatedTableEventSource eventSource = new ReplicatedTableEventSource();
+
+        public static void LogError(string format, params object[] args)
+        {
+            if (ReplicatedTableLogger.eventSource.IsEnabled())
+            {
+                ReplicatedTableLogger.eventSource.Error(string.Format(format, args));
+            }
+        }
+
+        public static void LogWarning(string format, params object[] args)
+        {
+            if (ReplicatedTableLogger.eventSource.IsEnabled())
+            {
+                ReplicatedTableLogger.eventSource.Warning(string.Format(format, args));
+            }
+        }
+
+        public static void LogInformational(string format, params object[] args)
+        {
+            if (ReplicatedTableLogger.eventSource.IsEnabled())
+            {
+                ReplicatedTableLogger.eventSource.Informational(string.Format(format, args));
+            }
+        }
+
[…]

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/ReplicatedTableStaleViewException.cs

@@ -1,0 +1,43 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using Microsoft.WindowsAzure.Storage;
+
+    public class ReplicatedTableStaleViewException : StorageException
+    {
+        public ReplicatedTableStaleViewException()
+        {
+        }
+
+        public ReplicatedTableStaleViewException(string message)
+            : base(message)
+        {
+        }
+
+        public ReplicatedTableStaleViewException(string message, Exception inner)
+            : base(message, inner)
+        {
+        }
+    }
+}

### 2015-09-20 migration/ArtifactServicesStress/CodeCoverage/CodeSnapshot/library/View.cs

@@ -1,0 +1,129 @@
+﻿// azure-rtable ver. 0.9
+//
+// Copyright (c) Microsoft Corporation
+//
+// All rights reserved. 
+//
+// MIT License
+//
+// Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files 
+// (the ""Software""), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, 
+// merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished 
+// to do so, subject to the following conditions:
+//
+// The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
+//
+// THE SOFTWARE IS PROVIDED *AS IS*, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF 
+// MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE 
+// FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+// WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+
+
+namespace Microsoft.Azure.Toolkit.Replication
+{
+    using System;
+    using System.Collections.Generic;
+    using Microsoft.WindowsAzure.Storage.Table;
+
+    public class View
+    {
+        public View()
+        {
+            this.Chain = new List<Tuple<ReplicaInfo, CloudTableClient>>();
+        }
+
+        public long ViewId { get; set; }
+
+        public List<Tuple<ReplicaInfo, CloudTableClient>> Chain;
+
+        public CloudTableClient this[int index]
+        {
+            get
+            {
+                return this.Chain[index].Item2;
+            }
+        }
+
+        public ReplicaInfo GetReplicaInfo(int index)
+        {
+            return this.Chain[index].Item1;
+        }
+
+        public int TailIndex
+        {
+            get { return this.Chain.Count - 1; }
+        }
+
+        public int ReadHeadIndex
+        {
+            get; set;
+        }
+
+        public int WriteHeadIndex
+        {
[…]

### 2015-09-20 intro.tex
some updates in intro and new citations

@@
-Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}.
-This challenge is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside components, unexpected node failures, data losses due to unreliable communication channels, and interaction with (human) clients.
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, leesatapornwongsa2014samc, laguna2015debugging, maddox2015test}.
+This challenge is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos, henry2009cloud}, such as unexpected node failures, the asynchronous interaction between system components, data losses due to unreliable communication channels, the use of multithreaded code to exploit multicore machines, and interaction with clients.
 All these sources of nondeterminism can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, corner-case bugs that are difficult to detect, diagnose and fix.
-A Heisenbug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic~\cite{amazon2012aws, google2014outage}.
+A Heisenbug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequences can be catastrophic~\cite{amazon2012aws, google2014outage}.
 
@@
 The gist of their approach is to extract the high-level logic from a production system, represent this logic as specifications in the expressive TLA+~\cite{lamport1994temporal} language, and finally verify the specifications using a model checker.
-While highly effective, as demonstrated by its use in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'', and the AWS team admits that ``we are not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
+While highly effective, as demonstrated by its use in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'', and the AWS team admits that it is ``not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
 

### 2015-09-20 motivation.tex
some updates in motivation

@@
 Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from 10s of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
+It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
 
 Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
-Similar to the current system, vNext employs containers, called {\em extents}, to store data.
-Extents are typically several gigabytes each, consisting of many data blocks, and replicated over multiple {\em Extent Nodes} (ENs).
-However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a completely \emph{distributed mapping}.
-In vNext, extents are divided into partitions, with each partition managed by a light-weight {\em Extent Manager} (ExtMgr).
+Similar to the current system, vNext employs containers, called \emph{extents}, to store data.
+Extents are typically several gigabytes each, consisting of many data blocks, and replicated over multiple \emph{Extent Nodes} (ENs).
+However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a \emph{distributed mapping}.
+In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr).
 This partitioning is illustrated in Figure~\ref{fig:vnext}.
@@
 The failure of an EN is detected by missing heartbeats.
-An ExtMgr also receives less frequent, but still periodic, {\em synchronization reports} from every EN.
+An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN.
 The sync reports list all the extents (and associated metadata) stored on the EN.
@@
 \noindent
-Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in the ExtMgr logic.
-All the unit test and integration test suites successfully pass every single time.
-However, the stress test suite fails {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired.
-The bug has proven to be difficult to identify, reproduce and troubleshoot.
-First, it takes very long executions to trigger.
-Second, an extent not being repaired is {\em not} a property that can be easily verified.
-In practice, the developers rely on very large time-out period to detect the bug.
+Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic.
+All the unit test suites and integration test suites successfully pass every single time.
+However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.
+This bug has proven to be difficult to identify, reproduce and troubleshoot.
+First, it requires very long executions to trigger.
+Second, an extent not being repaired is \emph{not} a property that can be easily verified.
+In practice, the developers rely on a very large time-out period to detect the bug.
 Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.

### 2015-09-21 overview.tex
updated title and some minor stuff

@@
 A \psharp program consists of multiple \emph{state machines} that communicate with each other \emph{asynchronously} by sending and receiving \emph{events}, which may contain an optional \emph{payload}.
-A \psharp machine is similar to a class in \csharp: it can contain any number of fields and methods, but it also contains an input event queue, and one or more states.
+A \psharp machine declaration is similar to a class declaration in \csharp: it can contain an arbitrary number of fields and methods.
+However, a machine declaration differs from a class declaration, as it contains an input event queue, and one or more \emph{states}.
 Each state can register \emph{actions} to handle incoming events.
-\psharp machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking the registered action.
+\psharp machine instances run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking the registered action.
 This action might access a field, call a method, transition the machine to a new state, create a new machine, or send an event to another machine.

### 2015-09-21 overview.tex
various fixes

--- 2015-09-20 overview.tex
@@
 A \psharp program consists of multiple \emph{state machines} that communicate with each other \emph{asynchronously} by sending and receiving \emph{events}, which may contain an optional \emph{payload}.
-A \psharp machine is similar to a class in \csharp: it can contain any number of fields and methods, but it also contains an input event queue, and one or more states.
+A \psharp machine declaration is similar to a class declaration in \csharp: it can contain an arbitrary number of fields and methods.
+However, a machine declaration differs from a class declaration, as it also contains an input event queue, and one or more \emph{states}.
 Each state can register \emph{actions} to handle incoming events.
+
 \psharp machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking the registered action.
 This action might access a field, call a method, transition the machine to a new state, create a new machine, or send an event to another machine.
-In \psharp, a send operation is non-blocking; the event is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.
+In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the input queue of the target machine, which will dequeue and handle the event concurrently.
 All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
@@
 Our approach of using \psharp to test distributed systems, implemented in .NET, requires the developer to perform the following three key modeling tasks.
-These tasks are illustrated in \S\ref{sec:method} using the Azure Storage vNext as a running example.
+These tasks are illustrated in \S\ref{sec:method} for the Azure Storage vNext system.
 
@@
 
-A hot state denotes a point in the execution where progress is required but has not happened yet, for example a node has failed but a new one has not come up yet.
+A hot state denotes a point in the execution where progress is required but has not happened yet; for example, a node has failed but a new one has not come up yet.
 A liveness monitor enters a hot state upon receiving notification of an event that requires the system to make progress.
@@
 
-\textbf{Using monitors}.
-To declare a safety or a liveness monitor, the programmer must inherit from the \psharp \texttt{Monitor} class.
+%\textbf{Using monitors}.
+%To declare a safety or a liveness monitor, the programmer must inherit from the \psharp \texttt{Monitor} class.
 \psharp monitors are singleton instances and, thus, an ordinary machine does not need a reference to a monitor to send it an event.
@@
 
-The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
-This testing process is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.
+The \psharp testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules.
 After a bug is found, the engine generates a trace that represents the buggy schedule.
-In contrast to logs typically generating during production, this log provides a global order of all communication events and, thus, is easier to debug.
+In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events and, thus, is easier to debug.
 
-We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
 It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.
-The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.
+The interface exposes callbacks that are invoked by the \psharp testing engine for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.
 
@@
 In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing.
-However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus unlikely to scale for large code bases.
+However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus infeasible for large code bases.
 To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.

### 2015-09-21 overview.tex
updates to address some of Cheng's comments

@@
 \begin{enumerate}
-\item
-The computation model underlying \psharp is communicating state machines.
-The environment of the system-under-test must be modeled using the \psharp APIs, while the real components of the system must be wrapped inside \psharp machines.
-We call this modeled environment the \psharp test harness.
+\item The real components of a distributed system must be wrapped inside \psharp machines.
+In addition, the environment of the system must be modeled using \psharp, so all external nondeterminism can be captured.
+We collectively refer to the environment model of a system and its wrapped components as the \psharp \emph{test harness}.
+This test harness can be used by \psharp to drive the system towards interesting behaviors.
 
-\item
-All the asynchrony due to message passing between system components must become explicit and be modeled using the \psharp event-sending APIs.
-Similarly, all other sources of nondeterminism (e.g.
-failures and timer expiration) must be modeled using appropriate \psharp APIs.
-This step allows the \psharp runtime to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing.
-
-\item
-The criteria for correctness of an execution must be specified.
-Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties.
+\item The criteria for correctness of an execution must be specified.
+Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:bg:bugs}).
 Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state.
 Liveness specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.
+
+\item All the asynchrony due to message passing in the system must be modeled using the \psharp event-sending APIs.
+Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs.
+This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:psharp:testing}).
 \end{enumerate}

### 2015-09-21 experience.tex
various fixes
updated for rest of Cheng's feedback

@@
 The MigratingTable modeling effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property; whereas the MigratingTable developers had to write many complex safety properties.
-MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp.
-To be able to test MigratingTable, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).
 
@@
 \includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Environmental model of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine).}
 \label{fig:mockedmigration}

### 2015-09-21 overview.tex
updated for rest of Cheng's feedback

@@
 \textbf{Handling \csharp 5.0 asynchrony}.
-The Live Azure Table Migration system (see \S\ref{sec:cases:migration}) heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+Our case studies heavily use the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
 An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread.
@@
 
+%The \texttt{async} and \texttt{await} \csharp 5.0 primitives were not originally supported in \psharp.
+To be able to test our case studies, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).
+
 Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
-Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon), any TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler.
-These tasks will be intercepted by our custom task scheduler and wrapped inside a special, available only internally, task-machine.
+%Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon),
+All TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler.
+These tasks will be intercepted by our custom task scheduler and wrapped inside a special (not available to the programmer) task-machine.
 When the \psharp testing engine schedules a task-machine for execution, the corresponding task is unwrapped and executed.

### 2015-09-21 experience.tex
Edits to MigratingTable case study description.

--- 2015-09-20 experience.tex
@@
 
-The Live Azure Table Migration (MigratingTable) is a library for \emph{transparently migrating} a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
-This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp model to increase confidence in the former's implementation.
-Indeed, this co-development process resulted in many tricky bugs discovered before the system went into production (see \S\ref{sec:eval}).
-The MigratingTable modeling effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property; whereas the MigratingTable developers had to write many complex safety properties.
+The Live Azure Table Migration library (MigratingTable) is designed to transparently migrate a data set between tables in an Azure Storage service while an application is accessing this data set.
+This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test to increase confidence in the system.
+Indeed, this co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}).
+The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface specification.
 MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp.
@@
 This interface is named \texttt{IChainTable}.
-The virtual table is backed by a pair of \emph{old} and \emph{new} tables.
+The virtual table is backed by a pair of \emph{old} and \emph{new} tables, which must also implement \texttt{IChainTable}.
 A background \emph{migrator} job is responsible for moving all data from the old table to the new table.
-Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol that guarantees linearizability of operations on the virtual table across multiple application processes, assuming that the backend tables respect their own linearizability guarantees.
-
-The testing goal was to ensure that when multiple application processes, implementing the \texttt{IChainTable} interface, issue read and write operations to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of \texttt{IChainTable} for the combined \emph{input history}.
+Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which is intended to guarantee that when multiple application processes issue \emph{input} operations to their own MigratingTable instances with the same backend tables, the output of the MigratingTable instances is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
+The goal of the \psharp test was to verify this property.
 
@@
 
+\textbf{Modeling and testing}.
 The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
-The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
+The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
+Instead they took advantage of the simplicity and near-determinism of the \texttt{IChainTable} specification to write an in-memory reference implementation, called \texttt{SpecTable}, to which the output of MigratingTable on an arbitrary input history can be compared.
+Then the test could sample from a distribution that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the test would have a positive probability of detecting the bug.
 
-\textbf{Modeling and testing}.
-Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history.
-This enabled sampling from a distribution that was defined over all possible input histories within certain bounds.
-The \psharp systematic testing engine automatically takes care of enumerating different input histories.
-The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call.
-Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.
-This makes it possible to check the correctness property as the model executes.
+The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table.
+This allows the test harness to keep a \emph{reference} \texttt{SpecTable} in sync with the virtual table by sending each operation to the reference table when the linearization point on the virtual table is reached.
+The output of the operation on the MigratingTable is then checked against its output on the reference table.
 
@@
 
-The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
-Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine.
-When MigratingTable reports the linearization point of an input call, the \texttt{Service} machine sends that input call to the reference table.
-When an input call completes, the \texttt{Service} machine checks that the results from the MigratingTable and the reference table agree.
+The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old and new backend tables and the reference table; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
+Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends backend operations to the \texttt{Tables} machine.
+The \texttt{Service} machine sends the operation to the reference table at the linearization point and compares the results as previously described.
 
-\psharp captures and controls the interleaving of the backend calls.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.
+% This paragraph can probably be removed if we are out of space. ~ Matt 2015-09-20
+\psharp captures and controls the interleaving of the backend operations.
+To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on the reference table has been completed.
 

### 2015-09-21 evaluation.tex
updates in related work

@@
 
-\subsection{Cost of environmental modeling}
+\subsection{Cost of environment modeling}
 \label{sec:eval:human_cost}
@@
 
-Environmental modeling is a core activity of using \psharp.
-It is required for \emph{closing the environment} of a system to make it amenable to systematic testing.
+Environment modeling is a core activity of using \psharp.
+It is required for \emph{closing} a system to make it amenable to systematic testing.
 Table~\ref{tab:stats} presents program statistics for the three case studies.
@@
 For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug.
-The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations.
+The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
 We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine.

### 2015-09-21 experience.tex

@@
 The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface specification.
-MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp.
-To be able to test MigratingTable, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).
 

### 2015-09-21 methodology.tex
minor updates

@@
 
-\texttt{RepairMonitor} is a \psharp liveness monitor machine (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
+The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
 Whenever an extent node fails, \texttt{RepairMonitor} is notified.
@@
 At the same time, there are only two \emph{true} replicas in the system, which is one less than the target.
-This execution sequence leads to one replica missing; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy.
+This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy.
 In production, such a bug can cause a very serious incident of customer data loss.

### 2015-09-21 evaluation.tex
Edits to MigratingTable evaluation.
Specifically, revise the discussion of the 4 bugs the default test
harness did not find.

@@
 
-Developing the Live Migration Table and its \psharp test harness took about five weeks.
+Developing MigratingTable and its \psharp test harness took about five weeks.
 The test harness was developed in parallel with the actual system.
@@
 
-Using \psharp we managed to uncover 12 serious bugs in our case studies \PDComment{update for Fabric}.
+Using \psharp we managed to uncover 8 serious bugs in our case studies \PDComment{update for Fabric}.
 As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting.
@@
 We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
-The CS column shows which case study corresponds to each bug: 1 is for the Azure Storage vNext; and 2 is for the Live Migration Table.
+The CS column shows which case study corresponds to each bug: 1 is for the Azure Storage vNext; and 2 is for MigratingTable.
 \PDComment{update with 3 for Fabric}
@@
 
-For the MigratingTable case study, the first seven bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history.
-For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug.
+For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on 11 ``bugs'', including 8 that actually occurred in development and 3 other code changes that we deemed interesting ways of making the code incorrect.
+The test found seven of the bugs, which are listed first in Table~\ref{tab:testing}.
 The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
-We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine.
-We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them.
-The design of these custom harnesses was influenced by the knowledge of these bugs; our objective here was to simply see if \psharp can reproduce these known bugs.
+We believe this is because the inputs and schedules that trigger them are too rare in our distribution.
+To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs, as shown.
+This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.
 

### 2015-09-21 methodology.tex

@@
 Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
-The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs.
+The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs.
 It is updated upon the periodic sync reports from the ENs.
@@
 Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth.
-The EN map records the latest heartbeat time from every EN.
+\texttt{ExtentNodeMap} records the latest heartbeat time from every EN.
 
-The ExtMgr runs internally a periodic {\em EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period, as well as cleaning up the corresponding records in the \texttt{ExtentCenter}.
-In addition, the ExtMgr runs a periodic {\em extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.
+The ExtMgr runs internally a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}.
+In addition, the ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.
 
@@
 Messages coming from the ENs do \emph{not} go through the mocked network engine.
-They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}.
+They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessMessage}.
 The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.
@@
       var rec = GetExtentRecord(this.Payload);
-      extCtr.AddOrUpdate(rec); // update Extent Center
+      extCtr.AddOrUpdate(rec); // update ExtentCenter
     }
@@
 The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
-Whenever an extent node fails, \texttt{RepairMonitor} is notified.
-As soon as the number of extent replicas falls below a specified target (three replicas in the current test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
-Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified.
+Whenever an EN fails, \texttt{RepairMonitor} is notified.
+As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
+Whenever a replica is repaired, \texttt{RepairMonitor} is notified.
 It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.
 
-In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should {\em always eventually} end up in the cold repaired state.
-Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long.
+In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should \emph{always eventually} end up in the cold state.
+Otherwise, \texttt{RepairMonitor} is stuck in the hot state for \emph{infinitely} long.
 This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug.
@@
 
-The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} fails one of the \texttt{ExtentNode} and launches a new one.
-The \texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.
+The liveness bug occurs in the second testing scenario, where the \texttt{TestingDriver} machine fails one of the \texttt{ExtentNode} machines and launches a new one.
+\texttt{RepairMonitor} transitions to the hot repairing state and is stuck in the state for infinitely long.
 
-The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from the EN map; (iii) the extent center is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3.
+The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3.
 This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task.

### 2015-09-21 evaluation.tex

@@
 
-For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on 11 ``bugs'', including 8 that actually occurred in development and 3 other code changes that we deemed interesting ways of making the code incorrect.
-The test found seven of the bugs, which are listed first in Table~\ref{tab:testing}.
+For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on 11 ``bugs'', including 8 \emph{organic} bugs that actually occurred in development and 3 \emph{notional} bugs (marked $*$), which are other code changes that we deemed interesting ways of making the code incorrect.
+The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
 The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).

### 2015-09-21 intro.tex

@@
 Developers of production distributed systems use many testing techniques, such as unit testing, integration testing, stress testing, and fault injection.
-In spite of extensive use of these testing methods, many bugs that arise from subtle combinations of concurrency and failure events are missed during testing and get exposed only after a system has been put into production.
-However, allowing bugs to reach production, before they are found and fixed, can cost organizations a lot of money~\cite{tassey2002economic} and lead to customer dissatisfaction~\cite{amazon2012aws, google2014outage}.
+In spite of extensive use of these testing methods, many bugs that arise from subtle combinations of concurrency and failure events are missed during testing and get exposed only in production.
+However, allowing serious bugs to reach production can cost organizations a lot of money~\cite{tassey2002economic} and lead to customer dissatisfaction~\cite{amazon2012aws, google2014outage}.
 
-We interviewed technical leaders and senior managers in the Microsoft Azure team regarding the top problems in distributed system development.
-The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes to production.
+We interviewed technical leaders and senior managers in Microsoft Azure regarding the top problems in distributed system development.
+The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} so that bugs can be uncovered \emph{during testing} and \emph{not in production}.
 The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems that are used in production.
 
-Recently, engineers in Amazon Web Services (AWS) described their use of formal methods ``to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
+Recently, the Amazon Web Services (AWS) team used formal methods ``to prevent serious but subtle bugs from reaching production''~\cite{newcombe2015aws}.
 The gist of their approach is to extract the high-level logic from a production system, represent this logic as specifications in the expressive TLA+~\cite{lamport1994temporal} language, and finally verify the specifications using a model checker.

### 2015-09-21 experience.tex
typo fix

@@
-The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system.
+The modeling and testing approach described in the earlier sections of the paper is not specific to the vNext system.
 \psharp is a generic framework for .NET distributed systems.
@@
 a big data-stream processing system. In prior work~\cite{deligiannis2015psharp}, 
-we used an earlier version of the Fabric model and reported bugs in sample serives.
+we used an earlier version of the Fabric model and reported bugs in sample services.
 Supporting CScale required significant additions to the model, making it much more

### 2015-09-21 paper.tex
update
updated title and some minor stuff
update in title
addressed last of Cheng's feedback

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)}
+\title{\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)}
+%\title{\Large \bf Uncovering Distributed Storage System Bugs during\\Testing (not in Production!)}
 
@@
 
-Our methodology has been applied to three production distributed storage systems in the Microsoft Azure platform.
+Our methodology has been applied to three distributed storage systems in the Microsoft Azure cloud computing platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.

### 2015-09-21 related.tex
addressed last of Cheng's feedback

@@
 Tools such as CrystalBall~\cite{yabandeh2009crystalball} and \textsc{DeMeter}~\cite{guo2011practical} employ techniques based on dynamic partial order reduction~\cite{lauterburg2010evaluating} to reduce the exploration state-space.
-The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing, but the focus is still on reducing the set of interleavings.
-Although the SAMC authors claim that their tool is orders of magnitude faster than previous testing techniques, they mention that their ``extensive evaluation exercised more than 100,000 executions and used approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
+The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing, but the focus is still on reducing the set of possible interleavings.
+Although SAMC is orders of magnitude faster than previous testing techniques, the authors mention that their ``extensive evaluation exercised more than 100,000 executions and used approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
 \psharp is able to explore a similar number of iterations in a matter of hours, and find serious bugs in seconds (see \S\ref{sec:eval}).

### 2015-09-21 evaluation.tex
updates

@@
 Table~\ref{tab:stats} presents program statistics for the three case studies.
-The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp model'' refer to the \psharp test harness.
+The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp Test Harness'' refer to the environment model written in \psharp.
 We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system-under-test (\#B); lines of \psharp code for the test harness (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
 
-Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two weeks of part-time developing.
-The \psharp model for testing this system is the smallest (in lines of code) from the three case studies.
+Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two person weeks of part-time developing.
+The \psharp test harness for this system is the smallest (in lines of code) from the three case studies.
 This was because this modeling exercise aimed to reproduce the particular liveness bug that was haunting the developers of vNext.
 
-Developing MigratingTable and its \psharp test harness took about five weeks.
-The test harness was developed in parallel with the actual system.
-This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and at a later stage of the development process.
+Developing both the MigratingTable system and its \psharp test harness took approximately five person weeks.
+The \psharp test harness was developed in parallel with the actual system.
+This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and after the development process had finished.
 
-Modeling Fabric required approximately five person months. Although this is a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.
+Modeling Fabric required approximately five person months, an effort undertaken by the authors of \psharp.
+In contrast the other two systems discussed in this paper were modeled and tested by their corresponding developers.
+Although modeling Fabric required a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.
 

### 2015-09-21 evaluation.tex
removed comment
update

@@
 
-Using \psharp we managed to uncover 8 serious bugs in our case studies \PDComment{update for Fabric}.
+Using \psharp we managed to uncover 8 serious bugs in our case studies.
 As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting.
@@
 We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
-The CS column shows which case study corresponds to each bug: 1 is for the Azure Storage vNext; and 2 is for MigratingTable.
-\PDComment{update with 3 for Fabric}
+The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
+We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:cases:fabric}).
 

### 2015-09-21 overview.tex
Edits

@@
 \item The real components of a distributed system must be wrapped inside \psharp machines.
-In addition, the environment of the system must be modeled using \psharp, so all external nondeterminism can be captured.
-We collectively refer to the environment model of a system and its wrapped components as the \psharp \emph{test harness}.
-This test harness can be used by \psharp to drive the system towards interesting behaviors.
+In addition, all the asynchrony due to message passing in the system must be modeled using the \psharp event-sending APIs.
+
+\item The environment of the system must be modeled using \psharp.
+\psharp offers APIs for obtaining non-deterministically chosen values, which can be used to simulate failures, timers, etc.
+(see \S\ref{sec:method:driver}).
+This step captures external sources of nondeterminism.
+\psharp's testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
+(see \S\ref{sec:psharp:testing}).
 
@@
 
-\item All the asynchrony due to message passing in the system must be modeled using the \psharp event-sending APIs.
-Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs.
+%\item Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs.
 This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:psharp:testing}).
@@
 \noindent
+We collectively refer to the environment model of a system and its wrapped components as the \psharp \emph{test harness}.
+This test harness can be used by \psharp to drive the system towards interesting behaviors.
 \psharp uses modern object-oriented language features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.

### 2015-09-21 related.tex
Edits

@@
 Tools such as CrystalBall~\cite{yabandeh2009crystalball} and \textsc{DeMeter}~\cite{guo2011practical} employ techniques based on dynamic partial order reduction~\cite{lauterburg2010evaluating} to reduce the exploration state-space.
-The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing, but the focus is still on reducing the set of possible interleavings.
-Although SAMC is orders of magnitude faster than previous testing techniques, the authors mention that their ``extensive evaluation exercised more than 100,000 executions and used approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
-\psharp is able to explore a similar number of iterations in a matter of hours, and find serious bugs in seconds (see \S\ref{sec:eval}).
-However, we do \emph{not} claim that \psharp is faster; as aforementioned, we focus on compositional testing, whereas prior work typically focuses on testing an entire system.
-We are just trying to solve a similar problem at a very different level.
+The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing to reduce the set of interleavings that they explore.
+These techniques are complementary to our approach: \psharp's systematic testing engine could use them for reducing the number of iterations it explores, just as the other tools can use language technnology (like \psharp) to write models and reduce the complexity of the system under test.
+
+% Further, note that we operate at a different scale by focusing on compositional testing.
+For instance, SAMC reports orders of magnitude speedup over previous testing techniques, but executing 100,000 runs still took ``approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
+In our case studies, we execute a similar number of iterations in a matter of hours, and find serious bugs in seconds (see \S\ref{sec:eval}).
+This does \emph{not} imply that \psharp is faster (nor we attempt to make that claim); but simply that we explore behaviors at a different level.
+
+%as aforementioned, we focus on compositional testing, whereas prior work typically focuses on testing an entire system. %We are just trying to solve a similar problem at a very different level.
 

### 2015-09-21 methodology.tex
Edits
remake in section 4

@@
-To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a testing harness in \psharp.
-The developers of vNext expected that it was more likely for the bug to occur in the ExtMgr logic, rather than the EN logic.
-Hence, the testing harness focuses on the real ExtMgr using mocked ENs.
+To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a test harness using \psharp.
+The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic.
+Hence, the developers focused on testing the real ExtMgr using modeled ENs.
+The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
+\begin{description}
+\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext (see \S\ref{sec:method:wrap_target}).
 
-The testing harness consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
-%\begin{enumerate}
-\begin{description}
-\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext.
+\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:mock_en}).
 
-\item[ExtentNode] represents a simplified mocked EN.
+\item[TestingDriver] is responsible for driving testing scenarios, relaying messages between machines, and injecting failures (see \S\ref{sec:method:driver}).
 
-\item[TestingDriver] communicates with all other machines, relays messages between the \texttt{ExtentManager} and the \texttt{ExtentNode} machines, and is responsible for driving testing scenarios.
+\item[Timer] exploits the nondeterministic choice generation available in \psharp to model timeouts (see \S\ref{sec:method:timer}).
 
-\item[RepairMonitor] collects states from all \texttt{ExtentNode} machines in the testing harness and checks desired properties, such as that an extent is replicated or repaired at the expected ENs.
+\item[RepairMonitor] collects EN-related state to check if the desired liveness property is satisfied (see \S\ref{sec:method:monitor}).
 \end{description}
@@
 \includegraphics[width=\linewidth]{img/mocked_vnext}
-\caption{Real Extent Manager with a mocked environment (each box represents one \psharp machine).}
+\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
 \label{fig:azurestoremodel}
+\vspace{-2mm}
 \end{figure}
@@
 
-The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.
+\begin{figure}[t]
+\begin{lstlisting}
+// wrapping the target vNext component in a P# machine
+class ExtentManagerMachine : Machine {
+  private ExtentManager ExtMgr; // real vNext code
+
+  void Init() {
+    ExtMgr = new ExtentManager();
+    ExtMgr.netEngine = new MockedNetEngine(); // mock network
+    ExtMgr.isMockingTimer = true;	 // disable internal timer
+  }
+
+  [OnEvent(ENMsgEvent, DeliverExtentNodeMessage)]
+  void DeliverExtentNodeMessage() {
+    var msg = (ExtentNodeMessage)this.Payload;
+    // relay messages from Extent Node to Extent Manager
+    ExtMgr.ProcessMessage(msg);
+  }
+
+  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
+  void ProcessExtentRepair() {
+    // extent repair loop driven by external timer
+    ExtMgr.ProcessEvent(new ExtentRepairEvent());
+  }
+}
+\end{lstlisting}
+\vspace{-3mm}
+\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
+\label{fig:wrap_target}
+%\vspace{-2mm}
+\end{figure}
+
+The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} machine, as illustrated in the code snippet of Figure~\ref{fig:wrap_target}.
 
@@
 \label{fig:extentmanager}
+\vspace{-3mm}
 \end{figure*}
@@
 \textbf{Internals of the real Extent Manager}.
-Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
+The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
 The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs.
@@
 Recall that the sync report from a particular EN lists all the extents stored at the EN.
-Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth.
+Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth.
 \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.
@@
 
-\textbf{Intercepting messages between machines}.
-The real ExtMgr uses a network engine to send messages to the ENs.
-The testing harness mocks the original network engine in vNext and overrides its interface.
-In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
+\textbf{Intercepting network messages}.
+The real ExtMgr uses a network engine to asynchronously send messages to the ENs.
+The \psharp test harness mocks the original network engine in vNext by overriding its interface.
+This enables the mocked network engine to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
 As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
-Conceivably, the mocked network engine could leverage the nondeterminism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.
-
-%Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution.
-There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.
-\csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class.
-This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.
-
-%Using method dispatch for modeling is straightforward.
-The system-under-test exposes a set of APIs as \emph{virtual methods}.
-The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.
-
-%\begin{figure}[t]
-%\begin{lstlisting}
-%// Public interface of the real network engine
-%class NetworkEngine {
-  %public virtual void SendMessage(Socket s, Message msg);
-  %public virtual void EnqueueMessage(Message msg);
-%}
-%
-%// The mocked network engine used during testing
-%class MockedNetEngine : NetworkEngine {
-  %ExtentManager EM; // Handle to actual system-under-test
-  %MachineId Env; // Handle to modeled environment
-  %
-  %public MockedNetEngine(ExtentManager em, MachineId env) {
-    %this.EM = em;
-    %this.Env = env;
-  %}
-  %
-  %public override void SendMessage(Socket s, Message msg) {
-    %PSharpRuntime.Send(this.Env, new MsgEvent(), s, msg);
-  %}
-  %
-  %public override void EnqueueMessage(Message msg) {
-    %this.EM.ProcessMessage(msg);
-  %}
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{The mocked network engine used for testing the Azure Storage vNext system.}
-%\label{fig:enginecode}
-%%\vspace{-2mm}
-%\end{figure}
+This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.
 
@@
     // intercept and relay Extent Manager messages
-    PSharpRuntime.Send(this.TestingDriver,
-      new MessageFromExtentManagerEvent(), s, msg);
+    PSharpRuntime.Send(TestingDriver,
+      new ExtentManagerMsgEvent(), s, msg);
   }
@@
 \end{lstlisting}
-\vspace{-2mm}
+\vspace{-3mm}
 \caption{Mocked network engine in vNext.}
 \label{fig:enginecode}
-%\vspace{-2mm}
+\vspace{-2mm}
 \end{figure}
 
-%We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}).
-The network engine is responsible for sending to and receiving messages from the various components of the system.
-During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication.
-For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs.
-We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing.
-In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload.
-Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.
+Intercepting all network messages and dispatching them through the \psharp runtime is important for two reasons.
+First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system.
+Second, the mocked network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).
 
-%For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines.
-When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.
-
-\textbf{Specifics of the ExtentManager machine}.
-Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine.
-\texttt{ExtentManager} is a thin wrapper of the real ExtMgr.
-The mocked network engine replaces the real one in ExtMgr, intercepts all the outbound messages from ExtMgr and relays them to \texttt{TestingDriver}.
-
-%A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
-The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
-The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
-This is discussed in detail in \S\ref{sec:method:model:dispath}.
-
-%As shown in Figure~\ref{fig:wrap_target}, the \texttt{ExtentManagerMachine} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code.
-A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested.
-The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class.
-The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives.
-This is discussed in detail in \S\ref{sec:method:model:dispath}.
-
-\begin{figure}[t]
-\begin{lstlisting}
-// Wrapping the target vNext component in a P# machine
-class ExtentManagerMachine : Machine {
-  private ExtentManager extMgr; // real vNext code
-
-  void Init() {
-    extMgr = new ExtentManager();
-    extMgr.netEngine = new MockedNetEngine(); // mock network
-    extMgr.isMockingTimer = true;	 // disable internal timer
-  }
-
-  [OnEvent(MessageFromExtentNode, DeliverExtentNodeMessage)]
-  void DeliverExtentNodeMessage() {
-    var msg = (ExtentNodeMessage)this.Payload;
-    // relay messages from Extent Node to Extent Manager
-    extMgr.ProcessMessage(msg);
-  }
-
-  [OnEvent(TimerTick), nameof(ProcessExtentRepair))]
-  void ProcessExtentRepair() {
-    // extent repair loop driven by external timer
-    extMgr.ProcessEvent(new ExtentRepairEvent());
-  }
-}
-\end{lstlisting}
-\vspace{-2mm}
-\caption{System-under-test: instance of the real Extent Manager wrapped inside the \texttt{ExtentManager} machine.}
-\label{fig:wrap_target}
-%\vspace{-2mm}
-\end{figure}
-
-Messages coming from the ENs do \emph{not} go through the mocked network engine.
-They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessMessage}.
-The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.
-
-System correctness should \emph{not} hinge on the frequency of any individual timer.
-Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
-To achieve this, all timers inside ExtMgr are disabled, and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
-The \psharp timers send nondeterministic timeout events that are controlled by the \psharp runtime.
-Hence, the \psharp testing engine has the freedom to schedule arbitrary interleavings between timeout events and all other regular system events.
+Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine.
+They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
+The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it running in a real distributed environment and communicating with real ENs.
 
@@
 
-The \texttt{ExtentNode} machine is a simplified version of the original EN.
-The machine omits much of the complex details of a real EN, and only mocks the logic necessary for the testing scenarios.
-This mocked logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to the \texttt{ExtentManager} machine.
-
-The testing harness leverages components of Azure Storage vNext whenever it is appropriate.
-For example, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, a component that is used inside a real EN for extent bookkeeping.
-
-In the mocked extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine.
-It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored.
-After receiving an \texttt{ExtentCopyResponse} event from the source, it updates the internal \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
-
-In the mocked EN sync logic, the machine is again driven by an external timer provided by \psharp.
-It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to \texttt{ExtentManager} with \texttt{PSharpRuntime.Send(...)}.
-
-%\begin{figure}[t]
-%\begin{lstlisting}
-%// Wrapping the target vNext component in P#
-%class ExtentNodeMachine : Machine {
-  %private ExtentNode.ExtentCenter extCtr; // real vNext component
-%
-	%[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
-	%[OnEventDoAction(typeof(DownloadExtentRequestEvent), nameof(ProcessDownloadExtentRequestEvent))]
-	%[OnEventDoAction(typeof(DownloadExtentResponseEvent), nameof(ProcessDownloadExtentResponseEvent))]
-	%[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSyncEvent))]
-%
-  %void ProcessRepairExtentRequest()
-  %{
-    %var source = GetRepairSourceMachine(this.Payload);
-    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
-  %}
-%
-  %void ProcessDownloadExtentRequestEvent()
-  %{
-	  %var destination = GetRepairDestinationMachine(this.Payload);
-    %var ext_id = GetExtentID(this.Payload);
-    %
-		%Extent ext;
-    %if (extCtr.TryGet(ext_id, out ext))
-      %PSharpRuntime.Send(destination, new DownloadExtentResponseEvent(), ext);
-  %}
-%
-  %void ProcessDownloadExtentResponseEvent()
-  %{
-    %var ext = (Extent)this.Payload;
-    %extCtr.AddExtent(ext);
-%
-    %this.Monitor<RepairMonitor>(new NotifyExtentRepairCompletion(), this.NodeID);
-  %}
-%
-  %void ProcessExtentNodeSyncEvent()
-  %{
-    %var sync = new ExtentNodeSync(this.NodeID);
-    %extCtr.Extents.ForEach(ext =>
-      %{
-        %var rec = new ExtentRecord(ext.ExtentID, ext.Length, ext.Sealed);
-        %sync.ExtentRecs.Add(rec);
-      %});
-    %PSharpRuntime.Send(this.NameNode, new NameNodeMessageEvent(), sync);
-  %}
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{Mocked Extent Node.}
-%\label{fig:mocked_en}
-%%\vspace{-2mm}
-%\end{figure}
-
-%\begin{figure}[t]
-%\begin{lstlisting}
-%// Mocking Extent Node in P#
-%class ExtentNodeMachine : Machine {
-  %// use real vNext component whenever appropriate
-  %private ExtentNode.ExtentCenter extCtr;
-%
-  %[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
-  %[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSync))]
-%
-  %// extent repair logic
-  %void ProcessRepairExtentRequest()
-  %{
-    %var source = GetRepairSourceMachine(this.Payload);
-    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
-  %}
-  %...
-%
-  %// extent node sync logic
-  %void ProcessExtentNodeSync()
-  %{
-    %var sync = new ExtentNodeSync(this.NodeID);
-    %extCtr.Extents.ForEach(ext =>
-      %{
-        %sync.ExtentRecords.Add(new ExtentRecord(ext));
-      %});
-    %PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
-  %}
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{Mocked Extent Node.}
-%\label{fig:mocked_en}
-%%\vspace{-2mm}
-%\end{figure}
-
 \begin{figure}[t]
 \begin{lstlisting}
-// Mocking Extent Node in P#
+// modeling Extent Node in P#
 class ExtentNodeMachine : Machine {
   // leverage real vNext component whenever appropriate
-  private ExtentNode.ExtentCenter extCtr;
+  private ExtentNode.ExtentCenter ExtCtr;
 
@@
   ...
-  [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
+  [OnEvent(ExtentCopyRespEvent, ProcessExtentCopyResponse)]
   void ProcessExtentCopyResponse() {
@@
       var rec = GetExtentRecord(this.Payload);
-      extCtr.AddOrUpdate(rec); // update ExtentCenter
+      ExtCtr.AddOrUpdate(rec); // update ExtentCenter
     }
@@
   // extent node sync logic
-  [OnEvent(TimerTick, ProcessExtentNodeSync)]
+  [OnEvent(TimerTickEvent, ProcessExtentNodeSync)]
   void ProcessExtentNodeSync() {
-    var sync = extCtr.GetSyncReport(); // prepare sync report
-    PSharpRuntime.Send(this.ExtentManagerMachine,
-      new MessageFromExtentNodeEvent(), sync);
+    var sync = ExtCtr.GetSyncReport(); // prepare sync report
+    this.Send(ExtentManagerMachine, new ENMsgEvent(), sync);
+  }
+
+  // extent node failure logic
+  [OnEvent(FailEvent, ProcessFailure)]
+  void ProcessFailure() {
+    // notifies the Extent Manager that this EN failed
+    this.Send(ExtentManagerMachine, new ENMsgEvent(),
+      new ExtentNodeDeletion(ExtentId));
+    // notifies the monitor that this EN failed
+    this.Monitor<RepairMonitor>(new ENFailedEvent(), this);
+    this.Halt(); // terminate this P# machine
   }
@@
 \end{lstlisting}
+\vspace{-3mm}
+\caption{The modeled EN in vNext.}
+\label{fig:mocked_en}
 \vspace{-2mm}
-\caption{The mocked EN in vNext.}
-\label{fig:mocked_en}
-%\vspace{-2mm}
 \end{figure}
+
+The \texttt{ExtentNode} machine is a simplified version of the original EN.
+The machine omits most of the complex details of a real EN, and only models the necessary logic for testing.
+This modeled logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to \texttt{ExtentManager}.
+
+The \psharp test harness leverages components of the real vNext system whenever it is appropriate.
+For example, \texttt{ExtentNode} re-uses the \texttt{ExtentCenter} data structure, which is used inside a real EN for extent bookkeeping.
+In the modeled extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine.
+It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored.
+After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
+
+In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}).
+It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharpRuntime.Send(...)}.
+The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.
 
@@
 
+\begin{figure}[t]
+\begin{lstlisting}
+// machine for driving testing scenarios in vNext
+class TestingDriver : Machine {
+  private HashSet<MachineId> ExtentNodes; // EN machines
+
+  // other test harness driving logic
+  ...
+  void InjectNodeFailure() {
+    // nondeterministically choose an EN using P#
+    var node = (MachineId)this.Nondet(ExtentNodes);    
+    this.Send(node, new FailEvent()); // fail the chosen EN
+  }
+}
+\end{lstlisting}
+\vspace{-3mm}
+\caption{The TestingDriver machine in vNext.}
+\label{fig:testing_driver}
+\vspace{-2mm}
+\end{figure}
+
 The \texttt{TestingDriver} machine drives two testing scenarios.
-In the first scenario, it launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs.
-It waits for the extent to be replicated at the other ENs.
-In the second scenario, it fails one of the \texttt{ExtentNode} machines and launches a new one.
-It waits for the extent to be repaired on the new EN.
-We omit illustrative code snippets due to space constraints.
+In the first scenario, \texttt{TestingDriver} launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs.
+\texttt{TestingDriver} then waits for the extent to be replicated at the remaining ENs.
+In the second testing scenario, \texttt{TestingDriver} fails one of the \texttt{ExtentNode} machines and launches a new one.
+\texttt{TestingDriver} then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.
 
-\subsection{The RepairMonitor machine}
+Figure~\ref{fig:testing_driver} illustrates how \psharp can be used to inject a nondeterministic failure.
+\texttt{TestingDriver} can call the \texttt{InjectNodeFailure} method, which invokes the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of available \texttt{ExtentNode} machines.
+\texttt{TestingDriver} then sends the chosen machine a \texttt{FailEvent} event.
+The chosen \texttt{ExtentNode} handles this event (see Figure~\ref{fig:mocked_en}) by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sends it to the \texttt{ExtentManager} machine.
+\texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for unit testing, which maximizes code re-use.
+Next, the chosen \texttt{ExtentNode} notifies the liveness monitor that an EN has failed.
+Finally, the chosen \texttt{ExtentNode} calls the \psharp method \texttt{Halt()}, which causes the machine to terminate.
+\psharp is able to not only enumerate interleavings of asynchronous event handlers, but also calls to \texttt{Nondet()}, and thus failure injections.
+
+\subsection{The Timer machine}
+\label{sec:method:timer}
+
+\begin{figure}[t]
+\begin{lstlisting}
+// modeling timer expiration in P#
+class Timer : Machine {
+  MachineId Target; // target machine
+
+  [OnEvent(RepeatEvent, GenerateTimeTick)]
+  void GenerateTimeTick() {
+    // nondeterministic choice controlled by P#
+    if (this.Nondet())
+      this.Send(this.Target, new TimerTickEvent());
+    this.Send(this, new RepeatEvent()); // loop
+  }
+}
+\end{lstlisting}
+\vspace{-3mm}
+\caption{Modeling timer expiration using \psharp.}
+\label{fig:mocked_timer}
+\vspace{-2mm}
+\end{figure}
+
+System correctness should \emph{not} hinge on the frequency of any individual timer.
+Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
+To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
+Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.
+
+Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext.
+\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime.
+Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine.
+The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.
+
+\subsection{The RepairMonitor liveness monitor}
 \label{sec:method:monitor}
 
-The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
-Whenever an EN fails, \texttt{RepairMonitor} is notified.
-As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired.
-Whenever a replica is repaired, \texttt{RepairMonitor} is notified.
-It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.
+\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
+Whenever an EN fails, the monitor is notified with an \texttt{ENFailedEvent} event.
+As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), the monitor transitions into the hot \emph{repairing} state, waiting for all missing replicas to be repaired.
+Whenever an extent replica is repaired, \texttt{RepairMonitor} is notified with an \texttt{ExtentRepairedEvent} event.
+When the replica number reaches again the target, the monitor transitions into the cold \emph{repaired} state, as illustrated in the code snippet of Figure~\ref{fig:monitor}.
 
@@
   // cold state: repaired
-  [OnEvent(NotifyEnFailure, ProcessExtentNodeFailure)]
+  [OnEvent(ENFailedEvent, ProcessExtentNodeFailure)]
   cold state Repaired {
@@
   // hot state: repairing
-  [OnEvent(NotifyExtentRepaired, ProcessExtentRepaired)]
+  [OnEvent(ExtentRepairedEvent, ProcessExtentRepaired)]
   hot state Repairing {
@@
 \end{lstlisting}
+\vspace{-3mm}
+\caption{The RepairMonitor liveness monitor.}
+\label{fig:monitor}
 \vspace{-2mm}
-\caption{The RepairMonitor machine.}
-\label{fig:monitor}
-%\vspace{-2mm}
 \end{figure}
@@
 In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit to be identified promptly, and the fix to be tested effectively and thoroughly, all of which have the potential to vastly increase the productivity of distributed storage system development.
-
-%%\subsection{Modeling the environment}
-%\subsection{Test Harness}
-%\label{sec:method:model}
-%
-%The environment of a distributed system might consist of other distributed systems and services, clients, operating system timers, as well as libraries for networking or other purposes.
-To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
-%
-%%\subsubsection{Creating a test harness using \psharp}
-%\subsubsection{Orchestrating Machines}
-%\label{sec:method:model:harness}
-%
-%\begin{figure}[t]
-%\centering
-%\includegraphics[width=\linewidth]{img/mocked_vnext}
-%\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
-%\label{fig:azurestoremodel}
-%\end{figure}
-%
-%Before being able to use \psharp to test an existing distributed system, the environment of the components that the developer wants to test have to be modeled using the \psharp state machine APIs.
-This involves creating mock classes that inherit from the \texttt{Machine} abstract class.
-The \texttt{Machine} class exposes methods for declaring a state machine (e.g.
-machine states and state transitions) as well as methods for sending events to other machines, declaring event handlers for received events, and accessing the payload from the latest received event.
-%
-%In the Azure Storage vNext case study, we created the following \psharp machines: \texttt{Environment}, \texttt{ExtentNode}, \texttt{ExtentManager}, and \texttt{Timer}.
-%
-%The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute.
-The programmer is free to create either a finite test harness (e.g.
-with a predefined number of node failures) or an infinite test harness (e.g.
-with an unbounded number of failures), as long as any nondeterminism (e.g.
-when to insert a failure) is captured using the available \psharp APIs.
-%
-%Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g.
-synchronization messages, heartbeats and repair logic loops).
-This is discussed in detail in \S\ref{sec:method:model:timers}.
-%
-%\subsubsection{Modeling and injecting failures}
-%\label{sec:method:model:failures}
-%
-%In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive.
-Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness.
-However, the Extent Manager logic relies on time intervals to detect node failures (see Figure~\ref{fig:expiration}).
-The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
-%
-%\begin{figure}[t]
-%\begin{lstlisting}
-%// Real code for detecting node expiration
-%public virtual bool IsNodeExpired(string id, DateTime time)
-%{
-  %return DateTime.Compare(time, DateTime.Now) <= 0;
-%}
-%
-%// Mocked code for detecting node expiration
-%public override bool IsNodeExpired(string id, DateTime time)
-%{
-  %return this.DeletedNodes.Contains(id);
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
-%\label{fig:expiration}
-%%\vspace{-2mm}
-%\end{figure}
-%
-%Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method.
-Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration.
-If it contains the id, then it means that the node has failed.
-The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
-%
-%\subsubsection{Entry point to a \psharp test harness}
-%\label{sec:method:model:entrypoint}
-%
-%A developer can specify the entry point of a \psharp test harness similar to how unit tests are typically written.
-Figure~\ref{fig:entrypoint} shows the source code of the entry point for the Azure Storage vNext \psharp test harness.
-The programmer can declare an entry point method using the \texttt{Microsoft.PSharp.Test} attribute.
-When the \psharp systematic testing engine is invoked, it will scan the binary and attempt to find a static method declared with this attribute.
-When such a method is found (in our example the \texttt{Execute()} method), the testing engine will invoke it and start testing the program.
-The systematic engine can optionally run multiple iterations of the test harness, each one potentially exploring a different schedule.
-In each iteration, the program state will be reset (any static fields must be explicitly reset by the programmer) and the entry point will be re-invoked.
-%
-%\begin{figure}[t]
-%\begin{lstlisting}
-%[Microsoft.PSharp.Test]
-%public static void Execute()
-%{
-  %PSharpRuntime.CreateMachine(typeof(Environment));
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{Static \csharp method acting as the entry point to the Azure Storage vNext \psharp test harness.}
-%\label{fig:entrypoint}
-%%\vspace{-2mm}
-%\end{figure}
-%
-%\subsubsection{Abstracting timers}
-%\label{sec:method:model:timers}
-%
-%Distributed systems are often using timers to determine when an event should be send from one component to another.
-For example, in the Azure Storage vNext system, each Extent Node is associated with a timer that fires of a synchronization message every 5 minutes and a heartbeat every 5 seconds.
-This timer is related to the liveness bug that we discovered: the synchronization message that gets fired every 5 minutes can potentially race with an Extent Node failure; if it arrives after the node failed, then the bug would manifest.
-Traditional testing techniques cannot easily find such a bug, due to the very infrequent occurrence of this race due to the timer.
-%
-%Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.
-This nondeterminism is introduced using the \psharp \texttt{Nondet()} method, which returns a nondeterministic boolean value that is controlled by the \psharp runtime during systematic testing.
-%
-%Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study.
-The Extend Manager, as well as each Extent Node in the harness, is associated with a unique \texttt{Timer} machine.
-When creating this machine, we pass as a payload the id of the machine that owns this timer.
-When the \texttt{Timer} machine is created, it stores this id in the \texttt{Owner} field and then transitions to the \texttt{Active} state.
-In this state, the \texttt{Timer} loops infinitely and nondeterministically (using \texttt{Nondet()}) sends a \texttt{TimerTickEvent} to \texttt{this.Owner}.
-When the Extent Node owner receives this event, it handles it by generating a synchronization message that is being send to the Extent Manager.
-Similarly, when the Extent Manager receives a \texttt{TimerTickEvent} from its own \texttt{Timer}, it handles it by nondeterministically invoking repair-related methods in the Extent Repair Center data structure.
-%
-%\begin{figure}[t]
-%\begin{lstlisting}
-%internal class Timer : Machine
-%{
-  %MachineId Owner; // Id of the owner machine
-%
-  %[Start]
-  %[OnEntry(nameof(InitOnEntryAction))]
-  %[OnEventGotoState(typeof(Unit), typeof(Active))]
-  %class Init : MachineState { }
-%
-  %void InitOnEntryAction()
-  %{
-    %this.Owner = (MachineId)this.Payload;
-    %// triggers state transition to Active
-    %this.Raise(new Unit());
-  %}
-%
-  %[OnEntry(nameof(ProcessTickEvent))]
-  %[OnEventGotoState(typeof(Unit), typeof(Active))]
-  %class Active : MachineState { }
-%
-  %void ProcessTickEvent()
-  %{
-    %// Nondeterministic boolean choice controlled by P#
-    %if (this.Nondet())
-      %// sends a timer tick event to the owner machine
-      %this.Send(this.Owner, new TimerTickEvent());
-    %// triggers state transition to Active
-    %this.Raise(new Unit());
-  %}
-%}
-%\end{lstlisting}
-%\vspace{-2mm}
-%\caption{Timers in Azure Storage vNext are modeled as nondeterministic \psharp machines.}
-%\label{fig:timer}
-%%\vspace{-2mm}
-%\end{figure}

### 2015-09-21 evaluation.tex

@@
 The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects the type of oversight that tends to occur as designs evolve.
-We do not discuss the details of the bug due to space constraints, but \psharp managed to discover this bug in a matter of seconds.
-The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces.
-The developers extended the logging of \psharp with additional trace information to understand the bug.
+This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key.
+The essential implementation idea is to perform streaming reads from both backend tables and merge the results.
+According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read.
+The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table.
+But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed.
+\psharp managed to discover this bug in a matter of seconds.
+The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable.
 \psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend and was done in all our case studies.

### 2015-09-21 methodology.tex
some updates

@@
 
+\item[Timer] exploits the nondeterministic choice generation available in \psharp to model timeouts (see \S\ref{sec:method:timer}).
+
 \item[TestingDriver] is responsible for driving testing scenarios, relaying messages between machines, and injecting failures (see \S\ref{sec:method:driver}).
-
-\item[Timer] exploits the nondeterministic choice generation available in \psharp to model timeouts (see \S\ref{sec:method:timer}).
 
@@
 \label{fig:extentmanager}
-\vspace{-3mm}
+\vspace{-5mm}
 \end{figure*}
@@
 \end{lstlisting}
-\vspace{-3mm}
+\vspace{-4mm}
 \caption{Mocked network engine in vNext.}
@@
 \end{lstlisting}
-\vspace{-3mm}
+\vspace{-4mm}
 \caption{The modeled EN in vNext.}
@@
 
+\subsection{The Timer machine}
+\label{sec:method:timer}
+
+\begin{figure}[t]
+\begin{lstlisting}
+// modeling timer expiration in P#
+class Timer : Machine {
+  MachineId Target; // target machine
+
+  [OnEvent(RepeatEvent, GenerateTimeTick)]
+  void GenerateTimeTick() {
+    // nondeterministic choice controlled by P#
+    if (this.Nondet())
+      this.Send(this.Target, new TimerTickEvent());
+    this.Send(this, new RepeatEvent()); // loop
+  }
+}
+\end{lstlisting}
+\vspace{-4mm}
+\caption{Modeling timer expiration using \psharp.}
+\label{fig:mocked_timer}
+\vspace{-2mm}
+\end{figure}
+
+System correctness should \emph{not} hinge on the frequency of any individual timer.
+Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
+To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
+Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.
+
+Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext.
+\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime.
+Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine.
+The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.
+
 \subsection{The TestingDriver machine}
@@
 \end{lstlisting}
-\vspace{-3mm}
+\vspace{-4mm}
 \caption{The TestingDriver machine in vNext.}
@@
 
-Figure~\ref{fig:testing_driver} illustrates how \psharp can be used to inject a nondeterministic failure.
-\texttt{TestingDriver} can call the \texttt{InjectNodeFailure} method, which invokes the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of available \texttt{ExtentNode} machines.
-\texttt{TestingDriver} then sends the chosen machine a \texttt{FailEvent} event.
-The chosen \texttt{ExtentNode} handles this event (see Figure~\ref{fig:mocked_en}) by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sends it to the \texttt{ExtentManager} machine.
+Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures.
+The \texttt{TestingDriver} machine can invoke the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode}machines.
+Let $n_f$ be this chosen machine.
+\texttt{TestingDriver} can emulate a failure by sending $n_f$ a \texttt{FailEvent} event.
+The $n_f$ machine handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}).
 \texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for unit testing, which maximizes code re-use.
-Next, the chosen \texttt{ExtentNode} notifies the liveness monitor that an EN has failed.
-Finally, the chosen \texttt{ExtentNode} calls the \psharp method \texttt{Halt()}, which causes the machine to terminate.
+Next, $n_f$ notifies the liveness monitor that it has failed.
+Finally, $n_f$ calls the \psharp method \texttt{Halt()}, which causes $n_f$ to terminate.
 \psharp is able to not only enumerate interleavings of asynchronous event handlers, but also calls to \texttt{Nondet()}, and thus failure injections.
-
-\subsection{The Timer machine}
-\label{sec:method:timer}
-
-\begin{figure}[t]
-\begin{lstlisting}
-// modeling timer expiration in P#
-class Timer : Machine {
-  MachineId Target; // target machine
-
-  [OnEvent(RepeatEvent, GenerateTimeTick)]
-  void GenerateTimeTick() {
-    // nondeterministic choice controlled by P#
-    if (this.Nondet())
-      this.Send(this.Target, new TimerTickEvent());
-    this.Send(this, new RepeatEvent()); // loop
-  }
-}
-\end{lstlisting}
-\vspace{-3mm}
-\caption{Modeling timer expiration using \psharp.}
-\label{fig:mocked_timer}
-\vspace{-2mm}
-\end{figure}
-
-System correctness should \emph{not} hinge on the frequency of any individual timer.
-Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
-To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
-Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.
-
-Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext.
-\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime.
-Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine.
-The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.
 
@@
 \end{lstlisting}
-\vspace{-3mm}
+\vspace{-4mm}
 \caption{The RepairMonitor liveness monitor.}

### 2015-09-21 methodology.tex
minor edits

--- 2015-09-21 methodology.tex
@@
 The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic.
-Hence, the developers focused on testing the real ExtMgr using modeled ENs.
+Hence, they focused on testing the real ExtMgr using modeled ENs.
 The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
@@
 The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
-The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs.
-It is updated upon the periodic sync reports from the ENs.
+The \texttt{ExtentCenter} data structure maintains the records that map from extents to their hosting ENs.
+It is updated upon receiving the periodic sync reports from the ENs.
 Recall that the sync report from a particular EN lists all the extents stored at the EN.

### 2015-09-21 methodology.tex
more edits

@@
 
-The ExtMgr runs internally a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}.
-In addition, the ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.
+ExtMgr internally runs a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}.
+In addition, ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.
 
@@
 The \psharp test harness mocks the original network engine in vNext by overriding its interface.
-This enables the mocked network engine to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
+This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
 As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
@@
 They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
-The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it running in a real distributed environment and communicating with real ENs.
+The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.
 
@@
 Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
-To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}.
+To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}.
 Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.
@@
 Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures.
-The \texttt{TestingDriver} machine can invoke the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode}machines.
-Let $n_f$ be this chosen machine.
-\texttt{TestingDriver} can emulate a failure by sending $n_f$ a \texttt{FailEvent} event.
-The $n_f$ machine handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}).
-\texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for unit testing, which maximizes code re-use.
-Next, $n_f$ notifies the liveness monitor that it has failed.
-Finally, $n_f$ calls the \psharp method \texttt{Halt()}, which causes $n_f$ to terminate.
-\psharp is able to not only enumerate interleavings of asynchronous event handlers, but also calls to \texttt{Nondet()}, and thus failure injections.
+The \texttt{TestingDriver} machine uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode} machines.
+If \texttt{n} is the chosen machine, \texttt{TestingDriver} emulates a failure by sending a \texttt{FailEvent} event to \texttt{n}.
+Machine \texttt{n} handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}).
+\texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for simulating failures in its unit tests; we simply re-use this functionality.
+Next, \texttt{n} notifies the liveness monitor that it has failed.
+Finally, machine \texttt{n} calls the \psharp method \texttt{Halt()}, which causes \texttt{n} to terminate.
+\psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.
 
@@
 
-However, the original \psharp trace did not include sufficient detail to allow the developers to identify the root cause of the problem.
-Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration.
+The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem.
+Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debuggin outputs in each iteration.
 After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug.
-Once the proposed solution was implemented, the developers ran again the test harness.
-No bugs were found during 100,000 iterations, a process that required only tens of minutes.
+Once the proposed solution was implemented, the developers ran the test harness again.
+No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.
 
@@
 At the same time, there are only two \emph{true} replicas in the system, which is one less than the target.
-This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy.
+This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr would still think that all replicas are healthy.
 In production, such a bug can cause a very serious incident of customer data loss.
 
-The culprit is in (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN.
-This may occur in \psharp due to arbitrary interleaving of events.
+The culprit is in step (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN.
+This interleaving is exposed quickly by \psharp's testing engine that has the control to arbitrarily interleave events.
 It may also occur, albeit much less frequently, during stress testing due to messages being delayed in the network.

### 2015-09-21 evaluation.tex
anonymization: azure -> mycloud, microsoft -> mycompany

@@
 \end{tabular}
-\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
+\caption{Statistics from modeling the environment of the three \Microsoft MyCloud-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
 \label{tab:stats}
@@
 
-Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two person weeks of part-time developing.
+Modeling the environment of the Extent Manager in the \Azure Storage vNext system required approximately two person weeks of part-time developing.
 The \psharp test harness for this system is the smallest (in lines of code) from the three case studies.
@@
 We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
-The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
+The CS column shows which case study corresponds to each bug: ``1'' is for the \Azure Storage vNext; and ``2'' is for MigratingTable.
 We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:cases:fabric}).

### 2015-09-21 experience.tex
anonymization: azure -> mycloud, microsoft -> mycompany

@@
 \psharp is a generic framework for .NET distributed systems.
-We showcase this capability by presenting developer experience of using \psharp to model and test two other systems used in production inside Microsoft.
+We showcase this capability by presenting developer experience of using \psharp to model and test two other systems used in production inside \Microsoft.
 
-\subsection{Live Azure Table Migration}
+\subsection{Live \Azure Table Migration}
 \label{sec:cases:migration}
 
-The Live Azure Table Migration library (MigratingTable) is designed to transparently migrate a data set between tables in an Azure Storage service while an application is accessing this data set.
+The Live \Azure Table Migration library (MigratingTable) is designed to transparently migrate a data set between tables in an \Azure Storage service while an application is accessing this data set.
 This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test to increase confidence in the system.
@@
 
-MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary Azure table.
+MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary \Azure table.
 This interface is named \texttt{IChainTable}.
@@
 
-\subsection{Azure Service Fabric}
+\subsection{\Azure Service Fabric}
 \label{sec:cases:fabric}
 
-Azure Service Fabric\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}} (or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
+\Azure Service Fabric
+%\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}}
+(or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
 %The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 

### 2015-09-21 intro.tex
anonymization: azure -> mycloud, microsoft -> mycompany

@@
 
-We interviewed technical leaders and senior managers in Microsoft Azure regarding the top problems in distributed system development.
+We interviewed technical leaders and senior managers in \Microsoft \Azure regarding the top problems in distributed system development.
 The consensus was that one of the most critical problems today is how to improve \emph{testing coverage} so that bugs can be uncovered \emph{during testing} and \emph{not in production}.
-The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems that are used in production.
+The need for better testing techniques is not specific to \Microsoft; other companies, such as Amazon and Google, have acknowledged~\cite{chandra2007paxos,newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems that are used in production.
 
@@
 Instead, we allow developers to test \emph{production code} by writing test harnesses in \csharp, a mainstream programming language.
-This significantly lowered the acceptance barrier for adoption by the Microsoft Azure team.
+This significantly lowered the acceptance barrier for adoption by the \Microsoft \Azure team.
 
@@
 
-We have applied \psharp to three distributed storage systems inside Microsoft: Azure Storage vNext; Azure Table Live Migration; and Azure Service Fabric.
+We have applied \psharp to three distributed storage systems inside \Microsoft: \Azure Storage vNext; \Azure Table Live Migration; and \Azure Service Fabric.
 %a distributed storage system; a system for migrating data across partitions; and a big-data computing system.

### 2015-09-21 methodology.tex
anonymization: azure -> mycloud, microsoft -> mycompany

@@
-To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a test harness using \psharp.
+To uncover the elusive extent repair bug in \Azure Storage vNext, its developers wrote a test harness using \psharp.
 The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic.
@@
 \includegraphics[width=.9\linewidth]{img/extent_manager}
-\caption{Internal components of the real Extent Manager in Azure Storage vNext.}
+\caption{Internal components of the real Extent Manager in \Azure Storage vNext.}
 \label{fig:extentmanager}
@@
 
-\subsection{Liveness Bug in Azure Storage vNext}
+\subsection{Liveness Bug in \Azure Storage vNext}
 \label{sec:method:azurestore}

### 2015-09-21 motivation.tex
updates to address some of Cheng's comments
anonymization: azure -> mycloud, microsoft -> mycompany

@@
-Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
+\Microsoft \Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
+It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
+%~\cite{greenberg2015keynote}.
 
-Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
+\Azure Storage vNext is the next generation storage system for \Microsoft \Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
 Similar to the current system, vNext employs containers, called \emph{extents}, to store data.
@@
 \includegraphics[width=\linewidth]{img/azurestore}
-\caption{Top-level components of vNext, a distributed extent management system for Microsoft Azure.}
+\caption{Top-level components for extent management in \Microsoft \Azure Storage vNext.}
 \label{fig:vnext}

### 2015-09-21 overview.tex
anonymization: azure -> mycloud, microsoft -> mycompany

@@
 Our approach of using \psharp to test distributed systems, implemented in .NET, requires the developer to perform the following three key modeling tasks.
-These tasks are illustrated in \S\ref{sec:method} for the Azure Storage vNext system.
+These tasks are illustrated in \S\ref{sec:method} for the \Azure Storage vNext system.
 
@@
 Programmers in industry are used to working with such features, and heavily use them in production for testing purposes.
-In our experience, this significantly lowers the bar for product teams inside Microsoft to embrace \psharp for testing.
+In our experience, this significantly lowers the bar for engineering teams inside \Microsoft to embrace \psharp for testing.
 

### 2015-09-21 methodology.tex
updated section 4

@@
 This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
-As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
+As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
 This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.
@@
     // intercept and relay Extent Manager messages
-    PSharpRuntime.Send(TestingDriver,
-      new ExtentManagerMsgEvent(), s, msg);
+    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
   }
@@
     var sync = ExtCtr.GetSyncReport(); // prepare sync report
-    this.Send(ExtentManagerMachine, new ENMsgEvent(), sync);
+    PSharp.Send(ExtentManagerMachine, 
+      new ENMsgEvent(), sync);
   }
@@
   // extent node failure logic
-  [OnEvent(FailEvent, ProcessFailure)]
+  [OnEvent(FailureEvent, ProcessFailure)]
   void ProcessFailure() {
-    // notifies the Extent Manager that this EN failed
-    this.Send(ExtentManagerMachine, new ENMsgEvent(),
-      new ExtentNodeDeletion(ExtentId));
     // notifies the monitor that this EN failed
-    this.Monitor<RepairMonitor>(new ENFailedEvent(), this);
-    this.Halt(); // terminate this P# machine
+    PSharp.Monitor<RepairMonitor>(new ENFailedEvent(), this);
+    PSharp.Halt(); // terminate this P# machine
   }
@@
 In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}).
-It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharpRuntime.Send(...)}.
+It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharp.Send(...)}.
 The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.
@@
 class Timer : Machine {
-  MachineId Target; // target machine
+  Machine Target; // target machine
 
-  [OnEvent(RepeatEvent, GenerateTimeTick)]
-  void GenerateTimeTick() {
+  [OnEvent(RepeatedEvent, GenerateTimerTick)]
+  void GenerateTimerTick() {
     // nondeterministic choice controlled by P#
-    if (this.Nondet())
-      this.Send(this.Target, new TimerTickEvent());
-    this.Send(this, new RepeatEvent()); // loop
+    if (PSharp.Nondet())
+      PSharp.Send(Target, new TimerTickEvent());
+    PSharp.Send(this, new RepeatedEvent()); // loop
   }
@@
 System correctness should \emph{not} hinge on the frequency of any individual timer.
-Instead, all nondeterminism due to timing-related events must be delegated to \psharp.
-To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}.
-Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.
+Hence, it makes sense to delegate all nondeterminism due to timing-related events to \psharp.
+To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp.
+%, an approach also used in previous work~\cite{desai2015building}.
+Similarly, \texttt{ExtentNode} machines do {\em not} have internal timers either. Their periodic heartbeats and sync reports are also driven by timers modeled in \psharp.
 
-Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext.
-\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime.
-Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine.
+Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine in the test harness.
+\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice controlled by the \psharp runtime.
+Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to its target (the \texttt{ExtentManager} or \texttt{ExtentNode} machines).
 The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.
@@
 
-  // other test harness driving logic
+  // other test harness logic
   ...
@@
     // nondeterministically choose an EN using P#
-    var node = (MachineId)this.Nondet(ExtentNodes);    
-    this.Send(node, new FailEvent()); // fail the chosen EN
+    var node = (Machine)PSharp.Nondet(ExtentNodes);    
+    PSharp.Send(node, new FailureEvent()); // fail chosen EN
   }
@@
 In the first scenario, \texttt{TestingDriver} launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs.
-\texttt{TestingDriver} then waits for the extent to be replicated at the remaining ENs.
+It then waits for the extent to be replicated at the remaining ENs.
 In the second testing scenario, \texttt{TestingDriver} fails one of the \texttt{ExtentNode} machines and launches a new one.
-\texttt{TestingDriver} then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.
+It then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.
 
-Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures.
-The \texttt{TestingDriver} machine uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode} machines.
-If \texttt{n} is the chosen machine, \texttt{TestingDriver} emulates a failure by sending a \texttt{FailEvent} event to \texttt{n}.
-Machine \texttt{n} handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}).
-\texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for simulating failures in its unit tests; we simply re-use this functionality.
-Next, \texttt{n} notifies the liveness monitor that it has failed.
-Finally, machine \texttt{n} calls the \psharp method \texttt{Halt()}, which causes \texttt{n} to terminate.
+Figure~\ref{fig:testing_driver} illustrates how \texttt{TestingDriver} leverages \psharp to inject nondeterministic failures.
+It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine and sends a \texttt{FailureEvent} to the chosen machine to emulate EN failure.
+As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent},
+%and instructs \texttt{ExtentManager} to expire itself by sending a \texttt{ExtentNodeDeletion} message.\texttt{ExtentNodeDeletion} is the same message that is used by the real \texttt{ExtMgr} to expire an EN.
+At the same time, \texttt{ExtentNode}
+notifies the liveness monitor of its failure and terminates itself by invoking \texttt{PSharp.Halt()}.
 \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.
@@
 class RepairMonitor : Monitor {
-  // true: EN has replica, false: EN has no replica
-  private Dictionary<Machine, bool> ExtentNode2Replica;
+  private HashSet<Machine> ExtentNodesWithReplica;
 
@@
         var node = GetExtentNode(this.Payload);
-        ExtentNode2Replica.Remove(node);
+        ExtentNodesWithReplica.Remove(node);
+        if (ReplicaCount < Harness.REPLICA_COUNT_TARGET)				
           goto Repairing;
@@
       var node = GetExtentNode(this.Payload);
-      ExtentNode2Replica[node] = true;
+      ExtentNodesWithReplica.Add(node);
       if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)

### 2015-09-22 methodology.tex
minor updates

@@
     // notifies the monitor that this EN failed
-    PSharp.Monitor<RepairMonitor>(new ENFailedEvent(), this);
+    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
     PSharp.Halt(); // terminate this P# machine
@@
 class TestingDriver : Machine {
-  private HashSet<MachineId> ExtentNodes; // EN machines
+  private HashSet<Machine> ExtentNodes; // EN machines
 
@@
 Figure~\ref{fig:testing_driver} illustrates how \texttt{TestingDriver} leverages \psharp to inject nondeterministic failures.
-It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine and sends a \texttt{FailureEvent} to the chosen machine to emulate EN failure.
-As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent},
-%and instructs \texttt{ExtentManager} to expire itself by sending a \texttt{ExtentNodeDeletion} message.\texttt{ExtentNodeDeletion} is the same message that is used by the real \texttt{ExtMgr} to expire an EN.
-At the same time, \texttt{ExtentNode}
-notifies the liveness monitor of its failure and terminates itself by invoking \texttt{PSharp.Halt()}.
+It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine, and then sends a \texttt{FailureEvent} to the chosen machine to emulate an EN failure.
+As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent}, notifies the liveness monitor of its failure (see \S\ref{sec:method:monitor}) and terminates itself by invoking the \psharp method \texttt{Halt()}.
 \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

### 2015-09-22 overview.tex
some fixes

@@
 
-\item The environment of the system must be modeled using \psharp.
-\psharp offers APIs for obtaining non-deterministically chosen values, which can be used to simulate failures, timers, etc.
-(see \S\ref{sec:method:driver}).
+\item The environment of the system must be modeled in \psharp, which offers APIs for obtaining non-deterministically chosen values.
+These values can be used to simulate failures, timers, etc.
+(see \S\ref{sec:method:timer} and \S\ref{sec:method:driver}).
 This step captures external sources of nondeterminism.
-\psharp's testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
+The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
 (see \S\ref{sec:psharp:testing}).

### 2015-09-22 related.tex
some fixes

@@
 Tools such as CrystalBall~\cite{yabandeh2009crystalball} and \textsc{DeMeter}~\cite{guo2011practical} employ techniques based on dynamic partial order reduction~\cite{lauterburg2010evaluating} to reduce the exploration state-space.
-The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing to reduce the set of interleavings that they explore.
-These techniques are complementary to our approach: \psharp's systematic testing engine could use them for reducing the number of iterations it explores, just as the other tools can use language technnology (like \psharp) to write models and reduce the complexity of the system under test.
+The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing to reduce the set of interleavings that the tool has to explore.
+These techniques are complementary to our approach: the \psharp systematic testing engine could use them for reducing the number of interleavings it explores, just as the other tools can use language technology (like \psharp) to write models and reduce the complexity of the system-under-test.
 
@@
 Verdi generates OCaml code from the verifying system, which can be used for execution.
-In contrast, \psharp performs bounded testing on a system already written in \csharp, which in our experience lowers the bar for adoption by product groups.
-Furthermore, Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed storage systems.
+In contrast, \psharp performs bounded testing on a system already written in \csharp, which in our experience lowers the bar for adoption by engineering teams.
+Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed storage systems.

### 2015-09-22 experience.tex
minor updates in case studies

@@
 
-\subsection{Live \Azure Table Migration}
+\subsection{Live Table Migration}
+%\subsection{Live \Azure Table Migration}
 \label{sec:cases:migration}
 
-The Live \Azure Table Migration library (MigratingTable) is designed to transparently migrate a data set between tables in an \Azure Storage service while an application is accessing this data set.
-This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test to increase confidence in the system.
+The Live Table Migration library (MigratingTable) is designed to \emph{transparently} migrate a data set between tables in a
+%The Live \Azure Table Migration library (MigratingTable) is designed to \emph{transparently} migrate a data set between tables in an
+\Azure Storage service \emph{while} an application is accessing this data set.
+This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test harness to increase confidence in the system.
 Indeed, this co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}).
-The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface specification.
+The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface (safety) specification.
 
@@
 
-MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary \Azure table.
+MigratingTable provides a \emph{virtual table} that has the same interface as an ordinary \Azure table.
 This interface is named \texttt{IChainTable}.
-The virtual table is backed by a pair of \emph{old} and \emph{new} tables, which must also implement \texttt{IChainTable}.
-A background \emph{migrator} job is responsible for moving all data from the old table to the new table.
+The virtual table is backed by a pair of \emph{old} and \emph{new} tables, which must implement \texttt{IChainTable}.
+A background \emph{migrator} job is responsible for transparently moving all data from the old table to the new table.
 Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which is intended to guarantee that when multiple application processes issue \emph{input} operations to their own MigratingTable instances with the same backend tables, the output of the MigratingTable instances is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
-The goal of the \psharp test was to verify this property.
+The goal of using \psharp was to systematically test this property.
+
+\textbf{Modeling and testing}.
+The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
+The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
+Instead they took advantage of the simplicity and near-determinism of the \texttt{IChainTable} specification to write an in-memory reference implementation, called \texttt{SpecTable}, to which the output of MigratingTable on an arbitrary input history can be compared.
+Then the test could sample from a distribution that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the test would have a positive probability of detecting the bug.
+
+The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table.
+This allows the test harness to keep a \emph{reference} \texttt{SpecTable} in sync with the virtual table by sending each operation to the reference table when the linearization point on the virtual table is reached.
+The output of the operation on the MigratingTable is then checked against its output on the reference table.
+
+The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old and new backend tables and the reference table; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
+Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends backend operations to the \texttt{Tables} machine.
+The \texttt{Service} machine sends the operation to the reference table at the linearization point and compares the results as previously described.
+
+% This paragraph can probably be removed if we are out of space. ~ Matt 2015-09-20
+\psharp captures and controls the interleaving of the backend operations.
+To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on the reference table has been completed.
 
@@
 % N.B. \texttt{SpecTable} = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
-
-\textbf{Modeling and testing}.
-The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
-The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
-Instead they took advantage of the simplicity and near-determinism of the \texttt{IChainTable} specification to write an in-memory reference implementation, called \texttt{SpecTable}, to which the output of MigratingTable on an arbitrary input history can be compared.
-Then the test could sample from a distribution that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the test would have a positive probability of detecting the bug.
-
-The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table.
-This allows the test harness to keep a \emph{reference} \texttt{SpecTable} in sync with the virtual table by sending each operation to the reference table when the linearization point on the virtual table is reached.
-The output of the operation on the MigratingTable is then checked against its output on the reference table.
 
@@
 This makes it possible to check the correctness property as the model executes.
-
-The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old and new backend tables and the reference table; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
-Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends backend operations to the \texttt{Tables} machine.
-The \texttt{Service} machine sends the operation to the reference table at the linearization point and compares the results as previously described.
-
-% This paragraph can probably be removed if we are out of space. ~ Matt 2015-09-20
-\psharp captures and controls the interleaving of the backend operations.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on the reference table has been completed.
 

### 2015-09-22 experience.tex
minor updates in case studies

@@
 
-MigratingTable provides a \emph{virtual table} that has the same interface as an ordinary \Azure table.
+MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary \Azure table.
 This interface is named \texttt{IChainTable}.
@@
 A background \emph{migrator} job is responsible for transparently moving all data from the old table to the new table.
-Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which is intended to guarantee that when multiple application processes issue \emph{input} operations to their own MigratingTable instances with the same backend tables, the output of the MigratingTable instances is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
+Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same old and new tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
 The goal of using \psharp was to systematically test this property.

### 2015-09-22 experience.tex
minor updates

@@
 
+\textbf{Specifics of the system}.
 MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary \Azure table.
 This interface is named \texttt{IChainTable}.
-The virtual table is backed by a pair of \emph{old} and \emph{new} tables, which must implement \texttt{IChainTable}.
-A background \emph{migrator} job is responsible for transparently moving all data from the old table to the new table.
+A virtual table is backed by an \emph{old} and a \emph{new} table, both of which implement \texttt{IChainTable}.
+A \emph{migrator} job is responsible for moving all data from the old table to the new table in the background.
 Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same old and new tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
@@
 
-\textbf{Modeling and testing}.
-The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent.
+\textbf{Testing approach}.
+There are two main challenges behind testing MigratingTable: the system is highly concurrent; and there are many possible input histories.
 The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
-Instead they took advantage of the simplicity and near-determinism of the \texttt{IChainTable} specification to write an in-memory reference implementation, called \texttt{SpecTable}, to which the output of MigratingTable on an arbitrary input history can be compared.
-Then the test could sample from a distribution that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the test would have a positive probability of detecting the bug.
+Instead, they created an in-memory reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, to which the MigratingTable output on an arbitrary input history can be compared.
 
+The \psharp test harness \emph{samples from a distribution} that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the \psharp systematic testing engine would have a positive probability of detecting the bug.
+
+\textbf{Modeling approach}.
 The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table.

### 2015-09-22 experience.tex
update in cite for double blind review
some fixes

@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine).}
-\label{fig:mockedmigration}
-\end{figure}
-
 \textbf{Specifics of the system}.
@@
 
-\textbf{Testing approach}.
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/mocked_migration}
+\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\label{fig:mockedmigration}
+\end{figure}
+
+\textbf{Testing challenges}.
 There are two main challenges behind testing MigratingTable: the system is highly concurrent; and there are many possible input histories.
 The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
-Instead, they created an in-memory reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, to which the MigratingTable output on an arbitrary input history can be compared.
+Instead, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, to which the MigratingTable output on an arbitrary input history can be compared.
+
+\textbf{Modeling and testing with \psharp}.
+Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable.
+The test harness consists of a collection of \texttt{Service} machines that contain identically configured instances of MigratingTable; a \texttt{Tables} machine, which contains the old and new tables, and an instance of \texttt{SpecTable}; and a \texttt{Migrator} machine that performs the background migration.
+Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which then sends these operations to the \texttt{Tables} machine.
+
+The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
+This allows the \psharp test harness to keep \texttt{SpecTable} in sync with the virtual table: the \texttt{Service} machine sends each input operation to \texttt{SpecTable} when the linearization point on the virtual table is reached.
+The output of the operation on the virtual table is then checked against its output on \texttt{SpecTable}.
 
@@
 
-\textbf{Modeling approach}.
-The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table.
-This allows the test harness to keep a \emph{reference} \texttt{SpecTable} in sync with the virtual table by sending each operation to the reference table when the linearization point on the virtual table is reached.
-The output of the operation on the MigratingTable is then checked against its output on the reference table.
-
-The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old and new backend tables and the reference table; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).
-Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends backend operations to the \texttt{Tables} machine.
-The \texttt{Service} machine sends the operation to the reference table at the linearization point and compares the results as previously described.
-
-% This paragraph can probably be removed if we are out of space. ~ Matt 2015-09-20
-\psharp captures and controls the interleaving of the backend operations.
-To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on the reference table has been completed.
+\psharp captures and controls all the interleaving in the system, and systematically explores them.
+To ensure that \texttt{SpecTable} is never observed to be out of sync with the old and new tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
 
@@
 The main system that we tested is \emph{CScale}~\cite{faleiro2012},
-a big data-stream processing system. In prior work~\cite{deligiannis2015psharp}, 
+a big data-stream processing system.
+In prior work~\cite{blind_review_1}, 
+%In prior work~\cite{deligiannis2015psharp}, 
 we used an earlier version of the Fabric model and reported bugs in sample services.

### 2015-09-22 evaluation.tex
some updates

@@
 
-For the vNext case study, the random scheduler was able to reproduce the bug in less than 10 seconds.
+For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug in less than 10 seconds.
 The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:psharp:testing} we leave the program to run for a long time before checking if the liveness property holds.
@@
 
-For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on 11 ``bugs'', including 8 \emph{organic} bugs that actually occurred in development and 3 \emph{notional} bugs (marked $*$), which are other code changes that we deemed interesting ways of making the code incorrect.
+For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect.
 The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
-The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
+The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
 We believe this is because the inputs and schedules that trigger them are too rare in our distribution.

### 2015-09-22 experience.tex
some updates

@@
 This interface is named \texttt{IChainTable}.
-A virtual table is backed by an \emph{old} and a \emph{new} table, both of which implement \texttt{IChainTable}.
+A virtual table is backed by a pair of \emph{old} and \emph{new} \emph{backend} tables, both of which implement \texttt{IChainTable}.
 A \emph{migrator} job is responsible for moving all data from the old table to the new table in the background.
-Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same old and new tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
+Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same backend tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
 The goal of using \psharp was to systematically test this property.
@@
 \psharp captures and controls all the interleaving in the system, and systematically explores them.
-To ensure that \texttt{SpecTable} is never observed to be out of sync with the old and new tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
+To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
 

### 2015-09-22 experience.tex
fix for anonymity

@@
 a big data-stream processing system.
-In prior work~\cite{blind_review_1}, 
 %In prior work~\cite{deligiannis2015psharp}, 
-we used an earlier version of the Fabric model and reported bugs in sample services.
+%we used an earlier version of the Fabric model and reported bugs in sample services.
 Supporting CScale required significant additions to the model, making it much more
@@
 However, we were able to find a \texttt{NullReferenceException}
-bug in CScale by running it on our model. The bug has been
+bug in CScale by running it against our Fabric model. The bug has been
 communicated to the developers of CScale, but we are still awaiting a

### 2015-09-22 paper.tex
updates
anonymization: azure -> mycloud, microsoft -> mycompany
update to include submission id and remove authors

@@
 
+\newcommand{\Azure}{MyCloud\xspace}
+\newcommand{\Microsoft}{MyCompany\xspace}
+
 \newcommand{\cmark}{\ding{51}}
@@
 \author{
-{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
-{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
-{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
-$^\dagger$Imperial College London, $^\diamond$Massachusetts Institute of Technology, $^\star$Microsoft\\
-} % end author
+{\rm Submission ID: \#127}} % end author
+%\author{
+%{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
+%{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
+%{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
+%$^\dagger$Imperial College London, $^\diamond$Massachusetts Institute of Technology, $^\star$Microsoft\\
+%} % end author
 
@@
 
-Our methodology has been applied to three distributed storage systems in the Microsoft Azure cloud computing platform.
+Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.
@@
 
-\section{Testing Azure Storage vNext with \psharp}
+\section{Testing \Azure Storage vNext with \psharp}
 \label{sec:method}

### 2015-09-29 methodology.tex
minor updates
updated vNext sample code

@@
     ExtMgr = new ExtentManager();
-    ExtMgr.netEngine = new MockedNetEngine(); // mock network
-    ExtMgr.isMockingTimer = true;	 // disable internal timer
+    ExtMgr.NetEngine = new MockedNetEngine(); // mock network
+    ExtMgr.IsMockingTimer = true;	 // disable internal timer
   }
 
-  [OnEvent(ENMsgEvent, DeliverExtentNodeMessage)]
-  void DeliverExtentNodeMessage() {
-    var msg = (ExtentNodeMessage)this.Payload;
+  [OnEvent(ExtentNodeMessageEvent, DeliverMessage)]
+  void DeliverMessage(ExtentNodeMessage msg) {
     // relay messages from Extent Node to Extent Manager
@@
   ...
-  [OnEvent(ExtentCopyRespEvent, ProcessExtentCopyResponse)]
-  void ProcessExtentCopyResponse() {
+  [OnEvent(ExtentCopyResponseEvent, ProcessCopyResponse)]
+  void ProcessCopyResponse(ExtentCopyResponse response) {
     // extent copy response from source replica
-    if (IsCopySucceeded(this.Payload)) {
-      var rec = GetExtentRecord(this.Payload);
+    if (IsCopySucceeded(response)) {
+      var rec = GetExtentRecord(response);
       ExtCtr.AddOrUpdate(rec); // update ExtentCenter
@@
     PSharp.Send(ExtentManagerMachine, 
-      new ENMsgEvent(), sync);
+      new ExtentNodeMessageEvent(), sync);
   }
@@
 Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine in the test harness.
-\texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice controlled by the \psharp runtime.
+\texttt{Timer} invokes the \psharp method \texttt{Nondet()}, which generates a nondeterministic choice controlled by the \psharp runtime.
 Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to its target (the \texttt{ExtentManager} or \texttt{ExtentNode} machines).
@@
   // cold state: repaired
-  [OnEvent(ENFailedEvent, ProcessExtentNodeFailure)]
   cold state Repaired {
-    void ProcessExtentNodeFailure() {
-        var node = GetExtentNode(this.Payload);
-        ExtentNodesWithReplica.Remove(node);
+    [OnEvent(ENFailedEvent, ProcessENFailure)]
+    void ProcessENFailure(ExtentNodeMachine en) {
+        ExtentNodesWithReplica.Remove(en);
         if (ReplicaCount < Harness.REPLICA_COUNT_TARGET)				
-          goto Repairing;
+          jumpto Repairing;
     }
@@
   // hot state: repairing
-  [OnEvent(ExtentRepairedEvent, ProcessExtentRepaired)]
   hot state Repairing {
-    void ProcessExtentRepaired() {
-      var node = GetExtentNode(this.Payload);
-      ExtentNodesWithReplica.Add(node);
+    [OnEvent(ExtentRepairedEvent, ProcessRepairCompletion)]
+    void ProcessRepairCompletion(ExtentNodeMachine en) {
+      ExtentNodesWithReplica.Add(en);
       if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
-        goto Repaired;
+        jumpto Repaired;
     }

### 2015-12-09 paper.tex
updated paper.tex

--- 2015-09-21 paper.tex
@@
 
+\newcommand{\Azure}{Azure\xspace}
+\newcommand{\Microsoft}{Microsoft\xspace}
+
 \newcommand{\cmark}{\ding{51}}
@@
 \definecolor{light-gray}{gray}{0.40}
-\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
-\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
-\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
-\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
-\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
-\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}
+%\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
+%\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
+%\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
+%\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
+%\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
+%\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}
 
@@
 \author{
+%{\rm Submission ID: \#127}} % end author
 {\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\

### 2015-12-16 evaluation.tex
Made some changes mostly relating to the first few comments from reviewer A.

@@
 \end{tabular}
-\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
+\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 \PTComment{executions [iterations]}.
 We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
@@
 We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-We configured the engine to perform 100,000 iterations.
+We configured the engine to perform 100,000 \PTComment{executions [iterations]}.
 The random seed for both schedulers was generated using the current time.
@@
 The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
-The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
+The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 \PTComment{executions [iterations]} (a process that required less than 30 minutes).
 We believe this is because the inputs and schedules that trigger them are too rare in our distribution.

### 2015-12-16 experience.tex
Made some changes mostly relating to the first few comments from reviewer A.

@@
 \includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine). \PTComment{Change Reference Table to SpecTable}}
 \label{fig:mockedmigration}
@@
 Instead, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, to which the MigratingTable output on an arbitrary input history can be compared.
+\PTComment{Suggested rephrasing: Instead, the developers wrote a more straightforward reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, which is backed by a single in-memory table.
+ After executing an arbitrary input history on both implementations, the outputs can be compared to ensure that MigratingTable produces the same output as \texttt{SpecTable}.}
 
@@
 
-The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
+\PTComment{This para is slightly unclear.} The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
 This allows the \psharp test harness to keep \texttt{SpecTable} in sync with the virtual table: the \texttt{Service} machine sends each input operation to \texttt{SpecTable} when the linearization point on the virtual table is reached.
@@
 
-The \psharp test harness \emph{samples from a distribution} that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the \psharp systematic testing engine would have a positive probability of detecting the bug.
+\PTComment{I re-worded this para.
+See what you think.} The \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
+Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set, then each execution of the test harness must have a non-negative probability of detecting the bug.
+\PTComment{Could re-add something about the small scope hypothesis, but it might just be noise.}
 
 \psharp captures and controls all the interleaving in the system, and systematically explores them.
-To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
+To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) \PTComment{I don't quite get this.
+Could you case-split here on whether the operation was or was not a linearization point?} the operation on \texttt{SpecTable} has been completed.
 
@@
 %\psharp takes control of the choice of input history, as well as the schedule, so both can be reproduced using a single random seed.
-Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
+Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each \PTComment{execution [iteration]} of the \psharp test.
 
@@
 %It was convenient to let \psharp control the choice of input history as well as the schedule so we could reproduce both using a single random seed.
-Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each iteration of the \psharp test.
+Then, under the \emph{small scope hypothesis} that any bug in MigratingTable leads to incorrect output for at least one input history in our distribution, we have a positive probability of detecting this incorrect output on each \PTComment{execution [iteration]} of the \psharp test.
 

### 2015-12-16 methodology.tex
Made some changes mostly relating to the first few comments from reviewer A.

@@
 
-Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine.
-They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
+Messages coming from the \texttt{ExtentNode} machines \PTComment{REMOVE?
+do \emph{not} go through the mocked network engine.
+They} are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
 The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.
@@
 The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem.
-Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debuggin outputs in each iteration.
+Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined \PTComment{fixed typo:} debugging outputs in each iteration.
 After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug.

### 2015-12-16 psharp.tex
Made some changes mostly relating to the first few comments from reviewer A.

@@
 The systematic testing engine serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler.
-The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure.
+The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of \PTComment{executions [iterations]} or time), or it hits an assertion failure.
 The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.

### 2015-12-17 experience.tex

@@
 \PTComment{I re-worded this para.
-See what you think.} The \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
-Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set, then each execution of the test harness must have a non-negative probability of detecting the bug.
-\PTComment{Could re-add something about the small scope hypothesis, but it might just be noise.}
+See what you think.} Conceptually, the \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
+Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set (the small scope hypothesis), then an execution of the test harness must have a non-negative probability of detecting the bug.
 

### 2015-12-17 overview.tex

@@
 
-The \psharp testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+The \psharp testing engine will repeatedly execute a program from start to completion \PTComment{(one execution)}, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of \PTComment{executions [iterations]} or time), or it hits a property violation.
 This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules.
@@
 
-We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+We have implemented two different schedulers \PTComment{which are responsible for making scheduling and other non-deterministic choices [REMOVE?
+inside the \psharp systematic testing engine]}: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
 It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.

### 2015-12-22 evaluation.tex
update

--- 2015-09-22 evaluation.tex
@@
 \end{tabular}
-\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations.
+\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 executions.
 We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
@@
 We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-We configured the engine to perform 100,000 iterations.
+We configured the engine to perform 100,000 executions.
 The random seed for both schedulers was generated using the current time.
@@
 The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
-The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes).
+The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes).
 We believe this is because the inputs and schedules that trigger them are too rare in our distribution.

### 2015-12-22 methodology.tex
update

--- 2015-09-29 methodology.tex
@@
 The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem.
-Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debuggin outputs in each iteration.
+Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration.
 After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug.

### 2015-12-22 motivation.tex
update

@@
 However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.
+
 This bug has proven to be difficult to identify, reproduce and troubleshoot.
-First, it requires very long executions to trigger.
-Second, an extent not being repaired is \emph{not} a property that can be easily verified.
-In practice, the developers rely on a very large time-out period to detect the bug.
+First, an extent never being repaired is \emph{not} a property that can be easily checked.
+Second, the bug seems to manifest only in very long executions.
 Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.
-
-To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing production-scale distributed storage systems.
-
-%Since this is a rather typical dilemma in the development of distributed storage systems, a general and systematic approach hopefully would be very helpful to the developers and greatly increase their productivity.
+To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.

### 2015-12-22 overview.tex
update

--- 2015-09-22 overview.tex
@@
-The goal of our work is to develop testing techniques for detecting and debugging Heisenbugs in distributed storage systems prior to deployment.
-To achieve our goal, we use \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream \csharp language that provides: (i) language support for \emph{modeling the environment} of distributed systems developed using Microsoft's .NET framework; and (ii) a \emph{testing engine} that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.
+The goal of our work is to detect distributed system bugs \emph{before} they reach production.
+Distributed systems typically consist of multiple components that interact with each other via \emph{message passing}.
+If these messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs.
+To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, failures and timeouts.
+If the \psharp testing engine identifies a sequence of events that triggers a bug (witnessed by an assertion failure), then exploration terminates, and a trace is reported.
+
+Due to the highly asynchronous nature of distributed storage systems, the number of possible states that these systems can reach is exponentially large.
+Tools such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug} focus on testing unmodified distributed systems, but this can easily lead to state-space explosion when trying to exhaustively explore the entire state-space of a production-scale system, such as the Azure Storage vNext.
+On the other hand, techniques such as TLA+~\cite{lamport1994temporal} have successfully verified specifications of distributed systems, but they are unable to verify the actual implementation.
+In this work, we are proposing a solution between the two extremes: systematically test the implementation of a distributed system component against its modeled by \psharp environment.
+The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space.
+Testing with our approach does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp.
+However, we argue that developers already have to build test suites and spend significant time testing a distributed system prior to deployment.
+Using \psharp might require a bit more effort and time, but pays dividends in found bugs (see \S\ref{sec:eval}).
+
+\subsection{A small example}
+\label{sec:overview:example}
+
+Draft.
+
+\subsection{Modeling with \psharp}
+\label{sec:overview:psharp}
 
@@
 The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
-(see \S\ref{sec:psharp:testing}).
+(see \S\ref{sec:overview:testing}).
 
 \item The criteria for correctness of an execution must be specified.
-Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:bg:bugs}).
+Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:bugs}).
 Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state.
@@
 %\item Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs.
-This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:psharp:testing}).
+This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:overview:testing}).
 \end{enumerate}
@@
 
-\subsection{Specifications}
-\label{sec:bg:bugs}
+\subsection{Writing specifications in \psharp}
+\label{sec:overview:bugs}
 
@@
 
-\subsection{Systematic testing}
-\label{sec:psharp:testing}
+\subsection{Systematic testing using \psharp}
+\label{sec:overview:testing}
 
 The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:bg:bugs}.
+A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:overview:bugs}.
 
@@
 
-The \psharp testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation.
+The \psharp testing engine will repeatedly execute a program from start to completion (one execution), each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of executions or time), or it hits a property violation.
 This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules.
@@
 
-We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
 It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.
@@
 
-\textbf{Handling \csharp 5.0 asynchrony}.
-Our case studies heavily use the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
+%\textbf{Handling \csharp 5.0 asynchrony}.
+%Our case studies heavily use the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
 An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread.
@@
 To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.
-
-%The \texttt{async} and \texttt{await} \csharp 5.0 primitives were not originally supported in \psharp.
-To be able to test our case studies, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).
-
-Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
-%Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon),
-All TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler.
+%
+%%The \texttt{async} and \texttt{await} \csharp 5.0 primitives were not originally supported in \psharp.
+To be able to test our case studies, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:overview:testing}).
+%
+%Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
+%%Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon),
+%All TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler.
 These tasks will be intercepted by our custom task scheduler and wrapped inside a special (not available to the programmer) task-machine.

### 2015-12-22 paper.tex
Made some changes mostly relating to the first few comments from reviewer A.
slight add in abstract
update

@@
 \definecolor{light-gray}{gray}{0.40}
+\definecolor{orange}{rgb}{1,0.5,0}
 %\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
@@
 %\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
-%\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}
+\newcommand{\PTComment}[1]{\textcolor{orange}{Paul: #1}}
 
@@
 \title{\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)}
-%\title{\Large \bf Uncovering Distributed Storage System Bugs during\\Testing (not in Production!)}
 
@@
 \subsection*{Abstract}
-Testing distributed storage systems is challenging due to multiple sources of nondeterminism.
-Conventional testing techniques, such as unit and stress testing, are ineffective in preventing serious but subtle bugs from reaching production.
+Testing distributed systems is challenging due to multiple sources of nondeterminism.
+Conventional testing techniques, such as unit, integration and stress testing, are ineffective in preventing serious but subtle bugs from reaching production.
 Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code.
-In this paper, we present a new methodology for testing distributed storage systems.
+In this paper, we present a new methodology for testing distributed systems.
 Our approach applies advanced systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.
 
-Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure platform.
+Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure cloud computing platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.
@@
 
-\section{Our Approach to Testing Distributed Systems}
+\section{Overview of our Approach}
+%\section{Our Approach to Testing Distributed Systems}
 \label{sec:overview}

### 2015-12-22 overview.tex
update

@@
 To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, failures and timeouts.
-If the \psharp testing engine identifies a sequence of events that triggers a bug (witnessed by an assertion failure), then exploration terminates, and a trace is reported.
+
+Modeling using \psharp involves two core activities.
+First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched using the \psharp \texttt{Send} method.
+Such modification is typically non-invasive, as it can be performed with dynamic dispatch, a \csharp language feature already used for unit testing.
+Second, the developer must write a \psharp test harness that nondeterministically triggers various distributed events, such as client requests, failures and timeouts.
+The purpose of these two modeling activities is to explicitly declare all possible sources of nondeterminism in the distributed system using \psharp.
+
+During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during the modeling process, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be evaluated.
+The \psharp systematic testing engine will serialize the system (in single-box setup), and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a bound (e.g.
+in number of executions or time), or it hits a safety or liveness property violation (see \S\ref{sec:overview:bugs}).
+This testing process is fully automatic, has no false-positives (assuming an accurate test harness), and can reproduce found bugs by replaying buggy schedules.
+After a bug is found, the \psharp testing engine generates a trace that represents the buggy schedule.
+In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events, and thus is easier to debug.
 

### 2015-12-22 overview.tex
update

@@
 Testing with our approach does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp.
-However, we argue that developers already have to build test suites and spend significant time testing a distributed system prior to deployment.
-Using \psharp might require a bit more effort and time, but pays dividends in found bugs (see \S\ref{sec:eval}).
+However, we argue that developers already spend significant time in building test suites and testing a distributed system prior to deployment.
+The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).
 

### 2015-12-23 overview.tex
update

@@
-The goal of our work is to detect distributed system bugs \emph{before} they reach production.
+The goal of our work is to detect distributed system bugs \emph{before} they reach production and effect users.
 Distributed systems typically consist of multiple components that interact with each other via \emph{message passing}.
-If these messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs.
-To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, failures and timeouts.
+If messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs.
+To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, client requests, failures and timeouts.
 
 Modeling using \psharp involves two core activities.
-First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched using the \psharp \texttt{Send} method.
-Such modification is typically non-invasive, as it can be performed with dynamic dispatch, a \csharp language feature already used for unit testing.
-Second, the developer must write a \psharp test harness that nondeterministically triggers various distributed events, such as client requests, failures and timeouts.
-The purpose of these two modeling activities is to explicitly declare all possible sources of nondeterminism in the distributed system using \psharp.
+First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched through the \psharp \texttt{Send} method.
+Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing.
+Second, the developer must write a \psharp test harness that nondeterministically triggers distributed events.
+The purpose of these two activities is to explicitly declare all possible sources of nondeterminism in the distributed system using \psharp.
 
@@
 Due to the highly asynchronous nature of distributed storage systems, the number of possible states that these systems can reach is exponentially large.
-Tools such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug} focus on testing unmodified distributed systems, but this can easily lead to state-space explosion when trying to exhaustively explore the entire state-space of a production-scale system, such as the Azure Storage vNext.
+Tools such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug} focus on testing unmodified distributed systems, but this can easily lead to state-space explosion when trying to exhaustively explore the entire state-space of a production-scale distributed storage system, such as the Azure Storage vNext.
 On the other hand, techniques such as TLA+~\cite{lamport1994temporal} have successfully verified specifications of distributed systems, but they are unable to verify the actual implementation.
-In this work, we are proposing a solution between the two extremes: systematically test the implementation of a distributed system component against its modeled by \psharp environment.
+
+In this work, we are proposing a solution between the above two extremes: systematically test the implementation of a distributed system component against its modeled by \psharp environment.
 The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space.

### 2015-12-23 experience.tex
update
update in case studies

@@
 
-\subsection{Live Table Migration}
-%\subsection{Live \Azure Table Migration}
+\subsection{Live \Azure Table Migration}
 \label{sec:cases:migration}
 
-The Live Table Migration library (MigratingTable) is designed to \emph{transparently} migrate a data set between tables in a
-%The Live \Azure Table Migration library (MigratingTable) is designed to \emph{transparently} migrate a data set between tables in an
-\Azure Storage service \emph{while} an application is accessing this data set.
-This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test harness to increase confidence in the system.
-Indeed, this co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}).
-The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface (safety) specification.
+Live \Azure Table Migration (MigratingTable) is a library designed to \emph{transparently} migrate a data set between tables in an \Azure Storage service \emph{while} an application is accessing this data set.
+This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness.
+This co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}).
+The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on a single liveness property, while the MigratingTable test harness checks complete compliance with an interface (safety) specification.
 
-\textbf{Specifics of the system}.
 MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary \Azure table.
@@
 \includegraphics[width=\linewidth]{img/mocked_migration}
-\caption{Environment model of MigratingTable (each box with a dotted line represents one \psharp machine). \PTComment{Change Reference Table to SpecTable}}
+\caption{The \psharp test harness of MigratingTable (each box with a dotted line represents one \psharp machine).}
 \label{fig:mockedmigration}
@@
 
-\textbf{Testing challenges}.
-There are two main challenges behind testing MigratingTable: the system is highly concurrent; and there are many possible input histories.
+There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) there are many possible input histories.
 The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
-Instead, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, to which the MigratingTable output on an arbitrary input history can be compared.
-\PTComment{Suggested rephrasing: Instead, the developers wrote a more straightforward reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, which is backed by a single in-memory table.
- After executing an arbitrary input history on both implementations, the outputs can be compared to ensure that MigratingTable produces the same output as \texttt{SpecTable}.}
+Instead, the developers created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, which is backed by a single in-memory table.
+After executing an arbitrary input history on both implementations, the outputs can be compared to ensure that MigratingTable produces the same output as \texttt{SpecTable}.
 
-\textbf{Modeling and testing with \psharp}.
 Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable.
-The test harness consists of a collection of \texttt{Service} machines that contain identically configured instances of MigratingTable; a \texttt{Tables} machine, which contains the old and new tables, and an instance of \texttt{SpecTable}; and a \texttt{Migrator} machine that performs the background migration.
+The harness consists of a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; a \texttt{Tables} machine, which contains the old and new tables, and an instance of \texttt{SpecTable}; and a \texttt{Migrator} machine that performs the background migration.
 Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which then sends these operations to the \texttt{Tables} machine.
@@
 
-\PTComment{I re-worded this para.
-See what you think.} Conceptually, the \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
-Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set (the small scope hypothesis), then an execution of the test harness must have a non-negative probability of detecting the bug.
+Conceptually, the \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
+Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set (the small scope hypothesis), then an execution of the test harness must have a non-zero probability of detecting the bug.
 

### 2015-12-23 evaluation.tex
various overall updates

@@
 Table~\ref{tab:stats} presents program statistics for the three case studies.
-The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp Test Harness'' refer to the environment model written in \psharp.
+The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp Test Harness'' refer to the test harness written in \psharp.
 We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system-under-test (\#B); lines of \psharp code for the test harness (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).
@@
 
-We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-We configured the engine to perform 100,000 executions.
+We performed all experiments on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
+We configured the \psharp testing engine to perform 100,000 executions.
 The random seed for both schedulers was generated using the current time.
@@
 For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug in less than 10 seconds.
-The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:psharp:testing} we leave the program to run for a long time before checking if the liveness property holds.
+The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:testing} we leave the program to run for a long time before checking if the liveness property holds.
 The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.

### 2015-12-23 overview.tex
various overall updates

@@
 
-The \psharp testing engine will repeatedly execute a program from start to completion (one execution), each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of executions or time), or it hits a property violation.
-This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules.
-After a bug is found, the engine generates a trace that represents the buggy schedule.
-In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events and, thus, is easier to debug.
-
 We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.

### 2015-12-23 related.tex
various overall updates

@@
-Model checking~\cite{godefroid1997verisoft} is a powerful technique that has been widely used in the past for finding tricky bugs in the actual implementation of distributed systems~\cite{killian2007life, yang2009modist, yabandeh2009crystalball, guerraoui2011model, guo2011practical, simsa2011dbug, leesatapornwongsa2014samc}.
+Most related to our work is model checking~\cite{godefroid1997verisoft}, a powerful technique that has been widely used in the past for finding Heisenbugs in the actual implementation of distributed systems~\cite{killian2007life, yang2009modist, yabandeh2009crystalball, guerraoui2011model, guo2011practical, simsa2011dbug, leesatapornwongsa2014samc}.
 
-State-of-the-art model checkers, such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug}, typically focus on testing entire, often unmodified, distributed systems, an approach that easily leads to state-space explosion.
-In contrast, we try to offer a more pragmatic approach for handling state-space explosion: \psharp provides the capability for creating models of individual components of a large system, which helps towards component isolation and enables compositional testing.
-Our approach aims to enhance unit and integration testing, techniques widely used in production, where only individual (or a small number of) components are tested at each time.
+State-of-the-art model checkers, such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug}, typically focus on testing entire, often \emph{unmodified}, distributed systems, an approach that easily leads to state-space explosion.
+\textsc{DeMeter}~\cite{guo2011practical}, built on top of \textsc{MoDist}, aims to reduce the state-space when exploring unmodified distributed systems.
+\textsc{DeMeter} explores individual components of a large system in isolation, and then dynamically extracts interface behavior between components to perform a global exploration.
+In contrast, we try to offer a more pragmatic approach for handling state-space explosion.
+We first \emph{partially} model a distributed system using \psharp.
+Then, we systematically test the actual implementation of each system component against its \psharp test harness.
+Our approach aims to enhance unit and integration testing, techniques widely used in production, where only individual or a small number of components are tested at each time.
 
-Tools such as CrystalBall~\cite{yabandeh2009crystalball} and \textsc{DeMeter}~\cite{guo2011practical} employ techniques based on dynamic partial order reduction~\cite{lauterburg2010evaluating} to reduce the exploration state-space.
-The SAMC~\cite{leesatapornwongsa2014samc} model checker offers a way of incorporating application-specific information during systematic testing to reduce the set of interleavings that the tool has to explore.
-These techniques are complementary to our approach: the \psharp systematic testing engine could use them for reducing the number of interleavings it explores, just as the other tools can use language technology (like \psharp) to write models and reduce the complexity of the system-under-test.
+SAMC~\cite{leesatapornwongsa2014samc} offers a way of incorporating application-specific information during systematic testing to reduce the set of interleavings that the tool has to explore.
+Such techniques based on partial-order reduction~\cite{godefroid1996partial, flanagan2005dynamic} are complementary to our approach: \psharp could use them to reduce the exploration state-space.
+Likewise, other tools can use language technology like \psharp to write models and reduce the complexity of the system-under-test.
 
-% Further, note that we operate at a different scale by focusing on compositional testing.
-For instance, SAMC reports orders of magnitude speedup over previous testing techniques, but executing 100,000 runs still took ``approximately 48 full machine days''~\cite{leesatapornwongsa2014samc}.
-In our case studies, we execute a similar number of iterations in a matter of hours, and find serious bugs in seconds (see \S\ref{sec:eval}).
-This does \emph{not} imply that \psharp is faster (nor we attempt to make that claim); but simply that we explore behaviors at a different level.
+\textsc{MaceMc}~\cite{killian2007life} is a model checker for distributed systems written in the \textsc{Mace}~\cite{killian2007mace} language.
+The focus of \textsc{MaceMc} is to find liveness property violations using an algorithm based on depth-bounded random walk and the use of heuristics.
+Because \textsc{MaceMc} can only test systems written in \textsc{Mace}, it cannot be easily used in an industrial setting.
+In contrast, \psharp can be applied on legacy code written in \csharp, a mainstream language.
 
-%as aforementioned, we focus on compositional testing, whereas prior work typically focuses on testing an entire system. %We are just trying to solve a similar problem at a very different level.
-
-\textsc{MaceMc}~\cite{killian2007life} is a model checker for distributed systems implemented in the \textsc{Mace} language.
-The focus of \textsc{MaceMc} is to find liveness property violations using an algorithm based on bounded random walk and the use of heuristics.
-\psharp differs from \textsc{MaceMc} in that it can be applied on legacy code written in a mainstream language, whereas a system to be tested with \textsc{MaceMc} has to be written in \textsc{Mace}, which makes \textsc{MaceMc} harder to be applied in an industrial setting.
-
-\textsc{Fate} and \textsc{Destini} is a framework for systematically injecting (combination of) failures in distributed systems~\cite{gunawi2011fate}.
-This framework focuses in exercising failure scenarios, whereas \psharp can be used for testing generic safety and liveness properties of distributed systems.
-\psharp also provides language extensions and modeling capabilities that further differentiate it from prior work.
+\textsc{Fate} and \textsc{Destini} is a framework for systematically injecting failures in distributed systems~\cite{gunawi2011fate}.
+This framework focuses in exercising various failure scenarios, whereas \psharp can be used for testing generic safety and liveness properties of distributed systems.
 

### 2015-12-23 motivation.tex
overall updates

@@
 Similar to the current system, vNext employs containers, called \emph{extents}, to store data.
-Extents are typically several gigabytes each, consisting of many data blocks, and replicated over multiple \emph{Extent Nodes} (ENs).
-However, in contrast to the current system, which employs a Paxos-based, centralized mapping from extents to ENs, vNext achieves its scalability target by employing a \emph{distributed mapping}.
+Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs).
+However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}.
 In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr).

### 2015-12-25 evaluation.tex
major update to address PC and shepherd comments

@@
 \end{tabular}
-\caption{Statistics from modeling the environment of the three \Microsoft MyCloud-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
+\caption{Statistics from modeling the environment of the three Microsoft MyCloud-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
 \label{tab:stats}
-%\vspace{-3mm}
+\vspace{-3mm}
 \end{table}
@@
 
-Modeling the environment of the Extent Manager in the \Azure Storage vNext system required approximately two person weeks of part-time developing.
+Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two person weeks of part-time developing.
 The \psharp test harness for this system is the smallest (in lines of code) from the three case studies.
@@
 
-Developing both the MigratingTable system and its \psharp test harness took approximately five person weeks.
-The \psharp test harness was developed in parallel with the actual system.
-This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and after the development process had finished.
+Developing both MigratingTable and its \psharp test harness took approximately five person weeks.
+The harness was developed in parallel with the actual system.
+This differs from the other two case studies, where the modeling activity occurred independently and after the development process had finished.
 
@@
 According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner.
-After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one to evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.
+After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one to evaluate the effectiveness of different testing strategies.
 
@@
 We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
-The CS column shows which case study corresponds to each bug: ``1'' is for the \Azure Storage vNext; and ``2'' is for MigratingTable.
-We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:cases:fabric}).
+The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
+We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:fabric}).
+
+We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.
+The interface exposes callbacks that are invoked by the \psharp testing engine for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.
 
@@
 For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug in less than 10 seconds.
-The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:testing} we leave the program to run for a long time before checking if the liveness property holds.
+The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
 The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.
 
-For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect.
+For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect.
 The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.

### 2015-12-25 intro.tex
major update to address PC and shepherd comments

--- 2015-09-21 intro.tex
@@
 Our methodology is different from prior approaches that required developers to either switch to an unfamiliar domain specific language~\cite{killian2007life, desai2015building}, or manually annotate and instrument their code~\cite{simsa2011dbug}.
-Instead, we allow developers to test \emph{production code} by writing test harnesses in \csharp, a mainstream programming language.
+Instead, we allow developers to systematically test \emph{production code} by writing test harnesses in \csharp, a mainstream programming language.
 This significantly lowered the acceptance barrier for adoption by the Microsoft Azure team.
 
-%We present a new method for testing an implementation of a distributed system and uncovering bugs before the system is released in the wild.
-Our methodology is based on \psharp~\cite{deligiannis2015psharp}, an extension of \csharp that provides support for modeling, specification, and systematic testing of distributed systems written in Microsoft's .NET framework.
-To test a distributed system with \psharp, the programmer augments the original system code with three artifacts -- a model of the nondeterministic execution environment of the system, a concurrent test harness that drives the system towards interesting behaviors, and safety or liveness specifications.
-The \psharp testing engine then systematically explores all behaviors exercised by the test harness and validates them against the provided specifications.
+Our testing methodology is based on \psharp~\cite{deligiannis2015psharp}, an extension of \csharp that provides support for modeling, specification, and systematic testing of distributed systems written in the Microsoft .NET framework.
+To use \psharp for testing, the programmer has to augment the original system with three artifacts: a model of the nondeterministic execution environment of the system; a test harness that drives the system towards interesting behaviors; and safety or liveness specifications.
+\psharp then systematically exercises the test harness and validates program behaviours against the provided specifications.
 
-We have applied \psharp to three distributed storage systems inside Microsoft: Azure Storage vNext; Azure Table Live Migration; and Azure Service Fabric.
-%a distributed storage system; a system for migrating data across partitions; and a big-data computing system.
-We uncovered numerous bugs in these systems, including, most notably, a subtle liveness bug that only intermittently manifested during stress testing for months without being fixed.
+The original \psharp paper~\cite{deligiannis2015psharp} discussed language design issues and data race detection for programs written in \psharp, whereas this work focuses on using \psharp to test three distributed storage systems inside Microsoft: Azure Storage vNext; Live Table Migration; and Azure Service Fabric.
+We uncovered numerous bugs in these systems, including a subtle liveness bug that only intermittently manifested during stress testing for months without being fixed.
 Our testing approach uncovered this bug in a very small setting, which made it easy for developers to examine traces, identify, and fix the problem.
-%Our experience demonstrates that bugs in production do not need a large setting to be reproduced; they can be reproduced in small settings with easy to understand traces.
 
@@
 \begin{itemize}
-\item We present a new methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms in \psharp.
-\item Because \psharp is an extension of \csharp, our infrastructure can test production code written in \csharp.
-\item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
-\item We show that \psharp can reproduce bugs in production-scale distributed systems in a small setting with easy to understand traces.
+\item We present a new methodology for modeling, specifying properties of correctness, and systematically testing real distributed systems with \psharp.
+\item We discuss our experience of using \psharp to test three distributed storage systems that are used inside Microsoft, finding subtle bugs that could not be found with traditional testing techniques.
+\item We evaluate the cost and benefits of using our approach, and show that \psharp can detect bugs in a small setting and with easy to understand traces.
 \end{itemize}

### 2015-12-25 migrating.tex
major update to address PC and shepherd comments

--- 2015-12-23 experience.tex
@@
-The modeling and testing approach described in the earlier sections of the paper is not specific to the vNext system.
-\psharp is a generic framework for .NET distributed systems.
-We showcase this capability by presenting developer experience of using \psharp to model and test two other systems used in production inside \Microsoft.
-
-\subsection{Live \Azure Table Migration}
-\label{sec:cases:migration}
-
-Live \Azure Table Migration (MigratingTable) is a library designed to \emph{transparently} migrate a data set between tables in an \Azure Storage service \emph{while} an application is accessing this data set.
+Live Table Migration (MigratingTable) is a library designed to \emph{transparently} migrate a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
 This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness.
@@
 
-MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary \Azure table.
+MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary Azure table.
 This interface is named \texttt{IChainTable}.
@@
 \centering
-\includegraphics[width=\linewidth]{img/mocked_migration}
+\includegraphics[width=\linewidth]{img/modeled_migration}
 \caption{The \psharp test harness of MigratingTable (each box with a dotted line represents one \psharp machine).}
@@
 
-\PTComment{This para is slightly unclear.} The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
+The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
 This allows the \psharp test harness to keep \texttt{SpecTable} in sync with the virtual table: the \texttt{Service} machine sends each input operation to \texttt{SpecTable} when the linearization point on the virtual table is reached.
@@
 \psharp captures and controls all the interleaving in the system, and systematically explores them.
-To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) \PTComment{I don't quite get this.
-Could you case-split here on whether the operation was or was not a linearization point?} the operation on \texttt{SpecTable} has been completed.
+To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
 
@@
 Thus, the use of these proxies as \texttt{IChainTable} backends is transparent to the MigratingTable library, thanks to dynamic dispatch.
-
-\subsection{\Azure Service Fabric}
-\label{sec:cases:fabric}
-
-\Azure Service Fabric
-%\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}}
-(or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
-%The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
-In order to make the user-written service \emph{reliable}, Fabric launches several \emph{replicas} (copies) of the service, where each replica runs as a separate process on a different node in the cluster.
-%\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
-One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}. The primary replicates state changes to the secondaries 
-%by sending them \emph{replication requests}.
-so that all replicas eventually have the same state. 
-If the primary fails,
-% (e.g.\ if the node on which the primary is running crashes), 
-Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive 
-%a full or partial copy (depending on whether persistent storage is used) 
-a copy
-of the state of the new primary in order to ``catch up'' with the other secondaries. 
-%Fabric provides a name-resolution service so that clients can always find the current primary.
-%The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
-User-written Fabric services are complex asynchronous and distributed applications, and are thus challenging to test.
-% They are interesting targets for systematic testing with \psharp{}.
-
-Our primary goal was to create a \psharp{} model of Fabric to allow
-thorough testing of services, where Fabric's asynchrony is controlled 
-by the \psharp{} runtime.
-The model is written once
-to include all behaviours of Fabric, including simulating failures and recovery,
-so that it can be reused repeatedly to test multiple Fabric services.
-This was the largest modeling exercise among the case studies but the cost
-was amortized across testing multiple servies. This is analogous to the work
-on driver verification \cite{ball2011slam} where the cost of developing a model
-of the Windows kernel was amortized across testing multiple device drivers.
-Note that we model the lowest Fabric API layer (\texttt{Fabric.dll})
-which is currently not documented for use externally.
-We target internally-developed services that invoke the lowest layer.
-
-\psharp's systematic testing was very helpful in debugging the model itself.
-% Without systematic testing, we feel even catching bugs in the model 
-% would have been difficult.
-In order to perform systematic testing on our model,
-we wrote a simple service in \psharp{}
-that runs on our Fabric model.
-%made up of a single machine
-% and executed it on the Fabric model,
-% thus yielding a pure \psharp{} system.
-We tested a scenario where the primary replica
-fails at some non-deterministic point
-within the execution.
-One such bug that we found
-involved
-the primary replica
-failing
-when a secondary replica
-was about to receive a copy of the state;
-the secondary was then elected to be the new primary
-and yet, because the secondary stopped waiting for
-a copy of the state, it was
-also promoted to be an \emph{active} secondary
-(one that has caught up with the other secondaries).
-This caused an assertion failure in our model
-allowing us to detect and fix this incorrect behaviour.
-
-%\PTComment{Cut: prior work created a Fabric model\ldots}
-% Prior work~\cite{deligiannis2015psharp}
-% created a model of Fabric with limited functionality;
-% it used a mixture of \csharp{} and \psharp{} internally,
-% only supported one in-flight replication request
-% (which restricts the asynchrony that can be tested),
-% and only supported one Fabric service.
-% Our new Fabric model was re-written to use only \psharp{}
-% internally,
-% support an arbitrary number of Fabric services and in-flight replication requests,
-% and in general be a more complete model of Fabric.
-% Note that \csharp{} code is still required to interface with the existing
-% \csharp{} service code.
-% We refer to our \csharp{} code
-% as the \emph{translation layer}
-% and
-%  the user-written \csharp{} service code as \emph{user code}.
-
-%\PTComment{Cut: Figure of model.}
-% \begin{figure}[thb]
-% \centering
-% \includegraphics[width=\linewidth]{img/fabricmodel}
-% \caption{Overview of the key machines and interfaces in our Fabric model.}
-% \label{fig:fabric_model}
-% \end{figure}
-
-%\PTComment{Cut: overview of model.}
-% An overview of our Fabric model is shown in Figure~\ref{fig:fabric_model}.
-% The \texttt{ClusterRuntime} machine 
-% handles the creation and management of 
-% one or more Fabric services,
-% as well as service resolution requests
-% which allows for client-service and inter-service communication
-% within the model.
-% Each Fabric service instance is managed by a \texttt{ServiceRuntime}
-% machine, which in turn manages 
-% several \texttt{ReplicaRuntime} machines.
-% Each \texttt{ReplicaRuntime} communicates with the user code
-% via several machines and interfaces from the translation layer
-% (only the translation layer for the primary is shown,
-% but every \texttt{ReplicaRuntime} has its own instance of the translation layer).
-% Note that communication between machines is hierachical;
-% thus, communication between \texttt{ReplicaRuntime}s
-% (such as the sending of replication requests)
-% is via the \texttt{ServiceRuntime} machine for that service.
-% This approach does not necessarily reflect how Fabric works in practice.
-% Instead, we chose an architecture
-% that keeps the model simple
-% while still allowing (what we believe to be) realistic
-% asynchrony and failure scenarios.
-
-%\PTComment{Cut: Description of how replication works in the model, including how we modeled at a fine granularity.}
-% \textbf{Replication example:} In order to replicate a state-mutating operation,
-% user code at the primary replica 
-% calls \texttt{IStateReplicator.ReplicateAsync},
-% passing the serialized operation
-% object.
-% The operation is sent to the \texttt{ServiceRuntime},
-% where it is assigned a \emph{logical sequence number} (LSN);
-% each operation is assigned a consecutive LSN
-% to track the total-order in which operations should be applied.
-% The LSN is sent back to the primary replica
-% where it is returned from the \texttt{ReplicateAsync} call,
-% along with a \texttt{Task} object that will ``complete'' once
-% the operation has been replicated to a majority of
-% secondaries;
-% thus, the user code can wait on the \texttt{Task}
-% before confirming to any clients that the request has been applied reliably.
-% The \texttt{ServiceRuntime} adds the operation to its list of in-flight
-% replication requests and sends $n$ events to itself to signal that the request
-% must be sent to a replica, where $n$ is the number of secondary replicas.
-% The reason for sending $n$ events to itself instead of simply sending events
-% directly to each secondary is so that the \texttt{ServiceRuntime}
-% can process a simulated failure event inbetween the sending of replica requests
-% to each secondary.
-% This is an example of where we carefully considered
-% the granularity of actions so that we could 
-% model failures appropriately.
-% The user code at a secondary receives the operation,
-% applies it and then calls \texttt{Acknowledge} on the operation object;
-% we implement this to send an event to the \texttt{ReplicaRuntime}
-% which forwards the acknowledgement to the \texttt{ServiceRuntime}.
-% Once a majority of secondaries have acknowledged, the
-% \texttt{ServiceRuntime} removes the replication request from its list
-% of replication requests
-% and sends an acknowledgement to the primary,
-% where the previously returned \texttt{Task} completes.
-
-%\PTComment{Cut: We reverse-engineered Fabric as needed.}
-% \subsubsection{Fabric model correctness}
-% Our model does not attempt to
-% simulate the internals of Fabric accurately,
-% as its purpose is to find bugs in user code
-% and not in Fabric itself (which we assume to be correct).   
-% However,
-% due to lack of documentation,
-% it is not always clear how Fabric should behave
-% in certain scenarios.
-% Thus,
-% we ran several variants of a simple Fabric service
-% that logs calls into the user code in order to
-% reverse-engineer the actual behaviour of Fabric;
-% we ensured that our model has the same behaviour,
-% although this is an ongoing process as we encounter
-% additional scenarios.
-
-%\PTComment{Cut: We used systematic testing to find bugs in our model!}
-% A further problem is that our model may contain bugs.
-% In order to find bugs in our model effectively,
-% we wrote a \psharp{} service made up of a single machine
-% which takes the place
-% of the user code and translation layer in Figure \ref{fig:fabric_model},
-% for each \texttt{ReplicaRuntime}.
-% Thus, we were able to run this pure \psharp{} system
-% under \psharp{}'s systematic testing mode
-% and uncover many assertion failures within our model.
-% We tested a scenario where the primary fails at some non-deterministic point
-% within the execution.
-
-%\PTComment{Cut: Example bug found in our model plus the fix.}
-% \textbf{Example Fabric model assertion failure:}
-% In the buggy trace,
-% the \texttt{ServiceRuntime}
-% sends an \texttt{EEpochInfo} event to the second
-% \texttt{ReplicaRuntime}
-% indicating that this is the first epoch and the
-% replica will be a secondary.
-% An \emph{epoch} represents a configuration of primary and
-% secondary replicas; when a different replica becomes the primary,
-% this indicates the start of a new epoch.
-% The \texttt{ReplicaRuntime} acknowledges that it has become
-% a secondary by responding with the same event type.
-% The \psharp{} service sends an \texttt{ESecondaryCopyContextOp};
-% this indicates what state the secondary has
-% and, thus, what the primary should send to this secondary so that it can catch
-% up.
-% The \texttt{ESecondaryCopyContextOp} event is forwarded to the
-% \texttt{ServiceRuntime}. 
-% The \texttt{ServiceRuntime} then receives and handles an \texttt{EKillPrimary}
-% event, which causes the second replica to become the new primary.
-% Thus,
-% the \texttt{ServiceRuntime} sends another \texttt{EEpochInfo}
-% to the second \texttt{ReplicaRuntime}
-% indicating that this is the second epoch
-% and the replica will be a primary.
-% As part of this change,
-% the \texttt{ReplicaRuntime}
-% sends an event to the \psharp{} service
-% indicating that it should stop waiting for
-% the state to be copied from the old primary to this replica,
-% which is acknowledged by sending an event to the \texttt{ReplicaRuntime}.
-% This event causes the \texttt{ReplicaRuntime}
-% to send an \texttt{ESecondaryCopyStateDone}
-% event to the \texttt{ServiceRuntime},
-% which unfortunately responds with an event indicating that the
-% \texttt{ReplicaRuntime} is now an \emph{active} secondary
-% (i.e.\ a secondary that has caught up with the primary).
-% However, this causes an assertion failure because
-% the \texttt{ReplicaRuntime} is becoming a primary and, thus,
-% cannot be a secondary.
-% Our fix to this bug was to ensure that the 
-% \texttt{ESecondaryCopyStateDone} event was marked as part of the first
-% epoch (as the \texttt{ReplicaRuntime} had not yet acknowledged the
-% change to primary); thus, 
-% the \texttt{ServiceRuntime} ignores the event and does not try to make the
-% replica an active secondary.
-
-% \subsubsection{CScale}
-
-The main system that we tested is \emph{CScale}~\cite{faleiro2012},
-a big data-stream processing system.
-%In prior work~\cite{deligiannis2015psharp}, 
-%we used an earlier version of the Fabric model and reported bugs in sample services.
-Supporting CScale required significant additions to the model, making it much more
-feature-complete.
-CScale chains multiple Fabric services, which communicate via remote procedure calls (RPCs).
-To close the system, we modeled RPCs 
-by sending and receiving \psharp{} events.
-This was enough to converted a distributed system
-that uses both Fabric and its own
-network communication protocol
-% that runs on the Fabric platform
-% and 
-% uses network communication
-into 
-a closed single process system. 
-% that uses multiple threads,
-% contains
-% the Fabric model
-% and the CScale services,
-% and does not use network communication.
-%\PTComment{Cut: we had to remove static fields.}
-% Additional changes were needed to
-% remove certain static (per-process) fields,
-% as these were inadvertently
-% being shared between services
-% after making the system a single process.
-A key challenge in our work
-was to test CScale despite the fact that it
-uses various synchronous and asynchronous APIs
-other than \psharp{}.
-This work is still in-progress.
-However, we were able to find a \texttt{NullReferenceException}
-bug in CScale by running it against our Fabric model. The bug has been
-communicated to the developers of CScale, but we are still awaiting a
-confirmation.

### 2015-12-25 overview.tex
major update to address PC and shepherd comments

@@
-The goal of our work is to detect distributed system bugs \emph{before} they reach production and effect users.
-Distributed systems typically consist of multiple components that interact with each other via \emph{message passing}.
+The goal of our work is to find distributed system Heisenbugs \emph{before} they reach production.
+Typical distributed systems consist of multiple components that interact with each other via \emph{message passing}.
 If messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs.
-To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, client requests, failures and timeouts.
+To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{specifying} properties of \emph{correctness}, and \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, client requests, failures and timeouts.
 
-Modeling using \psharp involves two core activities.
-First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched through the \psharp \texttt{Send} method.
+Modeling using \psharp involves three core activities.
+First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched through the \texttt{PSharp.Send(...)} method.
 Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing.
-Second, the developer must write a \psharp test harness that nondeterministically triggers distributed events.
-The purpose of these two activities is to explicitly declare all possible sources of nondeterminism in the distributed system using \psharp.
+Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various distributed events (see \S\ref{sec:overview:model}).
+The harness is essentially an environment model of the system.
+The purpose of these first two activities is to explicitly declare all sources of nondeterminism in the system using \psharp.
+Finally, the developer must specify the criteria for correctness of an execution of the system-under-test.
+Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}).
 
-During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during the modeling process, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be evaluated.
-The \psharp systematic testing engine will serialize the system (in single-box setup), and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a bound (e.g.
-in number of executions or time), or it hits a safety or liveness property violation (see \S\ref{sec:overview:bugs}).
+During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during modeling, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be taken.
+The \psharp testing engine will serialize (as single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g.
+in number of executions or time), or it hits a safety or liveness property violation.
 This testing process is fully automatic, has no false-positives (assuming an accurate test harness), and can reproduce found bugs by replaying buggy schedules.
-After a bug is found, the \psharp testing engine generates a trace that represents the buggy schedule.
+After a bug is discovered, \psharp generates a trace that represents the buggy schedule.
 In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events, and thus is easier to debug.
@@
 
-In this work, we are proposing a solution between the above two extremes: systematically test the implementation of a distributed system component against its modeled by \psharp environment.
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\linewidth]{img/example_code}
+\vspace{-7mm}
+\caption{Pseudocode of a simple distributed storage system that replicates data sent by a client.}
+\label{fig:example}
+\vspace{-2mm}
+\end{figure*}
+
+In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled by \psharp environment.
 The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space.
-Testing with our approach does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp.
+Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp.
 However, we argue that developers already spend significant time in building test suites and testing a distributed system prior to deployment.
@@
 
-\subsection{A small example}
+In principle, our modeling and testing methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
+
+\vspace{-2mm}
+\subsection{The \psharp programming model}
+\label{sec:overview:psharp}
+
+\psharp programs consist of multiple \emph{state machines} that communicate with each other \emph{asynchronously} by exchanging \emph{events}.
+In the case of distributed systems, \psharp events can be used to model regular messages between system components, failures or timeouts.
+A \psharp machine declaration is similar to a \csharp class declaration, but a machine also contains an event queue, and one or more \emph{states}.
+Each state can register \emph{actions} to handle incoming events.
+
+\psharp machines run concurrently with each other, each executing an event handling loop that dequeues the next event from the queue and handles it by invoking the registered action.
+An action might transition the machine to a new state, create a new machine, send an event to another machine, access a field or call a method.
+In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the queue of the target machine, which will dequeue and handle the event concurrently.
+All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+
+\vspace{-2mm}
+\subsection{An example distributed system}
 \label{sec:overview:example}
 
-Draft.
+Figure~\ref{fig:example} presents the pseudocode of a simple distributed storage system that was contrived for the purposes of explaining our testing methodology.
+The system consists of a client, a server and three storage nodes (SNs).
+The client sends the server a \texttt{ClientReq} message that contains data to be replicated, and then waits to get an acknowledgement before sending the next request.
+When the server receives \texttt{ClientReq}, it first stores the data locally, and then broadcasts a \texttt{ReplReq} message to all SNs.
+When an SN receives \texttt{ReplReq}, it handles the message by storing the data locally.
+Each SN has a timer installed, which sends periodic \texttt{Timeout} messages.
+Upon receiving \texttt{Timeout}, an SN sends a \texttt{Sync} message that contains the storage log.
+The server handles \texttt{Sync} by checking if the SN log is up-to-date.
+If it is not, the server sends a repeat \texttt{ReplReq} message to the outdated SN.
+If the SN log is up-to-date, then the server increments a replica counter by one.
+Finally, when there are three replicas available, the server sends an \texttt{Ack} message to the client.
 
-\subsection{Modeling with \psharp}
-\label{sec:overview:psharp}
+There are two bugs in the example.
+The first bug is that the server does not keep track of unique replicas.
+The replica counter increments upon each up-to-date \texttt{Sync}, even if the syncing SN is already considered a replica.
+This means that the server might send \texttt{Ack} when less than three replicas exist, which is erroneous behaviour.
+The second bug is that the server does not reset the replica counter to 0 upon sending an \texttt{Ack}.
+This means that when the client sends another \texttt{ClientReq}, it will never receive \texttt{Ack}, and thus block indefinitely.
 
-A \psharp program consists of multiple \emph{state machines} that communicate with each other \emph{asynchronously} by sending and receiving \emph{events}, which may contain an optional \emph{payload}.
-A \psharp machine declaration is similar to a class declaration in \csharp: it can contain an arbitrary number of fields and methods.
-However, a machine declaration differs from a class declaration, as it also contains an input event queue, and one or more \emph{states}.
-Each state can register \emph{actions} to handle incoming events.
+\vspace{-1mm}
+\subsection{Modeling the example system}
+\label{sec:overview:model}
 
-\psharp machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking the registered action.
-This action might access a field, call a method, transition the machine to a new state, create a new machine, or send an event to another machine.
-In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the input queue of the target machine, which will dequeue and handle the event concurrently.
-All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.
+To systematically test the example of Figure~\ref{fig:example}, the developer must first create a \psharp test harness, and then specify the correctness properties of the system.
+Figure~\ref{fig:example:model} illustrates a test harness that can find the two bugs discussed in \S\ref{sec:overview:example}.
+Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being send from one machine to another.
 
-Our approach of using \psharp to test distributed systems, implemented in .NET, requires the developer to perform the following three key modeling tasks.
-These tasks are illustrated in \S\ref{sec:method} for the \Azure Storage vNext system.
+We did not model the server and SN components, so we can test their actual implementation.
+Each real component is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component.
+We decided to model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
+We also modeled the timer so that \psharp takes control of all time-related nondeterminism in the system.
+This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.
 
-\begin{enumerate}
-\item The real components of a distributed system must be wrapped inside \psharp machines.
-In addition, all the asynchrony due to message passing in the system must be modeled using the \psharp event-sending APIs.
+\psharp uses object-oriented language features such as interfaces and virtual method dispatch to connect the real code with the modeled code.
+Programmers in industry are used to working with such features, and heavily employ them in production for testing purposes.
+In our experience, this significantly lowers the bar for engineering teams inside Microsoft to embrace \psharp for testing.
 
-\item The environment of the system must be modeled in \psharp, which offers APIs for obtaining non-deterministically chosen values.
-These values can be used to simulate failures, timers, etc.
-(see \S\ref{sec:method:timer} and \S\ref{sec:method:driver}).
-This step captures external sources of nondeterminism.
-The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
-(see \S\ref{sec:overview:testing}).
+In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct.
+The interested reader can refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.
+Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}), and read \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric} to learn more details about how \psharp was used to model and test real distributed storage systems in Microsoft.
 
-\item The criteria for correctness of an execution must be specified.
-Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:bugs}).
-Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state.
-Liveness specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.
+\vspace{-1mm}
+\subsection{Specifying safety properties in \psharp}
+\label{sec:overview:safety}
 
-%\item Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs.
-This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:overview:testing}).
-\end{enumerate}
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/example}
+\vspace{-5mm}
+\caption{\psharp test harness for the Figure~\ref{fig:example} example.}
+\label{fig:example:model}
+\vspace{-2mm}
+\end{figure}
 
-\noindent
-We collectively refer to the environment model of a system and its wrapped components as the \psharp \emph{test harness}.
-This test harness can be used by \psharp to drive the system towards interesting behaviors.
-\psharp uses modern object-oriented language features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
-Programmers in industry are used to working with such features, and heavily use them in production for testing purposes.
-In our experience, this significantly lowers the bar for engineering teams inside \Microsoft to embrace \psharp for testing.
-
-In principle, our modeling methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
-We also argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
-
-\subsection{Writing specifications in \psharp}
-\label{sec:overview:bugs}
-
-In  this section, we describe how safety and liveness specifications are expressed in \psharp.
-
-\textbf{Safety specifications}.
-In addition to the usual assertions for stating safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events.
+Safety property specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state.
+\psharp supports the usual assertions for specifying safety properties that are local to a \psharp machine, and also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, a special \psharp machine that can receive, but not send, events.
 
@@
 An erroneous global behavior is flagged via an assertion on the private state of the safety monitor.
-Thus, a \psharp monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).
+Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).
 
-\textbf{Liveness specifications}.
-Liveness property specifications are more difficult to encode since they are intended to capture program progress over infinite executions.
-Violation of a liveness property can only be demonstrated by an infinite execution.
-Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
+The first bug in the example of \S\ref{sec:overview:example} is a safety bug.
+To find it, the developer can write a safety monitor (see Figure~\ref{fig:example:model}) that contains a map from unique SN ids to a boolean value, which denotes if the SN is a replica or not.
+Each time an SN replicates the latest data, it notifies the monitor to update the map.
+Each time the server issues an \texttt{Ack}, it also notifies the monitor.
+If the monitor detects that an \texttt{Ack} was sent without three replicas actually existing, a safety violation is triggered.
+
+\vspace{-1mm}
+\subsection{Specifying liveness properties in \psharp}
+\label{sec:overview:liveness}
+
+Liveness property specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.
+Typically, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977, lamport1994temporal}.
 We take a different approach and allow the programmer to write a \emph{liveness monitor}~\cite{desai2015building}.
 Similar to a safety monitor, a liveness monitor can receive, but not send, events.
-Unlike a safety monitor, a liveness monitor contains two types of states: \emph{hot} and \emph{cold}.
 
+A liveness monitor contains two special types of states: \emph{hot} and \emph{cold}.
 A hot state denotes a point in the execution where progress is required but has not happened yet; for example, a node has failed but a new one has not come up yet.
-A liveness monitor enters a hot state upon receiving notification of an event that requires the system to make progress.
-The liveness monitor leaves the hot state and enters a cold state upon receiving another event notifying it of progress.
+A liveness monitor enters a hot state when it is notified that the system must make progress.
+The liveness monitor leaves the hot state and enters a cold state when it is notified that the system has progressed.
 An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
 Our liveness monitors can encode arbitrary temporal logic properties.
-A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
+%A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
 %The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events, such as a periodic request or a periodic timer expiration.
@@
 Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics.
-One such heuristic considers an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
-Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution, and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state graph of the program.
+In this work, we consider an execution longer than a large user-supplied depth-bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
+
+%Another heuristic maintains a cache of (pieces of) the state of the \psharp program obtained at each step in the execution, and reports an ``infinite'' execution when the latest step results in a cache hit, thus approximating a cycle in the state graph of the program.
+
+The second bug in the example of \S\ref{sec:overview:example} is a liveness bug.
+To detect it, the developer can write a liveness monitor (see Figure~\ref{fig:example:model}) that transitions from a hot state, which denotes that the client sent a \texttt{ClientReq} and waits for an \texttt{Ack}, to a cold state, which denotes that the server has sent an \texttt{Ack} in response to the last \texttt{ClientReq}.
+Each time a server receives a \texttt{ClientReq}, it notifies the monitor to transition to the hot state.
+Each time the server issues an \texttt{Ack}, it notifies the monitor to transition to the cold state.
+If the liveness monitor is in a hot state when the execution (either finite, or depth-bounded infinite) terminates, a liveness violation is triggered.
 
@@
 
-\subsection{Systematic testing using \psharp}
-\label{sec:overview:testing}
+%\subsection{Modeling and testing with \psharp}
+%\label{sec:overview:test}
 
-The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines.
-A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:overview:bugs}.
+%To detect the two bugs in the example system of Figure~\ref{fig:example}, the developer must first model the system using \psharp and then specify its correctness properties.
+Figure~\ref{fig:example:model} illustrates a test harness written in \psharp that can be used for finding the bugs in the example system of Figure~\ref{fig:example}.
+%
+%Our approach of using \psharp to find distributed system bugs requires the developer to perform the following three key modeling tasks.
+These tasks are illustrated in \S\ref{sec:method} for the Azure Storage vNext system.
+%
+%\begin{enumerate}
+%\item The real components of a distributed system must be wrapped inside \psharp machines.
+In addition, all the asynchrony due to message passing in the system must be modeled using the \psharp event-sending APIs.
+%
+%\item The environment of the system must be modeled in \psharp, which offers APIs for obtaining non-deterministically chosen values.
+These values can be used to simulate failures, timers, etc.
+(see \S\ref{sec:method:timer} and \S\ref{sec:method:driver}).
+This step captures external sources of nondeterminism.
+The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc.
+%\end{enumerate}
 
-When the \psharp runtime executes in bug-finding mode, an embedded \emph{systematic concurrency testing engine}~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
+%\noindent
+%We collectively refer to the environment model of a system and its wrapped components as the \psharp \emph{test harness}.
+This test harness can be used by \psharp to drive the system towards interesting behaviors.
+\psharp uses modern object-oriented language features such as \emph{interfaces} and \emph{virtual method dispatch} to connect the real code with the modeled code.
+Programmers in industry are used to working with such features, and heavily use them in production for testing purposes.
+In our experience, this significantly lowers the bar for engineering teams inside Microsoft to embrace \psharp for testing.
+
+%In principle, our modeling methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
+%We also argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing.
+
+%A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:overview:specs}.
+
+%When the \psharp runtime executes in bug-finding mode, an embedded \emph{systematic concurrency testing engine}~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime.
 In particular, the runtime is aware of nondeterminism due to the interleaving of event handlers in different machines.
@@
 
-We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+%We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
 It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.

### 2015-12-25 paper.tex
major update to address PC and shepherd comments

@@
 
-Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure cloud computing platform.
+Our methodology has been applied to three distributed storage systems in the Microsoft Azure cloud computing platform.
 In the process, numerous bugs were identified, reproduced, confirmed and fixed.
@@
 
+\vspace{-1mm}
 \section{Introduction}
 \label{sec:intro}
+\vspace{-1mm}
 
@@
 
-\section{Motivating Example}
-\label{sec:motivation}
+%\section{Motivating Example}
+%\label{sec:motivation}
+%
+%\input{motivation}
 
-\input{motivation}
-
-\section{Overview of our Approach}
-%\section{Our Approach to Testing Distributed Systems}
+\section{Testing Distributed Systems with \psharp}
 \label{sec:overview}
@@
 
-\section{Testing \Azure Storage vNext with \psharp}
-\label{sec:method}
+\section{Case Study: Azure Storage vNext}
+\label{sec:vnext}
 
-\input{methodology}
+\input{vnext}
 
-\section{Additional Case Studies}
-\label{sec:cases}
+\section{Case Study: Live Table Migration}
+\label{sec:migrating}
 
-\input{experience}
+\input{migrating}
+
+\section{Case Study: Azure Service Fabric}
+\label{sec:fabric}
+
+\input{fabric}
 
@@
 
-%\section{Acknowledgments}
+\section{Acknowledgments}
 
-%Draft.
+We thank our shepherd Haryadi Gunawi for his valuable guidance, and the reviewers for their constructive comments.
+We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions on various stages of this work.
+This research was partially supported by a gift from Intel Corporation.
 

### 2015-12-25 related.tex
major update to address PC and shepherd comments

@@
-Most related to our work is model checking~\cite{godefroid1997verisoft}, a powerful technique that has been widely used in the past for finding Heisenbugs in the actual implementation of distributed systems~\cite{killian2007life, yang2009modist, yabandeh2009crystalball, guerraoui2011model, guo2011practical, simsa2011dbug, leesatapornwongsa2014samc}.
+Most related to our work are model checking~\cite{godefroid1997verisoft} and systematic concurrency testing~\cite{musuvathi2008finding, emmi2011delay, thomson2014sct}, two powerful techniques that have been widely used in the past for finding Heisenbugs in the actual implementation of distributed systems~\cite{killian2007life, yang2009modist, yabandeh2009crystalball, guerraoui2011model, guo2011practical, simsa2011dbug, gunawi2011fate, leesatapornwongsa2014samc}.
 
@@
 
-\textsc{Fate} and \textsc{Destini} is a framework for systematically injecting failures in distributed systems~\cite{gunawi2011fate}.
+%\textsc{Fate} and \textsc{Destini} is a framework for systematically injecting failures in distributed systems~\cite{gunawi2011fate}.
 This framework focuses in exercising various failure scenarios, whereas \psharp can be used for testing generic safety and liveness properties of distributed systems.

### 2015-12-25 vnext.tex
major update to address PC and shepherd comments

--- 2015-12-22 methodology.tex
@@
-To uncover the elusive extent repair bug in \Azure Storage vNext, its developers wrote a test harness using \psharp.
+Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
+It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
+%~\cite{greenberg2015keynote}.
+
+Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
+Similar to the current system, vNext employs containers, called \emph{extents}, to store data.
+Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs).
+However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}.
+In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr).
+This partitioning is illustrated in Figure~\ref{fig:vnext}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/azurestore}
+\vspace{-5mm}
+\caption{Top-level components for extent management in Microsoft Azure Storage vNext.}
+\label{fig:vnext}
+\vspace{-2mm}
+\end{figure}
+
+One of the many responsibilities of an ExtMgr is to ensure that every extent maintains enough \emph{replicas} in the system.
+To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN.
+The failure of an EN is detected by missing heartbeats.
+An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN.
+The sync reports list all the extents (and associated metadata) stored on the EN.
+Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result.
+The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs.
+The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via their next periodic sync reports.
+All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
+
+To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing:
+\begin{enumerate}
+\item \emph{Unit testing}, in which emulated heartbeats and sync reports are sent to an ExtMgr. These tests check that the messages are processed as expected.
+
+\item \emph{Integration testing}, in which an ExtMgr is launched together with multiple ENs.
+An EN failure is subsequently injected.
+These tests check that the affected extents are eventually repaired.
+
+\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents.
+The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
+\end{enumerate}
+
+\noindent
+Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic.
+All the unit test suites and integration test suites successfully pass every single time.
+However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.
+
+This bug has proven to be difficult to identify, reproduce and troubleshoot.
+First, an extent never being repaired is \emph{not} a property that can be easily checked.
+Second, the bug seems to manifest only in very long executions.
+Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.
+%To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/modeled_vnext}
+\vspace{-5mm}
+\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
+\label{fig:azurestoremodel}
+\vspace{-2mm}
+\end{figure}
+
+To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a test harness using \psharp.
 The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic.
@@
 The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
+
 \begin{description}
@@
 
-\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:mock_en}).
+\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:model_en}).
 
@@
 \end{description}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/mocked_vnext}
-\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
-\label{fig:azurestoremodel}
-\vspace{-2mm}
-\end{figure}
 
@@
     ExtMgr = new ExtentManager();
-    ExtMgr.NetEngine = new MockedNetEngine(); // mock network
-    ExtMgr.IsMockingTimer = true;	 // disable internal timer
+    ExtMgr.NetEngine = new ModelNetEngine(); // model network
+    ExtMgr.IsModelingTimer = true;	 // disable internal timer
   }
@@
 \end{lstlisting}
-\vspace{-3mm}
+\vspace{-2mm}
 \caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
 \label{fig:wrap_target}
-%\vspace{-2mm}
 \end{figure}
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=.9\linewidth]{img/extent_manager}
-\caption{Internal components of the real Extent Manager in \Azure Storage vNext.}
+\includegraphics[width=.95\linewidth]{img/extent_manager}
+\vspace{-2mm}
+\caption{Internal components of the real Extent Manager in Microsoft Azure Storage vNext.}
 \label{fig:extentmanager}
-\vspace{-5mm}
-\end{figure*}
+\end{figure}
 
-\textbf{Internals of the real Extent Manager}.
-The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
-The \texttt{ExtentCenter} data structure maintains the records that map from extents to their hosting ENs.
+\textbf{Real Extent Manager}.
+The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
+The \texttt{ExtentCenter} maintains the records that map from extents to their hosting ENs.
 It is updated upon receiving the periodic sync reports from the ENs.
@@
 Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth.
-\texttt{ExtentNodeMap} records the latest heartbeat time from every EN.
+The \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.
 
@@
 \textbf{Intercepting network messages}.
-The real ExtMgr uses a network engine to asynchronously send messages to the ENs.
-The \psharp test harness mocks the original network engine in vNext by overriding its interface.
-This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines.
-As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}.
-This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.
+The real ExtMgr uses a network engine to asynchronously send messages to the \texttt{ExtentNode} machines.
+The \psharp test harness models the original network engine in vNext by overriding its interface.
+The modeled network engine (see Figure~\ref{fig:enginecode}) intercepts all outbound messages from the ExtMgr, and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding ENs.
+This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}.
 
@@
 
-// mocked engine for intercepting Extent Manager messages
-class MockedNetEngine : NetworkEngine {
+// modeled engine for intercepting Extent Manager messages
+class ModelNetEngine : NetworkEngine {
   public override void SendMessage(Socket s, Message msg) {
@@
 \vspace{-4mm}
-\caption{Mocked network engine in vNext.}
+\caption{Modeled network engine in vNext.}
 \label{fig:enginecode}
-\vspace{-2mm}
+\vspace{-3mm}
 \end{figure}
 
-Intercepting all network messages and dispatching them through the \psharp runtime is important for two reasons.
+Intercepting all network messages and dispatching them through \psharp is important for two reasons.
 First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system.
-Second, the mocked network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).
+Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).
 
-Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine.
+Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine.
 They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
@@
 \subsection{The ExtentNode machine}
-\label{sec:method:mock_en}
+\label{sec:method:model_en}
 
@@
 
-  // extent repair logic
-  ...
   [OnEvent(ExtentCopyResponseEvent, ProcessCopyResponse)]
@@
 \vspace{-4mm}
-\caption{The modeled EN in vNext.}
-\label{fig:mocked_en}
+\caption{Code snippet of the modeled EN in vNext.}
+\label{fig:modeled_en}
 \vspace{-2mm}
@@
 It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored.
-After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.
+After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:modeled_en}.
 
@@
 It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharp.Send(...)}.
-The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.
+The \texttt{ExtentNode} machine also includes failure-related logic (see \S\ref{sec:method:driver}).
 
@@
 \caption{Modeling timer expiration using \psharp.}
-\label{fig:mocked_timer}
+\label{fig:modeled_timer}
 \vspace{-2mm}
@@
 
-Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine in the test harness.
+Figure~\ref{fig:modeled_timer} shows the \texttt{Timer} machine in the test harness.
 \texttt{Timer} invokes the \psharp method \texttt{Nondet()}, which generates a nondeterministic choice controlled by the \psharp runtime.
@@
 
-  // other test harness logic
-  ...
   void InjectNodeFailure() {
@@
 \vspace{-4mm}
-\caption{The TestingDriver machine in vNext.}
+\caption{Code snippet of the TestingDriver machine.}
 \label{fig:testing_driver}
@@
 It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine, and then sends a \texttt{FailureEvent} to the chosen machine to emulate an EN failure.
-As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent}, notifies the liveness monitor of its failure (see \S\ref{sec:method:monitor}) and terminates itself by invoking the \psharp method \texttt{Halt()}.
+As shown in the earlier Figure~\ref{fig:modeled_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent}, notifies the liveness monitor of its failure (see \S\ref{sec:method:monitor}) and terminates itself by invoking the \psharp method \texttt{Halt()}.
 \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.
@@
 
-\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state.
+\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:overview:liveness}) that transitions between a cold and a hot state.
 Whenever an EN fails, the monitor is notified with an \texttt{ENFailedEvent} event.
@@
 
-\subsection{Liveness Bug in \Azure Storage vNext}
+\subsection{Liveness Bug in Azure Storage vNext}
 \label{sec:method:azurestore}

### 2015-12-25 evaluation.tex
added comment

@@
 
+\PDComment{TODO:}
 We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.

### 2015-12-28 overview.tex
minor edits

@@
-The goal of our work is to find distributed system Heisenbugs \emph{before} they reach production.
+The goal of our work is to find bugs in distributed system \emph{before} they reach production.
 Typical distributed systems consist of multiple components that interact with each other via \emph{message passing}.
@@
 Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing.
-Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various distributed events (see \S\ref{sec:overview:model}).
-The harness is essentially an environment model of the system.
+Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various events (see \S\ref{sec:overview:model}).
+The harness is essentially a model of the environment of the system.
 The purpose of these first two activities is to explicitly declare all sources of nondeterminism in the system using \psharp.
@@
 During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during modeling, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be taken.
-The \psharp testing engine will serialize (as single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g.
+The \psharp testing engine will serialize (in a single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g.
 in number of executions or time), or it hits a safety or liveness property violation.
@@
 
-In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled by \psharp environment.
+In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment.
 The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space.
-Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp.
-However, we argue that developers already spend significant time in building test suites and testing a distributed system prior to deployment.
+Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp.
+However, we argue that developers already spend significant time in building test suites for distributed systems prior to deployment.
 The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).
@@
 Figure~\ref{fig:example:model} illustrates a test harness that can find the two bugs discussed in \S\ref{sec:overview:example}.
-Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being send from one machine to another.
+Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being sent from one machine to another.
 
-We did not model the server and SN components, so we can test their actual implementation.
+We do not model the server and SN components since we want to test their actual implementation.
 Each real component is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component.
-We decided to model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
-We also modeled the timer so that \psharp takes control of all time-related nondeterminism in the system.
+We model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
+We also model the timer so that \psharp takes control of all time-related nondeterminism in the system.
 This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.
@@
 The first bug in the example of \S\ref{sec:overview:example} is a safety bug.
-To find it, the developer can write a safety monitor (see Figure~\ref{fig:example:model}) that contains a map from unique SN ids to a boolean value, which denotes if the SN is a replica or not.
+To find it, the developer can write a safety monitor (see Figure~\ref{fig:example:model}) that contains a map from unique SN ids to a Boolean value, which denotes if the SN is a replica or not.
 Each time an SN replicates the latest data, it notifies the monitor to update the map.

### 2016-01-01 evaluation.tex
Edit MigratingTable sections.

@@
 We believe this is because the inputs and schedules that trigger them are too rare in our distribution.
-To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs, as shown.
+To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
+% Be a little bit more transparent about the fact that the $\diamond$ rows do not
+% reflect results achievable via our main testing approach. ~ Matt 2015-12-31
+the table shows the results of these runs.
 This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.
 
-The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects the type of oversight that tends to occur as designs evolve.
+The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve.
 This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key.

### 2016-01-01 migrating.tex
Edit MigratingTable sections.

@@
-Live Table Migration (MigratingTable) is a library designed to \emph{transparently} migrate a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set.
+% Remove the \emph on ``transparently'' and ``while'' again.  Pantazis, I'm
+% imagining you were trying to emphasize what's interesting about the problem,
+% but I think suggesting that this idea is new to the reader is insulting.
+% ~ Matt 2015-12-31
+Live Table Migration (MigratingTable) is a library designed to transparently migrate a data set between tables in the Azure storage service while an application is accessing this data set.
 This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness.
@@
 
-MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary Azure table.
-This interface is named \texttt{IChainTable}.
-A virtual table is backed by a pair of \emph{old} and \emph{new} \emph{backend} tables, both of which implement \texttt{IChainTable}.
-A \emph{migrator} job is responsible for moving all data from the old table to the new table in the background.
-Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same backend tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table.
-The goal of using \psharp was to systematically test this property.
+MigratingTable provides a \emph{virtual table} that has an interface
+similar to an ordinary Azure table. This interface is
+named \texttt{IChainTable}. A virtual table is backed by a pair
+of \emph{old} and \emph{new} \emph{backend} tables, both of which
+implement \texttt{IChainTable}. A \emph{migrator} job is responsible
+for moving all data from the old table to the new table in the
+background. Meanwhile, each \emph{input} read and write operation
+issued to the virtual table is implemented via a sequence
+of reads and writes on the backend tables according to a protocol,
+which intends to guarantee that when multiple applications issue input
+operations to different MigratingTable instances with the same backend
+tables, the output of these operations complies with the specification
+of \texttt{IChainTable} for the combined \emph{input history} as if
+all the operations were done on a single table. The goal of
+using \psharp was to systematically test this property.
 
@@
 
-There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) there are many possible input histories.
-The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.
-Instead, the developers created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, which is backed by a single in-memory table.
-After executing an arbitrary input history on both implementations, the outputs can be compared to ensure that MigratingTable produces the same output as \texttt{SpecTable}.
+There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the input operations accept many parameters that affect the behavior in different ways.
+The developers could have chosen specific input histories to test, but they were not confident that they would cover the combinations of parameters that might trigger bugs.
+Instead, they used the \psharp API to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits.
+They created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, and compared the output of MigratingTable on the chosen input history to that of a \texttt{SpecTable} instance called the \emph{reference table}.
+The \texttt{SpecTable} implementation was reused for the backend tables of MigratingTable.
+% I'm removing the claim that each bug is caught with positive probability per
+% run because it's too awkward to explain here and it's nothing specific to
+% MigratingTable.  Also, we can't quite appeal to the small scope hypothesis in
+% its original form because I hand-picked which parameters to vary and what
+% limits to use.  The value of the input generator should be clear enough from
+% the above. ~ Matt 2015-12-31
 
 Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable.
-The harness consists of a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; a \texttt{Tables} machine, which contains the old and new tables, and an instance of \texttt{SpecTable}; and a \texttt{Migrator} machine that performs the background migration.
-Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which then sends these operations to the \texttt{Tables} machine.
+The harness consists of a \texttt{Tables} machine that contains the backend tables and the reference table and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration.
+Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the backend tables via \psharp events.
+The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single virtual table, so the test harness could perform the operation on the reference table at the same time.
+(More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table.
+This way, the rest of the system never observes the reference table to be out of sync with the virtual table.)
+% Note: If the tables were observed to be out of sync, it could cause both false
+% positives and false negatives. ~ Matt 2015-12-31
 
-The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table.
-This allows the \psharp test harness to keep \texttt{SpecTable} in sync with the virtual table: the \texttt{Service} machine sends each input operation to \texttt{SpecTable} when the linearization point on the virtual table is reached.
-The output of the operation on the virtual table is then checked against its output on \texttt{SpecTable}.
-
-Conceptually, the \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}.
-Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set (the small scope hypothesis), then an execution of the test harness must have a non-zero probability of detecting the bug.
-
-\psharp captures and controls all the interleaving in the system, and systematically explores them.
-To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.
+For further information about MigratingTable and its \psharp test harness, we refer the reader to the source repository~\cite{migratingtable-src}.
 

### 2016-01-17 evaluation.tex
fixes to the paper to prepare for camera ready

@@
 Although modeling Fabric required a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.
+Having modeled Fabric, we are now able to systematically test numerous services built on top of it.
 
@@
 According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner.
-After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one to evaluate the effectiveness of different testing strategies.
+After the developers fixed all the discovered bugs, we added flags to allow them to be individually reintroduced, for purposes of evaluation.
 
@@
 We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
-It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries.
-The interface exposes callbacks that are invoked by the \psharp testing engine for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.
 
@@
 The random seed for both schedulers was generated using the current time.
-The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps of 500.
+The PCT scheduler was further configured with a bug depth of 2 and a maximum number of scheduling steps of 500.
 All reported times are in seconds.
 
-For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug in less than 10 seconds.
+For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug within 10 seconds.
 The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
-The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger depth bound to be found.
+The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger bug depth bound to be found.
 
@@
 The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable.
-\psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend and was done in all our case studies.
+\psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend, and we extended trace information in a custom fashion for each of our case studies.
 

### 2016-01-17 intro.tex
fixes to the paper to prepare for camera ready

@@
-Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, leesatapornwongsa2014samc, laguna2015debugging, maddox2015test}.
-This challenge is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos, henry2009cloud}, such as unexpected node failures, the asynchronous interaction between system components, data losses due to unreliable communication channels, the use of multithreaded code to exploit multicore machines, and interaction with clients.
+Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, leesatapornwongsa2014samc, gunawi2014cbs, laguna2015debugging, maddox2015test}.
+This challenge is due to many sources of \emph{nondeterminism}~\cite{chandra2007paxos, henry2009cloud, leesatapornwongsa2016taxdc}, such as unexpected node failures, the asynchronous interaction between system components, data losses due to unreliable communication channels, the use of multithreaded code to exploit multicore machines, and interaction with clients.
 All these sources of nondeterminism can easily create \emph{Heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, corner-case bugs that are difficult to detect, diagnose and fix.
-A Heisenbug might hide deep inside one of these paths and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequences can be catastrophic~\cite{amazon2012aws, google2014outage}.
+These bugs might hide inside a code path that can only be triggered by a specific interleaving of concurrent events and only manifest under extremely rare conditions~\cite{gray1986computers, musuvathi2008finding}, but the consequences can be catastrophic~\cite{amazon2012aws, google2014outage}.
 
@@
 To use \psharp for testing, the programmer has to augment the original system with three artifacts: a model of the nondeterministic execution environment of the system; a test harness that drives the system towards interesting behaviors; and safety or liveness specifications.
-\psharp then systematically exercises the test harness and validates program behaviours against the provided specifications.
+\psharp then systematically exercises the test harness and validates program behaviors against the provided specifications.
 

### 2016-01-17 migrating.tex
fixes to the paper to prepare for camera ready

@@
 
-MigratingTable provides a \emph{virtual table} that has an interface
-similar to an ordinary Azure table. This interface is
-named \texttt{IChainTable}. A virtual table is backed by a pair
-of \emph{old} and \emph{new} \emph{backend} tables, both of which
+MigratingTable provides a \emph{virtual table} (VT) that has a similar interface
+to an ordinary Azure table. This interface is
+named \texttt{IChainTable}. Each VT is backed by a pair
+of \emph{old} and \emph{new} \emph{backend} tables (BTs), both of which
 implement \texttt{IChainTable}. A \emph{migrator} job is responsible
@@
 background. Meanwhile, each \emph{input} read and write operation
-issued to the virtual table is implemented via a sequence
-of reads and writes on the backend tables according to a protocol,
+issued to the VT is implemented via a sequence
+of reads and writes on the BTs according to a protocol,
 which intends to guarantee that when multiple applications issue input
-operations to different MigratingTable instances with the same backend
-tables, the output of these operations complies with the specification
-of \texttt{IChainTable} for the combined \emph{input history} as if
-all the operations were done on a single table. The goal of
+operations to different MigratingTable instances with the same BTs, the output of these operations complies with the specification
+of \texttt{IChainTable} for the combined \emph{input history}, as if
+all the operations were performed on a single table. The goal of
 using \psharp was to systematically test this property.
@@
 The developers could have chosen specific input histories to test, but they were not confident that they would cover the combinations of parameters that might trigger bugs.
-Instead, they used the \psharp API to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits.
+Instead, the developers used the \psharp nondeterministic choice generation APIs to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits.
 They created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, and compared the output of MigratingTable on the chosen input history to that of a \texttt{SpecTable} instance called the \emph{reference table}.
-The \texttt{SpecTable} implementation was reused for the backend tables of MigratingTable.
+The \texttt{SpecTable} implementation was reused for the BTs of MigratingTable.
 % I'm removing the claim that each bug is caught with positive probability per
@@
 
-Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable.
-The harness consists of a \texttt{Tables} machine that contains the backend tables and the reference table and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration.
-Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the backend tables via \psharp events.
-The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single virtual table, so the test harness could perform the operation on the reference table at the same time.
-(More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table.
-This way, the rest of the system never observes the reference table to be out of sync with the virtual table.)
+Figure~\ref{fig:mockedmigration} shows the \psharp test harness of MigratingTable.
+The test harness consists of a \texttt{Tables} machine, which contains the BTs and the reference table, and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration.
+Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the BTs via \psharp events.
+The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single VT, so the test harness could perform the operation on the reference table at the same time.
+More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table.
+This way, the rest of the system never observes the reference table to be out of sync with the VT.
 % Note: If the tables were observed to be out of sync, it could cause both false
@@
 
-For further information about MigratingTable and its \psharp test harness, we refer the reader to the source repository~\cite{migratingtable-src}.
+For further information about this case study, we refer the reader to the source code repository~\cite{migratingtable-src}.
 

### 2016-01-17 overview.tex
fixes to the paper to prepare for camera ready

@@
 Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing.
-Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various events (see \S\ref{sec:overview:model}).
+Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors by nondeterministically triggering various events (see \S\ref{sec:overview:model}).
 The harness is essentially a model of the environment of the system.
@@
 During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during modeling, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be taken.
-The \psharp testing engine will serialize (in a single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g.
-in number of executions or time), or it hits a safety or liveness property violation.
-This testing process is fully automatic, has no false-positives (assuming an accurate test harness), and can reproduce found bugs by replaying buggy schedules.
-After a bug is discovered, \psharp generates a trace that represents the buggy schedule.
+The \psharp testing engine will serialize (in a single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g.\ in number of executions or time), or it hits a safety or liveness property violation.
+This testing process is fully automatic and has no false-positives (assuming an accurate test harness).
+After a bug is discovered, \psharp generates a trace that represents the buggy schedule, which can then be replayed to reproduce the bug.
 In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events, and thus is easier to debug.
@@
 The client sends the server a \texttt{ClientReq} message that contains data to be replicated, and then waits to get an acknowledgement before sending the next request.
-When the server receives \texttt{ClientReq}, it first stores the data locally, and then broadcasts a \texttt{ReplReq} message to all SNs.
-When an SN receives \texttt{ReplReq}, it handles the message by storing the data locally.
+When the server receives \texttt{ClientReq}, it first stores the data locally (in the \texttt{Data} field), and then broadcasts a \texttt{ReplReq} message to all SNs.
+When an SN receives \texttt{ReplReq}, it handles the message by storing the received data locally (using the \texttt{store} method).
 Each SN has a timer installed, which sends periodic \texttt{Timeout} messages.
-Upon receiving \texttt{Timeout}, an SN sends a \texttt{Sync} message that contains the storage log.
-The server handles \texttt{Sync} by checking if the SN log is up-to-date.
+Upon receiving \texttt{Timeout}, an SN sends a \texttt{Sync} message to the server that contains the storage log.
+The server handles \texttt{Sync} by calling the \texttt{isUpToDate} method to check if the SN log is up-to-date.
 If it is not, the server sends a repeat \texttt{ReplReq} message to the outdated SN.
@@
 
-There are two bugs in the example.
+There are two bugs in the above example.
 The first bug is that the server does not keep track of unique replicas.
 The replica counter increments upon each up-to-date \texttt{Sync}, even if the syncing SN is already considered a replica.
-This means that the server might send \texttt{Ack} when less than three replicas exist, which is erroneous behaviour.
+This means that the server might send \texttt{Ack} when fewer than three replicas exist, which is erroneous behaviour.
 The second bug is that the server does not reset the replica counter to 0 upon sending an \texttt{Ack}.
@@
 Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being sent from one machine to another.
+We use three types of boxes: (i) a box with rounded corners and thick border denotes a real component wrapped inside a \psharp machine; (ii) a box with thin border denotes a modeled component in \psharp; and (iii) a box with dashed border denotes a special \psharp machine used for safety or liveness checking (see \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}).
 
-We do not model the server and SN components since we want to test their actual implementation.
-Each real component is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component.
-We model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
-We also model the timer so that \psharp takes control of all time-related nondeterminism in the system.
+We do not model the server component since we want to test its actual implementation.
+The server is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component.
+We model the SNs so that they store data in memory rather than on disk (which could be inefficient during testing).
+We also model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
+Finally, we model the timer so that \psharp takes control of all time-related nondeterminism in the system.
 This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.
@@
 In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct.
-The interested reader can refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.
-Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}), and read \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric} to learn more details about how \psharp was used to model and test real distributed storage systems in Microsoft.
+The interested reader can refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}), and read \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric} to learn more details about how \psharp was used to model and test real distributed storage systems in Microsoft.
 
@@
 A liveness monitor contains two special types of states: \emph{hot} and \emph{cold}.
-A hot state denotes a point in the execution where progress is required but has not happened yet; for example, a node has failed but a new one has not come up yet.
+A hot state denotes a point in the execution where progress is required but has not happened yet; e.g.\ a node has failed but a new one has not come up yet.
 A liveness monitor enters a hot state when it is notified that the system must make progress.
 The liveness monitor leaves the hot state and enters a cold state when it is notified that the system has progressed.
-An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
+An infinite execution is erroneous if the liveness monitor is in a hot state for an infinitely long time.
 Our liveness monitors can encode arbitrary temporal logic properties.
@@
 Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics.
-In this work, we consider an execution longer than a large user-supplied depth-bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
+In this work, we consider an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
+Note that fairness is not relevant when using this heuristic, due to our pragmatic use of a large bound instead of proper handling of infinite executions.
 
@@
 Each time the server issues an \texttt{Ack}, it notifies the monitor to transition to the cold state.
-If the liveness monitor is in a hot state when the execution (either finite, or depth-bounded infinite) terminates, a liveness violation is triggered.
+If the monitor is in a hot state when the bounded infinite execution terminates, a liveness violation is triggered.
 

### 2016-01-17 related.tex
fixes to the paper to prepare for camera ready

@@
 \textsc{MaceMc}~\cite{killian2007life} is a model checker for distributed systems written in the \textsc{Mace}~\cite{killian2007mace} language.
-The focus of \textsc{MaceMc} is to find liveness property violations using an algorithm based on depth-bounded random walk and the use of heuristics.
+The focus of \textsc{MaceMc} is to find liveness property violations using an algorithm based on bounded random walk, combined with heuristics.
 Because \textsc{MaceMc} can only test systems written in \textsc{Mace}, it cannot be easily used in an industrial setting.
@@
 Another formal approach is Verdi~\cite{wilcox2015verdi}, a framework for writing and verifying distributed systems in Coq~\cite{barras1997coq}.
-Verdi generates OCaml code from the verifying system, which can be used for execution.
+Verdi generates OCaml code from the verified Coq system, which can be then used for execution.
 In contrast, \psharp performs bounded testing on a system already written in \csharp, which in our experience lowers the bar for adoption by engineering teams.
-Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed storage systems.
+Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed systems.

### 2016-01-17 vnext.tex
fixes to the paper to prepare for camera ready

@@
 Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
-%~\cite{greenberg2015keynote}.
+It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
 
@@
 
-One of the many responsibilities of an ExtMgr is to ensure that every extent maintains enough \emph{replicas} in the system.
-To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN.
-The failure of an EN is detected by missing heartbeats.
-An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN.
+One of the responsibilities of an ExtMgr is to ensure that every managed extent maintains enough \emph{replicas} in the system.
+To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN that it manages.
+EN failure is detected by missing heartbeats.
+An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN that it manages.
 The sync reports list all the extents (and associated metadata) stored on the EN.
@@
 The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs.
-The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via their next periodic sync reports.
+The ENs then repair the extents from the existing replicas and lazily update the ExtMgr via their next periodic sync reports.
 All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
@@
 \noindent
-Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic.
-All the unit test suites and integration test suites successfully pass every single time.
-However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.
+Despite of the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic.
+All the unit test suites and integration test suites successfully passed on each test run.
+However, the stress test suite failed \emph{from time to time} after very long executions; in these cases, certain replicas of some extents failed without subsequently being repaired.
 
-This bug has proven to be difficult to identify, reproduce and troubleshoot.
+This bug proved difficult to identify, reproduce and troubleshoot.
 First, an extent never being repaired is \emph{not} a property that can be easily checked.
-Second, the bug seems to manifest only in very long executions.
-Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.
+Second, the bug appeared to manifest only in very long executions.
+Finally, by the time that the bug did manifest, very long execution traces had been collected, which made manual inspection tedious and ineffective.
 %To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.
@@
 \vspace{-5mm}
-\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
+\caption{Real Extent Manager with its \psharp test harness (each box represents one \psharp machine).}
 \label{fig:azurestoremodel}
@@
     ExtMgr.NetEngine = new ModelNetEngine(); // model network
-    ExtMgr.IsModelingTimer = true;	 // disable internal timer
+    ExtMgr.DisableTimer();	 // disable internal timer
   }
@@
 The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}.
-The \texttt{ExtentCenter} maintains the records that map from extents to their hosting ENs.
-It is updated upon receiving the periodic sync reports from the ENs.
-Recall that the sync report from a particular EN lists all the extents stored at the EN.
-Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth.
-The \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.
+The \texttt{ExtentCenter} maps extents to their hosting ENs.
+It is updated upon receiving a periodic sync reports from an EN.
+Recall that a sync report from a particular EN lists all the extents stored at the EN.
+Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth.
+The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.
 
@@
 \textbf{Intercepting network messages}.
-The real ExtMgr uses a network engine to asynchronously send messages to the \texttt{ExtentNode} machines.
-The \psharp test harness models the original network engine in vNext by overriding its interface.
+The real ExtMgr uses a network engine to asynchronously send messages to ENs.
+The \psharp test harness models the original network engine in vNext by overriding the original implementation.
 The modeled network engine (see Figure~\ref{fig:enginecode}) intercepts all outbound messages from the ExtMgr, and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding ENs.
-This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}.
+This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}).
 
@@
 \vspace{-4mm}
-\caption{Modeled network engine in vNext.}
+\caption{Modeled vNext network engine.}
 \label{fig:enginecode}
@@
 
-Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine.
-They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
-The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.
+Messages coming from \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine; they are instead delivered to the \texttt{ExtentManager} machine and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}).
+The benefit of this approach is that the real ExtMgr can be tested without modifying the original communication-related code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.
 
@@
 \vspace{-4mm}
-\caption{Code snippet of the modeled EN in vNext.}
+\caption{Code snippet of the modeled EN.}
 \label{fig:modeled_en}
@@
 Hence, it makes sense to delegate all nondeterminism due to timing-related events to \psharp.
-To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp.
-%, an approach also used in previous work~\cite{desai2015building}.
+To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}.\footnote{We had to make a minor change to the real ExtMgr code to facilitate modeling: we added the \texttt{DisableTimer} method, which disables the real ExtMgr timer so that it can be replaced with the \psharp timer.}
 Similarly, \texttt{ExtentNode} machines do {\em not} have internal timers either. Their periodic heartbeats and sync reports are also driven by timers modeled in \psharp.
@@
 Once the proposed solution was implemented, the developers ran the test harness again.
-No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.
+No bugs were found during 100,000 executions, a process that only took on the order of tens of minutes.
 

### 2016-01-17 evaluation.tex
fix

@@
 \end{tabular}
-\caption{Statistics from modeling the environment of the three Microsoft MyCloud-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
+\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
 \label{tab:stats}
@@
 Although modeling Fabric required a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.
-Having modeled Fabric, we are now able to systematically test numerous services built on top of it.
 

### 2016-01-17 evaluation.tex
updates in evaluation

@@
 
-\setlength{\tabcolsep}{.72em}
+\setlength{\tabcolsep}{.68em}
 \begin{table*}[t]
@@
 \end{tabular}
-\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 executions.
-We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
+\caption{Results from running the \psharp random and priority-based systematic testing schedulers for 100,000 executions.
+We report: time in seconds to find a bug (Time to Bug); number of nondeterministic choices made when a bug was found (\#NDC); and if a bug was found with a particular scheduler (BF?).}
 \label{tab:testing}
@@
 Using \psharp we managed to uncover 8 serious bugs in our case studies.
-As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting.
-According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner.
-After the developers fixed all the discovered bugs, we added flags to allow them to be individually reintroduced, for purposes of evaluation.
+As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover and reproduce them in a small setting.
+According to the developers, the \psharp traces were useful, as it allowed them to understand the bugs and fix them in a timely manner.
+After all the discovered bugs were fixed, we added flags to allow them to be individually reintroduced, for purposes of evaluation.
 
-Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using the random and PCT schedulers.
-We chose to use controlled random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
+We have implemented two different schedulers that are responsible for choosing the next \psharp machine to execute in each scheduling point: a \emph{random} scheduler, which randomly chooses a machine from a list of enabled\footnote{A \psharp machine is considered enabled when it has at least one event in its queue waiting to be dequeued and handled.} machines; and a \emph{randomized priority-based}~\cite{burckhardt2010pct} scheduler, which always schedules the highest priority enabled machine (these priorities change at random points during execution, based on a random distribution).
+We decided to only use random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
+
+Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using our two randomized schedulers.
 The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
-We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:fabric}).
-
-\PDComment{TODO:}
-We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler.
+We do not include Fabric, as we are awaiting confirmation of the found bug (see \S\ref{sec:fabric}).
 
@@
 The random seed for both schedulers was generated using the current time.
-The PCT scheduler was further configured with a bug depth of 2 and a maximum number of scheduling steps of 500.
+The priority-based scheduler was further configured with a bug depth of 2 and a maximum number of nondeterministic choices of 500.
 All reported times are in seconds.
@@
 For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug within 10 seconds.
-The reason that the number of scheduling steps to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
-The PCT scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger bug depth bound to be found.
+The reason that the number of nondeterministic choices to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
+The priority-based scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger bug depth bound to be found.
 

### 2016-01-18 evaluation.tex
updates towards camera ready

@@
 \label{tab:stats}
-\vspace{-3mm}
+\vspace{-4mm}
 \end{table}
@@
 \caption{Results from running the \psharp random and priority-based systematic testing schedulers for 100,000 executions.
-We report: time in seconds to find a bug (Time to Bug); number of nondeterministic choices made when a bug was found (\#NDC); and if a bug was found with a particular scheduler (BF?).}
+We report: time in seconds to find a bug (Time to Bug); number of nondeterministic choices made in the first execution that a bug was found (\#NDC); and if a bug was found (BF?) or not (\cmark\xspace denotes that a bug was found, \smark\xspace denotes that a bug was found only using a custom test harness, and \xmark\xspace denotes that a bug could not be found).}
 \label{tab:testing}
+%\vspace{-2mm}
 \end{table*}
@@
 According to the developers, the \psharp traces were useful, as it allowed them to understand the bugs and fix them in a timely manner.
-After all the discovered bugs were fixed, we added flags to allow them to be individually reintroduced, for purposes of evaluation.
+After all the discovered bugs were fixed, we added flags to allow them to be individually re-introduced, for purposes of evaluation.
+
+Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug.
+The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
+We do not include the Fabric case study, as we are awaiting confirmation of the found bug (see \S\ref{sec:fabric}).
+We performed all experiments on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
+We configured the \psharp systematic testing engine to perform 100,000 executions.
+All reported times are in seconds.
 
@@
 We decided to only use random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
+The random seed for both schedulers was generated using the current time. The priority-based scheduler was further configured with a budget of 2 random machine priority change switches per execution.
 
-Table~\ref{tab:testing} presents the results from running the \psharp systematic testing engine on each case study with a re-introduced bug using our two randomized schedulers.
-The CS column shows which case study corresponds to each bug: ``1'' is for the Azure Storage vNext; and ``2'' is for MigratingTable.
-We do not include Fabric, as we are awaiting confirmation of the found bug (see \S\ref{sec:fabric}).
+For the vNext case study, both schedulers were able to reproduce the ExtentNodeLivenessViolation bug within 11 seconds.
+The reason that the number of nondeterministic choices made in the buggy execution is much higher than the rest of the bugs is that ExtentNodeLivenessViolation is a liveness bug: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
 
-We performed all experiments on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit.
-We configured the \psharp testing engine to perform 100,000 executions.
-The random seed for both schedulers was generated using the current time.
-The priority-based scheduler was further configured with a bug depth of 2 and a maximum number of nondeterministic choices of 500.
-All reported times are in seconds.
-
-For the vNext case study, the random scheduler was able to reproduce the ExtentNodeLivenessViolation bug within 10 seconds.
-The reason that the number of nondeterministic choices to find the bug is much higher than the rest of the bugs in the table is that this bug is a liveness violation: as discussed in \S\ref{sec:overview:liveness} we leave the program to run for a long time before checking if the liveness property holds.
-The priority-based scheduler was unable to find the bug using the bug depth of 2, which suggests that the bug requires a larger bug depth bound to be found.
-
-For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect.
-The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
+For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven bugs, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect.
+The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
 The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes).
-We believe this is because the inputs and schedules that trigger them are too rare in our distribution.
+We believe this is because the inputs and schedules that trigger them are too rare in the used distribution.
 To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
@@
 the table shows the results of these runs.
-This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.
+Note that the random scheduler managed to only find seven of the MigratingTable bugs; we had to use the priority-based scheduler to discover the remaining bugs.
 

### 2016-01-18 intro.tex
updates towards camera ready

@@
 The gist of their approach is to extract the high-level logic from a production system, represent this logic as specifications in the expressive TLA+~\cite{lamport1994temporal} language, and finally verify the specifications using a model checker.
-While highly effective, as demonstrated by its use in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'', and the AWS team admits that it is ``not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
+While highly effective, as demonstrated by its use in AWS, this approach falls short of ``verifying that executable code correctly implements the high-level specification'''~\cite{newcombe2015aws}, and the AWS team admits that it is ``not aware of any such tools that can handle distributed systems as large and complex as those being built at Amazon''~\cite{newcombe2015aws}.
 

### 2016-01-20 paper.tex
added comment
Mark me as affiliated with Microsoft since most of my work was done there.
updates
updates towards camera ready
State that part of the work was done while interning at Microsoft.
The paper is now over length.  If necessary, I can remove material
later to fix that.

@@
 \newcommand{\xmark}{\ding{55}}
+\newcommand{\smark}{\ding{72}}
 
@@
 %\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
-%\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
+\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
 %\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
@@
 %{\rm Submission ID: \#127}} % end author
-{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
+{\rm Pantazis Deligiannis$^{\dagger1}$, Matt McCutchen$^{\diamond1}$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
 {\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
@@
 } % end author
-
 \maketitle
@@
 \thispagestyle{empty}
+
+\stepcounter{footnote}
+\footnotetext{Part of the work of these authors was done while they interned at Microsoft.}
 

### 2016-01-22 evaluation.tex
- Act on Daniel Jackson's feedback on my sections. - Some other edits to evaluation.   - Move "bug found" column before "time" and "NDC" because it     affects their interpretation, i.e., whether they reflect the     standard input generator or a custom test case.   - Now that we have stars in the "bug found" column (great!), the     diamonds on the bugs are redundant, so remove them.   - Replace star with diamond because diamond is easier to     distinguish from check and X.   - Reword caption (see included comment).

@@
 \caption{Results from running the \psharp random and priority-based systematic testing schedulers for 100,000 executions.
-We report: time in seconds to find a bug (Time to Bug); number of nondeterministic choices made in the first execution that a bug was found (\#NDC); and if a bug was found (BF?) or not (\cmark\xspace denotes that a bug was found, \smark\xspace denotes that a bug was found only using a custom test harness, and \xmark\xspace denotes that a bug could not be found).}
+We report: whether the bug was found (BF?) (\cmark\xspace means it was found, $\diamond$\xspace means it was found only using a custom test case, and \xmark\xspace means it was not found); time in seconds to find the bug (Time to Bug); and number of nondeterministic choices made in the first execution that found the bug (\#NDC).}
+% Changing ``a bug'' to ``the bug'' makes this a lot easier to read.  The risk
+% of this change is that we're claiming that all failures of the P# test in each
+% row represent the listed bug, rather than just ``a bug''.  I went to enough
+% trouble to verify this for MigratingTable that I'm comfortable making this
+% claim, although I don't have complete proof.  In any case, I don't think this
+% is a distinction we can reasonably hope to get across to the reader under our
+% space constraint. ~ Matt 2016-01-22
 \label{tab:testing}
@@
 The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
-The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes).
+The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes).
 We believe this is because the inputs and schedules that trigger them are too rare in the used distribution.
@@
 the table shows the results of these runs.
-Note that the random scheduler managed to only find seven of the MigratingTable bugs; we had to use the priority-based scheduler to discover the remaining bugs.
+Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining bugs.
+% The words ``find'' and ``discover'' are misleading for the diamond bugs.
+% ~ Matt 2016-01-21
 
-The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve.
+The QueryStreamedBackUpNewStream bug in MigratingTable, which was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve.
 This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key.
@@
 But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed.
-\psharp managed to discover this bug in a matter of seconds.
-The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable.
-\psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend, and we extended trace information in a custom fashion for each of our case studies.
+The \psharp test discovered this bug in a matter of seconds.
+The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces.
+(Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.)
 

### 2016-01-22 migrating.tex
updates towards camera ready
- Act on Daniel Jackson's feedback on my sections. - Some other edits to evaluation.   - Move "bug found" column before "time" and "NDC" because it     affects their interpretation, i.e., whether they reflect the     standard input generator or a custom test case.   - Now that we have stars in the "bug found" column (great!), the     diamonds on the bugs are redundant, so remove them.   - Replace star with diamond because diamond is easier to     distinguish from check and X.   - Reword caption (see included comment).

@@
 % ~ Matt 2015-12-31
-Live Table Migration (MigratingTable) is a library designed to transparently migrate a data set between tables in the Azure storage service while an application is accessing this data set.
-This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness.
-This co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}).
-The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on a single liveness property, while the MigratingTable test harness checks complete compliance with an interface (safety) specification.
+%
+% Delete ``safety'': the MigratingTable harness does have a crude liveness
+% check. ~ Matt 2016-01-20
+Live Table Migration (MigratingTable) is a library designed to transparently migrate a key-value data set between two Microsoft Azure tables~\cite{azure-table} (called the \emph{old} table and the \emph{new} table, or collectively the \emph{backend} tables) while an application is accessing this data set.
+The MigratingTable testing effort differs from the vNext effort in two significant ways: the \psharp test harness was developed along with the system rather than later, and it checks complete compliance with an interface specification rather than just a single liveness property.
+Indeed, the \psharp test caught bugs throughout the development process (see \S\ref{sec:eval}), and the developers did not bother with any other correctness tests.
 
-MigratingTable provides a \emph{virtual table} (VT) that has a similar interface
-to an ordinary Azure table. This interface is
-named \texttt{IChainTable}. Each VT is backed by a pair
-of \emph{old} and \emph{new} \emph{backend} tables (BTs), both of which
-implement \texttt{IChainTable}. A \emph{migrator} job is responsible
-for moving all data from the old table to the new table in the
-background. Meanwhile, each \emph{input} read and write operation
-issued to the VT is implemented via a sequence
-of reads and writes on the BTs according to a protocol,
-which intends to guarantee that when multiple applications issue input
-operations to different MigratingTable instances with the same BTs, the output of these operations complies with the specification
-of \texttt{IChainTable} for the combined \emph{input history}, as if
-all the operations were performed on a single table. The goal of
-using \psharp was to systematically test this property.
+During a migration, each application process creates its own MigratingTable instance (MT) referring to the same backend tables (BTs) and performs all data access via the MT.
+(The MT provides an interface named \texttt{IChainTable} similar to that of an Azure table, and it assumes the BTs provide the same interface via an adapter.) A \emph{migrator} job moves the data in the background.
+In the meantime, each \emph{logical} read or write operation issued to an MT is implemented via a sequence of \emph{backend} operations on the BTs according to a custom protocol.
+The protocol is designed to guarantee that the output of the logical operations complies with the \texttt{IChainTable} specification as if all the operations were performed on a single \emph{virtual table} (VT).
+The goal of using \psharp was to systematically test this property.
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{img/modeled_migration}
-\caption{The \psharp test harness of MigratingTable (each box with a dotted line represents one \psharp machine).}
-\label{fig:mockedmigration}
-\end{figure}
-
-There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the input operations accept many parameters that affect the behavior in different ways.
-The developers could have chosen specific input histories to test, but they were not confident that they would cover the combinations of parameters that might trigger bugs.
-Instead, the developers used the \psharp nondeterministic choice generation APIs to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits.
-They created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, and compared the output of MigratingTable on the chosen input history to that of a \texttt{SpecTable} instance called the \emph{reference table}.
-The \texttt{SpecTable} implementation was reused for the BTs of MigratingTable.
+There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the logical operations accept many parameters that affect the behavior in different ways.
+The developers could have chosen specific input sequences, but they were not confident that these sequences would cover the combinations of parameters that might trigger bugs.
+Instead, they used the \psharp nondeterministic choice API to choose all of the parameters independently within certain limits.
+They issued the same operations to the MTs and to a \emph{reference table} (RT) running a reference implementation of the \texttt{IChainTable} specification, and compared the output.
+This reference implementation was reused for the BTs, since the goal was not to test the real Azure tables.
 % I'm removing the claim that each bug is caught with positive probability per
@@
 
-Figure~\ref{fig:mockedmigration} shows the \psharp test harness of MigratingTable.
-The test harness consists of a \texttt{Tables} machine, which contains the BTs and the reference table, and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration.
-Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the BTs via \psharp events.
-The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single VT, so the test harness could perform the operation on the reference table at the same time.
-More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table.
-This way, the rest of the system never observes the reference table to be out of sync with the VT.
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{img/modeled_migration}
+\vspace{-6mm}
+\caption{The \psharp test environment of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\label{fig:mockedmigration}
+\vspace{-2mm}
+\end{figure}
+
+The complete test environment is shown in Figure~\ref{fig:mockedmigration}.
+It consists of a \texttt{Tables} machine, which contains the BTs and RT, and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured MTs; and a \texttt{Migrator} machine that performs the background migration.
+Each \texttt{Service} machine issues a random sequence of logical operations to its MT, which performs the backend operations on the BTs via \psharp events.
+The developers instrumented the MTs to report the \emph{linearization point} of each logical operation, i.e., the time at which it takes effect on the VT, so the test harness could perform the operation on the RT at the same time.
+(More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until the MT reports whether the backend operation was the linearization point and, if so, the logical operation has been completed on the RT.
+This way, the rest of the system never observes the RT to be out of sync with the VT.)
+% In my reading, the parentheses make it clear that we're amending the previous
+% sentence; without them, it looks more like a contradiction. ~ Matt 2016-01-21
 % Note: If the tables were observed to be out of sync, it could cause both false
 % positives and false negatives. ~ Matt 2015-12-31
-
-For further information about this case study, we refer the reader to the source code repository~\cite{migratingtable-src}.
+% Join paragraphs to save space
+For further implementation details of MigratingTable and its test harness, we refer the reader to the source code repository~\cite{migratingtable-src}.
 

### 2016-01-22 evaluation.tex
update

@@
 We report: whether the bug was found (BF?) (\cmark\xspace means it was found, $\diamond$\xspace means it was found only using a custom test case, and \xmark\xspace means it was not found); time in seconds to find the bug (Time to Bug); and number of nondeterministic choices made in the first execution that found the bug (\#NDC).}
-% Changing ``a bug'' to ``the bug'' makes this a lot easier to read.  The risk
-% of this change is that we're claiming that all failures of the P# test in each
-% row represent the listed bug, rather than just ``a bug''.  I went to enough
-% trouble to verify this for MigratingTable that I'm comfortable making this
-% claim, although I don't have complete proof.  In any case, I don't think this
-% is a distinction we can reasonably hope to get across to the reader under our
-% space constraint. ~ Matt 2016-01-22
 \label{tab:testing}
@@
 We believe this is because the inputs and schedules that trigger them are too rare in the used distribution.
-To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
-% Be a little bit more transparent about the fact that the $\diamond$ rows do not
-% reflect results achievable via our main testing approach. ~ Matt 2015-12-31
-the table shows the results of these runs.
-Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining bugs.
-% The words ``find'' and ``discover'' are misleading for the diamond bugs.
-% ~ Matt 2016-01-21
+To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs; the table shows the results of these runs.
+Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining four bugs.
 
@@
 The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces.
-(Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.)
+Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.
 

### 2016-01-22 migrating.tex
We may as well refer to the Nondet function by name in the MigratingTable section since it's explained before.
update

@@
-% Remove the \emph on ``transparently'' and ``while'' again.  Pantazis, I'm
-% imagining you were trying to emphasize what's interesting about the problem,
-% but I think suggesting that this idea is new to the reader is insulting.
-% ~ Matt 2015-12-31
-%
-% Delete ``safety'': the MigratingTable harness does have a crude liveness
-% check. ~ Matt 2016-01-20
 Live Table Migration (MigratingTable) is a library designed to transparently migrate a key-value data set between two Microsoft Azure tables~\cite{azure-table} (called the \emph{old} table and the \emph{new} table, or collectively the \emph{backend} tables) while an application is accessing this data set.
-The MigratingTable testing effort differs from the vNext effort in two significant ways: the \psharp test harness was developed along with the system rather than later, and it checks complete compliance with an interface specification rather than just a single liveness property.
-Indeed, the \psharp test caught bugs throughout the development process (see \S\ref{sec:eval}), and the developers did not bother with any other correctness tests.
+The MigratingTable testing effort differs from the vNext effort in two significant ways: the \psharp test harness was developed along with the system rather than later; and it checks complete compliance with an interface specification rather than just a single liveness property.
+Indeed, the \psharp systematic testing engine caught bugs throughout the development process (see \S\ref{sec:eval}).
 
-During a migration, each application process creates its own MigratingTable instance (MT) referring to the same backend tables (BTs) and performs all data access via the MT.
-(The MT provides an interface named \texttt{IChainTable} similar to that of an Azure table, and it assumes the BTs provide the same interface via an adapter.) A \emph{migrator} job moves the data in the background.
-In the meantime, each \emph{logical} read or write operation issued to an MT is implemented via a sequence of \emph{backend} operations on the BTs according to a custom protocol.
-The protocol is designed to guarantee that the output of the logical operations complies with the \texttt{IChainTable} specification as if all the operations were performed on a single \emph{virtual table} (VT).
+During migration, each application creates its own instance of MigratingTable (MT) that refers to the same backend tables (BTs).
+The application performs all data access via the MT, which provides an interface named \texttt{IChainTable} similar to that of an Azure table (the MT assumes that the BTs provide the same interface via an adapter).
+A \emph{migrator} job moves the data in the background.
+In the meantime, each \emph{logical} read and write operation issued to an MT is implemented via a sequence of \emph{backend} operations on the BTs according to a custom protocol.
+The protocol is designed to guarantee that the output of the logical operations complies with the \texttt{IChainTable} specification, as if all the operations were performed on a single \emph{virtual table} (VT).
 The goal of using \psharp was to systematically test this property.
@@
 The developers could have chosen specific input sequences, but they were not confident that these sequences would cover the combinations of parameters that might trigger bugs.
-Instead, they used the \psharp nondeterministic choice API to choose all of the parameters independently within certain limits.
+Instead, they used the \psharp \texttt{Nondet()} function to choose all of the parameters independently within certain limits.
 They issued the same operations to the MTs and to a \emph{reference table} (RT) running a reference implementation of the \texttt{IChainTable} specification, and compared the output.
 This reference implementation was reused for the BTs, since the goal was not to test the real Azure tables.
-% I'm removing the claim that each bug is caught with positive probability per
-% run because it's too awkward to explain here and it's nothing specific to
-% MigratingTable.  Also, we can't quite appeal to the small scope hypothesis in
-% its original form because I hand-picked which parameters to vary and what
-% limits to use.  The value of the input generator should be clear enough from
-% the above. ~ Matt 2015-12-31
 
@@
 \vspace{-6mm}
-\caption{The \psharp test environment of MigratingTable (each box with a dotted line represents one \psharp machine).}
+\caption{The test environment of MigratingTable (each box with a dotted line represents one \psharp machine).}
 \label{fig:mockedmigration}
@@
 (More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until the MT reports whether the backend operation was the linearization point and, if so, the logical operation has been completed on the RT.
-This way, the rest of the system never observes the RT to be out of sync with the VT.)
-% In my reading, the parentheses make it clear that we're amending the previous
-% sentence; without them, it looks more like a contradiction. ~ Matt 2016-01-21
-% Note: If the tables were observed to be out of sync, it could cause both false
-% positives and false negatives. ~ Matt 2015-12-31
-% Join paragraphs to save space
-For further implementation details of MigratingTable and its test harness, we refer the reader to the source code repository~\cite{migratingtable-src}.
+This way, the rest of the system never observes the RT to be out of sync with the VT.) For further implementation details of MigratingTable and its test harness, we refer the reader to the source code repository~\cite{migratingtable-src}.
 

### 2016-01-22 overview.tex
update

@@
-The goal of our work is to find bugs in distributed system \emph{before} they reach production.
+The goal of our work is to find bugs in distributed systems \emph{before} they reach production.
 Typical distributed systems consist of multiple components that interact with each other via \emph{message passing}.
@@
 
-Due to the highly asynchronous nature of distributed storage systems, the number of possible states that these systems can reach is exponentially large.
+Due to the highly asynchronous nature of distributed systems, the number of possible states that these systems can reach is exponentially large.
 Tools such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug} focus on testing unmodified distributed systems, but this can easily lead to state-space explosion when trying to exhaustively explore the entire state-space of a production-scale distributed storage system, such as the Azure Storage vNext.
-On the other hand, techniques such as TLA+~\cite{lamport1994temporal} have successfully verified specifications of distributed systems, but they are unable to verify the actual implementation.
+On the other hand, techniques such as TLA+~\cite{lamport1994temporal} have been successfully used in industry to verify specifications of complex distributed systems~\cite{newcombe2015aws}, but they are unable to verify the actual implementation.
 
@@
 The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).
-
-In principle, our modeling and testing methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
+In principle, our methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
 

### 2016-01-22 paper.tex
update

--- 2015-12-25 paper.tex
@@
 %\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
-\newcommand{\PTComment}[1]{\textcolor{orange}{Paul: #1}}
+%\newcommand{\PTComment}[1]{\textcolor{orange}{Paul: #1}}
 
@@
 %{\rm Submission ID: \#127}} % end author
-{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
+{\rm Pantazis Deligiannis$^{\dagger1}$, Matt McCutchen$^{\diamond1}$, Paul Thomson$^{\dagger1}$, Shuo Chen$^\star$}\\
 {\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
@@
 \thispagestyle{empty}
+
+\stepcounter{footnote}
+\footnotetext{Part of the work was done while interning at Microsoft.}
 
@@
 
-\vspace{-1mm}
+%\vspace{-4mm}
 \section{Introduction}
 \label{sec:intro}
-\vspace{-1mm}
+%\vspace{-1mm}
 

### 2016-01-22 migrating.tex
Revert a few small changes to the MigratingTable section.
See the added comments.

@@
 The MigratingTable testing effort differs from the vNext effort in two significant ways: the \psharp test harness was developed along with the system rather than later; and it checks complete compliance with an interface specification rather than just a single liveness property.
-Indeed, the \psharp systematic testing engine caught bugs throughout the development process (see \S\ref{sec:eval}).
+Indeed, the \psharp test caught bugs throughout the development process (see \S\ref{sec:eval}).
+% In this context, saying ``the \psharp systematic testing engine caught bugs''
+% sounded to me like it caught them by itself without a test harness.  In other
+% contexts, this wording may not be confusing. ~ Matt 2016-01-22
 
-During migration, each application creates its own instance of MigratingTable (MT) that refers to the same backend tables (BTs).
+% The first paragraph says the data set belongs to one application
+% The way I'm describing this, the data set belongs to one (logical)
+% application, which may consist of multiple processes (e.g., replicas of a web
+% service).  If we wanted to talk about multiple applications sharing a data
+% set, we'd need to introduce that earlier. ~ Matt 2016-01-22
+%
+% I want ``MT'' to refer to an instance.  With ``instance of MigratingTable
+% (MT)'', it's ambiguous whether MT refers to the instance or the library.
+% ~ Matt 2016-01-22
+During migration, each application process creates its own MigratingTable instance (MT) that refers to the same backend tables (BTs).
 The application performs all data access via the MT, which provides an interface named \texttt{IChainTable} similar to that of an Azure table (the MT assumes that the BTs provide the same interface via an adapter).

### 2016-01-23 evaluation.tex
update
updates

@@
 \label{tab:stats}
-\vspace{-4mm}
+\vspace{-2mm}
 \end{table}
@@
 The harness was developed in parallel with the actual system.
-This differs from the other two case studies, where the modeling activity occurred independently and after the development process had finished.
+This differs from the other two case studies, where the modeling activity occurred independently and after the development process.
 
@@
 
-We have implemented two different schedulers that are responsible for choosing the next \psharp machine to execute in each scheduling point: a \emph{random} scheduler, which randomly chooses a machine from a list of enabled\footnote{A \psharp machine is considered enabled when it has at least one event in its queue waiting to be dequeued and handled.} machines; and a \emph{randomized priority-based}~\cite{burckhardt2010pct} scheduler, which always schedules the highest priority enabled machine (these priorities change at random points during execution, based on a random distribution).
-We decided to only use random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
-The random seed for both schedulers was generated using the current time. The priority-based scheduler was further configured with a budget of 2 random machine priority change switches per execution.
+We implemented two different schedulers that are responsible for choosing the next \psharp machine to execute in each scheduling point: a \emph{random} scheduler, which randomly chooses a machine from a list of enabled\footnote{A \psharp machine is considered enabled when it has at least one event in its queue waiting to be dequeued and handled.} machines; and a \emph{randomized priority-based}~\cite{burckhardt2010pct} scheduler, which always schedules the highest priority enabled machine (these priorities change at random points during execution, based on a random distribution).
+We decided to use these two schedulers, because random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
+The random seed for the schedulers was generated using the current system time.
+The priority-based scheduler was configured with a budget of 2 random machine priority change switches per execution.
 
@@
 The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}.
-The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes).
+The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions.
 We believe this is because the inputs and schedules that trigger them are too rare in the used distribution.
@@
 But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed.
-The \psharp test discovered this bug in a matter of seconds.
+
+The \psharp test discovered the above bug in a matter of seconds.
 The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces.

### 2016-01-23 intro.tex
updates

@@
 \item We present a new methodology for modeling, specifying properties of correctness, and systematically testing real distributed systems with \psharp.
-\item We discuss our experience of using \psharp to test three distributed storage systems that are used inside Microsoft, finding subtle bugs that could not be found with traditional testing techniques.
+\item We discuss our experience of using \psharp to test three distributed storage systems built on top of Microsoft Azure, finding subtle bugs that could not be found with traditional testing techniques.
 \item We evaluate the cost and benefits of using our approach, and show that \psharp can detect bugs in a small setting and with easy to understand traces.

### 2016-01-23 overview.tex
updates

@@
 In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment.
-The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space.
+The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, and achieve this in a much reduced state-space.
 Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp.
-However, we argue that developers already spend significant time in building test suites for distributed systems prior to deployment.
+However, developers already spend significant time in building test suites for distributed systems prior to deployment.
 The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).
@@
 \psharp uses object-oriented language features such as interfaces and virtual method dispatch to connect the real code with the modeled code.
-Programmers in industry are used to working with such features, and heavily employ them in production for testing purposes.
+Developers in industry are used to working with such features, and heavily employ them in testing production systems.
 In our experience, this significantly lowers the bar for engineering teams inside Microsoft to embrace \psharp for testing.
@@
 In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct.
-The interested reader can refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}), and read \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric} to learn more details about how \psharp was used to model and test real distributed storage systems in Microsoft.
+The details of how \psharp was used to model and test real distributed storage systems in Microsoft are covered in \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric}.
+Interested readers can also refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}).
 
@@
 Typically, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977, lamport1994temporal}.
-We take a different approach and allow the programmer to write a \emph{liveness monitor}~\cite{desai2015building}.
+We take a different approach and allow the developers to write a \emph{liveness monitor}~\cite{desai2015building}.
 Similar to a safety monitor, a liveness monitor can receive, but not send, events.
@@
 A liveness monitor contains two special types of states: \emph{hot} and \emph{cold}.
-A hot state denotes a point in the execution where progress is required but has not happened yet; e.g.\ a node has failed but a new one has not come up yet.
+A hot state denotes a point in the execution where progress is required but has not happened yet; e.g.\ a node has failed but a new one has not launched yet.
 A liveness monitor enters a hot state when it is notified that the system must make progress.
-The liveness monitor leaves the hot state and enters a cold state when it is notified that the system has progressed.
-An infinite execution is erroneous if the liveness monitor is in a hot state for an infinitely long time.
+The liveness monitor leaves the hot state and enters the cold state when it is notified that the system has progressed.
+An infinite execution is erroneous if the liveness monitor is in the hot state for an infinitely long period of time.
 Our liveness monitors can encode arbitrary temporal logic properties.
@@
 A liveness violation is witnessed by an \emph{infinite} execution in which all concurrently executing \psharp machines are \emph{fairly} scheduled.
-Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time.
-Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics.
+Since it is impossible to generate an infinite execution by executing a program for a finite amount of time, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics.
 In this work, we consider an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}.
-Note that fairness is not relevant when using this heuristic, due to our pragmatic use of a large bound instead of proper handling of infinite executions.
+Note that checking for fairness is not relevant when using this heuristic, due to our pragmatic use of a large bound.
 

### 2016-01-23 paper.tex
updates

@@
 We thank our shepherd Haryadi Gunawi for his valuable guidance, and the reviewers for their constructive comments.
-We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions on various stages of this work.
-This research was partially supported by a gift from Intel Corporation.
+We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions at the various stages of this work.
+We acknowledge that this research was partially supported by a gift from Intel Corporation.
 

### 2016-01-23 vnext.tex
updates towards camera ready
updates

@@
 Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data.
-It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.
+It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored exceeding 60 trillion~\cite{greenberg2015keynote}.
 
-Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
+Azure Storage vNext is the next generation storage system currently being developed for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$.
 Similar to the current system, vNext employs containers, called \emph{extents}, to store data.
@@
 However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}.
-In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr).
+In vNext, extents are divided into partitions, with each partition managed by a lightweight \emph{Extent Manager} (ExtMgr).
 This partitioning is illustrated in Figure~\ref{fig:vnext}.
@@
 EN failure is detected by missing heartbeats.
-An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN that it manages.
+An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN.
 The sync reports list all the extents (and associated metadata) stored on the EN.
 Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result.
-The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs.
+The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the remaining ENs.
 The ENs then repair the extents from the existing replicas and lazily update the ExtMgr via their next periodic sync reports.
-All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
+All the communications between an ExtMgr and the ENs occur via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).
 
@@
 
-\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents.
-The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
+\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and many extents.
+The test keeps repeating the following process: injects an EN failure, launches a new EN and checks that the affected extents are eventually repaired.
 \end{enumerate}
@@
 \noindent
-Despite of the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic.
+Despite the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic.
 All the unit test suites and integration test suites successfully passed on each test run.
@@
 \end{lstlisting}
-\vspace{-2mm}
+\vspace{-3mm}
 \caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
 \label{fig:wrap_target}
+\vspace{-1mm}
 \end{figure}
@@
 \label{fig:extentmanager}
+\vspace{-2mm}
 \end{figure}
@@
 The \texttt{ExtentCenter} maps extents to their hosting ENs.
-It is updated upon receiving a periodic sync reports from an EN.
+It is updated upon receiving a periodic sync report from an EN.
 Recall that a sync report from a particular EN lists all the extents stored at the EN.
@@
 \label{fig:enginecode}
-\vspace{-3mm}
+\vspace{-2mm}
 \end{figure}
@@
 \end{lstlisting}
-\vspace{-4mm}
+\vspace{-5mm}
 \caption{Code snippet of the modeled EN.}
 \label{fig:modeled_en}
-\vspace{-2mm}
+\vspace{-4mm}
 \end{figure}
@@
 \end{lstlisting}
-\vspace{-4mm}
+\vspace{-5mm}
 \caption{Modeling timer expiration using \psharp.}
@@
 \label{fig:testing_driver}
-\vspace{-2mm}
+\vspace{-4mm}
 \end{figure}
@@
 \end{lstlisting}
-\vspace{-4mm}
+\vspace{-5mm}
 \caption{The RepairMonitor liveness monitor.}
@@
 Once the proposed solution was implemented, the developers ran the test harness again.
-No bugs were found during 100,000 executions, a process that only took on the order of tens of minutes.
+No bugs were found during 100,000 executions, a process that only took a few minutes.
 
@@
 This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr would still think that all replicas are healthy.
-In production, such a bug can cause a very serious incident of customer data loss.
+If released to production, this bug would have caused a very serious incident of customer data unavailability.
 

### 2016-01-24 paper.tex
updates

@@
 
+\vspace{-1mm}
 \section{Case Study: Azure Storage vNext}
 \label{sec:vnext}
+\vspace{-1mm}
 
@@
 
+\vspace{-1mm}
 \section{Conclusion}
@@
 
+\vspace{-1mm}
 \section{Acknowledgments}
 
-We thank our shepherd Haryadi Gunawi for his valuable guidance, and the reviewers for their constructive comments.
-We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions at the various stages of this work.
+We thank our shepherd Haryadi Gunawi for his valuable guidance that significantly improved the paper, and the anonymous reviewers for their constructive comments.
+We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions at various stages of this work.
 We acknowledge that this research was partially supported by a gift from Intel Corporation.

### 2016-01-24 related.tex
updates

@@
 Formal methods have been successfully used in industry to verify the correctness of distributed protocols.
-A recent notable example is the use of TLA+~\cite{lamport1994temporal} by the Amazon Web Services team~\cite{newcombe2015aws}.
+A notable example is the use of TLA+~\cite{lamport1994temporal} by the Amazon Web Services team~\cite{newcombe2015aws}.
 TLA+ is an expressive formal specification language that can be used to design and verify concurrent programs via model checking.
@@
 
-Another formal approach is Verdi~\cite{wilcox2015verdi}, a framework for writing and verifying distributed systems in Coq~\cite{barras1997coq}.
-Verdi generates OCaml code from the verified Coq system, which can be then used for execution.
-In contrast, \psharp performs bounded testing on a system already written in \csharp, which in our experience lowers the bar for adoption by engineering teams.
+More recent formal approaches include the Verdi~\cite{wilcox2015verdi} and IronFleet~\cite{hawblitzel2015ironfleet} frameworks.
+In Verdi, developers can write and verify distributed systems in Coq~\cite{barras1997coq}.
+After the system has been successfully verified, Verdi translates the Coq code to OCaml, which can be then compiled for execution.
 Verdi does not currently support detecting liveness property violations, an important class of bugs in distributed systems.
+In IronFleet, developers can build a distributed system using the Dafny~\cite{leino2010dafny} language and program verifier.
+Dafny verifies system correctness using the Z3~\cite{de2008z3} SMT solver, and finally compiles the verified system to a .NET executable.
+In contrast, \psharp performs bounded testing on a system already written in .NET, which in our experience lowers the bar for adoption by engineering teams.

### 2016-01-24 vnext.tex
updates

@@
 \includegraphics[width=\linewidth]{img/azurestore}
-\vspace{-5mm}
+\vspace{-6mm}
 \caption{Top-level components for extent management in Microsoft Azure Storage vNext.}
@@
 Finally, by the time that the bug did manifest, very long execution traces had been collected, which made manual inspection tedious and ineffective.
-%To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.
 
@@
 \includegraphics[width=\linewidth]{img/modeled_vnext}
-\vspace{-5mm}
+\vspace{-6mm}
 \caption{Real Extent Manager with its \psharp test harness (each box represents one \psharp machine).}
@@
 \label{fig:wrap_target}
-\vspace{-1mm}
+\vspace{-3mm}
 \end{figure}
@@
 \includegraphics[width=.95\linewidth]{img/extent_manager}
-\vspace{-2mm}
+\vspace{-3mm}
 \caption{Internal components of the real Extent Manager in Microsoft Azure Storage vNext.}
 \label{fig:extentmanager}
-\vspace{-2mm}
+\vspace{-3mm}
 \end{figure}
@@
 \end{lstlisting}
-\vspace{-4mm}
+\vspace{-5mm}
 \caption{Modeled vNext network engine.}
 \label{fig:enginecode}
-\vspace{-2mm}
+\vspace{-3mm}
 \end{figure}
@@
     var node = (Machine)PSharp.Nondet(ExtentNodes);
-    PSharp.Send(node, new FailureEvent()); // fail chosen EN
+    // fail chosen EN
+    PSharp.Send(node, new FailureEvent());
   }
@@
 \end{lstlisting}
-\vspace{-4mm}
+\vspace{-3mm}
 \caption{Code snippet of the TestingDriver machine.}
 \label{fig:testing_driver}
-\vspace{-4mm}
+\vspace{-1mm}
 \end{figure}
@@
 \end{lstlisting}
-\vspace{-5mm}
+\vspace{-4mm}
 \caption{The RepairMonitor liveness monitor.}

### 2016-01-24 overview.tex
updates

@@
 
-In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment.
-The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, and achieve this in a much reduced state-space.
-Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp.
+In this work, we are proposing a solution between the above two extremes: test the real implementation of one or more components against a modeled in \psharp environment.
+The benefit of our approach is that it can detect bugs in the \emph{actual} implementation by exploring a much reduced state-space.
+Note that testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp.
 However, developers already spend significant time in building test suites for distributed systems prior to deployment.
 The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).
-In principle, our methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
+In principle, our methodology is \emph{not} specific to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.
 
@@
 \psharp machines run concurrently with each other, each executing an event handling loop that dequeues the next event from the queue and handles it by invoking the registered action.
-An action might transition the machine to a new state, create a new machine, send an event to another machine, access a field or call a method.
+An action might transition the machine to a new state, create a new machine, send an event to a machine, access a field or call a method.
 In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the queue of the target machine, which will dequeue and handle the event concurrently.
@@
 We model the SNs so that they store data in memory rather than on disk (which could be inefficient during testing).
-We also model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}.
+We also model the client so that it can drive the system by repeatedly sending a nondeterministically generated \texttt{ClientReq}, and then waiting for an \texttt{Ack} message.
 Finally, we model the timer so that \psharp takes control of all time-related nondeterminism in the system.
@@
 
-In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct.
+In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example system is correct.
 The details of how \psharp was used to model and test real distributed storage systems in Microsoft are covered in \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric}.

### 2016-01-24 vnext.tex
updates

@@
 Recall that a sync report from a particular EN lists all the extents stored at the EN.
-Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth.
+Its purpose is to update the ExtMgr's possibly out-of-date view of the EN with the ground truth.
 The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.
@@
 First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system.
-Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).
+Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a nondeterministic fashion, in case emulating message loss is desirable (not shown in this example).
 

### 2016-01-25 vnext.tex
added new reference

@@
 Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs).
-However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}.
+However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs~\cite{calder2011windows}, vNext achieves scalability by using a \emph{distributed mapping}.
 In vNext, extents are divided into partitions, with each partition managed by a lightweight \emph{Extent Manager} (ExtMgr).
